### 2022-02-07 src/ms.tex
trying to merge files

% Define document class
\documentclass[twocolumn]{aastex631}

\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
% Begin!
\begin{document}

\title{Backprogating Gravitational wave events}

\pacs{}

\author{Kaze W. K. Wong} 
\email{kwong@flatironinstitute.org}
%\flatiron

\date{\today}

\begin{abstract}
In this
\end{abstract}

\maketitle

\section{Introduction}

\begin{enumerate}
\item We have more and more event
\item Problem with population synthesis -> Need for developing algorithm to explore the high dimensional space
\item We run the event backward in time
\end{enumerate}

While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
it is not necessarily the most efficient way to constraint the underlying physical model.
Forward modelling of a binary system start with a set of initial binary parameters,
then evolve it until some stopping criteria is met,
typically either a merger event or the binary system stays stable for a Hubble time.
The problem is, most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
This means if we want to evolve a binary to match the observation by pure trial and error, a large fraction of computation will be wasted in trying initial binary configurations that are not related to the observation.
To alleviate this inefficiency, the community often takes a population approach --- instead of trying to match binary systems to observations event by event,
one can draw systems from an initial population, forward model each member in that population then match it to the observed population properties.
However, this approach is essentially trading one inefficiency with another, since we still need to choose an initial distribution of binary parameters.
For relatively well-known source properties such as initial masses this is less of a problem,
but the distribution of some less well-studied properties such as common envelope efficiency is often heuristically chosen to be some simple distribution such as a delta function at some value, i.e. a fixed value for the entire population.
This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution,
or they should vary the family of distributions.

In this paper, instead of forward modelling the population of binary star to match observed GW events population,
we propose a method to "backward" model each GW event to its progenitor state.

This paper introduce a number of new perspectives in the study of binary evolution physics and gravitational-wave progenitors:
\begin{enumerate}
\item Instead of starting from a distribution 
\item Using derivative-based optimization algorithm, we are allowed to explore a moderate size parameter space 
\end{enumerate}

\section{Method}

We define the parameters that describe the progenitor's properties such as ZAMS masses and eccentricity as progenitor parameters $\bm{\theta'}$,
the parameters that control different physical prescriptions such as wind strength and common envelope efficiency as hyper-parameters $\bm{\lambda}$,
and random variables that affect certain stochastic process such as whether natal kick unbinds a binary as $\bm{X}$.
To avoid clutter in the following derivation,
we denote all the parameters related to mapping a particular progenitor system into a GW event collectively as evolutionary parameters $\bm{\Theta}$,
which includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

Once all the parameters including a particular draw of all the random variables are known,
the population synthesis code is a deterministic function that transforms the properties of the progenitors into the GW parameters.
Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\theta}'$ is given by

\begin{align}
    p(\bm{\theta}|\bm{\Theta}) &= \bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta}),
\end{align}

where $\bm{f}$ represents the population synthesis code.

While it is straightforward to write down the "forward" probability, the "backward" probability is more complicated since multiple evolutionary parameters can produce the same GW event.
Using Bayes theorem, the probability of an event with evolutionary parameters given the GW parameters can be derived as following

\begin{align}
    p(\bm{\Theta}|\bm{\theta}) &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{p(\bm{\theta})}, \nonumber \\
    &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{\int p(\bm{\theta}|\bm{\Theta}) p(\bm{\Theta}) d\bm{\Theta}}, \nonumber \\
    &= \frac{\bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}{\int d\bm{\theta} \bm{\delta}(f(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}.
\label{eq:backward_prob}
\end{align}

Using the result in eq.\ref{eq:backward_prob},
the probability of an event having evolutionary parameters given data can be obtained by marginalizing over the GW parameters

\begin{align}
    p(\bm{\Theta}|d) &= \int p(\bm{\Theta}|\bm{\theta}) p(\bm{\theta}|d) d\bm{\theta}\\
    &= \frac{1}{N}\sum_{i=1}^{N} p(\bm{\Theta}|\bm{\theta}_i). \label{eq:evolutionary_posterior}
\end{align}

The term $p(\bm{\theta}|d)$ is the inferred posterior probability for a particular event, which is usually presented by the LVK collaboration in the form of posterior samples.
In eq.\ref{eq:evolutionary_posterior}, we use importance sampling to turn the integration into a sum over the posterior samples.

For each posterior sample point, we can find the corresponding progenitor parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and progenitor parameters, i.e.

\begin{align}
\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2
\label{eq:loss}
\end{align}

In principle, we should only accept solutions that exactly reproduce the LVK posterior samples.
However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$.
To make sure we find the complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.

Ideally we would like to have explicit control over all evolutionary parameters.
In particular, having explicit control over variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicit,
i.e. the random variables are drawn within the program, instead of being arguments of the program.
This means the function we use to evolve the binary is not a deterministic function but a stochastic function instead. 
Hence, even if we can find all the roots that corresponds to the posterior samples,
forward propagating this set of roots does not gauruntee to reproduce the set of posterior samples due to the randomness of the evolution function.
To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,

\section{Result}

\begin{enumerate}
\item Raw posterior of successful events.
\item Trace back the event's $z_{\rm ZAMS}$.
\item Reproject to observable space
\item Evidence ratio.
\end{enumerate}

% \begin{figure}
% \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.png}
% \caption{Caption}
% \label{fig:GW150914_reprojection}
% \end{figure}

\section{Discussion}

Interpretation:
\begin{enumerate}
\item Highlight the fact that we turn hyper-parameters into parameters.
\item Discuss this is need for high dimensional exploration.
\end{enumerate}

Caveats: 
\begin{enumerate}
\item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
\item We only assume BSE
\item Caveat on marginalizing over random samples.
\item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
\item How next generation "Progenitor synthesis" code should be built.
\item Rate
\item Failure studies go into appendix
\end{enumerate}

\end{document}

### 2022-02-07 src/ms.tex
Working version
z

@@
 
-\title{Backprogating Gravitational wave events}
+\title{Backpropagating Gravitational wave events}
 
@@
 \email{kwong@flatironinstitute.org}
-%\flatiron
+\flatiron
 
@@
 
-\maketitle
+%\maketitle
 
@@
 
-This paper introduce a number of new perspectives in the study of binary evolution physics and gravitational-wave progenitors:
+This paper introduces a number of new perspectives in the study of binary evolution physics and gravitational-wave progenitors:
 \begin{enumerate}
@@
 
-For each posterior sample point, we can find the corresponding progenitor parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and progenitor parameters, i.e.
+For each posterior sample point, we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.
 
@@
 Ideally we would like to have explicit control over all evolutionary parameters.
-In particular, having explicit control over variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
-However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicit,
+In particular, having explicit control over random variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
+However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicitly,
 i.e. the random variables are drawn within the program, instead of being arguments of the program.
 This means the function we use to evolve the binary is not a deterministic function but a stochastic function instead. 
-Hence, even if we can find all the roots that corresponds to the posterior samples,
-forward propagating this set of roots does not gauruntee to reproduce the set of posterior samples due to the randomness of the evolution function.
+Hence, even if the root-finding algorithm performs perfectly,
+forward modelling this set of roots does not guarantee to reproduce the set of posterior samples due to randomness of the evolution function.
 To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,
+we forward model ("Reproject) the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
+We use KL divergence to measure the agreement between the two posterior distributions.
+A small KL divergence means \texttt{COSMIC} is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
+This is completely expected since \texttt{COSMIC} comes with its own assumptions and is expected to fail in reproducing a subset of the event in GWTC3, such as event with at least one component beyond the pair instability supernova massgap.
+On the other hand, the reprojection could subject to stochasticity in the evolution function we use.
+Instead of reprojecting the posterior only once, we reproject the posterior multiple times with different random seeds and check whether the KL divergence varies significantly.
+A varying KL divergence means this particular event is subjected to randomness in the evolution function, hence one should take extra caution in interpreting the result.
 

### 2022-02-07 src/figures/GW150914_reprojection.py
Add plot

@@ -1,0 +1,114 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
+                               AutoMinorLocator)
+from matplotlib.lines import Line2D
+
+params = {'axes.labelsize': 32,
+          'font.family': 'serif',
+          'font.serif': 'Computer Modern Raman',
+          'font.size': 32,
+          'axes.linewidth': 2,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 20,
+          'xtick.top': True,
+          'xtick.direction': "in",
+          'ytick.labelsize': 20,
+          'ytick.right': True,
+          'ytick.direction': "in",
+          'axes.grid' : False,
+          'text.usetex': True,
+          'savefig.dpi' : 100,
+          'lines.markersize' : 14,
+#          'axes.formatter.useoffset': False,
+          'axes.formatter.limits' : (-3,3)}
+
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+
+mpl.rcParams.update(params)
+
+def fill_corner(fig,data,ndim,axis,color='C0',true_val=None,zorder=1,lw=3,style='-',smooth1d = 0.01,no_fill_contours=False,fill_contours=True,alpha=1,levels=[0.68,0.95]):
+    for i in range(ndim):
+        for j in range(i+1):
+            ax = fig.axes[np.sum(np.arange(i+1))+j]
+            if i==j:
+                ax.hist(data[i],bins=axis[i],histtype='step',lw=lw,color=color,density=True,zorder=zorder)
+                ylim = ax.get_ylim()
+                ax.set_xlim(axis[i][0],axis[i][-1])
+                ax.set_ylim(ylim[0],ylim[1])
+                if true_val is not None:
+                    ax.vlines(true_val[i],ylim[0],ylim[1],color=color,lw=lw,zorder=zorder)
+            elif j<i:
[…]

### 2022-02-08 src/ms.tex
Method more or less done

@@
 \newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{bm}
+%\usepackage{align}
 % Begin!
@@
 To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,
-we forward model ("Reproject) the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
-We use KL divergence to measure the agreement between the two posterior distributions.
+we forward model ("Reproject") the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
+We use KL divergence to measure the agreement between the two posterior distributions, which is given by
+\begin{align}
+D_{KL}(P||Q) = \int P(\bm{x}) \log(P(\bm{x})/Q(\bm{x})) dx.
+\label{eq:KLdivergence}
+\end{align}
+
 A small KL divergence means \texttt{COSMIC} is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
@@
 
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.png}
-% \caption{Caption}
-% \label{fig:GW150914_reprojection}
-% \end{figure}
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{GW150914_reprojection.pdf}
+\caption{Caption}
+\label{fig:GW150914_reprojection}
+\end{figure}
 

### 2022-02-08 src/figures/GW150914_reprojection.py

@@ -1,11 +1,7 @@
 import numpy as np
 import h5py
 import matplotlib.pyplot as plt
-from scipy.interpolate import interp1d
-from scipy.integrate import cumtrapz
-from scipy.optimize import curve_fit
 import copy
-from astropy.cosmology import Planck18
 
 import matplotlib as mpl
 import corner
@@ -32,7 +28,7 @@ params = {'axes.labelsize': 32,
 #          'axes.formatter.useoffset': False,
           'axes.formatter.limits' : (-3,3)}
 
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+#mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
 
 mpl.rcParams.update(params)
 

### 2022-02-09 src/ms.tex
Add posterior script

@@
 \usepackage{bm}
-%\usepackage{align}
+\usepackage{xcolor}
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 % Begin!
 \begin{document}
-
 \title{Backpropagating Gravitational wave events}
@@
 a fixed value for the entire population.
-This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution,
+This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution with other independent studies,
 or they should vary the family of distributions.
+In general, it is unclear how one can cover all possible distribution that may agree with the observation in generating a suite of simulations.
+The problem worsen as we increase the dimensionality of the parameter space.
+Indeed, one of the most common skepticism population synthesis studies receives is a simulations set usually only cover a small subdomain of the entire parameter space,
+i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
 
-In this paper, instead of forward modelling the population of binary star to match observed GW events population,
+In this paper, instead of forward modelling a population of binary star to match observed GW events population,
 we propose a method to "backward" model each GW event to its progenitor state.
 
-This paper introduces a number of new perspectives in the study of binary evolution physics and gravitational-wave progenitors:
-\begin{enumerate}
-\item Instead of starting from a distribution 
-\item Using derivative-based optimization algorithm, we are allowed to explore a moderate size parameter space 
-\end{enumerate}
+This paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
+In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
+We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.
 
 \section{Method}
+\label{sec:method}
 
@@
 To make sure we find the complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
+\kw{Mention how we generate the initial guesses.}
 
@@
 \section{Result}
+\label{sec:result}
 
@@
 
+\begin{figure*}
+\includegraphics[width=0.98\textwidth]{GW150914_corner.pdf}
+\caption{The posterior for GW150914}
+\label{fig:GW150914_posterior}
+\end{figure*}
+
 \begin{figure}
 \includegraphics[width=0.49\textwidth]{GW150914_reprojection.pdf}
-\caption{Caption}
+\caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.}
 \label{fig:GW150914_reprojection}
@@
 
+We can see the reprojected posterior agrees very well with the original posterior in fig.\ref{fig:GW150914_reprojection}.
+\kw{Qoute the KL divergence.}
+
 \section{Discussion}
+\label{sec:discussion}
 
+We present a new pathway to understand binary evolution with GW events in this paper.
+Instead of forward modelling an assumed distribution of initial binary to the observed population,
+we solve a root-finding problem to find the best initial binary parameters that reproduce the posterior distribution for each event. 
+The most obvious advantage our method provides is it can return the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
+This enables a data-driven way to study the distribution of hyper-parameters,
+i.e.
+once we have a catalog of events, each has their own set of posterior in the hyper-parameter space, we can employ common technique such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters.
+
+While our method allows data-driven exploration of the hyper-parameters space for the first time, there are numerous improvements 
+In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
+For example, event with either of the component mass larger than the PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
+Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3.
+As pointed out in the literature, it is unlikely that all the observed events in GWTC3 can be explained by a single formation channel.
+As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.
+Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
+
+Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
+This introduces significant inefficiency in our root-finding algorithm.
+The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolution pathway of low mass events such as BNS events.
+The effect of natal kick is suppressed for heavier mass events due to fallback.
+This means events with the root-finding process for lighter masses are subject to stochasticity of the function,
+where the root-finding process for heavier events behaves as if the evolutionary function we use are deterministic.
+Indeed, we see this behavior in a couple of case studies.
+For heavy event such as GW150914, the reprojected posterior agrees with the data posterior.
+For light event such as GW170817, although during the root-finding process the convergence criteria is met,
+the reprojected posterior does not agree with the data posterior.
+Also, due to computational limitation, we only try 1000 different initial guesses per posterior sample in the root-finding process.
+This means any posterior sample that has a probability of merging rarer than 1 in 1000 could be missed.
+\kw{comment on the low mass ratio end.}
+
+\kw{Mention controlling the randomness is particular important for dynamical formation channel, which let different binary interact statisically.}
+
+We hope this letter would motivate the community to build the relevant infrastructure mentioned above,
+so in the near future we can explore the full parameter space of binary evolution models with the next-generation GW detectors network.
 Interpretation:
@@
 
+\section{Acknowledgement}
+
 \end{document}

### 2022-02-09 src/figures/GW150914_posterior.py

@@ -1,0 +1,135 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+import copy
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
+                               AutoMinorLocator)
+from matplotlib.lines import Line2D
+
+params = {'axes.labelsize': 32,
+          'font.family': 'serif',
+          'font.serif': 'Computer Modern Raman',
+          'font.size': 32,
+          'axes.linewidth': 2,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 20,
+          'xtick.top': True,
+          'xtick.direction': "in",
+          'ytick.labelsize': 20,
+          'ytick.right': True,
+          'ytick.direction': "in",
+          'axes.grid' : False,
+          'text.usetex': True,
+          'savefig.dpi' : 100,
+          'lines.markersize' : 14,
+#          'axes.formatter.useoffset': False,
+          'axes.formatter.limits' : (-3,3)}
+
+#mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+
+mpl.rcParams.update(params)
+
+def fill_corner(fig,data,ndim,axis,color='C0',true_val=None,zorder=1,lw=3,style='-',smooth1d = 0.01, smooth2d=0.01,no_fill_contours=False,fill_contours=True,alpha=1,levels=[0.68,0.95]):
+    for i in range(ndim):
+        for j in range(i+1):
+            ax = fig.axes[np.sum(np.arange(i+1))+j]
+            if i==j:
+                ax.hist(data[i],bins=axis[i],histtype='step',lw=lw,color=color,density=True,zorder=zorder)
+                ylim = ax.get_ylim()
+                ax.set_xlim(axis[i][0],axis[i][-1])
+                ax.set_ylim(ylim[0],ylim[1])
+                if true_val is not None:
+                    ax.vlines(true_val[i],ylim[0],ylim[1],color=color,lw=lw,zorder=zorder)
+            elif j<i:
[…]

### 2022-02-14 src/ms.tex
zz

@@
 
-\begin{figure*}
-\includegraphics[width=0.98\textwidth]{GW150914_corner.pdf}
-\caption{The posterior for GW150914}
-\label{fig:GW150914_posterior}
-\end{figure*}
+% \begin{figure*}
+% \includegraphics[width=0.98\textwidth]{GW150914_corner.pdf}
+% \caption{The posterior for GW150914}
+% \label{fig:GW150914_posterior}
+% \end{figure*}
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{GW150914_reprojection.pdf}
-\caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.}
-\label{fig:GW150914_reprojection}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=0.49\textwidth]{GW150914_reprojection.pdf}
+% \caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.}
+% \label{fig:GW150914_reprojection}
+% \end{figure}
 
@@
 we solve a root-finding problem to find the best initial binary parameters that reproduce the posterior distribution for each event. 
-The most obvious advantage our method provides is it can return the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
+The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
 This enables a data-driven way to study the distribution of hyper-parameters,
-i.e.
-once we have a catalog of events, each has their own set of posterior in the hyper-parameter space, we can employ common technique such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters.
+i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
+we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters.
+\kw{Talk about GW150914.}
 
+% Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
+Another perk of our method is the "rate" calculation is now independent of assumed initial condition such as a particular star formation rate model,
+which is a dominant factor of uncertainty for rate prediction in the past.
+Once we have pull back the GW event posterior from the observable space to the evolutionary parameter space,
+we can apply the same methods we use to account for the rate in the observable space.
+In the observable space, the overall rate is basically a normalization that is defined in the local universe i.e. the local merger rate. \kw{cite}
+In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
+\kw{Talk about prior are now in post-processing instead of pre-processing.}
+
+% Gradient decending in the evolutionary parameter space is much more efficent than rejection sampling.
+
+% Single formation channel and fundamental limitation on that.
 While our method allows data-driven exploration of the hyper-parameters space for the first time, there are numerous improvements 
@@
 
+% Randomness paragraph
 Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
@@
 \kw{comment on the low mass ratio end.}
+Obviously the problems which comes with the randomness can be alleviated by performing more tries per posterior sample,
+but this is not scalable in practice.
+On top of limitation in efficiency, some formation channels require explicit control of random variables by construction.
+For example, in a dynamical formation scenario such as binaries that form in a globular cluster,
+each binary has some probability of undergoing a multi-body encounter with another member in the cluster.
+These encounter probability distributions are either studied with direct N-body simulations or semi-analytical methods.
+In both cases, each member of the cluster is no longer completely independent of the other members, but coupled through the encounter probability distribution.
+By studying the encounter probability distribution, we can infer the properties of the environment which the binary lives in.
+This can only be done if we have explicit control over the random variables that characterize the encounter probability distribution.
 
-\kw{Mention controlling the randomness is particular important for dynamical formation channel, which let different binary interact statisically.}
+% Need for selection bias to recover intrinsic distribution
+
+% Need for full autodiff.
 

### 2022-03-03 src/ms.tex
update
Initial condition update

@@
 \end{enumerate}
+
+To understand the origin of gravitational wave (GW) events,
+modeller
 
@@
 
-In principle, we should only accept solutions that exactly reproduce the LVK posterior samples.
+In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the observable space.
 However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$.
-To make sure we find the complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
-\kw{Mention how we generate the initial guesses.}
+To make sure we find a reasonably complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
+As long as the solution fulfill the acceptance criteria, the root is as valid as all other roots that fulfill the same acceptance criteria, regardless to its initial guess.
+In other word, we have the freedom to choose the set of initial guesses however we think it might benefit the optimization.
+Most of the points in the evolutionary parameter space do not produce mergers, and for systems that do not merge, we set the binary masses to be 0.
+This means the gradient for systems that do not merge is 0, and the root finder would likely to get stuck in such state during the first step.\footnote{On a plateau of constant loss, the root finder performs no better than a random guess.
+In a high dimensional space, this means the root finder will need a long time to get close to a solution.}
+Therefore, it is beneficial to put initial guesses in region of the evolutionary parameter space that does produce merger.
+To construct a list of initial guesses, we evolve a set of points that we uniformly sampled from the evolutionary parameter space, then keep the points that merge within a Hubble time.
+%Note that this freedom in choosing the initial guess introduces a subtlety in interpreting the posterior in the evolutionary parameter space. \kw{continue explaining}
+
+% \kw{There is a caveat to this initialization.
+Imagine there are two clusters of posterior points in the evolutionary parameter space, one with N points and one with M points, do we treat each cluster as one point?
+Or do we give equal weight to each root?
+This is probably a choice at the end which matters a little bit, and depends on the prior choice, but we should clear the conceptual confusion.
+% Perhaps we can put a heuristic experiment result in appendix: we randomly sample the possible root for each posterior, then use that as the result, and we do it multiple times to show the result doesn't change dramatically.
+% Prior weight across posterior samples may help to alleviate the issue.}
+% \kw{The most robust thing one can do is sample the prior in the evolationary parameter space, pipe it through BSE then the GW PE code.
+I think it is still subject to the same multiple root problem, and N-M problem will still be there.
+% So in a sense, the only prior we have to take care is really the GW prior.}
+% \kw{Mention how we generate the initial guesses.}
 
@@
 
-A small KL divergence means \texttt{COSMIC} is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
+A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the observable space, hence it is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
 This is completely expected since \texttt{COSMIC} comes with its own assumptions and is expected to fail in reproducing a subset of the event in GWTC3, such as event with at least one component beyond the pair instability supernova massgap.
@@
 \item Reproject to observable space
-\item Evidence ratio.
+\item Evidence ratio?
+\item KL divergence?
 \end{enumerate}
@@
 i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
-we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters.
+we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters,
+without making specific assumption such as fixing the hyper-parameters for all types of event.
+
 \kw{Talk about GW150914.}
@@
 
-% Gradient decending in the evolutionary parameter space is much more efficent than rejection sampling.
-
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyper-parameters space for the first time, there are numerous improvements 
+While our method allows data-driven exploration of the hyper-parameters space for the first time, there are a number of improvements can be implemented in future studies.
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
@@
 Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3.
-As pointed out in the literature, it is unlikely that all the observed events in GWTC3 can be explained by a single formation channel.
+As pointed out in the literature \kw{Cite mike}, it is unlikely that all the observed events in GWTC3 can be explained by a single formation channel.
 As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.
@@
 % Randomness paragraph
-Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
+Due to thae implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
 This introduces significant inefficiency in our root-finding algorithm.
@@
 
+% Gradient decending in the evolutionary parameter space is much more efficent than rejection sampling.
+% Potential pitfall of interpreting this result.
 % Need for full autodiff.
+
+By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
+In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
+In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
+This greatly benefit the efficiency of the root-finding process.
+
+On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
+One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
+Since posterior samples in the observable space can have multiple roots in the evolutionary parameter space,
+the relatively weighting of these roots becomes ambiguous when the number of roots is large.
+For example, should we consider roots that are close to each other a single root or multiple roots?
+If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
+Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
+This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
+In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign each weight to each root,
+and we have checked the posterior samples produced by this method for events discussed in this paper agree with direct sampling using a uniform prior.
+This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the distribution in the observable space instead of multiple possible roots per posterior sample.
+We leave a detail comparison between this method and direct sampling to future studies.
+
+Another technical note is we use finite differencing to estimate the gradient of the objective function, which could be a significant source of error near transition points in the evolutionary parameter space.
+Also, finite differencing is increasing inefficiency as we increase the dimensionality of the problem.
+To improve the accuracy and efficiency in estimating the gradient of objective function, automatic differentiation is a promising feature that modeler should consider incorporating in their population synthesis code in the future.
 

### 2022-03-07 src/ms.tex
update introduction

@@
 
-\begin{enumerate}
-\item We have more and more event
-\item Problem with population synthesis -> Need for developing algorithm to explore the high dimensional space
-\item We run the event backward in time
-\end{enumerate}
+As gravitational-wave (GW) detectors sensitivity improve, the rapid increase of GW detectors shifts out 
+\kw{think about the first sentence.}
 
-To understand the origin of gravitational wave (GW) events,
-modeller
+A very common way to study the origin of detected GW events is through forward modelling of GW progenitor.
+Forward modelling of a binary system start with a set of initial binary parameters,
+then evolve it according to the physics of interest until some stopping criteria is met.
+This usually mean either a merger event or the binary system stays stable for a Hubble time.
 
@@
 it is not necessarily the most efficient way to constraint the underlying physical model.
-Forward modelling of a binary system start with a set of initial binary parameters,
-then evolve it until some stopping criteria is met,
-typically either a merger event or the binary system stays stable for a Hubble time.
-The problem is, most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
+Most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
 This means if we want to evolve a binary to match the observation by pure trial and error, a large fraction of computation will be wasted in trying initial binary configurations that are not related to the observation.
@@
 The problem worsen as we increase the dimensionality of the parameter space.
+
 Indeed, one of the most common skepticism population synthesis studies receives is a simulations set usually only cover a small subdomain of the entire parameter space,
 i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
+Even with recently development on using emulators to speed up the simulation process, training the emulators also require a decent amount of samples in the parameter space of interest.
+For example, say we want to emulate a model with 10 parameters, this simplest option for constructing a bank of simulations for training is to lay down a grid.
+Along each axis we put the barely minimum of $2$ simulations, this would still require $1024$ simulations.
+The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
+This means even training an emulator is impossible in high dimensional space.
 
-In this paper, instead of forward modelling a population of binary star to match observed GW events population,
+In this paper, instead of forward modelling a population of binary star to match the observed GW events population,
 we propose a method to "backward" model each GW event to its progenitor state.
+The idea is since we can compute the GW event properties of a guess by forward modelling its progenitor,
+we can define the difference between the GW event properties of this guess and the observed GW event properties.
+By minimizing this difference, i.e. solving a root finding problem, we can hope to recover the progenitor properties.
 
@@
 % Randomness paragraph
-Due to thae implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
+Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
 This introduces significant inefficiency in our root-finding algorithm.
@@
 This greatly benefit the efficiency of the root-finding process.
-
 On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
 One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
-Since posterior samples in the observable space can have multiple roots in the evolutionary parameter space,
+Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
 the relatively weighting of these roots becomes ambiguous when the number of roots is large.
@@
 This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
-In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign each weight to each root,
+In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root,
 and we have checked the posterior samples produced by this method for events discussed in this paper agree with direct sampling using a uniform prior.
-This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the distribution in the observable space instead of multiple possible roots per posterior sample.
+This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
 We leave a detail comparison between this method and direct sampling to future studies.
@@
 To improve the accuracy and efficiency in estimating the gradient of objective function, automatic differentiation is a promising feature that modeler should consider incorporating in their population synthesis code in the future.
+
+\kw{Mention potential complimentary studies of direct sampling, and how all these suggestions are still useful regardless of whether we are doing root-finding or not.}
 

### 2022-03-08 src/figures/GW150914_posterior.py
Update zenodo dependecies

@@ -3,8 +3,8 @@ import h5py
 import matplotlib.pyplot as plt
 import copy
 
-import matplotlib as mpl
 import corner
+import matplotlib as mpl
 from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
                                AutoMinorLocator)
 from matplotlib.lines import Line2D
@@ -73,27 +73,26 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-i = 0
-data = h5py.File('/mnt/home/wwong/ceph/GWProject/ProgenitorCatalogue/GWTC3_bse_event_1_merged.hdf5')
-GWTC = np.load('/mnt/ceph/users/wwong/GWProject/GWTC/Processed/Combined_GWTC_m1m2chieffz.npz')
-processed_data = h5py.File('/mnt/home/wwong/ceph/GWProject/ProgenitorCatalogue/GWTC3_bse_event_1_merged_process.hdf5')
-m1_obs = np.median(GWTC['m1'],axis=1)
-m2_obs = np.median(GWTC['m2'],axis=1)
-z_obs = np.median(GWTC['z'],axis=1)
+data = h5py.File('../data/GW150914_progenitor_posterior.hdf5')
+GWTC = np.load('../data/GW150914_posterior.npz')
+processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5')
+m1_obs = np.median(GWTC['m1'])
+m2_obs = np.median(GWTC['m2'])
+z_obs = np.median(GWTC['z'])
 q = m2_obs/m1_obs
 m1_pred = processed_data['m1']
 m2_pred = processed_data['m2']
 t_merge = processed_data['t_merge']
-observables = np.stack([GWTC['m1'],GWTC['m2'],GWTC['z']],axis=2)
+observables = np.stack([GWTC['m1'],GWTC['m2'],GWTC['z']],axis=1)
 posterior_array = []
 weight_array = []
 observable_array = []
 quantile_array = []
 
-a,b,w,index = get_posterior(data['result_value'][i],data['result_coord'][i],1e-1)
+a,b,w,index = get_posterior(data['result_value'][0],data['result_coord'][0],1e-1)
 posterior_array.append(a)
 weight_array.append(w)
-observable_array.append(observables[i][index[0]])
+observable_array.append(observables[index[0]])
[…]

### 2022-03-08 src/figures/GW150914_reprojection.py

@@ -72,18 +72,17 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-i = 0
-data = h5py.File('/mnt/home/wwong/ceph/GWProject/ProgenitorCatalogue/GWTC3_bse_event_1_merged.hdf5')
-GWTC = np.load('/mnt/ceph/users/wwong/GWProject/GWTC/Processed/Combined_GWTC_m1m2chieffz.npz')
-processed_data = h5py.File('/mnt/home/wwong/ceph/GWProject/ProgenitorCatalogue/GWTC3_bse_event_1_merged_process.hdf5')
-m1_obs = np.median(GWTC['m1'],axis=1)
-m2_obs = np.median(GWTC['m2'],axis=1)
-z_obs = np.median(GWTC['z'],axis=1)
+data = h5py.File('../data/GW150914_progenitor_posterior.hdf5')
+GWTC = np.load('../data/GW150914_posterior.npz')
+processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5')
+m1_obs = np.median(GWTC['m1'])
+m2_obs = np.median(GWTC['m2'])
+z_obs = np.median(GWTC['z'])
 q = m2_obs/m1_obs
 m1_pred = processed_data['m1']
 m2_pred = processed_data['m2']
 t_merge = processed_data['t_merge']
-observables = np.stack([GWTC['m1'],GWTC['m2'],GWTC['z']],axis=2)
+observables = np.stack([GWTC['m1'],GWTC['m2'],GWTC['z']],axis=1)
 posterior_array = []
 weight_array = []
 observable_array = []
@@ -92,14 +91,14 @@ quantile_array = []
 a,b,w,index = get_posterior(data['result_value'][0],data['result_coord'][0],1e-1)
 posterior_array.append(a)
 weight_array.append(w)
-observable_array.append(observables[i][index[0]])
+observable_array.append(observables[index[0]])
 quantile_array.append(np.quantile(a,[0.5-0.95/2,0.5,0.5+0.95/2],axis=0))
 quantile_array = np.array(quantile_array)
 plt.figure(figsize=(10,9))
 percentile = [0.7,0.9]
 corner.hist2d(m1_pred[()],m2_pred[()],weights=weight_array[0],color='C0',plot_density=False,labels='Reprojected',levels=percentile)
[…]

### 2022-03-08 src/figures/GW150914_posterior.py
Update plotting script and environment

@@ -73,9 +73,9 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-data = h5py.File('../data/GW150914_progenitor_posterior.hdf5')
+data = h5py.File('../data/GW150914_progenitor_posterior.hdf5','r')
+processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5','r')
 GWTC = np.load('../data/GW150914_posterior.npz')
-processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5')
 m1_obs = np.median(GWTC['m1'])
 m2_obs = np.median(GWTC['m2'])
 z_obs = np.median(GWTC['z'])

### 2022-03-08 src/figures/GW150914_reprojection.py

@@ -72,9 +72,9 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-data = h5py.File('../data/GW150914_progenitor_posterior.hdf5')
+data = h5py.File('../data/GW150914_progenitor_posterior.hdf5','r')
 GWTC = np.load('../data/GW150914_posterior.npz')
-processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5')
+processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5','r')
 m1_obs = np.median(GWTC['m1'])
 m2_obs = np.median(GWTC['m2'])
 z_obs = np.median(GWTC['z'])

### 2022-03-10 src/ms.tex
Update zenodo dependecies
Add result paragraph

@@
 
-% \begin{figure*}
-% \includegraphics[width=0.98\textwidth]{GW150914_corner.pdf}
-% \caption{The posterior for GW150914}
-% \label{fig:GW150914_posterior}
-% \end{figure*}
+We show our main result in figure \ref{fig:GW150914_posterior}. 
+\kw{Recover ratio of the posterior sample goes here.}
 
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{GW150914_reprojection.pdf}
-% \caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.}
-% \label{fig:GW150914_reprojection}
-% \end{figure}
+\begin{figure*}
+\includegraphics[width=0.98\textwidth]{static/GW150914_corner.pdf}
+\caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
+$M_1$ and $M_2$ are the progenitors' masses. $\log{t_{orb}}, e, \log{Z}$ are the progenitor binary's log orbital period, eccentricity and log metallicity at ZAMS.
+$\alpha$ is the common envelope efficiency.
+$f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
+$q_{\rm crit 3,4}$ are the ciritical mass during \kw{Katie fill in},
+$\sigma$ is the mean magnitude of the Maxwellian distribution characterizing natal kick strength.
+}
+\label{fig:GW150914_posterior}
+\end{figure*}
+
+To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
+We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
+For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} on average.
+The KL divergence stays stable for reprojecting with different random seeds.
+
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{static/GW150914_reprojection.pdf}
+\caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
+The blue contour is the posterior reprojected using the samples in the evolutionary parameter space.
+The orange contour is the posterior samples from LVK that have corresponding samples in the evolutionary parameter space.
+The green contour is the posterior plotted using all of LVK posterior samples.
+}
+\label{fig:GW150914_reprojection}
+\end{figure}
 
@@
 \kw{Qoute the KL divergence.}
+
+\kw{Add figure of stable mass transfer vs common envelope}
 
@@
 so in the near future we can explore the full parameter space of binary evolution models with the next-generation GW detectors network.
+
 Interpretation:

### 2022-03-14 src/ms.tex
Add figures

@@
 
-We show our main result in figure \ref{fig:GW150914_posterior}. 
+We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
 \kw{Recover ratio of the posterior sample goes here.}
+
+The full posterior is available in appendix \kw{Or we should w=put a link to the dataset?}
 
@@
 
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{static/GW150914_CE_SMT.pdf}
+\caption{}
+\label{fig:GW150914_CE_SMT}
+\end{figure}
 \kw{Add figure of stable mass transfer vs common envelope}
+
+\begin{figure}
+\caption{}
+\label{fig:GW150914_M1_M2_q3}
+\end{figure}
+
+The most important point in this paper is illustrated in figure \ref{fig:GW150914_M1_M2_q3}.
+We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $q_3$,
+this means if we fixed the value of $q_3$ instead of allowing it to jointly vary with all other parameters,
+we would miss the samples that do not match the fixed value of $q_3$.
+Comparing our result with 
+\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters incorrectly.}
 

### 2022-03-14 src/ms.tex
Add paragraph on companion paper

@@
 this means if we fixed the value of $q_3$ instead of allowing it to jointly vary with all other parameters,
-we would miss the samples that do not match the fixed value of $q_3$.
-Comparing our result with 
-\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters incorrectly.}
+we would miss the samples that do not match the chosen value of $q_3$.
+By folding in the hyper-parameters into the root-finding algorithm,
+we can capture the correlation between 
+\kw{Dartboard comparsion} Comparing our result with 
+\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}
 
@@
 we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters,
-without making specific assumption such as fixing the hyper-parameters for all types of event.
+without making specific assumptions such as fixing the hyper-parameters for all types of event.
 
-\kw{Talk about GW150914.}
+In this work we showcase the power of the proposed method with two events, GW150914 and GW191129.
+We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
+In particular, to the authors knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is shown.
+Our work is not only a more efficient way to study GW events' progenitors,
+but also allows the possibility of measuring the astrophysics related the binary evolution.
 
@@
 Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
+In this paper, our main focus is to illustrate the concept of "back-propagating" GW events posterior sample in this work,
+highlighting the capability of our method and motivating the community to build the next generation population synthesis code that can work with our method.
+To avoid cluttering of focus, we discuss the physical implication of the result presented in this work under our specific assumption (i.e.
+using \texttt{COSMIC} as our evolutionary function) in a companion paper.
 

### 2022-03-14 src/ms.tex
Add paragraphs

@@
 Our work is not only a more efficient way to study GW events' progenitors,
-but also allows the possibility of measuring the astrophysics related the binary evolution.
+but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.
 
@@
 
+To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
+We "back propagate" the posterior in the observable space to the evolutionary parameter space,
+thus allowing us to study the correlation between progenitor parameters and hyper-parameters.
+We demonstrate the capability of this method with two events, GW150914 and GW191129, while employing the binary population synthesis model \texttt{COSMIC}.
+We 
 We hope this letter would motivate the community to build the relevant infrastructure mentioned above,

### 2022-03-17 src/ms.tex
Add m1 m2 alpha figure. I don't think we need the CE vs SMT figure here
typo fixing

@@
 
-Indeed, one of the most common skepticism population synthesis studies receives is a simulations set usually only cover a small subdomain of the entire parameter space,
+Indeed, one of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
 i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
@@
 $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-$q_{\rm crit 3,4}$ are the ciritical mass during \kw{Katie fill in},
+$q_{\rm crit 3,4}$ are the cirtical mass during \kw{Katie fill in},
 $\sigma$ is the mean magnitude of the Maxwellian distribution characterizing natal kick strength.
@@
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{static/GW150914_CE_SMT.pdf}
-\caption{}
-\label{fig:GW150914_CE_SMT}
-\end{figure}
-\kw{Add figure of stable mass transfer vs common envelope}
+% \begin{figure}
+% \includegraphics[width=0.49\textwidth]{static/GW150914_CE_SMT.pdf}
+% \caption{}
+% \label{fig:GW150914_CE_SMT}
+% \end{figure}
+% \kw{Add figure of stable mass transfer vs common envelope}
 
 \begin{figure}
+\includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
 \caption{}

### 2022-03-17 src/ms.tex
Update discussion

@@
 
-The full posterior is available in appendix \kw{Or we should w=put a link to the dataset?}
+The full posterior is available in appendix \kw{Or we should put a link to the dataset?}
 
@@
 \caption{}
-\label{fig:GW150914_M1_M2_q3}
+\label{fig:GW150914_m1_m2_alpha}
 \end{figure}
 
-The most important point in this paper is illustrated in figure \ref{fig:GW150914_M1_M2_q3}.
-We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $q_3$,
-this means if we fixed the value of $q_3$ instead of allowing it to jointly vary with all other parameters,
-we would miss the samples that do not match the chosen value of $q_3$.
+The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
+We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
+this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
+we would miss the samples that do not match the chosen value of $\alpha$.
 By folding in the hyper-parameters into the root-finding algorithm,
@@
 To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
+We point out hyper-parameters in the usual population synthesis simulation context are not actually parameters related to the population,
+but parameters about the evolutionary function.
+This means the binary evolution functions can be constrained on an individual event basis.
 We "back propagate" the posterior in the observable space to the evolutionary parameter space,
-thus allowing us to study the correlation between progenitor parameters and hyper-parameters.
-We demonstrate the capability of this method with two events, GW150914 and GW191129, while employing the binary population synthesis model \texttt{COSMIC}.
-We 
-We hope this letter would motivate the community to build the relevant infrastructure mentioned above,
-so in the near future we can explore the full parameter space of binary evolution models with the next-generation GW detectors network.
+thus allowing us to study hyper-parameters and its correlation with progenitor parameters in a data-driven manner.
+Our method makes less assumption than the traditional forward modelling approach,
+which often fix the hyper-parameters across the entire population.
+Since we are not limited to the fix hyper-parameters assumption, we can explore the behavior of the hyper-parameters across the population much more efficiently.
+While our work lay down a data analysis pathway to understand the population of GW events,
+no physics can be learned without a comprehensive complimentary physical model.
+We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:
+first, they should have explicit control over the random variables so marginalizing over random variables can be done more precisely;
+second, they should be as automatically differentiable as much as possible so exploring the evolutionary parameter space is efficient.
+Combining this work and the next generation population synthesis code,
+we can explore the full parameter space of binary evolution models with the next-generation GW detectors network in the near future.
 
-Interpretation:
-\begin{enumerate}
-\item Highlight the fact that we turn hyper-parameters into parameters.
-\item Discuss this is need for high dimensional exploration.
-\end{enumerate}
+% Interpretation:
+% \begin{enumerate}
+% \item Highlight the fact that we turn hyper-parameters into parameters.
+% \item Discuss this is need for high dimensional exploration.
+% \end{enumerate}
 
-Caveats: 
-\begin{enumerate}
-\item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
-\item We only assume BSE
-\item Caveat on marginalizing over random samples.
-\item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
-\item How next generation "Progenitor synthesis" code should be built.
-\item Rate
-\item Failure studies go into appendix
-\end{enumerate}
+% Caveats: 
+% \begin{enumerate}
+% \item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
+% \item We only assume BSE
+% \item Caveat on marginalizing over random samples.
+% \item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
+% \item How next generation "Progenitor synthesis" code should be built.
+% \item Rate
+% \item Failure studies go into appendix
+% \end{enumerate}
 

### 2022-03-17 src/ms.tex
typo fix

@@
 $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-$q_{\rm crit 3,4}$ are the cirtical mass during \kw{Katie fill in},
+$q_{\rm crit 3,4}$ are the critical mass during \kw{Katie fill in},
 $\sigma$ is the mean magnitude of the Maxwellian distribution characterizing natal kick strength.

### 2022-03-21 src/ms.tex
Add table. Need formatting

@@
 In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the observable space.
-However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$.
+However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$ to be our threshold.
 To make sure we find a reasonably complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
@@
 To construct a list of initial guesses, we evolve a set of points that we uniformly sampled from the evolutionary parameter space, then keep the points that merge within a Hubble time.
-%Note that this freedom in choosing the initial guess introduces a subtlety in interpreting the posterior in the evolutionary parameter space. \kw{continue explaining}
-
-% \kw{There is a caveat to this initialization.
-Imagine there are two clusters of posterior points in the evolutionary parameter space, one with N points and one with M points, do we treat each cluster as one point?
-Or do we give equal weight to each root?
-This is probably a choice at the end which matters a little bit, and depends on the prior choice, but we should clear the conceptual confusion.
-% Perhaps we can put a heuristic experiment result in appendix: we randomly sample the possible root for each posterior, then use that as the result, and we do it multiple times to show the result doesn't change dramatically.
-% Prior weight across posterior samples may help to alleviate the issue.}
-% \kw{The most robust thing one can do is sample the prior in the evolationary parameter space, pipe it through BSE then the GW PE code.
-I think it is still subject to the same multiple root problem, and N-M problem will still be there.
-% So in a sense, the only prior we have to take care is really the GW prior.}
-% \kw{Mention how we generate the initial guesses.}
 
@@
 
-We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+\begin{table*}[h!]
+    \begin{center}
+    \begin{tabular}{ l l l }
+    \hline
+    \hline
+    Parameters &  Description & Optimization range\\
+    \hline
+    \hline
+    Observables &\ &\  \\
+    \hline
+    \hline
+    $M_{\rm GW,1}$ & primary mass of the GW event & NA \\
+    $M_{\rm GW,2}$ & secondary mass of the GW event  & NA\\
+    \hline
+    \hline
+    Progenitor parameters &\ &\  \\
+    \hline
+    \hline
+    $M_{\rm ZAMS,1}$ & primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $M_{\rm ZAMS,2}$ & secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $t_{\rm orb}$ & orbital period at ZAMS & $[5,5000]$ days\\
+    $e$ & eccentricity at ZAMS & $[0,1]$\\
+    $Z$ & metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
+    $z$ & redshift at ZAMS & NA\\
+    \hline
+    \hline
+    Hyper-parameters &\ &\ \\
+    \hline
+    \hline
+
+    $\alpha$ & common envelope efficiency & $[0.1,10]$\\
+    $f_{\rm acc}$ & fraction of accretion during stable mass transfer & $[0,1]$\\
+    $q_{\rm crit,3}$ & & $[0.1,10]$\\
+    $\sigma$ & mean natal kick magnitude & $[10,300]\ {\rm km/h}$\\
+
+    \hline
+    \hline
+    \end{tabular}
+    \caption{A list of parameters we used in this study.}
+    \label{tab:parameters}
+    \end{center}
+\end{table*}
+
+We apply the method described in section \ref{sec:method} to the first GW events GW150914.
+The parameters involved in the analysis are tabulated in table \ref{tab:parameters}.
+
+and we show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+
 \kw{Recover ratio of the posterior sample goes here.}
@@
 
-\begin{figure*}
-\includegraphics[width=0.98\textwidth]{static/GW150914_corner.pdf}
+\begin{figure*}[h!]
+\includegraphics[width=0.98\textwidth]{static/GW150914_corner_zoomed.pdf}
 \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
-$M_1$ and $M_2$ are the progenitors' masses. $\log{t_{orb}}, e, \log{Z}$ are the progenitor binary's log orbital period, eccentricity and log metallicity at ZAMS.
+$M_1$ and $M_2$ are the progenitors' masses. $\log{Z}$ is log metallicity at ZAMS.
+$z$ is the redshift at ZAMS. 
 $\alpha$ is the common envelope efficiency.
 $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-$q_{\rm crit 3,4}$ are the critical mass during \kw{Katie fill in},
-$\sigma$ is the mean magnitude of the Maxwellian distribution characterizing natal kick strength.
+$q_{\rm crit 3}$ are the critical mass during \kw{Katie fill in},
+Note that the redshift is not fitted during the root finding process.
+Once we have find the evolutionary parameters, we add the time it would take the progenitor to merge to the lookback time of the observed posterior sample,
+then from the total lookback time we can compute the redshift at ZAMS.
 }

### 2022-03-23 src/ms.tex
Giving up on formatting. Moving on to editing
The magic is hbt!

@@
 \newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+% \setlength{\floatsep}{0cm}
+% \setlength{\textfloatsep}{0cm}
+% \setlength{\belowcaptionskip}{-5pt}
+% \setlength{\abovecaptionskip}{-5pt}
 % Begin!
@@
 
-\begin{table*}[h!]
+\begin{table*}[hbt!]
     \begin{center}
@@
     \hline
-    $M_{\rm GW,1}$ & primary mass of the GW event & NA \\
-    $M_{\rm GW,2}$ & secondary mass of the GW event  & NA\\
+    $M_{\rm GW,1}$ & Primary mass of the GW event & NA \\
+    $M_{\rm GW,2}$ & Secondary mass of the GW event  & NA\\
     \hline
@@
     \hline
-    $M_{\rm ZAMS,1}$ & primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
-    $M_{\rm ZAMS,2}$ & secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
-    $t_{\rm orb}$ & orbital period at ZAMS & $[5,5000]$ days\\
-    $e$ & eccentricity at ZAMS & $[0,1]$\\
-    $Z$ & metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
-    $z$ & redshift at ZAMS & NA\\
+    $M_{\rm ZAMS,1}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $M_{\rm ZAMS,2}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
+    $e$ & Eccentricity at ZAMS & $[0,1]$\\
+    $Z$ & Metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
+    $z$ & Redshift at ZAMS & NA\\
     \hline
@@
 
-    $\alpha$ & common envelope efficiency & $[0.1,10]$\\
-    $f_{\rm acc}$ & fraction of accretion during stable mass transfer & $[0,1]$\\
+    $\alpha$ & Common envelope efficiency & $[0.1,10]$\\
+    $f_{\rm acc}$ & Fraction of accretion during stable mass transfer & $[0,1]$\\
     $q_{\rm crit,3}$ & & $[0.1,10]$\\
-    $\sigma$ & mean natal kick magnitude & $[10,300]\ {\rm km/h}$\\
+    $\sigma$ & & $[10,300]\ {\rm km/h}$\\
 
@@
 We apply the method described in section \ref{sec:method} to the first GW events GW150914.
-The parameters involved in the analysis are tabulated in table \ref{tab:parameters}.
-
-and we show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+The parameters involved in the analysis and the range we allow for finding the root are tabulated in table \ref{tab:parameters}.
+\kw{@Katie state why not spin here.}
+Therefore, we solve the rootfinding problem with only the two component masses in the observable space.
+For progenitor parameters, we characterize each guess with five parameters: the two component masses $M_{\rm ZAMS,1}$ and $M_{\rm ZAMS,2}$, the orbital period $t_{\rm orb}$, the eccentricity $e$, the metallicity $Z$ at ZAMS.
+Note that the progenitor formation redshift is not obtained through the root-finding process.
+This is because the formation redshift does not affect the evolution of the binary, so we do not need to include in the optimization.
+Instead, we compute the time it takes for the binary to evolve from formation to the merger, then we add this time back to the lookback time when this event is observed.
+Using the total lookback time, we can compute the redshift of the binary at formation.
+We adopt the cosmological model in \kw{cite Planck 18} to compute the redshift of the binary at formation.
+For hyper-parameters, we choose parameters that affect binary evolution in \texttt{COSMIC} mostly significantly, including the common envelope efficiency $\alpha$, the fraction of accretion during stable mass transfer $f_{\rm acc}$, the critical mass fraction $q_{\rm crit,3}$, and the accretion rate $\sigma$.
+We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
 
@@
 
-\begin{figure*}[h!]
-\includegraphics[width=0.98\textwidth]{static/GW150914_corner_zoomed.pdf}
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed.pdf}
 \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
@@
 \includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
-\caption{}
+\caption{The posterior samples of GW150914 in the progenitor masses space, colorcoded by the common envelope efficiency $\alpha$ for that sample.}
 \label{fig:GW150914_m1_m2_alpha}

### 2022-03-28 src/ms.tex
Will minor comment on introduction.

@@
 \newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
 % \setlength{\floatsep}{0cm}
@@
 While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
-it is not necessarily the most efficient way to constraint the underlying physical model.
+it is not necessarily the most efficient way to constrain the underlying physical model.
 Most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
@@
 The problem worsen as we increase the dimensionality of the parameter space.
+\wf{This says what you want to say, but it's a bit wordy right now---I'll return to it, but it would be good to tighten up the language a lot.}
 

### 2022-03-29 src/ms.tex
WIP: edits to Methods Section

@@
 the population synthesis code is a deterministic function that transforms the properties of the progenitors into the GW parameters.
-Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\theta}'$ is given by
+Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\Theta}$ is given by
 
@@
 
-where $\bm{f}$ represents the population synthesis code.
+where $\bm{f}$ represents the evolution encoded by the population synthesis model with fixed random parameters $\bm{X}$.
 

### 2022-04-12 src/ms.tex
Add MCMC paragraph in discussion

--- 2022-03-23 src/ms.tex
@@
 A varying KL divergence means this particular event is subjected to randomness in the evolution function, hence one should take extra caution in interpreting the result.
+
+To ensure
 
@@
 This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
-In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root,
-and we have checked the posterior samples produced by this method for events discussed in this paper agree with direct sampling using a uniform prior.
-This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
+In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.
+On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
+The two results agree well with each other.
+This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.
+The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
+We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
+We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
+This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
 We leave a detail comparison between this method and direct sampling to future studies.

### 2022-04-13 src/ms.tex

--- 2022-03-29 src/ms.tex
@@
 We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.
+
+\section{Method Rewrite}
+\label{sec:method-rewrite}
+
+\wf{Re-writing the methods section with a new viewpoint, gonna try to eventually merge the content.}
+
+We define the parameters that describe the progenitor's properties such as ZAMS
+masses and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
+that control different physical prescriptions such as wind strength and common
+envelope efficiency as hyper-parameters $\bm{\lambda}$, and random variables
+that affect certain stochastic process such as whether natal kick unbinds a
+binary as $\bm{X}$. To avoid clutter in the following derivation, we denote all
+the parameters related to mapping a particular progenitor system into a GW event
+collectively as evolutionary parameters $\bm{\Theta}$, which includes
+$\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.
+
+Once all the parameters including a particular draw of all the random variables
+represented by $\bm{X}$ are known, the population synthesis code is a
+deterministic function that transforms the properties of the progenitors into
+the GW parameters $\bm{\theta}$:
+\begin{equation}
+    \bm{\theta} = \bm{f}\left( \bm{\Theta} \right).
+\end{equation}
+The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ is many-to-one.  
+In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
+need to be able to evaluate the likelihood of that data at fixed progenitor
+parameters, 
+\begin{equation}
+    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right),
+\end{equation}
+because the likelihood depends only on the gravitational waveform generated by
+parameters $\bm{\theta}$ to which the progenitor parameters map.  In principle
+this likelihood could be computed at arbitrary vaues of $\bm{\Theta}$ using the
+same machinery that is used to estimate source parameters in gravitational wave
+catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
+of progenitor parameters, hyper-parameters, and random variables in
+$\bm{\Theta}$ from a posterior density.
+
+However, it is very expensive to couple progenitor evolution to stochastic
+sampling in this way.  Because the gravitational wave likelihood function
+ultimately depends only on the GW parameters $\theta$, and catalogs have already
+produced samples over these parameters from a posterior density 
+\begin{equation}
+    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) p_\mathrm{GW} \left( \bm{\theta} \right),
+\end{equation}
+where $p_\mathrm{GW}$ is the prior density used for sampling, it is much more
+efficient to \wf{WIP}.
 
@@
 
+\bibliography{bib}
+
 \end{document}

### 2022-04-13 src/ms.tex
Move root finding around

@@
 
+On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
+The two results agree well with each other.
+This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.less
 \section{Result}
@@
 This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
-In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root,
-and we have checked the posterior samples produced by this method for events discussed in this paper agree with direct sampling using a uniform prior.
-This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
+In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.
+
+The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
+We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
+We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
+This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
 We leave a detail comparison between this method and direct sampling to future studies.

### 2022-04-14 src/ms.tex
Update first paragraph

@@
 
-As gravitational-wave (GW) detectors sensitivity improve, the rapid increase of GW detectors shifts out 
-\kw{think about the first sentence.}
-
-A very common way to study the origin of detected GW events is through forward modelling of GW progenitor.
-Forward modelling of a binary system start with a set of initial binary parameters,
-then evolve it according to the physics of interest until some stopping criteria is met.
+As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{},
+understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
+A very common way to study the origin of detected GW events is through population synthesis simulation \kw{cite},
+which evolves a population of binaries to form a population of GW events.
+Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
+such as the initial mass function or star-formation history,
+then evolve each individual members according to the physics of interest until some stopping criteria is met.
 This usually mean either a merger event or the binary system stays stable for a Hubble time.
+After this, one can collect all the GW emitting GW events in the synthetic population and compare to the observed population.
 
@@
     $f_{\rm acc}$ & Fraction of accretion during stable mass transfer & $[0,1]$\\
-    $q_{\rm crit,3}$ & & $[0.1,10]$\\
-    $\sigma$ & & $[10,300]\ {\rm km/h}$\\
+    $q_{\rm crit,3}$ & \kw{Katie fill in} & $[0.1,10]$\\
+    $\sigma$ & \kw{Katie fill in}& $[10,300]\ {\rm km/h}$\\
 

### 2022-04-14 src/ms.tex
Commit before reworking

@@
 This usually mean either a merger event or the binary system stays stable for a Hubble time.
-After this, one can collect all the GW emitting GW events in the synthetic population and compare to the observed population.
+After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
 
+\kw{Shorten}
 While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,

### 2022-04-16 src/ms.tex
update

@@
 \end{equation}
-The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ is many-to-one.  
+The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one.  
 In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
 need to be able to evaluate the likelihood of that data at fixed progenitor
-parameters, 
+parameters, namely
 \begin{equation}
-    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right),
+    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right).
 \end{equation}
-because the likelihood depends only on the gravitational waveform generated by
-parameters $\bm{\theta}$ to which the progenitor parameters map.  In principle
-this likelihood could be computed at arbitrary vaues of $\bm{\Theta}$ using the
+The equality holds because the likelihood depends only on the gravitational waveform generated by
+parameters $\bm{\theta}$ to which the progenitor parameters map onto.  In principle
+this likelihood could be computed at arbitrary values of $\bm{\Theta}$ using the
 same machinery that is used to estimate source parameters in gravitational wave
@@
 
-However, it is very expensive to couple progenitor evolution to stochastic
-sampling in this way.  Because the gravitational wave likelihood function
-ultimately depends only on the GW parameters $\theta$, and catalogs have already
+\kw{Talk about the MCMC}
+
+Since the gravitational wave likelihood function
+ultimately depends only on the GW parameters $\theta$, and GWTC3 have already
 produced samples over these parameters from a posterior density 
 \begin{equation}
-    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) p_\mathrm{GW} \left( \bm{\theta} \right),
+    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) \pi_\mathrm{GW} \left( \bm{\theta} \right),
 \end{equation}
-where $p_\mathrm{GW}$ is the prior density used for sampling, it is much more
-efficient to \wf{WIP}.
+where $\pi_\mathrm{GW}$ is the prior density used for sampling,
+we can save the computation cost in evaluating the GW likelihood by rewriting the posterior density of $\Theta$ in terms of the posterior density of $\theta$, namely
+
+\begin{align}
+    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta)}.
+\end{align}
+
+Including hyper-parameters and random variables in the sampling drastically increase the dimensionality of the problem,
+which makes the sampling process more computationally expensive and harder to converge.
+To speed up the convergence, we
 

### 2022-04-18 src/ms.tex
add abstract

@@
 \begin{abstract}
-In this
+Population synthesis simulation is the main workhorse for extracting physical information from a population of gravitational wave (GW) observations.
+Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimenionsnailty.
+In addition, it is very common to assume a particular distribution for "hyper-parameters" for the entire population (often a fixed value), such as common envelope efficiency.
+This limits the flexibility of the model hence potential to discover complex relations between hyper-parameters.
+In this work, we introduce a new paradiagm to extract physical information from GW observations.
+We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters for each observation in GWTC3.
+We showcase the effectiveness of this pipeline with GW150914, as well as its flexibility in studying a population of GW observations. 
 \end{abstract}
@@
 
-\kw{Talk about the MCMC}
-
 Since the gravitational wave likelihood function
-ultimately depends only on the GW parameters $\theta$, and GWTC3 have already
-produced samples over these parameters from a posterior density 
+ultimately depends only on the GW parameters $\theta$, and GWTC3 already has samples over these parameters from a posterior density 
 \begin{equation}
@@
 \begin{align}
-    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta)}.
+    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta)},
 \end{align}
+which we applies a kernel density estimator to the posterior samples released in GWTC3 to estimate $p(\theta(\Theta)|d)$.
 
@@
 which makes the sampling process more computationally expensive and harder to converge.
-To speed up the convergence, we
+To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density,
+then we use those points to initialize a set of chains in the MCMC process.
 
-\section{Method}
-\label{sec:method}
-
-We define the parameters that describe the progenitor's properties such as ZAMS masses and eccentricity as progenitor parameters $\bm{\theta'}$,
-the parameters that control different physical prescriptions such as wind strength and common envelope efficiency as hyper-parameters $\bm{\lambda}$,
-and random variables that affect certain stochastic process such as whether natal kick unbinds a binary as $\bm{X}$.
-To avoid clutter in the following derivation,
-we denote all the parameters related to mapping a particular progenitor system into a GW event collectively as evolutionary parameters $\bm{\Theta}$,
-which includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.
-
-Once all the parameters including a particular draw of all the random variables are known,
-the population synthesis code is a deterministic function that transforms the properties of the progenitors into the GW parameters.
-Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\Theta}$ is given by
-
-\begin{align}
-    p(\bm{\theta}|\bm{\Theta}) &= \bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta}),
-\end{align}
-
-where $\bm{f}$ represents the evolution encoded by the population synthesis model with fixed random parameters $\bm{X}$.
-
-While it is straightforward to write down the "forward" probability, the "backward" probability is more complicated since multiple evolutionary parameters can produce the same GW event.
-Using Bayes theorem, the probability of an event with evolutionary parameters given the GW parameters can be derived as following
-
-\begin{align}
-    p(\bm{\Theta}|\bm{\theta}) &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{p(\bm{\theta})}, \nonumber \\
-    &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{\int p(\bm{\theta}|\bm{\Theta}) p(\bm{\Theta}) d\bm{\Theta}}, \nonumber \\
-    &= \frac{\bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}{\int d\bm{\theta} \bm{\delta}(f(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}.
-\label{eq:backward_prob}
-\end{align}
-
-Using the result in eq.\ref{eq:backward_prob},
-the probability of an event having evolutionary parameters given data can be obtained by marginalizing over the GW parameters
-
-\begin{align}
-    p(\bm{\Theta}|d) &= \int p(\bm{\Theta}|\bm{\theta}) p(\bm{\theta}|d) d\bm{\theta}\\
-    &= \frac{1}{N}\sum_{i=1}^{N} p(\bm{\Theta}|\bm{\theta}_i). \label{eq:evolutionary_posterior}
-\end{align}
-
-The term $p(\bm{\theta}|d)$ is the inferred posterior probability for a particular event, which is usually presented by the LVK collaboration in the form of posterior samples.
-In eq.\ref{eq:evolutionary_posterior}, we use importance sampling to turn the integration into a sum over the posterior samples.
-
-For each posterior sample point, we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.
+The root-finding procedure is described as followed:
+For each posterior sample point in the GW observables space,
+we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.
 
@@
 
-On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
-The two results agree well with each other.
-This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.less
 \section{Result}
@@
 
+\kw{Add population figure?}
+
 \section{Discussion}
@@
 Instead of forward modelling an assumed distribution of initial binary to the observed population,
-we solve a root-finding problem to find the best initial binary parameters that reproduce the posterior distribution for each event. 
+we find the corresponding evolutionary parameters for each GW events. 
 The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
@@
 
-In this work we showcase the power of the proposed method with two events, GW150914 and GW191129.
+In this work we showcase the power of the proposed method with GW150914.
 We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
@@
 Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
-This introduces significant inefficiency in our root-finding algorithm.
+This introduces significant inefficiency in our root-finding and sampling algorithm.
 The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolution pathway of low mass events such as BNS events.
 The effect of natal kick is suppressed for heavier mass events due to fallback.
-This means events with the root-finding process for lighter masses are subject to stochasticity of the function,
-where the root-finding process for heavier events behaves as if the evolutionary function we use are deterministic.
-Indeed, we see this behavior in a couple of case studies.
-For heavy event such as GW150914, the reprojected posterior agrees with the data posterior.
-For light event such as GW170817, although during the root-finding process the convergence criteria is met,
-the reprojected posterior does not agree with the data posterior.
-Also, due to computational limitation, we only try 1000 different initial guesses per posterior sample in the root-finding process.
+This means events with lighter masses are subject to stochasticity of the function,
+where the sampling process for heavier events behaves as if the evolutionary function we use are deterministic.
+% Indeed, we see this behavior in a couple of case studies.
+% For heavy event such as GW150914, the reprojected posterior agrees with the data posterior.
+% For light event such as GW170817, although during the root-finding process the convergence criteria is met,
+% the reprojected posterior does not agree with the data posterior.
+Due to computational limitation, we only try 1000 different initial guesses per posterior sample in the root-finding process.
 This means any posterior sample that has a probability of merging rarer than 1 in 1000 could be missed.
-\kw{comment on the low mass ratio end.}
 Obviously the problems which comes with the randomness can be alleviated by performing more tries per posterior sample,
@@
 
-By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
-In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
-In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
-This greatly benefit the efficiency of the root-finding process.
-On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
-One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
-Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
-the relatively weighting of these roots becomes ambiguous when the number of roots is large.
-For example, should we consider roots that are close to each other a single root or multiple roots?
-If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
-Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
-This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
-In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.
+% By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
+% In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
+% In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
+% This greatly benefit the efficiency of the root-finding process.
+% On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
+% One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
+% Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
+% the relatively weighting of these roots becomes ambiguous when the number of roots is large.
+% For example, should we consider roots that are close to each other a single root or multiple roots?
+% If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
+% Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
+% This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
+% In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.
 
-The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
-We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
-We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
-This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
-We leave a detail comparison between this method and direct sampling to future studies.
+% The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
+% We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
+% We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
+% This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
+% We leave a detail comparison between this method and direct sampling to future studies.
 
@@
 To improve the accuracy and efficiency in estimating the gradient of objective function, automatic differentiation is a promising feature that modeler should consider incorporating in their population synthesis code in the future.
-
-\kw{Mention potential complimentary studies of direct sampling, and how all these suggestions are still useful regardless of whether we are doing root-finding or not.}
 

### 2022-04-19 src/ms.tex
update

@@
 Population synthesis simulation is the main workhorse for extracting physical information from a population of gravitational wave (GW) observations.
-Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimenionsnailty.
+Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimensionality.
 In addition, it is very common to assume a particular distribution for "hyper-parameters" for the entire population (often a fixed value), such as common envelope efficiency.
@@
 
-As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{},
+As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{cite},
 understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
@@
 
-\section{Method Rewrite}
-\label{sec:method-rewrite}
+\section{Method}
+\label{sec:method}
 
@@
 \begin{align}
-    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta)},
+    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta(\Theta))},
 \end{align}
@@
 \label{sec:result}
-
-\begin{enumerate}
-\item Raw posterior of successful events.
-\item Trace back the event's $z_{\rm ZAMS}$.
-\item Reproject to observable space
-\item Evidence ratio?
-\item KL divergence?
-\end{enumerate}
 
@@
 we find the corresponding evolutionary parameters for each GW events. 
+In this work we first showcase the power of the proposed method with GW150914.
+We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
+In particular, to the authors' knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is explored.
+Our work is not only a more efficient way to study GW events' progenitors,
+but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.
+
+To further illustrate the power of the method, we show the inferred common envelope efficiency depends on the progenitor masses.
 The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
@@
 without making specific assumptions such as fixing the hyper-parameters for all types of event.
-
-In this work we showcase the power of the proposed method with GW150914.
-We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
-In particular, to the authors knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is shown.
-Our work is not only a more efficient way to study GW events' progenitors,
-but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.
 
@@
 In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
-\kw{Talk about prior are now in post-processing instead of pre-processing.}
 
@@
 
-% Interpretation:
-% \begin{enumerate}
-% \item Highlight the fact that we turn hyper-parameters into parameters.
-% \item Discuss this is need for high dimensional exploration.
-% \end{enumerate}
-
-% Caveats: 
-% \begin{enumerate}
-% \item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
-% \item We only assume BSE
-% \item Caveat on marginalizing over random samples.
-% \item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
-% \item How next generation "Progenitor synthesis" code should be built.
-% \item Rate
-% \item Failure studies go into appendix
-% \end{enumerate}
-
 \section{Acknowledgement}

### 2022-04-19 src/ms.tex
update

@@
 While our work lay down a data analysis pathway to understand the population of GW events,
-no physics can be learned without a comprehensive complimentary physical model.
+no physics can be learned without a comprehensive physical model.
 We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:

### 2022-04-20 src/ms.tex
Remove some comments
update

@@
 \label{sec:method}
-
-\wf{Re-writing the methods section with a new viewpoint, gonna try to eventually merge the content.}
 
@@
 \end{figure}
-
 We can see the reprojected posterior agrees very well with the original posterior in fig.\ref{fig:GW150914_reprojection}.
-\kw{Qoute the KL divergence.}
 
@@
 
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{static/GWTC3_alpha_mass.pdf}
+\caption{This figure is ugly as fucked now. Let's change it in later iteration}
+\label{fig:GWTC3_alpha_mass}
+\end{figure}
+
+\kw{Talk about implication on single event posterior.}
 The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
@@
 \kw{Dartboard comparsion} Comparing our result with
-\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}
 
-\kw{Add population figure?}
+\kw{Discuss implication in population.}
+To illustrate the potential benefit of our method in the population, we perform the same analysis for most of the events in GWTC3.
+
+Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint of $f_{\rm acc}$ for all the events included in the analysis sorted by the median of their inferred primary mass in the observable space.
+Note that some events on the extreme of the mass spectrum did not pass the KL divergence test we proposed in section \ref{sec:method}.
+\kw{Double check}On the low mass end, the events are more subjected to random fluctuation because natal kick is less modulated by fallback in lighter system, so the reprojected 
+On the high mass end, \texttt{COSMIC} struggles to produce events in the pair instability supernova mass gap,
+so the posterior in the evolutionary parameter space only correspond part of the posterior in the observable space, therefore events beyond the PISN mass gap has higher KL divergence.
+
+The point of figure \ref{fig:GWTC3_alpha_mass} is to show this method could reveal the correlation between the primary mass in the BBH and the common envelope efficiency.
+A careful treatment of each event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_alpha_mass} is beyond the scope of this paper,
+therefore we defer a detail study of the physics related to the population of GW events in the future.
 

### 2022-04-21 src/ms.tex
update

@@
 \email{kwong@flatironinstitute.org}
+\flatiron
+
+\author{Katelyn Breivik}
+\flatiron
+
+\author{Will Farr}
 \flatiron
@@
 
-\kw{Shorten}
 While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
 it is not necessarily the most efficient way to constrain the underlying physical model.
-Most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
-This means if we want to evolve a binary to match the observation by pure trial and error, a large fraction of computation will be wasted in trying initial binary configurations that are not related to the observation.
-To alleviate this inefficiency, the community often takes a population approach --- instead of trying to match binary systems to observations event by event,
-one can draw systems from an initial population, forward model each member in that population then match it to the observed population properties.
-However, this approach is essentially trading one inefficiency with another, since we still need to choose an initial distribution of binary parameters.
-For relatively well-known source properties such as initial masses this is less of a problem,
-but the distribution of some less well-studied properties such as common envelope efficiency is often heuristically chosen to be some simple distribution such as a delta function at some value, i.e.
-a fixed value for the entire population.
-This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution with other independent studies,
-or they should vary the family of distributions.
-In general, it is unclear how one can cover all possible distribution that may agree with the observation in generating a suite of simulations.
-The problem worsen as we increase the dimensionality of the parameter space.
-\wf{This says what you want to say, but it's a bit wordy right now---I'll return to it, but it would be good to tighten up the language a lot.}
-
-Indeed, one of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
-i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
+One of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
+i.e. only vary a number of hyper-parameters (e.g. common envelope efficiency) while holding all others fixed to some "default" values.
 Even with recently development on using emulators to speed up the simulation process, training the emulators also require a decent amount of samples in the parameter space of interest.
@@
 The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
-This means even training an emulator is impossible in high dimensional space.
+This means even training an emulator is impossible in a high dimensional space.
+To make the problem even worse, there is no gaurantee the entire population should share the same hyper-parameters (i.e. a delta distribution at a fixed value.),
+which means there could be a large (possibly infinite) family of hyper-parameters distribution one need to test to compare the simulations to the observed data.
 
-In this paper, instead of forward modelling a population of binary star to match the observed GW events population,
-we propose a method to "backward" model each GW event to its progenitor state.
-The idea is since we can compute the GW event properties of a guess by forward modelling its progenitor,
-we can define the difference between the GW event properties of this guess and the observed GW event properties.
-By minimizing this difference, i.e. solving a root finding problem, we can hope to recover the progenitor properties.
-
+However, this issue is not a fundamental issue, but an apparent obstacle because we are approach the problem from an infinitely inefficient angle.
+Since each binary system could in principle come from a different environment, hence they can have different progenitor parameters as well as hyper-parameters.
+What one should do is to ask what is the posterior density in the progenitor and hyper-parameter joint space for each individual binary system,
+then perform population analysis in that space.
+In this paper, we propose a method to "backward" model each GW event to its progenitor state.
+The method is a two stages process:
+first we solve a root finding problem to obtain guesses that are likely to produce the desire system properties in the joint space,
+then we sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the root found in the previous stage.
 This paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.

### 2022-04-24 src/ms.tex
update

@@
 This limits the flexibility of the model hence potential to discover complex relations between hyper-parameters.
-In this work, we introduce a new paradiagm to extract physical information from GW observations.
+In this work, we introduce a new paradigm to extract physical information from GW observations.
 We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters for each observation in GWTC3.
@@
 We apply the method described in section \ref{sec:method} to the first GW events GW150914.
-The parameters involved in the analysis and the range we allow for finding the root are tabulated in table \ref{tab:parameters}.
+The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
 \kw{@Katie state why not spin here.}
-Therefore, we solve the rootfinding problem with only the two component masses in the observable space.
+Therefore, we only consider the two component masses in the observable space in the both the root-finding stage and the MCMC stage..
 For progenitor parameters, we characterize each guess with five parameters: the two component masses $M_{\rm ZAMS,1}$ and $M_{\rm ZAMS,2}$, the orbital period $t_{\rm orb}$, the eccentricity $e$, the metallicity $Z$ at ZAMS.
 Note that the progenitor formation redshift is not obtained through the root-finding process.
-This is because the formation redshift does not affect the evolution of the binary, so we do not need to include in the optimization.
+This is because the formation redshift does not affect the evolution of the binary explicitly.
 Instead, we compute the time it takes for the binary to evolve from formation to the merger, then we add this time back to the lookback time when this event is observed.
@@
 We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
-
-\kw{Recover ratio of the posterior sample goes here.}
-
-The full posterior is available in appendix \kw{Or we should put a link to the dataset?}
 
@@
 The KL divergence stays stable for reprojecting with different random seeds.
+In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior agrees very well with the original posterior.
+Compared to the result shown in a previous study, our result covers almost all the posterior of GW150914 in the masses space,
+which means our result is more consistent with the observation.
+\kw{Do we want to draw comparison with Jeff's result?}
 
@@
 \end{figure}
-We can see the reprojected posterior agrees very well with the original posterior in fig.\ref{fig:GW150914_reprojection}.
-
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{static/GW150914_CE_SMT.pdf}
-% \caption{}
-% \label{fig:GW150914_CE_SMT}
-% \end{figure}
-% \kw{Add figure of stable mass transfer vs common envelope}
 
@@
 \includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
-\caption{The posterior samples of GW150914 in the progenitor masses space, colorcoded by the common envelope efficiency $\alpha$ for that sample.}
+\caption{The posterior samples of GW150914 in the progenitor masses space, color-coded by the common envelope efficiency $\alpha$ for that sample.}
 \label{fig:GW150914_m1_m2_alpha}
@@
 \includegraphics[width=0.49\textwidth]{static/GWTC3_alpha_mass.pdf}
-\caption{This figure is ugly as fucked now. Let's change it in later iteration}
+\caption{This figure is ugly as fucked now. Let's change it in later iteration.
+Also replace only with events that pass the KL divergence test.}
 \label{fig:GWTC3_alpha_mass}
@@
 
-\kw{Talk about implication on single event posterior.}
-The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
+% \kw{Talk about implication on single event posterior.}
+Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
 We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
@@
 we would miss the samples that do not match the chosen value of $\alpha$.
-By folding in the hyper-parameters into the root-finding algorithm,
-we can capture the correlation between 
-\kw{Dartboard comparsion} Comparing our result with
+Also, our method shows it is capable to find mode that is hard to find by previous methods.
+\kw{Talk about the high mass branch that is missed by Jeff's paper}
 
-\kw{Discuss implication in population.}
-To illustrate the potential benefit of our method in the population, we perform the same analysis for most of the events in GWTC3.
+% \kw{Discuss implication in population.}
+To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
+Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint on $f_{\rm acc}$ for events passing the KL divergence test sorted by the median of their primary mass in the observable space.
+We can see $f_{\rm acc}$ increases as the primary mass of the event increases.
+This means the stable mass transfer (SMT) phase of a light binary is preferably non-conservative, whereas the heavier the events are, the more likely their SMT is conservative.
 
-Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint of $f_{\rm acc}$ for all the events included in the analysis sorted by the median of their inferred primary mass in the observable space.
-Note that some events on the extreme of the mass spectrum did not pass the KL divergence test we proposed in section \ref{sec:method}.
-\kw{Double check}On the low mass end, the events are more subjected to random fluctuation because natal kick is less modulated by fallback in lighter system, so the reprojected 
-On the high mass end, \texttt{COSMIC} struggles to produce events in the pair instability supernova mass gap,
+Note that some events did not pass the KL divergence test we proposed in section \ref{sec:method}, which we do not include in figure \ref{fig:GWTC3_alpha_mass}.
+Low mass events are more subjected to random fluctuation because natal kick is less modulated by fallback in lighter system,
+so the reprojected posterior may not agree with the original posterior in the observable space consistently when varying the random seed.
+On the other hand, \texttt{COSMIC} struggles to produce events above the pair instability supernova mass cutoff \kw{State cutoff},
 so the posterior in the evolutionary parameter space only correspond part of the posterior in the observable space, therefore events beyond the PISN mass gap has higher KL divergence.
+Events with more extreme mass ratios are not likely to be recovered by the method. \kw{@Katie gives reason?}
 
-The point of figure \ref{fig:GWTC3_alpha_mass} is to show this method could reveal the correlation between the primary mass in the BBH and the common envelope efficiency.
-A careful treatment of each event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_alpha_mass} is beyond the scope of this paper,
+The point of figure \ref{fig:GWTC3_alpha_mass} is to show this method could in principal reveal the correlation between progenitor parameters and hyper-parameters on a population level.
+A careful treatment of all event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_alpha_mass} is beyond the scope of this paper,
 therefore we defer a detail study of the physics related to the population of GW events in the future.
+Also, despite some events do not pass the KL divergence test, we can still see the correlation in the region where the events are neither affected by randomness nor the PISN mass gap.
 
@@
 
-To further illustrate the power of the method, we show the inferred common envelope efficiency depends on the progenitor masses.
+To further illustrate the power of the method, we show the inferred fraction of accretion during SMT depends on the progenitor masses.
 The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
@@
 In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
+This also means one can study the progenitor-formation rate in a data-driven manner instead of assuming a particular model.
 

### 2022-04-26 src/ms.tex
add annotated corner plot
reworking a bit of the introduction

@@
 \newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
+\newcommand{\kb}[1]{\textcolor{pink}{KB: #1}}
 % \setlength{\floatsep}{0cm}
@@
 Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimensionality. 
-In addition, it is very common to assume a particular distribution for "hyper-parameters" for the entire population (often a fixed value), such as common envelope efficiency.
-This limits the flexibility of the model hence potential to discover complex relations between hyper-parameters.
+\kb{I think we need to substantiate the previous sentence with citations that show this.}
+In addition, it is very common to assume a \kb{set of assumptions, or "hyper-parameters", for how uncertain binary interactions proceed }
+for the entire population (often a fixed value), such as common envelope efficiency.
+This limits the flexibility of the model \kb{and hence the potential} to discover complex relations between hyper-parameters.
 In this work, we introduce a new paradigm to extract physical information from GW observations.
-We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters for each observation in GWTC3.
+We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters 
+\kb{associated with uncertainties in binary evolution} for each observation in GWTC3.
 We showcase the effectiveness of this pipeline with GW150914, as well as its flexibility in studying a population of GW observations. 
@@
 \section{Introduction}
+{\kb{As the detection rate of gravitation waves (GWs) from merging double-compact-object (DCO) binaries increases 
+with the sensitivty of the ground-based GW detector network, the ability to disentangle the astrophysical processes 
+which shape the evolution of GW progenitor populations is becoming increasingly attainable. One of the most common ways
+to study progenitor populations of GW mergers is through population synthesis simulations for a host of formation environments
+including isolated binary stars, triple systems, globular clusters, nuclear star clusters, AGN disks, and more 
+cite Mandel \& Broekgaarden. In these types of studies, the population is forward modeled based on random samples from
+empirically or theoretically motivated distributions of initial parameters which define the age, metallicity, and intrinsic 
+properties of the population like masses, orbital separations, and eccentricites. These initial populations are then evolved
+using a set of assumptions for how astrophysical processes impact the evolution of the DCO merger progenitors, ultimately producing
+a synthetic catalog of mergers which can be compared of observations from ground-based GW detectors.
+}}
 
-As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{cite},
-understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
-A very common way to study the origin of detected GW events is through population synthesis simulation \kw{cite},
-which evolves a population of binaries to form a population of GW events.
-Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
-such as the initial mass function or star-formation history,
-then evolve each individual members according to the physics of interest until some stopping criteria is met.
-This usually mean either a merger event or the binary system stays stable for a Hubble time.
-After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
+%As the rate of detection of gravitational-wave (GW) increases with the sensitivity of the \kb{ground-based} GW detector network \kw{cite},
+%understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}. 
+%\kb{I think we can reword the end of this previous sentence -- I should come back to it.}
+%A very common way to study the origin of detected GW events is through population synthesis simulations \kw{cite} \kb{cite Mandel \& Broekgaarden review},
+%which evolve populations of various kinds of binaries to form populations of GW events. \kb{Do we want to do isolated only here? or other channels?}
+%\kb{The process to forward model a population of GW events relies on random samples }
+%Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
+%such as the initial mass function or star-formation history,
+%then evolve each individual members according to the physics of interest until some stopping criteria is met.
+%This usually mean either a merger event or the binary system stays stable for a Hubble time.
+%After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
 
@@
 
+\begin{figure*}
+    \includegraphics[width=0.98\textwidth]{static/compare_fixed_variable.png}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
+     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
+     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
+    \label{fig:compare_fixed_variable}
+    \end{figure*}
+
 \begin{figure}

### 2022-04-26 src/ms.tex
updates to section 2

@@
 
-While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
-it is not necessarily the most efficient way to constrain the underlying physical model.
-One of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
-i.e. only vary a number of hyper-parameters (e.g. common envelope efficiency) while holding all others fixed to some "default" values.
-Even with recently development on using emulators to speed up the simulation process, training the emulators also require a decent amount of samples in the parameter space of interest.
-For example, say we want to emulate a model with 10 parameters, this simplest option for constructing a bank of simulations for training is to lay down a grid.
-Along each axis we put the barely minimum of $2$ simulations, this would still require $1024$ simulations.
-The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
-This means even training an emulator is impossible in a high dimensional space.
-To make the problem even worse, there is no gaurantee the entire population should share the same hyper-parameters (i.e. a delta distribution at a fixed value.),
-which means there could be a large (possibly infinite) family of hyper-parameters distribution one need to test to compare the simulations to the observed data.
+While forward modelling \kb{entire populations} has been the canonical pathway to \kb{compare simulations and GW observations,}%simulate binary systems and compare to GW observation,
+it is not necessarily the most efficient way to constrain the underlying physical \kb{processes, or "hyper parameters", which drive the evolution of the population}.
+One of the most common skepticisms of population synthesis studies \kb{is that these studies employ} simulation sets which often 
+cover a small subdomain of the entire parameter space \kb{of uncertain physics, or only vary one uncertain process per simulation}
+%i.e. only vary a number of hyper-parameters 
+(e.g. common envelope efficiency) while holding all others fixed to some "default" values \kb{which have not been calibrated to data.}
+Even with recent developments in using emulators to speed up the simulation process, training the emulators 
+\kb{still requires a significant amount of computation to span the wide range of uncertainty.} %require a decent amount of samples in the parameter space of interest.
+\kb{As an} example, \kb{to} emulate a model with $10$ parameters, the simplest \kb{way to construct a training set of simulations is to }
+\kb{run a grid which varies combinations of uncertain physics.} %option for constructing a bank of simulations for training is to lay down a grid.
+\kb{For 10 uncertain physical processes, if we consider $2$ variations }%each axis we put the barely minimum of $2$ simulations, 
+this would still require $1024$ simulations.
+{}\kb{Because population sythesis simulations require hundreds to thousands of cpu hours, training an emulator which spans 
+a high dimensional uncertainty space remains an impossibility at present.}}%The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
+%This means even training an emulator is impossible in a high dimensional space.
+%To make the problem even worse, 
+\kb{This compounds with the fact that} there is no gaurantee the entire population should share the same fixed hyper parameters, 
+\kb{as is often assumed in most population synthesis studies.}
+% (i.e. a delta distribution at a fixed value.),
+%which means 
+\kb{This suggests that} there could be a large (possibly infinite) family of hyper-parameter distributions %one need to test to compare the simulations to the observed data.
+\kb{required to compare data to simulations which span the full set of uncertain physical models.}
+However, this issue is not a fundamental issue, but an apparent obstacle %because 
+\kb{which arises from an inneficient approach to the problem.} %we are approach the problem from an infinitely inefficient angle.
+Furthermore, each binary system could in principle come from a different formation environment and hence %they can 
+have different progenitor parameters as well as hyper parameters \kb{which define uncertain physics associated with each environment}.
+{\kb{In this case, it is advantageous to determine the posterier density in the progenitor and hyper-parameter joint space 
+for each individual binary system, then perform population in analyses in that space.}}
+%What one should do is to ask what is the posterior density in the progenitor and hyper-parameter joint space for each individual binary system,
+%then perform population analysis in that space.
+In this paper, we propose a method to "backward" model each GW event to its progenitor state 
+\kb{while allowing the hyper parameters to vary across the full range of physical uncertainty}
+following a two-stage process.
+First, we solve a root finding problem to obtain guesses that are likely to produce the desired system properties in the joint progenitor--hyper-parameter space.
+We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
 
-However, this issue is not a fundamental issue, but an apparent obstacle because we are approach the problem from an infinitely inefficient angle.
-Since each binary system could in principle come from a different environment, hence they can have different progenitor parameters as well as hyper-parameters.
-What one should do is to ask what is the posterior density in the progenitor and hyper-parameter joint space for each individual binary system,
-then perform population analysis in that space.
-In this paper, we propose a method to "backward" model each GW event to its progenitor state.
-The method is a two stages process:
-first we solve a root finding problem to obtain guesses that are likely to produce the desire system properties in the joint space,
-then we sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the root found in the previous stage.
-This paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
+The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
 In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
@@
 
-We define the parameters that describe the progenitor's properties such as ZAMS
+We define the parameters that describe the progenitor's zero age main sequence (ZAMS) properties
 masses and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
-that control different physical prescriptions such as wind strength and common
+that control different uncertain physical prescriptions such as wind strength and common
 envelope efficiency as hyper-parameters $\bm{\lambda}$, and random variables
-that affect certain stochastic process such as whether natal kick unbinds a
+that affect certain stochastic process such as whether CO natal kick unbinds a
 binary as $\bm{X}$. To avoid clutter in the following derivation, we denote all
@@
 \end{equation}
-The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one.  
-In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
-need to be able to evaluate the likelihood of that data at fixed progenitor
+The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one
+{\kb{due to degeneracies in the different physical processes and initial parameters  
+of the GW event progenitors}}.
+In order to draw inferences about $\bm{\Theta}$ from GW data, we
+must be able to evaluate the likelihood of that data at fixed progenitor
 parameters, namely
@@
 this likelihood could be computed at arbitrary values of $\bm{\Theta}$ using the
-same machinery that is used to estimate source parameters in gravitational wave
+same machinery that is used to estimate source parameters in GW
 catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
@@
 
-Since the gravitational wave likelihood function
+Since the GW likelihood function
 ultimately depends only on the GW parameters $\theta$, and GWTC3 already has samples over these parameters from a posterior density 
@@
 where $\pi_\mathrm{GW}$ is the prior density used for sampling,
-we can save the computation cost in evaluating the GW likelihood by rewriting the posterior density of $\Theta$ in terms of the posterior density of $\theta$, namely
+we can save the computation cost associated with evaluating the GW likelihood by rewriting the posterior 
+density of $\Theta$ in terms of the posterior density of $\theta$, namely
 
@@
 
-Including hyper-parameters and random variables in the sampling drastically increase the dimensionality of the problem,
-which makes the sampling process more computationally expensive and harder to converge.
-To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density,
+Including hyper-parameters and random variables %in the sampling 
+drastically increases the dimensionality of the problem,
+which makes the sampling process more computationally expensive and \kb{thus more} to converge.
+To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density of a GW event,
 then we use those points to initialize a set of chains in the MCMC process.
 
-The root-finding procedure is described as followed:
-For each posterior sample point in the GW observables space,
-we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.
+%The root-finding procedure is described as follows:
+For each posterior sample point in the GW-observable space,
+we can find the corresponding evolutionary parameters by solving an optimization problem that minimizes 
+the mean square difference between the GW event parameters and evolutionary parameters as:
 
 \begin{align}
-\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2
+\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2.
 \label{eq:loss}
@@
 
-In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the observable space.
-However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$ to be our threshold.
-To make sure we find a reasonably complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
-As long as the solution fulfill the acceptance criteria, the root is as valid as all other roots that fulfill the same acceptance criteria, regardless to its initial guess.
-In other word, we have the freedom to choose the set of initial guesses however we think it might benefit the optimization.
-Most of the points in the evolutionary parameter space do not produce mergers, and for systems that do not merge, we set the binary masses to be 0.
-This means the gradient for systems that do not merge is 0, and the root finder would likely to get stuck in such state during the first step.\footnote{On a plateau of constant loss, the root finder performs no better than a random guess.
-In a high dimensional space, this means the root finder will need a long time to get close to a solution.}
-Therefore, it is beneficial to put initial guesses in region of the evolutionary parameter space that does produce merger.
-To construct a list of initial guesses, we evolve a set of points that we uniformly sampled from the evolutionary parameter space, then keep the points that merge within a Hubble time.
+In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the GW-observable space.
+However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.\ref{eq:loss} 
+to a small acceptance threshold. \kb{For this study, we picked a threshold of $10^{-2}$}. To make sure we find a reasonably 
+complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses 
+in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all 
+other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
+\kb{This means}, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
+
+Most of the points in the evolutionary parameter space do not produce DCO mergers in a Hubble time. For systems that do not merge, 
+we set the binary masses to be $0$ such that the gradient for the same systems is also $0$ and leads the root finder to get stuck on 
+its first step.{\footnote{On a plateau of constant loss, the root finder performs no better than a random guess. 
+In a high dimensional space, this means the root finder will need a long time to get close to a solution.}}
+Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter space that \kb{are very likely to} 
+produce DCO mergers. To construct a list of initial guesses, we evolve a set of \kb{ZAMS parameters} uniformly sampled from the 
+\kb{initial binary} parameter space, then keep the points that merge within a Hubble time.
 

### 2022-04-27 src/ms.tex
Paginated abstract.
Added a bunch of detector references.

--- 2022-04-24 src/ms.tex
@@
 \newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
+\newcommand{\stonybrook}{\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, USA}}
 \usepackage{amsmath}
@@
 
-\author{Will Farr}
+\author{Will M. Farr}
+\email{will.farr@stonybrook.edu}
 \flatiron
+\stonybrook
 
@@
 \begin{abstract}
-Population synthesis simulation is the main workhorse for extracting physical information from a population of gravitational wave (GW) observations.
-Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimensionality.
-In addition, it is very common to assume a particular distribution for "hyper-parameters" for the entire population (often a fixed value), such as common envelope efficiency.
-This limits the flexibility of the model hence potential to discover complex relations between hyper-parameters.
-In this work, we introduce a new paradigm to extract physical information from GW observations.
-We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters for each observation in GWTC3.
-We showcase the effectiveness of this pipeline with GW150914, as well as its flexibility in studying a population of GW observations. 
+Population synthesis simulation is the main workhorse for extracting physical
+information from a population of gravitational wave (GW) observations.
+Unfortunately, current pathways for comparing simulations to observations are
+not scalable to high dimensionality. In addition, it is very common to assume a
+particular distribution for "hyper-parameters" for the entire population (often
+a fixed value), such as common envelope efficiency. This limits the flexibility
+of the model hence potential to discover complex relations between
+hyper-parameters. In this work, we introduce a new paradigm to extract physical
+information from GW observations. We propose an inference pipeline to jointly
+infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and
+hyper-parameters for each observation in GWTC3. We showcase the effectiveness of
+this pipeline with GW150914, as well as its flexibility in studying a population
+of GW observations. 
 \end{abstract}
@@
 
-As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{cite},
-understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
-A very common way to study the origin of detected GW events is through population synthesis simulation \kw{cite},
-which evolves a population of binaries to form a population of GW events.
-Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
-such as the initial mass function or star-formation history,
-then evolve each individual members according to the physics of interest until some stopping criteria is met.
-This usually mean either a merger event or the binary system stays stable for a Hubble time.
-After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
+As the rate of detection of gravitational-wave (GW) increase with the
+sensitivity of the global GW detector network
+\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
+understanding astrophysical processes related to progenitors of GW events
+through analyzing a population of GW events has become more popular \kw{cite}. A
+very common way to study the origin of detected GW events is through population
+synthesis simulation \kw{cite}, which evolves a population of binaries to form a
+population of GW events. Forward modelling of a population of GW events start
+with sampling a set of initial binary parameters from some distributions related
+to the progenitor population, such as the initial mass function or
+star-formation history, then evolve each individual members according to the
+physics of interest until some stopping criteria is met. This usually mean
+either a merger event or the binary system stays stable for a Hubble time. After
+this, one can collect all the GW emitting binaries in the synthetic population
+and compare to the observed population.
 
@@
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed.pdf}
+\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed_annotated.pdf}
 \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.

### 2022-04-29 src/ms.tex

--- 2022-04-26 src/ms.tex
@@
 \newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
-\newcommand{\kb}[1]{\textcolor{pink}{KB: #1}}
+\newcommand{\kb}[1]{\textcolor{pink}{#1}}
 % \setlength{\floatsep}{0cm}
@@
 \kb{run a grid which varies combinations of uncertain physics.} %option for constructing a bank of simulations for training is to lay down a grid.
-\kb{For 10 uncertain physical processes, if we consider $2$ variations }%each axis we put the barely minimum of $2$ simulations, 
+\kb{For 10 uncertain physical processes, if we consider $2$ variations} %each axis we put the barely minimum of $2$ simulations, 
 this would still require $1024$ simulations.
-{}\kb{Because population sythesis simulations require hundreds to thousands of cpu hours, training an emulator which spans 
-a high dimensional uncertainty space remains an impossibility at present.}}%The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
+\kb{Because population sythesis simulations require hundreds to thousands of cpu hours, }
+\kb{training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.}
+
+%The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
 %This means even training an emulator is impossible in a high dimensional space.
 %To make the problem even worse,
+
 \kb{This compounds with the fact that} there is no gaurantee the entire population should share the same fixed hyper parameters, 
@@
 {\kb{In this case, it is advantageous to determine the posterier density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space.}}
+for each individual binary system, then perform population in analyses in that space. \citet{Andrews2021} (herafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
+to deterine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
+set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
+\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.}}
+
 %What one should do is to ask what is the posterior density in the progenitor and hyper-parameter joint space for each individual binary system,
@@
 In this paper, we propose a method to "backward" model each GW event to its progenitor state 
-\kb{while allowing the hyper parameters to vary across the full range of physical uncertainty}
+\kb{while allowing the hyper parameters to vary across their full range of physical uncertainty}
 following a two-stage process.
@@
 In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the GW-observable space.
-However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.\ref{eq:loss} 
+However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.~\ref{eq:loss} 
 to a small acceptance threshold. \kb{For this study, we picked a threshold of $10^{-2}$}. To make sure we find a reasonably 
@@
 its first step.{\footnote{On a plateau of constant loss, the root finder performs no better than a random guess. 
-In a high dimensional space, this means the root finder will need a long time to get close to a solution.}}
-Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter space that \kb{are very likely to} 
-produce DCO mergers. To construct a list of initial guesses, we evolve a set of \kb{ZAMS parameters} uniformly sampled from the 
-\kb{initial binary} parameter space, then keep the points that merge within a Hubble time.
+Because of this, a root finder in a high dimensional space will require a significant amount of computation
+to get close to a solution.}} Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter 
+space that \kb{are very likely to} produce DCO mergers. To construct a list of initial guesses, we evolve a set of 
+\kb{ZAMS parameters} uniformly sampled from the \kb{initial binary} parameter space, then keep the systems that 
+merge within a Hubble time.
 
-Ideally we would like to have explicit control over all evolutionary parameters.
-In particular, having explicit control over random variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
-However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicitly,
-i.e. the random variables are drawn within the program, instead of being arguments of the program.
-This means the function we use to evolve the binary is not a deterministic function but a stochastic function instead. 
-Hence, even if the root-finding algorithm performs perfectly,
-forward modelling this set of roots does not guarantee to reproduce the set of posterior samples due to randomness of the evolution function.
-To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,
-we forward model ("Reproject") the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
-We use KL divergence to measure the agreement between the two posterior distributions, which is given by
+%Ideally, we would like to have explicit control over all evolutionary parameters.
+\kb{Retaining} explicit control over random variables allows us to marginalize over their contribution, 
+\kb{and thus allows us to} focus on progenitor parameters and hyper-parameters.
+However, in practice most population synthesis codes (including \texttt{COSMIC} which we use in this study) 
+define their random variables implicitly \kb{such that} the random variables are drawn within the program instead of 
+being \kb{passed as arguments.} This means the function we use to evolve the binary is not fully deterministic. % function but a stochastic function instead. 
+\kb{Thus}, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not 
+guarantee that the \kb{simulated population} reproduces \kb{exactly} the set of posterior samples due to 
+randomness in the evolution \kb{of each binary}. To assure \kb{that} the recovered progenitor and hyper-parameters 
+robustly correspond to the posterior in the GW-observable space, we forward model (or "reproject") the recovered 
+evolutionary parameters to the GW-observable space to check whether the reprojected posterior agrees 
+with the posterior given by the LVK collaboration. We use KL divergence to measure the agreement between the two posterior 
+distributions as:
 \begin{align}
@@
 
-A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the observable space, hence it is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
-This is completely expected since \texttt{COSMIC} comes with its own assumptions and is expected to fail in reproducing a subset of the event in GWTC3, such as event with at least one component beyond the pair instability supernova massgap.
-On the other hand, the reprojection could subject to stochasticity in the evolution function we use.
-Instead of reprojecting the posterior only once, we reproject the posterior multiple times with different random seeds and check whether the KL divergence varies significantly.
-A varying KL divergence means this particular event is subjected to randomness in the evolution function, hence one should take extra caution in interpreting the result.
+A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the 
+observable space, \kb{and is thus} a viable channel for that specific event. Otherwise it either means \texttt{COSMIC} 
+can only explain part of the posterior or cannot explain the event at all. This is completely expected behaviour since 
+\texttt{COSMIC}, \kb{and isolated binary evolution more generally}, carries its own assumptions and is expected to 
+fail in reproducing a subset of the event in GWTC3. \kb{One example of this are GW events} with at least one component 
+\kb{with masses above the} the pair instability supernova mass limit \kb{cite Farmer, and others}.
+\kb{The reprojection could also be} subject to stochasticity in the evolution \kb{of each binary}.
+\kb{To account for this}, we reproject the posterior multiple times with different random seeds and 
+check whether the KL divergence varies significantly. A varying KL divergence means a particular GW event is subject 
+to randomness in the evolution \kb{of the progenitor}, and extra caution should be used when interpreting the result.
 
@@
     $f_{\rm acc}$ & Fraction of accretion during stable mass transfer & $[0,1]$\\
-    $q_{\rm crit,3}$ & \kw{Katie fill in} & $[0.1,10]$\\
-    $\sigma$ & \kw{Katie fill in}& $[10,300]\ {\rm km/h}$\\
+    $q_{\rm crit,3}$ & Critical mass ratio on the Hertzsprung Gap & $[0.1,10]$\\
+    $\sigma$ & Root mean square of Maxwellian natal kick distribution& $[10,300]\ {\rm km/s}$\\
 
@@
 
-We apply the method described in section \ref{sec:method} to the first GW events GW150914.
+We apply the method described in section \ref{sec:method} to the first GW event GW150914.
 The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-\kw{@Katie state why not spin here.}
-Therefore, we only consider the two component masses in the observable space in the both the root-finding stage and the MCMC stage..
-For progenitor parameters, we characterize each guess with five parameters: the two component masses $M_{\rm ZAMS,1}$ and $M_{\rm ZAMS,2}$, the orbital period $t_{\rm orb}$, the eccentricity $e$, the metallicity $Z$ at ZAMS.
-Note that the progenitor formation redshift is not obtained through the root-finding process.
-This is because the formation redshift does not affect the evolution of the binary explicitly.
-Instead, we compute the time it takes for the binary to evolve from formation to the merger, then we add this time back to the lookback time when this event is observed.
-Using the total lookback time, we can compute the redshift of the binary at formation.
-We adopt the cosmological model in \kw{cite Planck 18} to compute the redshift of the binary at formation.
-For hyper-parameters, we choose parameters that affect binary evolution in \texttt{COSMIC} mostly significantly, including the common envelope efficiency $\alpha$, the fraction of accretion during stable mass transfer $f_{\rm acc}$, the critical mass fraction $q_{\rm crit,3}$, and the accretion rate $\sigma$.
-We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+{\kb{The spin of BHs formed in isolated binaries depends strongly on the angular 
+momentum transport in massive stars which is still highly uncertain (see however Bavera+, fuller+),}}
+therefore we only consider the two component masses in the observable space in the both the 
+root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with 
+five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$, 
+the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not 
+obtained through the root-finding process because the formation redshift does not \kb{intrinsically} affect 
+the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from 
+formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total 
+time to merger to by less than $13.7\,\rm{Gyr}$.
+Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
+\kw{cite Planck 18} cosmological model. For hyper-parameters, we choose parameters that affect 
+binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 
+efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$, 
+the critical mass ratio which determines whether Roche-overflow mass transfer 
+proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap 
+$q_{\rm crit,3}$, and the root mean square of the Maxwellian distribution
+employed for CO natal kicks, $\sigma$.
+
+{\kb{Once we obtain a set of binaries which successfuly map the ZAMS parameters and hyper parameters to BBH merger masses, 
+we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
+to explore how keeping a fixed model which only varies one hyper parameter affects our analysis. 
+Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
+each with a different $\alpha$ but the same hyper parameters as A21, 
+as well as the results of our sampling which allows our hyper parameters to vary (fourth column). We find that holding the accretion 
+efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
+that produces GW150914-like mergers. In contrast, by allowing the hyper parameters to fully span the model uncertainty, we 
+find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
+fail to produce GW150914-like mergers \emph{regardless of our hyper parameter choice.}}}
+
+\kb{Need to describe the formation channels and gaps here.}
+
+\begin{figure*}
+    \includegraphics[width=0.98\textwidth]{static/compare_fixed_variable.png}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
+     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
+     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
+    \label{fig:compare_fixed_variable}
+\end{figure*}
+
+We show a portion of the joint posterior of the progenitor parameters 
+and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+\kb{We need to describe the interpretation of this as well.}
 
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed.pdf}
+\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed_annotated.pdf}
 \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
@@
 Note that the redshift is not fitted during the root finding process.
-Once we have find the evolutionary parameters, we add the time it would take the progenitor to merge to the lookback time of the observed posterior sample,
+Once we have find the evolutionary parameters, we add the time it would take the progenitor 
+to merge to the lookback time of the observed posterior sample,
 then from the total lookback time we can compute the redshift at ZAMS.
@@
 The KL divergence stays stable for reprojecting with different random seeds. 
+\kb{Can we show this in the figure somehow? Maybe like KL divergence as a function of reprojection number as a lower small panel?}
 In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior agrees very well with the original posterior.
@@
 \kw{Do we want to draw comparison with Jeff's result?}
+\kb{Yes -- but I think we need to point to Jeff's result in the intro as well.}
 
@@
 \end{figure}
-
-\begin{figure*}
-    \includegraphics[width=0.98\textwidth]{static/compare_fixed_variable.png}
-    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
-     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
-     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
-    \label{fig:compare_fixed_variable}
-    \end{figure*}
 

### 2022-05-04 src/ms.tex
Add new figure 1
Commented on comments

@@
 Population synthesis simulation is the main workhorse for extracting physical information from a population of gravitational wave (GW) observations.
-Unfortunately, current pathways for comparing simulations to observations are not scalable to high dimensionality. 
-\kb{I think we need to substantiate the previous sentence with citations that show this.}
-In addition, it is very common to assume a \kb{set of assumptions, or "hyper-parameters", for how uncertain binary interactions proceed }
+\kw{Unfortunately, it is difficult to follow current pathways for comparing simulations to observations when handling problems with a high dimensionality.}
+\kb{I think we need to substantiate the previous sentence with citations that show this.}\kw{We do that in the introduction, is that enough? I reword the statement to be softer.}
+In addition, it is very common to assume a set of assumptions, or "hyper-parameters", for how uncertain binary interactions proceed
 for the entire population (often a fixed value), such as common envelope efficiency.
-This limits the flexibility of the model \kb{and hence the potential} to discover complex relations between hyper-parameters.
+This limits the flexibility of the model and hence the potential to discover complex relations between hyper-parameters.
 In this work, we introduce a new paradigm to extract physical information from GW observations.
 We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters 
-\kb{associated with uncertainties in binary evolution} for each observation in GWTC3.
+\kb{associated with uncertainties in binary evolution} \kw{I don't think we need this addition?} for each observation in GWTC3.
 We showcase the effectiveness of this pipeline with GW150914, as well as its flexibility in studying a population of GW observations. 
@@
 \section{Introduction}
-{\kb{As the detection rate of gravitation waves (GWs) from merging double-compact-object (DCO) binaries increases 
-with the sensitivty of the ground-based GW detector network, the ability to disentangle the astrophysical processes 
+As the detection rate of gravitation waves (GWs) from merging double-compact-object (DCO) binaries increases 
+with the sensitivity of the ground-based GW detector network, the ability to disentangle the astrophysical processes 
 which shape the evolution of GW progenitor populations is becoming increasingly attainable. One of the most common ways
@@
 empirically or theoretically motivated distributions of initial parameters which define the age, metallicity, and intrinsic 
-properties of the population like masses, orbital separations, and eccentricites. These initial populations are then evolved
+properties of the population like masses, orbital separations, and eccentricities. These initial populations are then evolved
 using a set of assumptions for how astrophysical processes impact the evolution of the DCO merger progenitors, ultimately producing
 a synthetic catalog of mergers which can be compared of observations from ground-based GW detectors.
-}}
 
@@
 
-While forward modelling \kb{entire populations} has been the canonical pathway to \kb{compare simulations and GW observations,}%simulate binary systems and compare to GW observation,
-it is not necessarily the most efficient way to constrain the underlying physical \kb{processes, or "hyper parameters", which drive the evolution of the population}.
-One of the most common skepticisms of population synthesis studies \kb{is that these studies employ} simulation sets which often 
-cover a small subdomain of the entire parameter space \kb{of uncertain physics, or only vary one uncertain process per simulation}
-%i.e. only vary a number of hyper-parameters 
-(e.g. common envelope efficiency) while holding all others fixed to some "default" values \kb{which have not been calibrated to data.}
+While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
+it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyper-parameters", which drive the evolution of the population.
+One of the most common skepticism of population synthesis studies is that these studies employ simulation sets which often 
+cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
+(e.g. common envelope efficiency) while holding all others fixed to some "default" values.
+On top of this, these values are often not calibrated to data.
 Even with recent developments in using emulators to speed up the simulation process, training the emulators 
-\kb{still requires a significant amount of computation to span the wide range of uncertainty.} %require a decent amount of samples in the parameter space of interest.
-\kb{As an} example, \kb{to} emulate a model with $10$ parameters, the simplest \kb{way to construct a training set of simulations is to }
-\kb{run a grid which varies combinations of uncertain physics.} %option for constructing a bank of simulations for training is to lay down a grid.
-\kb{For 10 uncertain physical processes, if we consider $2$ variations} %each axis we put the barely minimum of $2$ simulations, 
+still requires a significant amount of computation to cover a wide range of uncertainty.
+As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
+run a grid which varies combinations of uncertain physics.
+For 10 uncertain physical processes, if we consider $2$ variations for each physical process 
 this would still require $1024$ simulations.
-\kb{Because population sythesis simulations require hundreds to thousands of cpu hours, }
-\kb{training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.}
+Because population synthesis simulations require hundreds to thousands of cpu hours,
+training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
 
-%The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
-%This means even training an emulator is impossible in a high dimensional space.
-%To make the problem even worse,
+\kw{Reorganizing discussion related to Jeff's work such that it doesn't seem like our method is just an extension of DartBoard.}
 
-\kb{This compounds with the fact that} there is no gaurantee the entire population should share the same fixed hyper parameters, 
-\kb{as is often assumed in most population synthesis studies.}
-% (i.e. a delta distribution at a fixed value.),
-%which means 
-\kb{This suggests that} there could be a large (possibly infinite) family of hyper-parameter distributions %one need to test to compare the simulations to the observed data.
-\kb{required to compare data to simulations which span the full set of uncertain physical models.}
-However, this issue is not a fundamental issue, but an apparent obstacle %because 
-\kb{which arises from an inneficient approach to the problem.} %we are approach the problem from an infinitely inefficient angle.
-Furthermore, each binary system could in principle come from a different formation environment and hence %they can 
-have different progenitor parameters as well as hyper parameters \kb{which define uncertain physics associated with each environment}.
-{\kb{In this case, it is advantageous to determine the posterier density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space. \citet{Andrews2021} (herafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
-to deterine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
-set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
-\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.}}
+This compounds with the fact that there is no guarantee the entire population should share the same fixed hyper-parameters, 
+as is often assumed in most population synthesis studies.
+This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
+required to compare data to simulations which span the full set of uncertain physical models.
+However, this issue is not a fundamental issue, but an apparent obstacle which arises from an inefficient approach to the problem.
+Furthermore, each binary system could in principle come from a different formation environment and hence 
+have different progenitor parameters as well as hyper-parameters, which define uncertain physics associated with each environment.
+In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
+for each individual binary system, then perform population in analyses in that space.
+\citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
+to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
+set of hyper-parameters which define assumptions for how isolated binary-star interactions proceed using 
+\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
+However, the method was limited to a single event instead of the population.
+More importantly, the method did not vary the hyper-parameters while sampling the posterior density of the ZAMS parameters,
+which limit its capability to be applied to understand a population of binary systems.
 
-%What one should do is to ask what is the posterior density in the progenitor and hyper-parameter joint space for each individual binary system,
-%then perform population analysis in that space.
 In this paper, we propose a method to "backward" model each GW event to its progenitor state 
-\kb{while allowing the hyper parameters to vary across their full range of physical uncertainty}
+while allowing the hyper-parameters to vary across their full range of physical uncertainty
 following a two-stage process.
@@
 We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
-
 The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
@@
 The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one
-{\kb{due to degeneracies in the different physical processes and initial parameters  
-of the GW event progenitors}}.
+due to degeneracies in the different physical processes and initial parameters  
+of the GW event progenitors.
 In order to draw inferences about $\bm{\Theta}$ from GW data, we
@@
 \end{align}
-which we applies a kernel density estimator to the posterior samples released in GWTC3 to estimate $p(\theta(\Theta)|d)$.
+which we apply a kernel density estimator to the posterior samples released in GWTC3 to estimate $p(\theta(\Theta)|d)$.
 
@@
 drastically increases the dimensionality of the problem,
-which makes the sampling process more computationally expensive and \kb{thus more} to converge.
+which makes the sampling process more computationally expensive to converge.
 To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density of a GW event,
@@
 However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.~\ref{eq:loss} 
-to a small acceptance threshold. \kb{For this study, we picked a threshold of $10^{-2}$}. To make sure we find a reasonably 
+to a small acceptance threshold. For this study, we picked a threshold of $10^{-2}$. To make sure we find a reasonably 
 complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses 
@@
 other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
-\kb{This means}, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
+This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
 
@@
 to get close to a solution.}} Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter 
-space that \kb{are very likely to} produce DCO mergers. To construct a list of initial guesses, we evolve a set of 
-\kb{ZAMS parameters} uniformly sampled from the \kb{initial binary} parameter space, then keep the systems that 
+space that are likely to produce DCO mergers. To construct a list of initial guesses, we evolve a set of 
+ZAMS parameters uniformly sampled from the initial binary parameter space, then keep the systems that 
 merge within a Hubble time.
 
-%Ideally, we would like to have explicit control over all evolutionary parameters.
-\kb{Retaining} explicit control over random variables allows us to marginalize over their contribution, 
-\kb{and thus allows us to} focus on progenitor parameters and hyper-parameters.
+Retaining explicit control over random variables allows us to marginalize over their contribution, 
+and thus allows us to focus on progenitor parameters and hyper-parameters.
 However, in practice most population synthesis codes (including \texttt{COSMIC} which we use in this study) 
-define their random variables implicitly \kb{such that} the random variables are drawn within the program instead of 
-being \kb{passed as arguments.} This means the function we use to evolve the binary is not fully deterministic. % function but a stochastic function instead. 
-\kb{Thus}, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not 
-guarantee that the \kb{simulated population} reproduces \kb{exactly} the set of posterior samples due to 
-randomness in the evolution \kb{of each binary}. To assure \kb{that} the recovered progenitor and hyper-parameters 
+define their random variables implicitly such that the random variables are drawn within the program instead of 
+being passed as arguments. This means the function we use to evolve the binary is not fully deterministic.
+Thus, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not 
+guarantee that the simulated population reproduces exactly the set of posterior samples due to 
+randomness in the evolution of each binary. To assure that the recovered progenitor and hyper-parameters 
 robustly correspond to the posterior in the GW-observable space, we forward model (or "reproject") the recovered 
@@
 A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the 
-observable space, \kb{and is thus} a viable channel for that specific event. Otherwise it either means \texttt{COSMIC} 
-can only explain part of the posterior or cannot explain the event at all. This is completely expected behaviour since 
-\texttt{COSMIC}, \kb{and isolated binary evolution more generally}, carries its own assumptions and is expected to 
-fail in reproducing a subset of the event in GWTC3. \kb{One example of this are GW events} with at least one component 
-\kb{with masses above the} the pair instability supernova mass limit \kb{cite Farmer, and others}.
-\kb{The reprojection could also be} subject to stochasticity in the evolution \kb{of each binary}.
-\kb{To account for this}, we reproject the posterior multiple times with different random seeds and 
+observable space, and is thus a viable channel for that specific event. Otherwise, it either means \texttt{COSMIC} 
+can only explain part of the posterior or cannot explain the event at all. This is completely expected behavior since 
+\texttt{COSMIC}, and isolated binary evolution more generally, carries its own assumptions and is expected to 
+fail in reproducing a subset of the event in GWTC3. One example of this are GW events with at least one component 
+with mass above the pair instability supernova mass limit \kb{cite Farmer, and others}.
+The reprojection could also be subject to stochasticity in the evolution of each binary
+To account for this, we reproject the posterior multiple times with different random seeds and 
 check whether the KL divergence varies significantly. A varying KL divergence means a particular GW event is subject 
-to randomness in the evolution \kb{of the progenitor}, and extra caution should be used when interpreting the result.
+to randomness in the evolution of the progenitor, and extra caution should be used when interpreting the result.
 
@@
 The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-{\kb{The spin of BHs formed in isolated binaries depends strongly on the angular 
-momentum transport in massive stars which is still highly uncertain (see however Bavera+, fuller+),}}
+The spin of BHs formed in isolated binaries depends strongly on the angular 
+momentum transport in massive stars which is still highly uncertain \kb{(see however Bavera+, fuller+)},
 therefore we only consider the two component masses in the observable space in the both the 
@@
 the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not 
-obtained through the root-finding process because the formation redshift does not \kb{intrinsically} affect 
+obtained through the root-finding process because the formation redshift does not intrinsically affect 
 the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from 
@@
 
-{\kb{Once we obtain a set of binaries which successfuly map the ZAMS parameters and hyper parameters to BBH merger masses, 
+Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
 we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
@@
 Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
-each with a different $\alpha$ but the same hyper parameters as A21, 
-as well as the results of our sampling which allows our hyper parameters to vary (fourth column). We find that holding the accretion 
+each with a different $\alpha$ but the same hyper-parameters as A21, 
+as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
 efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
-that produces GW150914-like mergers. In contrast, by allowing the hyper parameters to fully span the model uncertainty, we 
+that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
 find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
-fail to produce GW150914-like mergers \emph{regardless of our hyper parameter choice.}}}
+fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
 
@@
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed_annotated.pdf}
+\includegraphics[width=\textwidth]{static/GW150914_corner_zoomed.png}
 \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
@@
 \kb{Can we show this in the figure somehow? Maybe like KL divergence as a function of reprojection number as a lower small panel?}
+\kw{The problem with that is it is basically one dot or a flat line, which is very unentertaining.
+I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
 In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior agrees very well with the original posterior.

### 2022-05-04 src/ms.tex
Reorganize result

@@
 proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap 
-$q_{\rm crit,3}$, and the root mean square of the Maxwellian distribution
+$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
 employed for CO natal kicks, $\sigma$.
-
-Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
-we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
-to explore how keeping a fixed model which only varies one hyper parameter affects our analysis. 
-Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
-each with a different $\alpha$ but the same hyper-parameters as A21, 
-as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
-efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
-that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
-find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
-fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
-
-\kb{Need to describe the formation channels and gaps here.}
-
-\begin{figure*}
-    \includegraphics[width=0.98\textwidth]{static/compare_fixed_variable.png}
-    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
-     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
-     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
-    \label{fig:compare_fixed_variable}
-\end{figure*}
 
@@
 
+% \kw{Talk about implication on single event posterior.}
+Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
+we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
+to explore how keeping a fixed model which only varies one hyper-parameter affects our analysis. 
+Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
+each with a different $\alpha$ but the same hyper-parameters as A21, 
+as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
+efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
+that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
+find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
+fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
+\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+
+\kb{Need to describe the formation channels and gaps here.}
+\kw{I think this can be very brief. Save the main explanation to the physics paper.}
+
+\begin{figure*}
+    \includegraphics[width=0.98\textwidth]{static/compare_fixed_variable.png}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
+     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
+     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
+    \label{fig:compare_fixed_variable}
+\end{figure*}
+
 To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
 We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
-For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} on average.
+In particular, we compute KL divergence of the samples obtained from root-finding alone and after MCMC sampling.
+For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later}logits on average if we use the root-finding result alone,
+whereas the MCMC result has a higher KL divergence at \kw{fill} logits.
 The KL divergence stays stable for reprojecting with different random seeds. 
@@
 I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
-In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior agrees very well with the original posterior.
+In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-fidining process alone and the MCMC result agrees well with the original posterior,
+even though the agreement between the MCMC result and the original posterior is not perfect. 
+This degradation of the agreement comes from the fact that we are using a KDE to estimate the posterior density in the observable space.
+Note that the result from root-finding alone and the result from MCMC agrees well in the evolutionary parameter space.
 Compared to the result shown in a previous study, our result covers almost all the posterior of GW150914 in the masses space,
 which means our result is more consistent with the observation.
-\kw{Do we want to draw comparison with Jeff's result?}
-\kb{Yes -- but I think we need to point to Jeff's result in the intro as well.}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
-\caption{The posterior samples of GW150914 in the progenitor masses space, color-coded by the common envelope efficiency $\alpha$ for that sample.}
-\label{fig:GW150914_m1_m2_alpha}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{static/GWTC3_alpha_mass.pdf}
-\caption{This figure is ugly as fucked now. Let's change it in later iteration.
-Also replace only with events that pass the KL divergence test.}
-\label{fig:GWTC3_alpha_mass}
-\end{figure}
-
-% \kw{Talk about implication on single event posterior.}
-Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
-We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
-this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
-we would miss the samples that do not match the chosen value of $\alpha$.
-Also, our method shows it is capable to find mode that is hard to find by previous methods.
-\kw{Talk about the high mass branch that is missed by Jeff's paper}
-
-% \kw{Discuss implication in population.}
 To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
@@
 Also, despite some events do not pass the KL divergence test, we can still see the correlation in the region where the events are neither affected by randomness nor the PISN mass gap.
+
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{static/GWTC3_alpha_mass.pdf}
+\caption{This figure is ugly as fucked now. Let's change it in later iteration.
+Also replace only with events that pass the KL divergence test.}
+\label{fig:GWTC3_alpha_mass}
+\end{figure}
+
+% Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
+% We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
+% this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
+% we would miss the samples that do not match the chosen value of $\alpha$.
+% Also, our method shows it is capable to find mode that is hard to find by previous methods.
+% \kw{Talk about the high mass branch that is missed by Jeff's paper}
+
+% \kw{Discuss implication in population.}
 

### 2022-05-04 src/ms.tex
update

@@
 \kb{We need to describe the interpretation of this as well.}
+\kw{I think we can just hightlight the fact that we have three kinds of correlation, progenitor-hyper, progenitor-hyper, hyper-hyper,
+since the implication is related to figure @Katie made.}
+\kb{Need to describe the formation channels and gaps here.}
 
@@
 
-\kb{Need to describe the formation channels and gaps here.}
-\kw{I think this can be very brief. Save the main explanation to the physics paper.}
-
 \begin{figure*}
@@
 In particular, we compute KL divergence of the samples obtained from root-finding alone and after MCMC sampling.
-For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later}logits on average if we use the root-finding result alone,
-whereas the MCMC result has a higher KL divergence at \kw{fill} logits.
+For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later}nat on average if we use the root-finding result alone,
+whereas the MCMC result has a higher KL divergence at \kw{fill} nats.
 The KL divergence stays stable for reprojecting with different random seeds. 
@@
 \end{figure}
-
-% Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
-% We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
-% this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
-% we would miss the samples that do not match the chosen value of $\alpha$.
-% Also, our method shows it is capable to find mode that is hard to find by previous methods.
-% \kw{Talk about the high mass branch that is missed by Jeff's paper}
-
-% \kw{Discuss implication in population.}
 

### 2022-05-06 src/ms.tex

@@
 As the detection rate of gravitation waves (GWs) from merging double-compact-object (DCO) binaries increases 
-with the sensitivity of the ground-based GW detector network, the ability to disentangle the astrophysical processes 
+with the sensitivity of the ground-based GW detector network \citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021}, the ability to disentangle the astrophysical processes
 which shape the evolution of GW progenitor populations is becoming increasingly attainable. One of the most common ways

### 2022-05-07 src/ms.tex
Abstract improvements?

@@
 \begin{abstract}
-Population synthesis simulation is the main workhorse for extracting physical information from a population of gravitational wave (GW) observations.
-\kw{Unfortunately, it is difficult to follow current pathways for comparing simulations to observations when handling problems with a high dimensionality.}
-\kb{I think we need to substantiate the previous sentence with citations that show this.}\kw{We do that in the introduction, is that enough? I reword the statement to be softer.}
-In addition, it is very common to assume a set of assumptions, or "hyper-parameters", for how uncertain binary interactions proceed
-for the entire population (often a fixed value), such as common envelope efficiency.
-This limits the flexibility of the model and hence the potential to discover complex relations between hyper-parameters.
-In this work, we introduce a new paradigm to extract physical information from GW observations.
-We propose an inference pipeline to jointly infer the posterior density of both progenitor parameters (e.g. ZAMS masses) and hyper-parameters 
-\kb{associated with uncertainties in binary evolution} \kw{I don't think we need this addition?} for each observation in GWTC3.
-We showcase the effectiveness of this pipeline with GW150914, as well as its flexibility in studying a population of GW observations. 
+One promising way to extract information about stellar astrophysics from
+gravitational wave catalogs is to compare the catalog to the outputs of
+population synthesis modeling with varying physical assumptions.  The parameter
+space of physical assumptions in population synthesis is high-dimensional and
+the choice of parameters that best represents the evolution of a binary system
+may depend in an as-yet-to-be-determined way on the system's properties; indeed,
+measuring the variation of population synthesis parameter settings with binary
+properties is one of the principal ways to extract information about binary
+astrophysics from catalogs. Here we propose an inference pipeline to
+simultaneously infer the posterior density of stellar progenitor parameters
+(e.g. zero-age main sequence masses) and population synthesis parameter settings
+controlling modeled binary evolution from gravitational wave observations.  Our
+pipeline can efficiently explore the high-dimensional space of population
+synthesis settings and progenitor system properties for each system in a catalog
+of gravitational wave observations.  We apply our pipeline to observations in
+the third LIGO Virgo Kagra Gravitational Wave Transient Catalog (GWTC3).  We
+showcase the effectiveness of this pipeline with a detailed study of the
+progenitor properties and population synthesis settings that produce mergers
+like the observed GW150914.
 \end{abstract}

### 2022-05-07 src/ms.tex
Will starts to work on intro.

@@
 \section{Introduction}
-As the detection rate of gravitation waves (GWs) from merging double-compact-object (DCO) binaries increases 
-with the sensitivity of the ground-based GW detector network \citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021}, the ability to disentangle the astrophysical processes
-which shape the evolution of GW progenitor populations is becoming increasingly attainable. One of the most common ways
-to study progenitor populations of GW mergers is through population synthesis simulations for a host of formation environments
-including isolated binary stars, triple systems, globular clusters, nuclear star clusters, AGN disks, and more 
-cite Mandel \& Broekgaarden. In these types of studies, the population is forward modeled based on random samples from
-empirically or theoretically motivated distributions of initial parameters which define the age, metallicity, and intrinsic 
-properties of the population like masses, orbital separations, and eccentricities. These initial populations are then evolved
-using a set of assumptions for how astrophysical processes impact the evolution of the DCO merger progenitors, ultimately producing
-a synthetic catalog of mergers which can be compared of observations from ground-based GW detectors.
+As the detection rate of gravitation waves (GWs) from merging
+double-compact-object (DCO) binaries increases with the sensitivity of the
+ground-based GW detector network
+\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
+we are beginning to constrain the astrophysical processes which shape the
+evolution of GW progenitor populations. One of the most common ways to study
+progenitor populations of GW mergers is through population synthesis simulations
+for a host of formation environments including isolated binary stars, triple
+systems, globular clusters, nuclear star clusters, AGN disks, and more
+\citep{Mandel2022}. \wf{Moar citations here} In these types of studies, the
+population is forward modeled based on random samples from empirically or
+theoretically motivated distributions of initial parameters which define the
+age, metallicity, and intrinsic properties of the population like masses,
+orbital separations, and eccentricities. These initial populations are then
+evolved using a set of assumptions for how astrophysical processes impact the
+evolution of the DCO merger progenitors, ultimately producing a synthetic
+catalog of mergers which can be compared of observations from ground-based GW
+detectors.
 

### 2022-05-10 src/ms.tex
Change figures 4 to blob plot
Add text related to the new figure 4

@@
 While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyper-parameters", which drive the evolution of the population.
+it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyperparameters", which drive the evolution of the population.
 One of the most common skepticism of population synthesis studies is that these studies employ simulation sets which often 
@@
 
-This compounds with the fact that there is no guarantee the entire population should share the same fixed hyper-parameters, 
+This compounds with the fact that there is no guarantee the entire population should share the same fixed hyperparameters, 
 as is often assumed in most population synthesis studies.
@@
 Furthermore, each binary system could in principle come from a different formation environment and hence 
-have different progenitor parameters as well as hyper-parameters, which define uncertain physics associated with each environment.
+have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
 In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
@@
 to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
-set of hyper-parameters which define assumptions for how isolated binary-star interactions proceed using 
+set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
 \texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
 However, the method was limited to a single event instead of the population.
-More importantly, the method did not vary the hyper-parameters while sampling the posterior density of the ZAMS parameters,
+More importantly, the method did not vary the hyperparameters while sampling the posterior density of the ZAMS parameters,
 which limit its capability to be applied to understand a population of binary systems.
@@
 In this paper, we propose a method to "backward" model each GW event to its progenitor state 
-while allowing the hyper-parameters to vary across their full range of physical uncertainty
+while allowing the hyperparameters to vary across their full range of physical uncertainty
 following a two-stage process.
@@
 that control different uncertain physical prescriptions such as wind strength and common
-envelope efficiency as hyper-parameters $\bm{\lambda}$, and random variables
+envelope efficiency as hyperparameters $\bm{\lambda}$, and random variables
 that affect certain stochastic process such as whether CO natal kick unbinds a
@@
 catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
-of progenitor parameters, hyper-parameters, and random variables in
+of progenitor parameters, hyperparameters, and random variables in
 $\bm{\Theta}$ from a posterior density.
@@
 
-Including hyper-parameters and random variables %in the sampling 
+Including hyperparameters and random variables %in the sampling 
 drastically increases the dimensionality of the problem,
@@
 Retaining explicit control over random variables allows us to marginalize over their contribution, 
-and thus allows us to focus on progenitor parameters and hyper-parameters.
+and thus allows us to focus on progenitor parameters and hyperparameters.
 However, in practice most population synthesis codes (including \texttt{COSMIC} which we use in this study) 
@@
 guarantee that the simulated population reproduces exactly the set of posterior samples due to 
-randomness in the evolution of each binary. To assure that the recovered progenitor and hyper-parameters 
+randomness in the evolution of each binary. To assure that the recovered progenitor and hyperparameters 
 robustly correspond to the posterior in the GW-observable space, we forward model (or "reproject") the recovered 
@@
     \hline
-    $M_{\rm GW,1}$ & Primary mass of the GW event & NA \\
-    $M_{\rm GW,2}$ & Secondary mass of the GW event  & NA\\
+    $m_{\rm 1, GW}$ & Primary mass of the GW event & NA \\
+    $m_{\rm 2, GW}$ & Secondary mass of the GW event  & NA\\
     \hline
@@
     \hline
-    $M_{\rm ZAMS,1}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
-    $M_{\rm ZAMS,2}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $m_{\rm 1, ZAMS}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
+    $m_{\rm 2, ZAMS}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
     $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
@@
     \hline
-    Hyper-parameters &\ &\ \\
+    hyperparameters &\ &\ \\
     \hline
@@
 Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
-\kw{cite Planck 18} cosmological model. For hyper-parameters, we choose parameters that affect 
+\kw{cite Planck 18} cosmological model. For hyperparameters, we choose parameters that affect 
 binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 
@@
 We show a portion of the joint posterior of the progenitor parameters 
-and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+and hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}.
 \kb{We need to describe the interpretation of this as well.}
@@
 \begin{figure*}[h]
-    \includegraphics[width=\textwidth]{static/GW150914_corner_zoomed.png}
+    \includegraphics[width=\textwidth]{static/GW150914_corner_zoomed_lowres.png}
     \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
@@
     $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-    $q_{\rm crit 3}$ are the critical mass during \kw{Katie fill in},
-    Note that the redshift is not fitted during the root finding process.
-    Once we have find the evolutionary parameters, we add the time it would take the progenitor 
-    to merge to the lookback time of the observed posterior sample,
+    $q_{\rm crit 3}$ is the critical mass ratio on the Hertzsprung Gap,
+    Note that the redshift is not fitted during the root finding process or the MCMC process.
+    Once we have find the evolutionary parameters, we add the delay time to the lookback time of the observed posterior sample,
     then from the total lookback time we can compute the redshift at ZAMS.
+    We highlight three panels in the corner plots to show the fine structure of the set of posterior samples in the evolutionary parameter space.
+    We also color the posterior in a particular panel according to the type of parameters involved in the corner.
+    Blue denotes panels that include only progenitor parameters,
+    green denotes panels that include a mix of progenitor and hyperparameters,
+    and orange denotes panels that include only hyperparameters.
     }
@@
 % \kw{Talk about implication on single event posterior.}
-Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
+Once we obtain a set of binaries which successfully map the ZAMS parameters and hyperparameters to BBH merger masses, 
 we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
@@
 Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
-each with a different $\alpha$ but the same hyper-parameters as A21, 
-as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
+each with a different $\alpha$ but the same hyperparameters as A21, 
+as well as the results of our sampling which allows our hyperparameters to vary (fourth column). We find that holding the accretion 
 efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
-that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
+that produces GW150914-like mergers. In contrast, by allowing the hyperparameters to fully span the model uncertainty, we 
 find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
@@
 
+% Validity paragraph
 To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
@@
 In particular, we compute KL divergence of the samples obtained from root-finding alone and after MCMC sampling.
-For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later}nat on average if we use the root-finding result alone,
+For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} nats on average if we use the root-finding result alone,
 whereas the MCMC result has a higher KL divergence at \kw{fill} nats.
@@
 I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
-In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-fidining process alone and the MCMC result agrees well with the original posterior,
+In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-finding process alone and the MCMC result agrees well with the original posterior,
 even though the agreement between the MCMC result and the original posterior is not perfect. 
@@
 \caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
-The blue contour is the posterior reprojected using the samples in the evolutionary parameter space.
-The orange contour is the posterior samples from LVK that have corresponding samples in the evolutionary parameter space.
-The green contour is the posterior plotted using all of LVK posterior samples.
+The blue contour is the reprojected posterior after the root-finding procedure.
+The orange contour is the reprojected posterior after the MCMC procedure.
+The green contour is the posterior plotted using the original LVK posterior samples.
 }
@@
 
+% Paragraph discussing the potential application on the population side.
 To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
-Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint on $f_{\rm acc}$ for events passing the KL divergence test sorted by the median of their primary mass in the observable space.
-We can see $f_{\rm acc}$ increases as the primary mass of the event increases.
+Plotted in figure \ref{fig:GWTC3_f_acc_mass} is the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
+Each contour represents the $68\%$ credible interval of the posterior density for that particular event.
+Although it does not conclusively suggest a dependency between $f_{\rm acc}$ and $m_{1,{\rm GW}}$,
+there is a suggestive trend showing $f_{\rm acc}$ could possibly increase as the mass of the progenitor increases.
 This means the stable mass transfer (SMT) phase of a light binary is preferably non-conservative, whereas the heavier the events are, the more likely their SMT is conservative.
 
-Note that some events did not pass the KL divergence test we proposed in section \ref{sec:method}, which we do not include in figure \ref{fig:GWTC3_alpha_mass}.
+Note that some events did not pass the KL divergence test we proposed in section \ref{sec:method}, which we do not include in figure \ref{fig:GWTC3_f_acc_mass}.
 Low mass events are more subjected to random fluctuation because natal kick is less modulated by fallback in lighter system,
-so the reprojected posterior may not agree with the original posterior in the observable space consistently when varying the random seed.
-On the other hand, \texttt{COSMIC} struggles to produce events above the pair instability supernova mass cutoff \kw{State cutoff},
+which can lead to a final system either with or without a merger event randomly,
+so the reprojected posterior may not agree with the original posterior in the observable space consistently, hence yielding a higher KL divergence on average.
+On the other hand, \texttt{COSMIC} struggles to produce events above the pair instability supernova mass cutoff $m_{1,{\rm GW}}$,
 so the posterior in the evolutionary parameter space only correspond part of the posterior in the observable space, therefore events beyond the PISN mass gap has higher KL divergence.
@@
 
-The point of figure \ref{fig:GWTC3_alpha_mass} is to show this method could in principal reveal the correlation between progenitor parameters and hyper-parameters on a population level.
-A careful treatment of all event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_alpha_mass} is beyond the scope of this paper,
+The point of figure \ref{fig:GWTC3_f_acc_mass} is to show this method could in principal reveal the correlation between progenitor parameters and hyperparameters on a population level.
+A careful treatment of all event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_f_acc_mass} is beyond the scope of this paper,
 therefore we defer a detail study of the physics related to the population of GW events in the future.
@@
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{static/GWTC3_alpha_mass.pdf}
-\caption{This figure is ugly as fucked now. Let's change it in later iteration.
-Also replace only with events that pass the KL divergence test.}
-\label{fig:GWTC3_alpha_mass}
+\includegraphics[width=0.49\textwidth]{static/GWTC3_f_acc_mass.pdf}
+\caption{
+    The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
+    Each contour is the $68\%$ credible interval of the posterior density for a particular event. \kw{Check interval number}
+    At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair instability supernova mechanism prevents \texttt{COSMIC} from producing events that are more massive than this cutoff.
+    Therefore, events with the majority of posterior support above this cutoff is not compatible with \texttt{COSMIC}, hence has a large KL divergence and excluded from this figure.
+    On the low mass end, neutron star binaries or neutron star-black hole binaries are subjected to randomness induced by the natal kick,
+    also resulting in a larger and fluctuating KL divergence, therefore they are also excluded from the analysis.
+}
+\label{fig:GWTC3_f_acc_mass}
 \end{figure}
@@
 In this work we first showcase the power of the proposed method with GW150914.
-We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
-In particular, to the authors' knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is explored.
+We show the joint posterior of the two events' progenitor parameters and hyperparameters.
+In particular, to the authors' knowledge, this is the first time the correlation between progenitor parameters and hyperparameters is explored.
 Our work is not only a more efficient way to study GW events' progenitors,
-but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.
+but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyperparameters.
 
 To further illustrate the power of the method, we show the inferred fraction of accretion during SMT depends on the progenitor masses.
-The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
-This enables a data-driven way to study the distribution of hyper-parameters,
+The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyperparameters for each event.
+This enables a data-driven way to study the distribution of hyperparameters,
 i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
-we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters,
-without making specific assumptions such as fixing the hyper-parameters for all types of event.
+we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyperparameters,
+without making specific assumptions such as fixing the hyperparameters for all types of event.
 
@@
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyper-parameters space for the first time, there are a number of improvements can be implemented in future studies.
+While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements can be implemented in future studies.
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
@@
 To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
-We point out hyper-parameters in the usual population synthesis simulation context are not actually parameters related to the population,
+We point out hyperparameters in the usual population synthesis simulation context are not actually parameters related to the population,
 but parameters about the evolutionary function.
@@
 We "back propagate" the posterior in the observable space to the evolutionary parameter space,
-thus allowing us to study hyper-parameters and its correlation with progenitor parameters in a data-driven manner.
+thus allowing us to study hyperparameters and its correlation with progenitor parameters in a data-driven manner.
 Our method makes less assumption than the traditional forward modelling approach,
-which often fix the hyper-parameters across the entire population.
-Since we are not limited to the fix hyper-parameters assumption, we can explore the behavior of the hyper-parameters across the population much more efficiently.
+which often fix the hyperparameters across the entire population.
+Since we are not limited to the fix hyperparameters assumption, we can explore the behavior of the hyperparameters across the population much more efficiently.
 While our work lay down a data analysis pathway to understand the population of GW events,

### 2022-05-10 src/ms.tex
text update

@@
 Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
-\kw{cite Planck 18} cosmological model. For hyperparameters, we choose parameters that affect 
+\kw{cite Planck 18} cosmological model.
+The main benefit of post-processing the redshift in this way is we do not rely on a particular assumption
+of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
+This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
+by reweighting the posterior distribution according to a particular SFR model,
+therefore we do not have to rerun the inference everytime we change the SFR model.
+For hyperparameters, we choose parameters that affect 
 binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 

### 2022-05-13 src/ms.tex
adding in extra text

--- 2022-05-07 src/ms.tex
@@
 While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyper-parameters", which drive the evolution of the population.
-One of the most common skepticism of population synthesis studies is that these studies employ simulation sets which often 
+it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyper-parameters", which drive the evolution of the population.
+One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
 cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-(e.g. common envelope efficiency) while holding all others fixed to some "default" values.
+(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
 On top of this, these values are often not calibrated to data.
@@
 run a grid which varies combinations of uncertain physics.
-For 10 uncertain physical processes, if we consider $2$ variations for each physical process 
+For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
 this would still require $1024$ simulations.
@@
 
-In this paper, we propose a method to "backward" model each GW event to its progenitor state 
+In this paper, we propose a method to ``backward" model each GW event to its progenitor state 
 while allowing the hyper-parameters to vary across their full range of physical uncertainty
@@
 
-We define the parameters that describe the progenitor's zero age main sequence (ZAMS) properties
-masses and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
+We define the progenitor's zero age main sequence (ZAMS) properties like
+mass, orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
 that control different uncertain physical prescriptions such as wind strength and common
@@
 as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
-efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
+efficiency, $f_{\rm{acc}}$, and common envelope ejection efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space 
 that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
 find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
-fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
-\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses between 
+$\sim95-135\,M_{\odot}$ and secondary ZAMS masses between $\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
+efficiency and common envelope ejection efficiency which produces merging BBHs that have masses consistent with GW150914's posteriors.
+In this region, if BBHs with masses consistent with GW150914's masses are produced through a combination of $\alpha$ and 
+$f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely, in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
+which does produce GW150914-like mergers, the combination of $\alpha$, $f_{\rm{acc}}$, and the ZAMS masses and orbital separations 
+balance to produce BHs with the correct masses that merge within a Hubble time.
+%\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+%\kb{done!}
 

### 2022-05-13 src/ms.tex
actually fixing conflicts

--- 2022-05-10 src/ms.tex
@@
 While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyperparameters", which drive the evolution of the population.
-One of the most common skepticism of population synthesis studies is that these studies employ simulation sets which often 
+it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
+One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
 cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-(e.g. common envelope efficiency) while holding all others fixed to some "default" values.
+(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
 On top of this, these values are often not calibrated to data.
@@
 
-In this paper, we propose a method to "backward" model each GW event to its progenitor state 
-while allowing the hyperparameters to vary across their full range of physical uncertainty
+In this paper, we propose a method to ``backward" model each GW event to its progenitor state 
+while allowing the hyper-parameters to vary across their full range of physical uncertainty
 following a two-stage process.
@@
 Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
-each with a different $\alpha$ but the same hyperparameters as A21, 
-as well as the results of our sampling which allows our hyperparameters to vary (fourth column). We find that holding the accretion 
-efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
-that produces GW150914-like mergers. In contrast, by allowing the hyperparameters to fully span the model uncertainty, we 
+each with a different $\alpha$ but the same hyper-parameters as A21, 
+as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
+efficiency, $f_{\rm{acc}}$, and common envelope ejection efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space 
+that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
 find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
-fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
-\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses between 
+$\sim95-135\,M_{\odot}$ and secondary ZAMS masses between $\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
+efficiency and common envelope ejection efficiency which produces merging BBHs that have masses consistent with GW150914's posteriors.
+In this region, if BBHs with masses consistent with GW150914's masses are produced through a combination of $\alpha$ and 
+$f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely, in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
+which does produce GW150914-like mergers, the combination of $\alpha$, $f_{\rm{acc}}$, and the ZAMS masses and orbital separations 
+balance to produce BHs with the correct masses that merge within a Hubble time.
+%\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+%\kb{done!}
 

### 2022-05-20 src/ms.tex
Will works on intro and abstract.

--- 2022-05-10 src/ms.tex
@@
 astrophysics from catalogs. Here we propose an inference pipeline to
-simultaneously infer the posterior density of stellar progenitor parameters
-(e.g. zero-age main sequence masses) and population synthesis parameter settings
-controlling modeled binary evolution from gravitational wave observations.  Our
-pipeline can efficiently explore the high-dimensional space of population
-synthesis settings and progenitor system properties for each system in a catalog
-of gravitational wave observations.  We apply our pipeline to observations in
-the third LIGO Virgo Kagra Gravitational Wave Transient Catalog (GWTC3).  We
-showcase the effectiveness of this pipeline with a detailed study of the
-progenitor properties and population synthesis settings that produce mergers
-like the observed GW150914.
+simultaneously infer stellar progenitor parameters (e.g. zero-age main sequence
+masses) and population synthesis parameter settings controlling modeled binary
+evolution from individual gravitational wave observations of merging compact
+binaries.  Our pipeline can efficiently explore the high-dimensional space of
+population synthesis settings and progenitor system properties for each system
+in a catalog of gravitational wave observations.  We apply our pipeline to
+observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog
+(GWTC3).  We showcase the effectiveness of this pipeline with a detailed study
+of the progenitor properties and population synthesis settings that produce
+mergers like the observed GW150914.
 \end{abstract}
@@
 systems, globular clusters, nuclear star clusters, AGN disks, and more
-\citep{Mandel2022}. \wf{Moar citations here} In these types of studies, the
-population is forward modeled based on random samples from empirically or
-theoretically motivated distributions of initial parameters which define the
-age, metallicity, and intrinsic properties of the population like masses,
-orbital separations, and eccentricities. These initial populations are then
-evolved using a set of assumptions for how astrophysical processes impact the
-evolution of the DCO merger progenitors, ultimately producing a synthetic
-catalog of mergers which can be compared of observations from ground-based GW
-detectors.
+\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
+population is studied through a Monte-Carlo approach.  Using theoretically
+motivated distributions of initial parameters (age, metallicity, masses, orbital
+separations, eccentricities, etc), initial populations are evolved with fixed
+astrophysical assumptions (population synthesis parameter settings)to produce a
+synthetic catalog of mergers which can be compared of observations from
+ground-based GW detectors.
 
@@
 
-While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or "hyperparameters", which drive the evolution of the population.
-One of the most common skepticism of population synthesis studies is that these studies employ simulation sets which often 
-cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-(e.g. common envelope efficiency) while holding all others fixed to some "default" values.
-On top of this, these values are often not calibrated to data.
-Even with recent developments in using emulators to speed up the simulation process, training the emulators 
-still requires a significant amount of computation to cover a wide range of uncertainty.
-As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
-run a grid which varies combinations of uncertain physics.
-For 10 uncertain physical processes, if we consider $2$ variations for each physical process 
-this would still require $1024$ simulations.
-Because population synthesis simulations require hundreds to thousands of cpu hours,
-training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
+While forward modelling entire populations has been the canonical pathway to
+compare simulations and GW observations, it is not necessarily the most
+efficient way to constrain the underlying physical processes, represented by
+population synthesis ``hyperparameters,'' which drive the evolution of the
+population.  \wf{I would not turn to ``skepticism'' here; instead, I would say
+something like [but worded better] ``it is likely that the hyperparameter
+settings that best represent the evolutionary physics vary across progenitor
+parameter space in unknown ways, which makes the modeling very expensive if you
+want to produce an entire population with one hyperparameter setting.
+Additionally, comparing at the level of populations is not as efficient as
+comparing event-by-event.  Ideally, we would like to \emph{hierarchically}
+combine several observations to directly constrain the progenitor population
+\emph{and} hyperparameter settings all at once from the observed catalog.''  I
+see that these points come in the next paragraph, so one thing you could do is
+dramatically shorten this paragraph and use it to lead into the next one.} One
+of the most common skepticism of population synthesis studies is that these
+studies employ simulation sets which often cover a small subdomain of the entire
+parameter space of uncertain physics, or only vary one uncertain process per
+simulation (e.g. common envelope efficiency) while holding all others fixed to
+some "default" values. On top of this, these values are often not calibrated to
+data. Even with recent developments in using emulators to speed up the
+simulation process, training the emulators still requires a significant amount
+of computation to cover a wide range of uncertainty. As an example, to emulate a
+model with $10$ parameters, the simplest way to construct a training set of
+simulations is to run a grid which varies combinations of uncertain physics. For
+10 uncertain physical processes, if we consider $2$ variations for each physical
+process this would still require $1024$ simulations. Because population
+synthesis simulations require hundreds to thousands of cpu hours, training an
+emulator which spans a high dimensional uncertainty space remains an
+impossibility at present.
 
@@
 
-This compounds with the fact that there is no guarantee the entire population should share the same fixed hyperparameters, 
-as is often assumed in most population synthesis studies.
-This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
-required to compare data to simulations which span the full set of uncertain physical models.
-However, this issue is not a fundamental issue, but an apparent obstacle which arises from an inefficient approach to the problem.
-Furthermore, each binary system could in principle come from a different formation environment and hence 
-have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
-In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space.
-\citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
-to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
-set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
-\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
-However, the method was limited to a single event instead of the population.
-More importantly, the method did not vary the hyperparameters while sampling the posterior density of the ZAMS parameters,
-which limit its capability to be applied to understand a population of binary systems.
+This compounds with the fact that there is no guarantee the entire population
+should share the same fixed hyperparameters, as is often assumed in most
+population synthesis studies. This suggests that there could be a large
+(possibly infinite) family of hyper-parameter distributions required to compare
+data to simulations which span the full set of uncertain physical models.
+However, this issue is not a fundamental issue, but an apparent obstacle which
+arises from an inefficient approach to the problem. Furthermore, each binary
+system could in principle come from a different formation environment and hence
+have different progenitor parameters as well as hyperparameters, which define
+uncertain physics associated with each environment. In this case, it is
+advantageous to determine the posterior density in the progenitor and
+hyper-parameter joint space for each individual binary system, then perform
+population in analyses in that space. \citet{Andrews2021} (hereafter A21) used
+\texttt{DartBoard} \citep{Andrews2018} to determine the ZAMS parameters which
+produce GW150914-like BBH merger based on posterior samples for GW150914 and a
+fixed set of hyperparameters which define assumptions for how isolated
+binary-star interactions proceed using \texttt{COSMIC}, a binary population
+synthesis code \citep{Breivik2020}. Crucially, the method did not vary the
+hyperparameters while sampling the posterior density of the ZAMS parameters,
+which limit its capability to be applied to understand a population of binary
+systems.  \wf{This sounds sort of aggressive against DartBoard, and we probably
+don't want to do that.  Instead, maybe re-focus the 'graph on
+\emph{distinguishing} our method from DB; less DB sux and more ``here is how we
+are not DB.''}
 
-In this paper, we propose a method to "backward" model each GW event to its progenitor state 
-while allowing the hyperparameters to vary across their full range of physical uncertainty
-following a two-stage process.
-First, we solve a root finding problem to obtain guesses that are likely to produce the desired system properties in the joint progenitor--hyper-parameter space.
-We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
-The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
-In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
-We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.
+In this paper, we propose a method to "backward" model each GW event to its
+progenitor state while allowing the hyperparameters to vary across their full
+range of physical uncertainty following a two-stage process. First, we solve a
+root finding problem to obtain guesses that are likely to produce the desired
+system properties in the joint progenitor--hyper-parameter space. We then sample
+the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm
+that is initialized using the roots found in the previous stage; the
+initialization step makes the burnin and convergence of the MCMC stage
+considerably faster than initializing with random parameter choices
+\citep{Andrews2018,Andrews2021}. The rest of the paper is structured as follows:
+We detail our method in Sec. ~\ref{sec:method}. In Sec. ~\ref{sec:result}, we
+show the results of applying our method to a number of real events. We discuss
+the implications of this work and future direction in Sec.
+~\ref{sec:discussion}.
 

### 2022-05-20 src/ms.tex
Comment about the dtd.

--- 2022-05-13 src/ms.tex
@@
 \author{Katelyn Breivik}
+\email{kbreivik@flatironinstitute.org}
 \flatiron
@@
 
-\section{Result}
+\section{Results}
 \label{sec:result}
@@
 formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total 
-time to merger to by less than $13.7\,\rm{Gyr}$.
+time to merger to be less than $13.7\,\rm{Gyr}$.
 Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
-\kw{cite Planck 18} cosmological model.
-The main benefit of post-processing the redshift in this way is we do not rely on a particular assumption
+Planck 2018 cosmological model \citep{Planck2018}.
+The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
 of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
@@
 by reweighting the posterior distribution according to a particular SFR model,
-therefore we do not have to rerun the inference everytime we change the SFR model.
+\kb{and} therefore do not have to rerun the inference everytime we change the SFR model. This is especially helpful
+since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep{Broekgaarden2021}.
 For hyperparameters, we choose parameters that affect 
@@
 and hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. 
-\kb{We need to describe the interpretation of this as well.}
-\kw{I think we can just hightlight the fact that we have three kinds of correlation, progenitor-hyper, progenitor-hyper, hyper-hyper,
-since the implication is related to figure @Katie made.}
-\kb{Need to describe the formation channels and gaps here.}
+Each panel is colored based on the comparison between progenitor parameters (blue), hyperparameters (orange), or 
+a mix of the two (green). 
+We find that the choice for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not unexpected 
+since we use the \citet{Fryer2012} delayed model which reduces the natal kick strength at BH formation based on the amount
+of fallback onto the proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive enough that the 
+fallback reduces the natal kick to zero. Because of this we do not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}.
+Similarly, we do not include the ZAMS orbital period and eccentricity are correlated with one another, but not strongly with the other progenitor
+parameters or hyperparameters. We find strong correlations in the ZAMS masses, as well as 
+strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$ and $f_{\rm{acc}}$ and $\alpha$. We show scatter points of each of these 
+combinations in the upper-right inset of the figure.
+
+The correlations between the ZAMS masses are similar to those found by A21 with the majority of the population preferring primary 
+and secondary masses between $60--90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to $150\,M_{\odot}$ which 
+is the limit imposed by our assumptions. We explore the formation scenarios which lead to successful GW150914-like mergers and those 
+which fail to produce GW150914-like mergers below in Figure~\ref{fig:compare_fixed_variable}.
+
+The correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm acc}$ illustrate the variety of ways which GW150914-like mergers can be produced.
+For binaries with $M_{2,\rm{ZAMS}} < 70\,M_{\odot}$, we find that accretion efficiencies of $f_{\rm acc} > 0.5$ 
+are preferred. One reason for this is based on the requirement that the total mass in the binary must remain above the 
+total mass of GW150914 and strongly non-conservative mass transfer (i.e. $f_{\rm acc} > 0.5$) reduces the total 
+mass of the system. A compounding factor is that as mass leaves the binary due to non-conservative mass transfer, the 
+evolution of the binary's orbit is less dramatic and leads to wider binaries on average and thus fewer mergers in a Hubble time.
+For binaries with $M_{2,\rm{ZAMS}} \sim 80\,M_{\odot}$, $f_{\rm acc}$ is less constrained. This is because of a preference for 
+these secondaries to also have primary masses near $ 80\,M_{\odot}$ and thus enter a double-core common envelope evolution in 
+which both stars' envelopes are ejected, leaving behind two stripped helium cores in a tight orbit. In this case, $f_{\rm acc}$ 
+does not affect the binary evolution and is thus unconstrained. This effect can also be seen in the ZAMS mass and $q_{\rm crit 3}$ 
+panels for masses near $80\,M_{\odot}$ because double-core common-envelope evolution is triggered in \texttt{COSMIC} when the
+radii of the two stars touch due to the rapid expansion of the primary star upon helium ignition. In this case, the choice for 
+Roche-overflow mass transfer to proceed stably (as prescribed by $f_{\rm{acc}}$) or unstably (as prescribed by $\alpha$) using 
+$q_{\rm crit, 3}$ is totally irrelevant, and thus unconstrained. Finally, for binaries with $M_{2,\rm{ZAMS}} > 80$, we find that 
+$f_{\rm acc}$ is correlated to decrease with increasing mass. This is due to limits on the total mass which must be ejected from 
+the binary to produce BH masses that match GW150914. Binaries with both component masses near $150\,M_{\odot}$ must either go 
+through a common envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very non-conservative mass transfer 
+(where $f_{\rm acc}$) is low, to produce BBHs with the proper mass.
+
+Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated, though there exists independent trends in each hyper parameter.
+This is not totally unexpected since the physical processes which are described by each hyperparameter, i.e. stable Roche-lobe overflow and 
+common envelope evolution are two independent channels. Generally, we find that GW150914-like BBHs tend to prefer larger accretion efficiencies 
+and have common envelope ejection efficiencies peaking near $\alpha=1$. One shortcoming of our method implementation is the application of a single 
+hyper parameter for $f_acc$ and $\alpha$ for each binary, while both the primary and the secondary star could in principle be defined by their own 
+hyperparameters that prescribe the outcomes for when each star fills it's Roche lobe. We reserve full treatment of this for future work but note 
+that this improvement could reveal correlations between hyperparameters that are not present in our current analysis.
+%\kb{We need to describe the interpretation of this as well.}
+%\kw{I think we can just hightlight the fact that we have three kinds of correlation, progenitor-hyper, progenitor-hyper, hyper-hyper,
+%since the implication is related to figure @Katie made.}
+%\kb{Need to describe the formation channels and gaps here.}
 
@@
 we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
-to explore how keeping a fixed model which only varies one hyper-parameter affects our analysis. 
+to explore how keeping a fixed model which only varies one hyper-parameter contrasts to our results. 
 Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
@@
 The KL divergence stays stable for reprojecting with different random seeds. 
-\kb{Can we show this in the figure somehow? Maybe like KL divergence as a function of reprojection number as a lower small panel?}
-\kw{The problem with that is it is basically one dot or a flat line, which is very unentertaining.
-I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
 In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-finding process alone and the MCMC result agrees well with the original posterior,
@@
 so the posterior in the evolutionary parameter space only correspond part of the posterior in the observable space, therefore events beyond the PISN mass gap has higher KL divergence.
-Events with more extreme mass ratios are not likely to be recovered by the method. \kw{@Katie gives reason?}
+Events with more extreme mass ratios are not likely to be recovered by the method. This could be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as discussed 
+in Section~\ref{sec:result}.
 
@@
 but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyperparameters. 
+In future work, we will extend our method to applying hyperparameters to individual binary components in order to better capture correlations between 
+hyperparameters and progenitor parameters.
 
@@
 thus allowing us to study hyperparameters and its correlation with progenitor parameters in a data-driven manner.
-Our method makes less assumption than the traditional forward modelling approach,
+Our method makes less assumptions than the traditional forward modelling approach,
 which often fix the hyperparameters across the entire population.
-Since we are not limited to the fix hyperparameters assumption, we can explore the behavior of the hyperparameters across the population much more efficiently.
-While our work lay down a data analysis pathway to understand the population of GW events,
+Since we are not limited to the fixed hyperparameters assumption, we can explore the behavior of the hyperparameters across the population much more efficiently.
+While our work lays down a data analysis pathway to understand the population of GW events,
 no physics can be learned without a comprehensive physical model.

### 2022-05-20 src/ms.tex
Merging in Katie's edits and Will's line-breaks.

@@
 astrophysics from catalogs. Here we propose an inference pipeline to
-    simultaneously infer the posterior density of stellar progenitor parameters
-    (e.g. zero-age main sequence masses) and population synthesis parameter settings
-    controlling modeled binary evolution from gravitational wave observations.  Our
-    pipeline can efficiently explore the high-dimensional space of population
-    synthesis settings and progenitor system properties for each system in a catalog
-    of gravitational wave observations.  We apply our pipeline to observations in
-    the third LIGO Virgo Kagra Gravitational Wave Transient Catalog (GWTC3).  We
-    showcase the effectiveness of this pipeline with a detailed study of the
-    progenitor properties and population synthesis settings that produce mergers
-    like the observed GW150914.
+simultaneously infer stellar progenitor parameters (e.g. zero-age main sequence
+masses) and population synthesis parameter settings controlling modeled binary
+evolution from individual gravitational wave observations of merging compact
+binaries.  Our pipeline can efficiently explore the high-dimensional space of
+population synthesis settings and progenitor system properties for each system
+in a catalog of gravitational wave observations.  We apply our pipeline to
+observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog
+(GWTC3).  We showcase the effectiveness of this pipeline with a detailed study
+of the progenitor properties and population synthesis settings that produce
+mergers like the observed GW150914.
 \end{abstract}
@@
 systems, globular clusters, nuclear star clusters, AGN disks, and more
-    \citep{Mandel2022}. \wf{Moar citations here} In these types of studies, the
-    population is forward modeled based on random samples from empirically or
-    theoretically motivated distributions of initial parameters which define the
-    age, metallicity, and intrinsic properties of the population like masses,
-    orbital separations, and eccentricities. These initial populations are then
-    evolved using a set of assumptions for how astrophysical processes impact the
-    evolution of the DCO merger progenitors, ultimately producing a synthetic
-    catalog of mergers which can be compared of observations from ground-based GW
-    detectors.
+\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
+population is studied through a Monte-Carlo approach.  Using theoretically
+motivated distributions of initial parameters (age, metallicity, masses, orbital
+separations, eccentricities, etc), initial populations are evolved with fixed
+astrophysical assumptions (population synthesis parameter settings) to produce a
+synthetic catalog of mergers which can be compared to observations from
+GW detectors.
 
@@
 
-While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
-One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
-cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
-On top of this, these values are often not calibrated to data.
-Even with recent developments in using emulators to speed up the simulation process, training the emulators 
-still requires a significant amount of computation to cover a wide range of uncertainty.
-As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
-run a grid which varies combinations of uncertain physics.
-For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
-this would still require $1024$ simulations.
-Because population synthesis simulations require hundreds to thousands of cpu hours,
-training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
+While forward modelling entire populations has been the canonical pathway to
+compare simulations and GW observations, it is not necessarily the most
+efficient way to constrain the underlying physical processes, represented by
+population synthesis ``hyperparameters,'' which drive the evolution of the
+population.  \wf{I would not turn to ``skepticism'' here; instead, I would say
+something like [but worded better] ``it is likely that the hyperparameter
+settings that best represent the evolutionary physics vary across progenitor
+parameter space in unknown ways, which makes the modeling very expensive if you
+want to produce an entire population with one hyperparameter setting.
+Additionally, comparing at the level of populations is not as efficient as
+comparing event-by-event.  Ideally, we would like to \emph{hierarchically}
+combine several observations to directly constrain the progenitor population
+\emph{and} hyperparameter settings all at once from the observed catalog.''  I
+see that these points come in the next paragraph, so one thing you could do is
+dramatically shorten this paragraph and use it to lead into the next one.} One
+of the most common skepticism of population synthesis studies is that these
+studies employ simulation sets which often cover a small subdomain of the entire
+parameter space of uncertain physics, or only vary one uncertain process per
+simulation (e.g. common envelope efficiency) while holding all others fixed to
+some "default" values. On top of this, these values are often not calibrated to
+data. Even with recent developments in using emulators to speed up the
+simulation process, training the emulators still requires a significant amount
+of computation to cover a wide range of uncertainty. As an example, to emulate a
+model with $10$ parameters, the simplest way to construct a training set of
+simulations is to run a grid which varies combinations of uncertain physics. For
+10 uncertain physical processes, if we consider $2$ variations for each physical
+process this would still require $1024$ simulations. Because population
+synthesis simulations require hundreds to thousands of cpu hours, training an
+emulator which spans a high dimensional uncertainty space remains an
+impossibility at present.
 
@@
 
-This compounds with the fact that there is no guarantee the entire population should share the same fixed hyperparameters, 
-as is often assumed in most population synthesis studies.
-This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
-required to compare data to simulations which span the full set of uncertain physical models.
-However, this issue is not a fundamental issue, but an apparent obstacle which arises from an inefficient approach to the problem.
-Furthermore, each binary system could in principle come from a different formation environment and hence 
-have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
-In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space.
-\citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
-to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
-set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
-\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
-However, the method was limited to a single event instead of the population.
-More importantly, the method did not vary the hyperparameters while sampling the posterior density of the ZAMS parameters,
-which limit its capability to be applied to understand a population of binary systems.
+This compounds with the fact that there is no guarantee the entire population
+should share the same fixed hyperparameters, as is often assumed in most
+population synthesis studies. This suggests that there could be a large
+(possibly infinite) family of hyper-parameter distributions required to compare
+data to simulations which span the full set of uncertain physical models.
+However, this issue is not a fundamental issue, but an apparent obstacle which
+arises from an inefficient approach to the problem. Furthermore, each binary
+system could in principle come from a different formation environment and hence
+have different progenitor parameters as well as hyperparameters, which define
+uncertain physics associated with each environment. In this case, it is
+advantageous to determine the posterior density in the progenitor and
+hyper-parameter joint space for each individual binary system, then perform
+population in analyses in that space. \citet{Andrews2021} (hereafter A21) used
+\texttt{DartBoard} \citep{Andrews2018} to determine the ZAMS parameters which
+produce GW150914-like BBH merger based on posterior samples for GW150914 and a
+fixed set of hyperparameters which define assumptions for how isolated
+binary-star interactions proceed using \texttt{COSMIC}, a binary population
+synthesis code \citep{Breivik2020}. Crucially, the method did not vary the
+hyperparameters while sampling the posterior density of the ZAMS parameters,
+which limit its capability to be applied to understand a population of binary
+systems.  \wf{This sounds sort of aggressive against DartBoard, and we probably
+don't want to do that.  Instead, maybe re-focus the 'graph on
+\emph{distinguishing} our method from DB; less DB sux and more ``here is how we
+are not DB.''}
 
-In this paper, we propose a method to ``backward" model each GW event to its progenitor state 
-while allowing the hyper-parameters to vary across their full range of physical uncertainty
-following a two-stage process.
-First, we solve a root finding problem to obtain guesses that are likely to produce the desired system properties in the joint progenitor--hyper-parameter space.
-We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
-The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
-In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
-We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.
+In this paper, we propose a method to "backward" model each GW event to its
+progenitor state while allowing the hyperparameters to vary across their full
+range of physical uncertainty following a two-stage process. First, we solve a
+root finding problem to obtain guesses that are likely to produce the desired
+system properties in the joint progenitor--hyper-parameter space. We then sample
+the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm
+that is initialized using the roots found in the previous stage; the
+initialization step makes the burnin and convergence of the MCMC stage
+considerably faster than initializing with random parameter choices
+\citep{Andrews2018,Andrews2021}. The rest of the paper is structured as follows:
+We detail our method in Sec. ~\ref{sec:method}. In Sec. ~\ref{sec:result}, we
+show the results of applying our method to a number of real events. We discuss
+the implications of this work and future direction in Sec.
+~\ref{sec:discussion}.
 
@@
 % Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
-Another perk of our method is the "rate" calculation is now independent of assumed initial condition such as a particular star formation rate model,
-which is a dominant factor of uncertainty for rate prediction in the past.
-Once we have pull back the GW event posterior from the observable space to the evolutionary parameter space,
-we can apply the same methods we use to account for the rate in the observable space.
-In the observable space, the overall rate is basically a normalization that is defined in the local universe i.e. the local merger rate. \kw{cite}
-In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
-This also means one can study the progenitor-formation rate in a data-driven manner instead of assuming a particular model.
+Another perk of our method is the "rate" calculation is now independent of
+assumed initial condition such as a particular star formation rate model, which
+is a dominant factor of uncertainty for rate prediction in the past. Once we
+have pull back the GW event posterior from the observable space to the
+evolutionary parameter space, we can apply the same methods we use to account
+for the rate in the observable space. In the observable space, the overall rate
+is basically a normalization that is defined in the local universe i.e. the
+local merger rate. \kw{cite} In the evolutionary parameter space, we can instead
+fit for the production rate of merger events at some redshift. This also means
+one can study the progenitor-formation rate in a data-driven manner instead of
+assuming a particular model. \wf{Emphasize this more---build up the SFR directly
+from a population of GW events, with realistic assumptions about the dtd.}
 

### 2022-05-20 src/ms.tex
Minor typo fixes in results section.

--- 2022-05-20 src/ms.tex
@@
 
-We show a portion of the joint posterior of the progenitor parameters 
-and hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. 
-Each panel is colored based on the comparison between progenitor parameters (blue), hyperparameters (orange), or 
-a mix of the two (green). 
-We find that the choice for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not unexpected 
-since we use the \citet{Fryer2012} delayed model which reduces the natal kick strength at BH formation based on the amount
-of fallback onto the proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive enough that the 
-fallback reduces the natal kick to zero. Because of this we do not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}.
-Similarly, we do not include the ZAMS orbital period and eccentricity are correlated with one another, but not strongly with the other progenitor
-parameters or hyperparameters. We find strong correlations in the ZAMS masses, as well as 
-strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$ and $f_{\rm{acc}}$ and $\alpha$. We show scatter points of each of these 
-combinations in the upper-right inset of the figure.
+We show a portion of the joint posterior of the progenitor parameters and
+hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
+is colored based on the comparison between progenitor parameters (blue),
+hyperparameters (orange), or a mix of the two (green). We find that the choice
+for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
+unexpected since we use the \citet{Fryer2012} delayed model which reduces the
+natal kick strength at BH formation based on the amount of fallback onto the
+proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
+enough that the fallback reduces the natal kick to zero. Because of this we do
+not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
+not include the ZAMS orbital period and eccentricity, which are correlated with
+one another, but not strongly with the other progenitor parameters or
+hyperparameters. We find strong correlations between the ZAMS masses, as well as
+strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
+$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
+in the upper-right inset of the figure.
 

### 2022-05-20 src/ms.tex
Collaborative edits to discussion.

@@
 
-The correlations between the ZAMS masses are similar to those found by A21 with the majority of the population preferring primary 
-and secondary masses between $60--90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to $150\,M_{\odot}$ which 
-is the limit imposed by our assumptions. We explore the formation scenarios which lead to successful GW150914-like mergers and those 
-which fail to produce GW150914-like mergers below in Figure~\ref{fig:compare_fixed_variable}.
+The correlations between the ZAMS masses are similar to those found by A21 with
+the majority of the population preferring primary and secondary masses between
+$60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
+$150\,M_{\odot}$ which is the limit imposed by our assumptions. We explore the
+formation scenarios which lead to successful GW150914-like mergers and those
+which fail to produce GW150914-like mergers below in
+Figure~\ref{fig:compare_fixed_variable}.
 
-The correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm acc}$ illustrate the variety of ways which GW150914-like mergers can be produced.
-For binaries with $M_{2,\rm{ZAMS}} < 70\,M_{\odot}$, we find that accretion efficiencies of $f_{\rm acc} > 0.5$ 
-are preferred. One reason for this is based on the requirement that the total mass in the binary must remain above the 
-total mass of GW150914 and strongly non-conservative mass transfer (i.e. $f_{\rm acc} > 0.5$) reduces the total 
-mass of the system. A compounding factor is that as mass leaves the binary due to non-conservative mass transfer, the 
-evolution of the binary's orbit is less dramatic and leads to wider binaries on average and thus fewer mergers in a Hubble time.
-For binaries with $M_{2,\rm{ZAMS}} \sim 80\,M_{\odot}$, $f_{\rm acc}$ is less constrained. This is because of a preference for 
-these secondaries to also have primary masses near $ 80\,M_{\odot}$ and thus enter a double-core common envelope evolution in 
-which both stars' envelopes are ejected, leaving behind two stripped helium cores in a tight orbit. In this case, $f_{\rm acc}$ 
-does not affect the binary evolution and is thus unconstrained. This effect can also be seen in the ZAMS mass and $q_{\rm crit 3}$ 
-panels for masses near $80\,M_{\odot}$ because double-core common-envelope evolution is triggered in \texttt{COSMIC} when the
-radii of the two stars touch due to the rapid expansion of the primary star upon helium ignition. In this case, the choice for 
-Roche-overflow mass transfer to proceed stably (as prescribed by $f_{\rm{acc}}$) or unstably (as prescribed by $\alpha$) using 
-$q_{\rm crit, 3}$ is totally irrelevant, and thus unconstrained. Finally, for binaries with $M_{2,\rm{ZAMS}} > 80$, we find that 
-$f_{\rm acc}$ is correlated to decrease with increasing mass. This is due to limits on the total mass which must be ejected from 
-the binary to produce BH masses that match GW150914. Binaries with both component masses near $150\,M_{\odot}$ must either go 
-through a common envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very non-conservative mass transfer 
-(where $f_{\rm acc}$) is low, to produce BBHs with the proper mass.
+The correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm acc}$ illustrate the
+variety of ways which GW150914-like mergers can be produced. For binaries with
+$M_{2,\rm{ZAMS}} < 70\,M_{\odot}$, we find that accretion efficiencies of
+$f_{\rm acc} > 0.5$ are preferred. One reason for this is based on the
+requirement that the total mass in the binary must remain above the total mass
+of GW150914 and strongly non-conservative mass transfer (i.e. $f_{\rm acc} <
+0.5$) reduces the total mass of the system. A compounding factor is that as mass
+leaves the binary due to non-conservative mass transfer, the evolution of the
+binary's orbit is less dramatic and leads to wider binaries on average and thus
+fewer mergers in a Hubble time. For binaries with $M_{2,\rm{ZAMS}} \sim
+80\,M_{\odot}$, $f_{\rm acc}$ is less constrained. This is because of a
+preference for these secondaries to also have primary masses near $
+80\,M_{\odot}$ and thus enter a double-core common envelope evolution in which
+both stars' envelopes are ejected, leaving behind two stripped helium cores in a
+tight orbit. In this case, $f_{\rm acc}$ does not affect the binary evolution
+and is thus unconstrained. This effect can also be seen in the ZAMS mass and
+$q_{\rm crit 3}$ panels for masses near $80\,M_{\odot}$ because double-core
+common-envelope evolution is triggered in \texttt{COSMIC} when the radii of the
+two stars touch due to the rapid expansion of the primary star upon helium
+ignition. In this case, the choice for Roche-overflow mass transfer to proceed
+stably (as prescribed by $f_{\rm{acc}}$) or unstably (as prescribed by $\alpha$)
+using $q_{\rm crit, 3}$ is totally irrelevant, and thus unconstrained. Finally,
+for binaries with $M_{2,\rm{ZAMS}} > 80$, we find that $f_{\rm acc}$ is
+correlated to decrease with increasing mass. This is due to limits on the total
+mass which must be ejected from the binary to produce BH masses that match
+GW150914, the reverse situation to $M_{2,\rm{ZAMS}} < 70 \, M_\odot$. Binaries
+with both component masses near $150\,M_{\odot}$ must either go through a common
+envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very
+non-conservative mass transfer (where $f_{\rm acc}$) is low, to produce BBHs
+with the proper mass.
 
-Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated, though there exists independent trends in each hyper parameter.
-This is not totally unexpected since the physical processes which are described by each hyperparameter, i.e. stable Roche-lobe overflow and 
-common envelope evolution are two independent channels. Generally, we find that GW150914-like BBHs tend to prefer larger accretion efficiencies 
-and have common envelope ejection efficiencies peaking near $\alpha=1$. One shortcoming of our method implementation is the application of a single 
-hyper parameter for $f_acc$ and $\alpha$ for each binary, while both the primary and the secondary star could in principle be defined by their own 
-hyperparameters that prescribe the outcomes for when each star fills it's Roche lobe. We reserve full treatment of this for future work but note 
-that this improvement could reveal correlations between hyperparameters that are not present in our current analysis.
+Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
+though there exist independent trends in each hyper parameter. This is not
+totally unexpected since the physical processes which are described by each
+hyperparameter, i.e. stable Roche-lobe overflow and common envelope evolution
+are two independent channels. Generally, we find that GW150914-like BBHs tend to
+prefer larger accretion efficiencies and have common envelope ejection
+efficiencies peaking near $\alpha=1$. One shortcoming of our method
+implementation is the application of a single hyper parameter for
+$f_{\mathrm{acc}}$ and $\alpha$ for each binary, while both the primary and the
+secondary star could in principle be defined by their own hyperparameters that
+prescribe the outcomes for when each star fills it's Roche lobe. We reserve full
+treatment of this for future work but note that this improvement could reveal
+correlations between hyperparameters that are not present in our current
+analysis.
 %\kb{We need to describe the interpretation of this as well.}
@@
 % \kw{Talk about implication on single event posterior.}
-Once we obtain a set of binaries which successfully map the ZAMS parameters and hyperparameters to BBH merger masses, 
-we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
-to explore how keeping a fixed model which only varies one hyper-parameter contrasts to our results. 
-Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
-each with a different $\alpha$ but the same hyper-parameters as A21, 
-as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
-efficiency, $f_{\rm{acc}}$, and common envelope ejection efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space 
-that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
-find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
-fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses between 
-$\sim95-135\,M_{\odot}$ and secondary ZAMS masses between $\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
-efficiency and common envelope ejection efficiency which produces merging BBHs that have masses consistent with GW150914's posteriors.
-In this region, if BBHs with masses consistent with GW150914's masses are produced through a combination of $\alpha$ and 
-$f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely, in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
-which does produce GW150914-like mergers, the combination of $\alpha$, $f_{\rm{acc}}$, and the ZAMS masses and orbital separations 
-balance to produce BHs with the correct masses that merge within a Hubble time.
+Once we obtain a set of binaries which successfully map the ZAMS parameters and
+hyperparameters to BBH merger masses, we re-evolve the set of ZAMS parameters
+with the same physical assumptions as A21, but vary the common envelope
+efficiency to explore how keeping a fixed model which only varies one
+hyper-parameter contrasts to our results.
+Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses
+for three models (first through third columns), each with a different $\alpha$
+but the same hyper-parameters as A21, as well as the results of our sampling
+which allows our hyper-parameters to vary (fourth column). We find that holding
+the accretion efficiency, $f_{\rm{acc}}$, and common envelope ejection
+efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space
+that produces GW150914-like mergers. In contrast, by allowing the
+hyper-parameters to fully span the model uncertainty, we find that there are
+distinct ZAMS parameters which produce GW150914-like mergers.
+
+Because we can explore the full evolutionary parameter space, we can also
+determine that there are ZAMS parameters which fail to produce GW150914-like
+mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses
+between $\sim95-135\,M_{\odot}$ and secondary ZAMS masses between
+$\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
+efficiency and common envelope ejection efficiency which produces merging BBHs
+that have masses consistent with GW150914's posteriors. In this region, if BBHs
+with masses consistent with GW150914's masses are produced through a combination
+of $\alpha$ and $f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely,
+in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
+which does produce GW150914-like mergers, the combination of $\alpha$,
+$f_{\rm{acc}}$, and the ZAMS masses and orbital separations balance to produce
+BHs with the correct masses that merge within a Hubble time.
 %\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
@@
 % Validity paragraph
-To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
-We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
-In particular, we compute KL divergence of the samples obtained from root-finding alone and after MCMC sampling.
-For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} nats on average if we use the root-finding result alone,
-whereas the MCMC result has a higher KL divergence at \kw{fill} nats.
-The KL divergence stays stable for reprojecting with different random seeds. 
-In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-finding process alone and the MCMC result agrees well with the original posterior,
-even though the agreement between the MCMC result and the original posterior is not perfect. 
-This degradation of the agreement comes from the fact that we are using a KDE to estimate the posterior density in the observable space.
-Note that the result from root-finding alone and the result from MCMC agrees well in the evolutionary parameter space.
-Compared to the result shown in a previous study, our result covers almost all the posterior of GW150914 in the masses space,
-which means our result is more consistent with the observation.
+\wf{What if we made this a subsection of ``methods'', and then just one sentence here about the values of the KL divergence for GW150914?}
+To check the validity of the recovered posterior over evolutionary parameters,
+it is essential to check whether it is consistent with the original posterior in
+the observable space. We reproject the recovered posterior to the observable
+space, and compare the KL divergence between the original posterior and the
+reprojected posterior. In particular, we compute KL divergence of the samples
+obtained from root-finding alone and after MCMC sampling. \wf{Clean up wording.}
+For GW150914, the KL divergence between the reprojected posterior and the
+original posterior is \kw{fill later} nats on average if we use the root-finding
+result alone, whereas the MCMC result has a higher KL divergence at \kw{fill}
+nats. The KL divergence stays stable for reprojecting with different random
+seeds. In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected
+posterior from both the root-finding process alone and the MCMC result agrees
+well with the original posterior, even though the agreement between the MCMC
+result and the original posterior is not perfect. This degradation of the
+agreement comes from the fact that we are using a KDE to estimate the posterior
+density in the observable space. Note that the result from root-finding alone
+and the result from MCMC agrees well in the evolutionary parameter space.
+Compared to the result shown in a previous study, our result covers almost all
+the posterior of GW150914 in the masses space, which means our result is more
+consistent with the observation.
 
@@
 % Paragraph discussing the potential application on the population side.
-To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
-Plotted in figure \ref{fig:GWTC3_f_acc_mass} is the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
-Each contour represents the $68\%$ credible interval of the posterior density for that particular event.
-Although it does not conclusively suggest a dependency between $f_{\rm acc}$ and $m_{1,{\rm GW}}$,
-there is a suggestive trend showing $f_{\rm acc}$ could possibly increase as the mass of the progenitor increases.
-This means the stable mass transfer (SMT) phase of a light binary is preferably non-conservative, whereas the heavier the events are, the more likely their SMT is conservative.
+To illustrate the potential benefit of our method on the population level, we
+perform the same analysis for all events in GWTC3. Plotted in figure
+\ref{fig:GWTC3_f_acc_mass} is the posterior density in the $m_{1,{\rm
+GW}}-f_{\rm acc}$ space for most of the events in GWTC3. Each contour represents
+the $68\%$ credible interval of the posterior density for that particular event.
+There is a suggestive trend showing $f_{\rm acc}$ could increase as the mass of
+the progenitor increases. Such a trend implies the stable mass transfer (SMT)
+phase of a less massive binary would be preferentially non-conservative, with
+more conservative mass transfer in more massive binary systems.
 
-Note that some events did not pass the KL divergence test we proposed in section \ref{sec:method}, which we do not include in figure \ref{fig:GWTC3_f_acc_mass}.
-Low mass events are more subjected to random fluctuation because natal kick is less modulated by fallback in lighter system,
-which can lead to a final system either with or without a merger event randomly,
-so the reprojected posterior may not agree with the original posterior in the observable space consistently, hence yielding a higher KL divergence on average.
-On the other hand, \texttt{COSMIC} struggles to produce events above the pair instability supernova mass cutoff $m_{1,{\rm GW}}$,
-so the posterior in the evolutionary parameter space only correspond part of the posterior in the observable space, therefore events beyond the PISN mass gap has higher KL divergence.
-Events with more extreme mass ratios are not likely to be recovered by the method. This could be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as discussed 
-in Section~\ref{sec:result}.
+Note that some events did not pass the KL divergence test we proposed in section
+\ref{sec:method}; we do not include these systems in figure
+\ref{fig:GWTC3_f_acc_mass}. Binaries that form low-mass compact objects have
+lower amounts of fallback and tend to have correspondingly larger variance in
+their random natal kicks. Larger kicks can unbind the progenitor binary, or lead
+to wide binaries that do not merge within a Hubble time.  In these cases, the
+extra variance during the reprojection can produce a posterior that may not
+agree with the original posterior, hence yielding a higher KL divergence.
 
-The point of figure \ref{fig:GWTC3_f_acc_mass} is to show this method could in principal reveal the correlation between progenitor parameters and hyperparameters on a population level.
-A careful treatment of all event in the catalog and discussion related the physical implication of figure \ref{fig:GWTC3_f_acc_mass} is beyond the scope of this paper,
-therefore we defer a detail study of the physics related to the population of GW events in the future.
-Also, despite some events do not pass the KL divergence test, we can still see the correlation in the region where the events are neither affected by randomness nor the PISN mass gap.
+For high-mass binaries, \texttt{COSMIC} struggles to produce events above the
+pair instability supernova mass cutoff $m_{1,{\rm GW}}$, so the posterior in the
+evolutionary parameter space only corresponds to part of the posterior in the
+observable space below this cutoff, and therefore events beyond the PISN mass
+gap also have higher KL divergence.
+
+Events with more extreme mass ratios are hard to produce with \texttt{COSMIC},
+and therefore less likely to be accurately recovered by our method. This could
+be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as
+discussed in Section~\ref{sec:result}.
+
+Merger observations for which \texttt{COSMIC} struggles are likely to be very
+informative about formation channels and binary evolution physics, and are
+therefore likely worthy of close study.  We anticipate that future work will
+follow up these events in detail.
+
+Figure \ref{fig:GWTC3_f_acc_mass} shows that our method could in principle
+reveal the correlation between progenitor parameters and hyperparameters on a
+population level. A careful treatment of all events in the catalog and
+discussion related to the detailed physical implications of figure
+\ref{fig:GWTC3_f_acc_mass} is beyond the scope of this paper.  We defer a
+detailed study of the physics related to the population of GW events to future
+work.
 
@@
 
-We present a new pathway to understand binary evolution with GW events in this paper.
-Instead of forward modelling an assumed distribution of initial binary to the observed population,
-we find the corresponding evolutionary parameters for each GW events. 
-In this work we first showcase the power of the proposed method with GW150914.
-We show the joint posterior of the two events' progenitor parameters and hyperparameters.
-In particular, to the authors' knowledge, this is the first time the correlation between progenitor parameters and hyperparameters is explored.
-Our work is not only a more efficient way to study GW events' progenitors,
-but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyperparameters. 
-In future work, we will extend our method to applying hyperparameters to individual binary components in order to better capture correlations between 
-hyperparameters and progenitor parameters.
+We present a new pathway to understand binary evolution with GW events in this
+paper. Instead of forward modelling an assumed distribution of initial binaries
+to the observed population, we find the corresponding evolutionary parameters
+event-by-event. In this first work, we showcase the power of the proposed method
+with an application to GW150914. We show the joint posterior of the event's
+progenitor parameters and hyperparameters.  Our work is both a more efficient
+way to study GW event progenitors, and also allows the possibility of
+constraining the astrophysics related to binary evolution, especially by
+capturing the correlation between hyperparameters in different systems.
 
-To further illustrate the power of the method, we show the inferred fraction of accretion during SMT depends on the progenitor masses.
-The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyperparameters for each event.
-This enables a data-driven way to study the distribution of hyperparameters,
-i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
-we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyperparameters,
-without making specific assumptions such as fixing the hyperparameters for all types of event.
+Our results suggest that the accretion efficiency during stable mass transfer
+may depend on the primary black hole mass.  In general, our method returns the
+joint posterior distribution of progenitor parameters and hyperparameters for
+each event, which enables a data-driven way to study the distribution of
+hyperparameters.  That is, once we have a catalog of events, each has their own
+posterior in the hyperparameter space, and we can employ well-known techniques
+such as hierarchical Bayesian analysis to fit a population model to the
+distribution of hyperparameters; this avoids making overly-specific assumptions
+like fixing the hyperparameters for all types of event.
 
 % Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
-Another perk of our method is the "rate" calculation is now independent of assumed initial condition such as a particular star formation rate model,
-which is a dominant factor of uncertainty for rate prediction in the past.
-Once we have pull back the GW event posterior from the observable space to the evolutionary parameter space,
-we can apply the same methods we use to account for the rate in the observable space.
-In the observable space, the overall rate is basically a normalization that is defined in the local universe i.e. the local merger rate. \kw{cite}
-In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
-This also means one can study the progenitor-formation rate in a data-driven manner instead of assuming a particular model.
+Working event-by-event as we do permits post-processing application of
+ arbitrarily complicated models of star formation and metalliticy evolution by
+ re-weighting the samples we obtain \wf{Cite at least Lieke here}. Once we have
+ pulled the GW event posterior back from the observable space to the
+ evolutionary parameter space, we can apply the same hierarchical inference
+ methods currently used to infer the compact binary population to our progenitor
+ population.  In particular, we can use our progenitor population to infer the
+ formation rate of compact binary population progenitor systems over cosmic time
+ without needing to make artifical and simplistic assumptions about the delay
+ time distribution or the metallicity-specific star formation rate \wf{again,
+ cite Lieke, and also Vitale, Farr, Ng, Rodriguez?}.  Comparing the formation
+ rate of progenitors to the star formation rate provides an avenue to check our
+ understanding of binary evolution and the relative rate of other formation
+ channels.
 

### 2022-05-23 src/tex/ms.tex
upgrade to v0.3

% Define document class
\documentclass[twocolumn]{aastex631}

\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{xcolor}
\usepackage{showyourwork}
\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
\newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
% \setlength{\floatsep}{0cm}
% \setlength{\textfloatsep}{0cm}
% \setlength{\belowcaptionskip}{-5pt}
% \setlength{\abovecaptionskip}{-5pt}
% Begin!
\begin{document}
\title{Backpropagating Gravitational wave events}

\pacs{}

\author{Kaze W. K. Wong} 
\email{kwong@flatironinstitute.org}
\flatiron

\date{\today}

\begin{abstract}
In this
\end{abstract}

%\maketitle

\section{Introduction}

As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{},
understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
A very common way to study the origin of detected GW events is through population synthesis simulation \kw{cite},
which evolves a population of binaries to form a population of GW events.
Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
such as the initial mass function or star-formation history,
then evolve each individual members according to the physics of interest until some stopping criteria is met.
This usually mean either a merger event or the binary system stays stable for a Hubble time.
After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.

\kw{Shorten}
While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
it is not necessarily the most efficient way to constrain the underlying physical model.
Most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
This means if we want to evolve a binary to match the observation by pure trial and error, a large fraction of computation will be wasted in trying initial binary configurations that are not related to the observation.
To alleviate this inefficiency, the community often takes a population approach --- instead of trying to match binary systems to observations event by event,
one can draw systems from an initial population, forward model each member in that population then match it to the observed population properties.
However, this approach is essentially trading one inefficiency with another, since we still need to choose an initial distribution of binary parameters.
For relatively well-known source properties such as initial masses this is less of a problem,
but the distribution of some less well-studied properties such as common envelope efficiency is often heuristically chosen to be some simple distribution such as a delta function at some value, i.e. a fixed value for the entire population.
This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution with other independent studies,
or they should vary the family of distributions.
In general, it is unclear how one can cover all possible distribution that may agree with the observation in generating a suite of simulations.
The problem worsen as we increase the dimensionality of the parameter space.
\wf{This says what you want to say, but it's a bit wordy right now---I'll return to it, but it would be good to tighten up the language a lot.}

Indeed, one of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
Even with recently development on using emulators to speed up the simulation process, training the emulators also require a decent amount of samples in the parameter space of interest.
For example, say we want to emulate a model with 10 parameters, this simplest option for constructing a bank of simulations for training is to lay down a grid. Along each axis we put the barely minimum of $2$ simulations, this would still require $1024$ simulations.
The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
This means even training an emulator is impossible in high dimensional space.

In this paper, instead of forward modelling a population of binary star to match the observed GW events population,
we propose a method to "backward" model each GW event to its progenitor state.
The idea is since we can compute the GW event properties of a guess by forward modelling its progenitor,
we can define the difference between the GW event properties of this guess and the observed GW event properties.
By minimizing this difference, i.e. solving a root finding problem, we can hope to recover the progenitor properties.

This paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.

\section{Method Rewrite}
\label{sec:method-rewrite}

\wf{Re-writing the methods section with a new viewpoint, gonna try to eventually merge the content.}

We define the parameters that describe the progenitor's properties such as ZAMS
masses and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
that control different physical prescriptions such as wind strength and common
envelope efficiency as hyper-parameters $\bm{\lambda}$, and random variables
that affect certain stochastic process such as whether natal kick unbinds a
binary as $\bm{X}$. To avoid clutter in the following derivation, we denote all
the parameters related to mapping a particular progenitor system into a GW event
collectively as evolutionary parameters $\bm{\Theta}$, which includes
$\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

Once all the parameters including a particular draw of all the random variables
represented by $\bm{X}$ are known, the population synthesis code is a
deterministic function that transforms the properties of the progenitors into
the GW parameters $\bm{\theta}$:
\begin{equation}
    \bm{\theta} = \bm{f}\left( \bm{\Theta} \right).
\end{equation}
The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ is many-to-one.  
In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
need to be able to evaluate the likelihood of that data at fixed progenitor
parameters, 
\begin{equation}
    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right),
\end{equation}
because the likelihood depends only on the gravitational waveform generated by
parameters $\bm{\theta}$ to which the progenitor parameters map.  In principle
this likelihood could be computed at arbitrary vaues of $\bm{\Theta}$ using the
same machinery that is used to estimate source parameters in gravitational wave
catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
of progenitor parameters, hyper-parameters, and random variables in
$\bm{\Theta}$ from a posterior density.

However, it is very expensive to couple progenitor evolution to stochastic
sampling in this way.  Because the gravitational wave likelihood function
ultimately depends only on the GW parameters $\theta$, and catalogs have already
produced samples over these parameters from a posterior density 
\begin{equation}
    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) p_\mathrm{GW} \left( \bm{\theta} \right),
\end{equation}
where $p_\mathrm{GW}$ is the prior density used for sampling, it is much more
efficient to \wf{WIP}.

\section{Method}
\label{sec:method}

We define the parameters that describe the progenitor's properties such as ZAMS masses and eccentricity as progenitor parameters $\bm{\theta'}$,
the parameters that control different physical prescriptions such as wind strength and common envelope efficiency as hyper-parameters $\bm{\lambda}$,
and random variables that affect certain stochastic process such as whether natal kick unbinds a binary as $\bm{X}$.
To avoid clutter in the following derivation,
we denote all the parameters related to mapping a particular progenitor system into a GW event collectively as evolutionary parameters $\bm{\Theta}$,
which includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

Once all the parameters including a particular draw of all the random variables are known,
the population synthesis code is a deterministic function that transforms the properties of the progenitors into the GW parameters.
Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\Theta}$ is given by

\begin{align}
    p(\bm{\theta}|\bm{\Theta}) &= \bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta}),
\end{align}

where $\bm{f}$ represents the evolution encoded by the population synthesis model with fixed random parameters $\bm{X}$.

While it is straightforward to write down the "forward" probability, the "backward" probability is more complicated since multiple evolutionary parameters can produce the same GW event.
Using Bayes theorem, the probability of an event with evolutionary parameters given the GW parameters can be derived as following

\begin{align}
    p(\bm{\Theta}|\bm{\theta}) &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{p(\bm{\theta})}, \nonumber \\
    &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{\int p(\bm{\theta}|\bm{\Theta}) p(\bm{\Theta}) d\bm{\Theta}}, \nonumber \\
    &= \frac{\bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}{\int d\bm{\theta} \bm{\delta}(f(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}.
\label{eq:backward_prob}
\end{align}

Using the result in eq.\ref{eq:backward_prob},
the probability of an event having evolutionary parameters given data can be obtained by marginalizing over the GW parameters

\begin{align}
    p(\bm{\Theta}|d) &= \int p(\bm{\Theta}|\bm{\theta}) p(\bm{\theta}|d) d\bm{\theta}\\
    &= \frac{1}{N}\sum_{i=1}^{N} p(\bm{\Theta}|\bm{\theta}_i). \label{eq:evolutionary_posterior}
\end{align}

The term $p(\bm{\theta}|d)$ is the inferred posterior probability for a particular event, which is usually presented by the LVK collaboration in the form of posterior samples.
In eq.\ref{eq:evolutionary_posterior}, we use importance sampling to turn the integration into a sum over the posterior samples.

For each posterior sample point, we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.

\begin{align}
\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2
\label{eq:loss}
\end{align}

In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the observable space.
However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$ to be our threshold.
To make sure we find a reasonably complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
As long as the solution fulfill the acceptance criteria, the root is as valid as all other roots that fulfill the same acceptance criteria, regardless to its initial guess.
In other word, we have the freedom to choose the set of initial guesses however we think it might benefit the optimization.
Most of the points in the evolutionary parameter space do not produce mergers, and for systems that do not merge, we set the binary masses to be 0.
This means the gradient for systems that do not merge is 0, and the root finder would likely to get stuck in such state during the first step.\footnote{On a plateau of constant loss, the root finder performs no better than a random guess. In a high dimensional space, this means the root finder will need a long time to get close to a solution.}
Therefore, it is beneficial to put initial guesses in region of the evolutionary parameter space that does produce merger.
To construct a list of initial guesses, we evolve a set of points that we uniformly sampled from the evolutionary parameter space, then keep the points that merge within a Hubble time.

Ideally we would like to have explicit control over all evolutionary parameters.
In particular, having explicit control over random variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicitly,
i.e. the random variables are drawn within the program, instead of being arguments of the program.
This means the function we use to evolve the binary is not a deterministic function but a stochastic function instead. 
Hence, even if the root-finding algorithm performs perfectly,
forward modelling this set of roots does not guarantee to reproduce the set of posterior samples due to randomness of the evolution function.
To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,
we forward model ("Reproject") the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
We use KL divergence to measure the agreement between the two posterior distributions, which is given by
\begin{align}
D_{KL}(P||Q) = \int P(\bm{x}) \log(P(\bm{x})/Q(\bm{x})) dx.
\label{eq:KLdivergence}
\end{align}

A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the observable space, hence it is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
This is completely expected since \texttt{COSMIC} comes with its own assumptions and is expected to fail in reproducing a subset of the event in GWTC3, such as event with at least one component beyond the pair instability supernova massgap.
On the other hand, the reprojection could subject to stochasticity in the evolution function we use.
Instead of reprojecting the posterior only once, we reproject the posterior multiple times with different random seeds and check whether the KL divergence varies significantly.
A varying KL divergence means this particular event is subjected to randomness in the evolution function, hence one should take extra caution in interpreting the result.

On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
The two results agree well with each other.
This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.less
\section{Result}
\label{sec:result}

\begin{enumerate}
\item Raw posterior of successful events.
\item Trace back the event's $z_{\rm ZAMS}$.
\item Reproject to observable space
\item Evidence ratio?
\item KL divergence?
\end{enumerate}

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l }
    \hline
    \hline
    Parameters &  Description & Optimization range\\
    \hline
    \hline
    Observables &\ &\  \\
    \hline
    \hline
    $M_{\rm GW,1}$ & Primary mass of the GW event & NA \\
    $M_{\rm GW,2}$ & Secondary mass of the GW event  & NA\\
    \hline
    \hline
    Progenitor parameters &\ &\  \\
    \hline
    \hline
    $M_{\rm ZAMS,1}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $M_{\rm ZAMS,2}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
    $e$ & Eccentricity at ZAMS & $[0,1]$\\
    $Z$ & Metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
    $z$ & Redshift at ZAMS & NA\\
    \hline
    \hline
    Hyper-parameters &\ &\ \\
    \hline
    \hline

    $\alpha$ & Common envelope efficiency & $[0.1,10]$\\
    $f_{\rm acc}$ & Fraction of accretion during stable mass transfer & $[0,1]$\\
    $q_{\rm crit,3}$ & \kw{Katie fill in} & $[0.1,10]$\\
    $\sigma$ & \kw{Katie fill in}& $[10,300]\ {\rm km/h}$\\

    \hline
    \hline
    \end{tabular}
    \caption{A list of parameters we used in this study.}
    \label{tab:parameters}
    \end{center}
\end{table*}

We apply the method described in section \ref{sec:method} to the first GW events GW150914.
The parameters involved in the analysis and the range we allow for finding the root are tabulated in table \ref{tab:parameters}.
\kw{@Katie state why not spin here.}
Therefore, we solve the rootfinding problem with only the two component masses in the observable space.
For progenitor parameters, we characterize each guess with five parameters: the two component masses $M_{\rm ZAMS,1}$ and $M_{\rm ZAMS,2}$, the orbital period $t_{\rm orb}$, the eccentricity $e$, the metallicity $Z$ at ZAMS.
Note that the progenitor formation redshift is not obtained through the root-finding process.
This is because the formation redshift does not affect the evolution of the binary, so we do not need to include in the optimization.
Instead, we compute the time it takes for the binary to evolve from formation to the merger, then we add this time back to the lookback time when this event is observed.
Using the total lookback time, we can compute the redshift of the binary at formation.
We adopt the cosmological model in \kw{cite Planck 18} to compute the redshift of the binary at formation.
For hyper-parameters, we choose parameters that affect binary evolution in \texttt{COSMIC} mostly significantly, including the common envelope efficiency $\alpha$, the fraction of accretion during stable mass transfer $f_{\rm acc}$, the critical mass fraction $q_{\rm crit,3}$, and the accretion rate $\sigma$.
We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.

\kw{Recover ratio of the posterior sample goes here.}

The full posterior is available in appendix \kw{Or we should put a link to the dataset?}

\begin{figure*}[h]
\includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed.pdf}
\caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
$M_1$ and $M_2$ are the progenitors' masses. $\log{Z}$ is log metallicity at ZAMS.
$z$ is the redshift at ZAMS. 
$\alpha$ is the common envelope efficiency.
$f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
$q_{\rm crit 3}$ are the critical mass during \kw{Katie fill in},
Note that the redshift is not fitted during the root finding process.
Once we have find the evolutionary parameters, we add the time it would take the progenitor to merge to the lookback time of the observed posterior sample,
then from the total lookback time we can compute the redshift at ZAMS.
}
\label{fig:GW150914_posterior}
\end{figure*}

To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} on average.
The KL divergence stays stable for reprojecting with different random seeds.

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
\caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
The blue contour is the posterior reprojected using the samples in the evolutionary parameter space.
The orange contour is the posterior samples from LVK that have corresponding samples in the evolutionary parameter space.
The green contour is the posterior plotted using all of LVK posterior samples.
}
\label{fig:GW150914_reprojection}
\end{figure}

We can see the reprojected posterior agrees very well with the original posterior in fig.\ref{fig:GW150914_reprojection}.
\kw{Qoute the KL divergence.}

% \begin{figure}
% \includegraphics[width=0.49\textwidth]{figures/GW150914_CE_SMT.pdf}
% \caption{}
% \label{fig:GW150914_CE_SMT}
% \end{figure}
% \kw{Add figure of stable mass transfer vs common envelope}

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GW150914_m1m2alpha.png}
\caption{The posterior samples of GW150914 in the progenitor masses space, colorcoded by the common envelope efficiency $\alpha$ for that sample.}
\label{fig:GW150914_m1_m2_alpha}
\end{figure}

The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
we would miss the samples that do not match the chosen value of $\alpha$.
By folding in the hyper-parameters into the root-finding algorithm,
we can capture the correlation between 
\kw{Dartboard comparsion} Comparing our result with 
\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}

\section{Discussion}
\label{sec:discussion}

We present a new pathway to understand binary evolution with GW events in this paper.
Instead of forward modelling an assumed distribution of initial binary to the observed population,
we solve a root-finding problem to find the best initial binary parameters that reproduce the posterior distribution for each event. 
The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
This enables a data-driven way to study the distribution of hyper-parameters,
i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters,
without making specific assumptions such as fixing the hyper-parameters for all types of event.

In this work we showcase the power of the proposed method with two events, GW150914 and GW191129.
We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
In particular, to the authors knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is shown.
Our work is not only a more efficient way to study GW events' progenitors,
but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.

% Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
Another perk of our method is the "rate" calculation is now independent of assumed initial condition such as a particular star formation rate model,
which is a dominant factor of uncertainty for rate prediction in the past.
Once we have pull back the GW event posterior from the observable space to the evolutionary parameter space,
we can apply the same methods we use to account for the rate in the observable space.
In the observable space, the overall rate is basically a normalization that is defined in the local universe i.e. the local merger rate. \kw{cite}
In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
\kw{Talk about prior are now in post-processing instead of pre-processing.}

% Single formation channel and fundamental limitation on that.
While our method allows data-driven exploration of the hyper-parameters space for the first time, there are a number of improvements can be implemented in future studies.
In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
For example, event with either of the component mass larger than the PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3.
As pointed out in the literature \kw{Cite mike}, it is unlikely that all the observed events in GWTC3 can be explained by a single formation channel.
As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.
Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
In this paper, our main focus is to illustrate the concept of "back-propagating" GW events posterior sample in this work,
highlighting the capability of our method and motivating the community to build the next generation population synthesis code that can work with our method.
To avoid cluttering of focus, we discuss the physical implication of the result presented in this work under our specific assumption (i.e. using \texttt{COSMIC} as our evolutionary function) in a companion paper.

% Randomness paragraph
Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
This introduces significant inefficiency in our root-finding algorithm.
The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolution pathway of low mass events such as BNS events.
The effect of natal kick is suppressed for heavier mass events due to fallback.
This means events with the root-finding process for lighter masses are subject to stochasticity of the function,
where the root-finding process for heavier events behaves as if the evolutionary function we use are deterministic.
Indeed, we see this behavior in a couple of case studies.
For heavy event such as GW150914, the reprojected posterior agrees with the data posterior.
For light event such as GW170817, although during the root-finding process the convergence criteria is met,
the reprojected posterior does not agree with the data posterior.
Also, due to computational limitation, we only try 1000 different initial guesses per posterior sample in the root-finding process.
This means any posterior sample that has a probability of merging rarer than 1 in 1000 could be missed.
\kw{comment on the low mass ratio end.}
Obviously the problems which comes with the randomness can be alleviated by performing more tries per posterior sample,
but this is not scalable in practice.
On top of limitation in efficiency, some formation channels require explicit control of random variables by construction.
For example, in a dynamical formation scenario such as binaries that form in a globular cluster,
each binary has some probability of undergoing a multi-body encounter with another member in the cluster.
These encounter probability distributions are either studied with direct N-body simulations or semi-analytical methods.
In both cases, each member of the cluster is no longer completely independent of the other members, but coupled through the encounter probability distribution.
By studying the encounter probability distribution, we can infer the properties of the environment which the binary lives in.
This can only be done if we have explicit control over the random variables that characterize the encounter probability distribution.

% Need for selection bias to recover intrinsic distribution

% Gradient decending in the evolutionary parameter space is much more efficent than rejection sampling.
% Potential pitfall of interpreting this result.
% Need for full autodiff.

By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
This greatly benefit the efficiency of the root-finding process.
On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
the relatively weighting of these roots becomes ambiguous when the number of roots is large.
For example, should we consider roots that are close to each other a single root or multiple roots?
If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.

The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
We leave a detail comparison between this method and direct sampling to future studies.

Another technical note is we use finite differencing to estimate the gradient of the objective function, which could be a significant source of error near transition points in the evolutionary parameter space.
Also, finite differencing is increasing inefficiency as we increase the dimensionality of the problem.
To improve the accuracy and efficiency in estimating the gradient of objective function, automatic differentiation is a promising feature that modeler should consider incorporating in their population synthesis code in the future.

\kw{Mention potential complimentary studies of direct sampling, and how all these suggestions are still useful regardless of whether we are doing root-finding or not.}

To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
We point out hyper-parameters in the usual population synthesis simulation context are not actually parameters related to the population,
but parameters about the evolutionary function.
This means the binary evolution functions can be constrained on an individual event basis.
We "back propagate" the posterior in the observable space to the evolutionary parameter space,
thus allowing us to study hyper-parameters and its correlation with progenitor parameters in a data-driven manner.
Our method makes less assumption than the traditional forward modelling approach,
which often fix the hyper-parameters across the entire population.
Since we are not limited to the fix hyper-parameters assumption, we can explore the behavior of the hyper-parameters across the population much more efficiently.
While our work lay down a data analysis pathway to understand the population of GW events,
no physics can be learned without a comprehensive complimentary physical model.
We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:
first, they should have explicit control over the random variables so marginalizing over random variables can be done more precisely;
second, they should be as automatically differentiable as much as possible so exploring the evolutionary parameter space is efficient.
Combining this work and the next generation population synthesis code,
we can explore the full parameter space of binary evolution models with the next-generation GW detectors network in the near future.

% Interpretation:
% \begin{enumerate}
% \item Highlight the fact that we turn hyper-parameters into parameters.
% \item Discuss this is need for high dimensional exploration.
% \end{enumerate}

% Caveats: 
% \begin{enumerate}
% \item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
% \item We only assume BSE
% \item Caveat on marginalizing over random samples.
% \item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
% \item How next generation "Progenitor synthesis" code should be built.
% \item Rate
% \item Failure studies go into appendix
% \end{enumerate}

\section{Acknowledgement}

\bibliography{bib}

\end{document}

### 2022-05-23 src/scripts/GW150914_posterior.py
figures working

@@ -2,7 +2,7 @@ import numpy as np
 import h5py
 import matplotlib.pyplot as plt
 import copy
-
+import paths
 import corner
 import matplotlib as mpl
 from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
@@ -73,9 +73,9 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-data = h5py.File('../data/GW150914_progenitor_posterior.hdf5','r')
-processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5','r')
-GWTC = np.load('../data/GW150914_posterior.npz')
+data = h5py.File(paths.data / 'GW150914_progenitor_posterior.hdf5','r')
+processed_data = h5py.File(paths.data / 'GW150914_progenitor_posterior_reproject.hdf5','r')
+GWTC = np.load(paths.data / 'GW150914_posterior.npz')
 m1_obs = np.median(GWTC['m1'])
 m2_obs = np.median(GWTC['m2'])
 z_obs = np.median(GWTC['z'])
@@ -131,4 +131,4 @@ for i in range(Ndim):
         ax.tick_params(axis="both",direction="in",which='minor',width=1.5,length=3,right=True,top=True)
         ax.xaxis.set_minor_locator(AutoMinorLocator())
         ax.yaxis.set_minor_locator(AutoMinorLocator())
-fig.savefig('./GW150914_corner.pdf',bbox_inches='tight')
+fig.savefig(paths.figures / 'GW150914_corner.pdf',bbox_inches='tight')

### 2022-05-23 src/scripts/GW150914_reprojection.py

@@ -2,7 +2,7 @@ import numpy as np
 import h5py
 import matplotlib.pyplot as plt
 import copy
-
+import paths
 import matplotlib as mpl
 import corner
 from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
@@ -72,9 +72,9 @@ def get_posterior(value, coord,cut = 1e-2):
     index = np.where((b<criteria))
     return a[index],b[index],weight[index],index
 
-data = h5py.File('../data/GW150914_progenitor_posterior.hdf5','r')
-GWTC = np.load('../data/GW150914_posterior.npz')
-processed_data = h5py.File('../data/GW150914_progenitor_posterior_reproject.hdf5','r')
+data = h5py.File(paths.data / 'GW150914_progenitor_posterior.hdf5','r')
+GWTC = np.load(paths.data / 'GW150914_posterior.npz')
+processed_data = h5py.File(paths.data / 'GW150914_progenitor_posterior_reproject.hdf5','r')
 m1_obs = np.median(GWTC['m1'])
 m2_obs = np.median(GWTC['m2'])
 z_obs = np.median(GWTC['z'])
@@ -106,4 +106,4 @@ custom_lines = [Line2D([0], [0], color='C0', lw=4),
                 Line2D([0], [0], color='C1', lw=4),
                 Line2D([0], [0], color='C2', lw=4)]
 plt.legend(custom_lines, ['Reprojected','Corresponding','Data'],fontsize=20)
-plt.savefig('GW150914_reprojection.pdf')
+plt.savefig(paths.figures / 'GW150914_reprojection.pdf')

### 2022-05-23 src/figures/GW150914_posterior.py
add plotting script

@@ -1,134 +1,0 @@
-import numpy as np
-import h5py
-import matplotlib.pyplot as plt
-import copy
-
-import corner
-import matplotlib as mpl
-from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
-                               AutoMinorLocator)
-from matplotlib.lines import Line2D
-
-params = {'axes.labelsize': 32,
-          'font.family': 'serif',
-          'font.serif': 'Computer Modern Raman',
-          'font.size': 32,
-          'axes.linewidth': 2,
-          'legend.fontsize': 20,
-          'xtick.labelsize': 20,
-          'xtick.top': True,
-          'xtick.direction': "in",
-          'ytick.labelsize': 20,
-          'ytick.right': True,
-          'ytick.direction': "in",
-          'axes.grid' : False,
-          'text.usetex': True,
-          'savefig.dpi' : 100,
-          'lines.markersize' : 14,
-#          'axes.formatter.useoffset': False,
-          'axes.formatter.limits' : (-3,3)}
-
-#mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
-
-mpl.rcParams.update(params)
-
-def fill_corner(fig,data,ndim,axis,color='C0',true_val=None,zorder=1,lw=3,style='-',smooth1d = 0.01, smooth2d=0.01,no_fill_contours=False,fill_contours=True,alpha=1,levels=[0.68,0.95]):
-    for i in range(ndim):
-        for j in range(i+1):
-            ax = fig.axes[np.sum(np.arange(i+1))+j]
-            if i==j:
-                ax.hist(data[i],bins=axis[i],histtype='step',lw=lw,color=color,density=True,zorder=zorder)
-                ylim = ax.get_ylim()
-                ax.set_xlim(axis[i][0],axis[i][-1])
-                ax.set_ylim(ylim[0],ylim[1])
-                if true_val is not None:
-                    ax.vlines(true_val[i],ylim[0],ylim[1],color=color,lw=lw,zorder=zorder)
-            elif j<i:
[…]

### 2022-05-23 src/figures/GW150914_reprojection.py

@@ -1,109 +1,0 @@
-import numpy as np
-import h5py
-import matplotlib.pyplot as plt
-import copy
-
-import matplotlib as mpl
-import corner
-from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
-                               AutoMinorLocator)
-from matplotlib.lines import Line2D
-
-params = {'axes.labelsize': 32,
-          'font.family': 'serif',
-          'font.serif': 'Computer Modern Raman',
-          'font.size': 32,
-          'axes.linewidth': 2,
-          'legend.fontsize': 20,
-          'xtick.labelsize': 20,
-          'xtick.top': True,
-          'xtick.direction': "in",
-          'ytick.labelsize': 20,
-          'ytick.right': True,
-          'ytick.direction': "in",
-          'axes.grid' : False,
-          'text.usetex': True,
-          'savefig.dpi' : 100,
-          'lines.markersize' : 14,
-#          'axes.formatter.useoffset': False,
-          'axes.formatter.limits' : (-3,3)}
-
-#mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
-
-mpl.rcParams.update(params)
-
-def fill_corner(fig,data,ndim,axis,color='C0',true_val=None,zorder=1,lw=3,style='-',smooth1d = 0.01,no_fill_contours=False,fill_contours=True,alpha=1,levels=[0.68,0.95]):
-    for i in range(ndim):
-        for j in range(i+1):
-            ax = fig.axes[np.sum(np.arange(i+1))+j]
-            if i==j:
-                ax.hist(data[i],bins=axis[i],histtype='step',lw=lw,color=color,density=True,zorder=zorder)
-                ylim = ax.get_ylim()
-                ax.set_xlim(axis[i][0],axis[i][-1])
-                ax.set_ylim(ylim[0],ylim[1])
-                if true_val is not None:
-                    ax.vlines(true_val[i],ylim[0],ylim[1],color=color,lw=lw,zorder=zorder)
-            elif j<i:
[…]

### 2022-05-23 src/figures/figure1_cornerPlot.py

@@ -1,0 +1,235 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
+                               AutoMinorLocator)
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+
+
+params = {'axes.labelsize': 32,
+          'font.family': 'serif',
+          'font.serif': 'Computer Modern Raman',
+          'font.size': 32,
+          'axes.linewidth': 2,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 20,
+          'xtick.top': True,
+          'xtick.direction': "in",
+          'ytick.labelsize': 20,
+          'ytick.right': True,
+          'ytick.direction': "in",
+          'axes.grid' : False,
+          'text.usetex': True,
+          'savefig.dpi' : 100,
+          'lines.markersize' : 14,
+          'axes.formatter.limits' : (-3,3)}
+
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+
+mpl.rcParams.update(params)
+
+def fill_corner(fig,data,ndim,axis,color='C0',true_val=None,zorder=1,lw=3,style='-',smooth1d = 0.01, smooth2d=0.01,no_fill_contours=False,fill_contours=True,alpha=1,levels=[0.68,0.95]):
+    for i in range(ndim):
+        for j in range(i+1):
+            ax = fig.axes[np.sum(np.arange(i+1))+j]
+            if i==j:
+                ax.hist(data[i],bins=axis[i],histtype='step',lw=lw,color=color,density=True,zorder=zorder)
+                ylim = ax.get_ylim()
+                ax.set_xlim(axis[i][0],axis[i][-1])
+                ax.set_ylim(ylim[0],ylim[1])
+                if true_val is not None:
[…]

### 2022-05-23 src/figures/figure1_cornerPlot.py
Update figure 1 and add figure 3

@@ -95,17 +95,17 @@ def sample_z(kde,m1m2):
     return kde_mass[2,np.argmin(distance,axis=0)]
     
 i = 0
-processed_data = h5py.File('../data/GWTC3_processed.hdf5','r')
+processed_data = h5py.File('../data/GWTC3_mcmc_processed.hdf5','r')
 
 m1_pred = processed_data['m1'][i]
 m2_pred = processed_data['m2'][i]
-t_merge = processed_data['t_merge'][0][()]/1000
+t_merge = processed_data['t_merge'][i][()]/1000
 
 posterior_array = []
 observable_array = []
 quantile_array = []
 
-posterior_array.append(processed_data['coord_array'][0][()].T)
+posterior_array.append(processed_data['coord_array'][i][()].T)
 m1m2_array = np.stack([m1_pred[()],m2_pred[()]]).T
 obs_kde = gaussian_kde(observables[0].T)
 

### 2022-05-23 src/figures/figure3_sanityCheck.py

@@ -1,0 +1,109 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
+                               AutoMinorLocator)
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+
+
+params = {'axes.labelsize': 32,
+          'font.family': 'serif',
+          'font.serif': 'Computer Modern Raman',
+          'font.size': 32,
+          'axes.linewidth': 2,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 20,
+          'xtick.top': True,
+          'xtick.direction': "in",
+          'ytick.labelsize': 20,
+          'ytick.right': True,
+          'ytick.direction': "in",
+          'axes.grid' : False,
+          'text.usetex': True,
+          'savefig.dpi' : 100,
+          'lines.markersize' : 14,
+#          'axes.formatter.useoffset': False,
+          'axes.formatter.limits' : (-3,3)}
+
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+
+mpl.rcParams.update(params)
+
+def get_posterior(value, coord,cut = 1e-2):
+    a = copy.deepcopy(coord)
+    a[...,0] = 10**a[...,0]
+    a[...,1] = 10**a[...,1]
+
+    a[...,[1,0]] = np.sort(a[...,[0,1]],axis=2)
+    a[...,2] = a[...,2]
+    a[...,4] = a[...,4]
+    a[...,7] = a[...,7]
+    b = value
+
+    criteria = cut
+
+    count = np.sum((b<criteria).astype(int),axis=1)
+    weight = np.ones(b.shape)/count[:,None]
+    weight[np.invert(np.isfinite(weight))] = 0
+
+    index = np.where((b<criteria))
+    return a[index],b[index],weight[index],index
+
+
+GWTC = np.load('../data/GWTC3_LVK_posterior.npz')
[…]

### 2022-05-23 src/figures/figure4_population.py
add figure 4

@@ -1,0 +1,67 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
+                               AutoMinorLocator)
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+
+
+params = {'axes.labelsize': 32,
+          'font.family': 'serif',
+          'font.serif': 'Computer Modern Raman',
+          'font.size': 32,
+          'axes.linewidth': 2,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 20,
+          'xtick.top': True,
+          'xtick.direction': "in",
+          'ytick.labelsize': 20,
+          'ytick.right': True,
+          'ytick.direction': "in",
+          'axes.grid' : False,
+          'text.usetex': True,
+          'savefig.dpi' : 100,
+          'lines.markersize' : 14,
+#          'axes.formatter.useoffset': False,
+          'axes.formatter.limits' : (-3,3)}
+
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
+
+mpl.rcParams.update(params)
+
+posterior_array = []
+data = h5py.File('../data/GWTC3_mcmc_processed.hdf5','r')
+for i in range(data['coord_array'].shape[0]):
+        evo_params = data['coord_array'][i][6]
+        joint_dist = np.stack([data['m1'][i][()],evo_params],axis=-1)
+        posterior_array.append(joint_dist)
+
+posterior_array = np.array(posterior_array)
+evo_params_array = np.delete(posterior_array,29,axis=0)
+fig, ax = plt.subplots(1,1,figsize=(10,9))
+
+for i in range(data['coord_array'].shape[0]):
+    try:
+        if i != 12: # Event number 12 has significant instability in KL divergence, that it is removed from the analysis. 
[…]

### 2022-05-23 src/figures/figure2_varyAlphaFacc.py
add figure2

@@ -1,0 +1,62 @@
+import pandas as pd
+import numpy as np
+
+import matplotlib.pyplot as plt
+from matplotlib import colors
+import seaborn as sns
+
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["mathtext.fontset"] = "dejavuserif"
+
+models = ['A22_02', 'A22', 'A22_5', 'KW']
+
+def get_data(model):
+    if model == 'KW':
+        bpp = pd.read_hdf('{}_rerun.h5'.format(model), key='bpp')
+        initC = pd.read_hdf('{}_rerun.h5'.format(model), key='initC')
+    else:    
+        bpp = pd.read_hdf('{}_rerun.hdf'.format(model), key='bpp')
+        initC = pd.read_hdf('{}_rerun.hdf'.format(model), key='initC')
+    
+    mergers = bpp.loc[(bpp.kstar_1 == 14) & (bpp.kstar_2 == 14) & (bpp.evol_type == 3)]
+    formation = bpp.loc[(bpp.bin_num.isin(mergers.bin_num)) & (bpp.kstar_1 == 14) & (bpp.kstar_2 == 14)].groupby('bin_num').first().reset_index()   
+    bpp = []
+    return initC.loc[initC.bin_num.isin(mergers.bin_num)], mergers, formation
+
+initC_list = []
+merger_list = []
+formation_list =[]
+for m in models:
+    dat = get_data(model=m)
+    initC_list.append(dat[0])
+    merger_list.append(dat[1])
+    formation_list.append(dat[2])
+
+labels = [r'$\alpha=0.2$, $f_{\rm{acc}}=0.5$', 
+          r'$\alpha=1$, $f_{\rm{acc}}=0.5$', 
+          r'$\alpha=5$, $f_{\rm{acc}}=0.5$', 
+          r'$\alpha=\rm{var}$, $f_{\rm{acc}}=\rm{var}$']
+
+fig, axes = plt.subplots(nrows=1, ncols=4, figsize=(16, 3))
+
+for ax, initC, label, ii in zip(axes, initC_list, labels, range(len(labels))):
+    p, x_edges, y_edges, im = ax.hist2d(x=initC.mass_1.values, y=initC.mass_2.values, 
+                                        bins=[np.linspace(50, 150, 50), np.linspace(35, 150, 50)], 
[…]

### 2022-05-23 src/scripts/GW150914_posterior.py
disable tex

@@ -22,7 +22,7 @@ params = {'axes.labelsize': 32,
           'ytick.right': True,
           'ytick.direction': "in",
           'axes.grid' : False,
-          'text.usetex': True,
+          'text.usetex': False,
           'savefig.dpi' : 100,
           'lines.markersize' : 14,
 #          'axes.formatter.useoffset': False,

### 2022-05-23 src/ms.tex
citation explosion

@@
 progenitor populations of GW mergers is through population synthesis simulations
-for a host of formation environments including isolated binary stars, triple
-systems, globular clusters, nuclear star clusters, AGN disks, and more
-\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
+for a host of formation environments. In the case of isolated binary evolution, merging 
+double compact objects can form from massive binary stars through the
+standard stable mass transfer or common envelope channels 
+\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022} 
+as well as through chemically homogenous evolution \citep{Mandel2016, Marchant2016, deMink2016} 
+or with population III stars \citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017, Tanikawa2021, Tanikawa2022}.
+Merging DCO binaries can also originate from a wide variety of dynamically active environments including 
+triple (or higher multiple) systems \citep[e.g. ][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021}, 
+as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017, DiCarlo2020, Chattopadhyay2022}, 
+globular clusters \citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014, Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019}, 
+nuclear star clusters \citep[][]{Miller2009, Antonini2016}, or a mix of all three \citep[e.g.][]{Mapelli2022}.
+Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions 
+\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black hole formation \textbf{KB add these citations} can lead 
+to merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 
+formation envrionment and channels within, see \citet{Mandel2022}. Traditionally in these studies, the
 population is studied through a Monte-Carlo approach.  Using theoretically
@@
 it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
-One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
+One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often 
 cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation

### 2022-05-23 src/figures/figure3_sanityCheck.py
update to new version of syw [skip ci]

@@ -1,109 +1,0 @@
-import numpy as np
-import h5py
-import matplotlib.pyplot as plt
-from scipy.interpolate import interp1d
-from scipy.integrate import cumtrapz
-from scipy.optimize import curve_fit
-import copy
-from astropy.cosmology import Planck18
-
-import matplotlib as mpl
-import corner
-from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
-                               AutoMinorLocator)
-from matplotlib.lines import Line2D
-from scipy.stats import gaussian_kde
-from scipy.spatial.distance import pdist, cdist
-
-
-params = {'axes.labelsize': 32,
-          'font.family': 'serif',
-          'font.serif': 'Computer Modern Raman',
-          'font.size': 32,
-          'axes.linewidth': 2,
-          'legend.fontsize': 20,
-          'xtick.labelsize': 20,
-          'xtick.top': True,
-          'xtick.direction': "in",
-          'ytick.labelsize': 20,
-          'ytick.right': True,
-          'ytick.direction': "in",
-          'axes.grid' : False,
-          'text.usetex': True,
-          'savefig.dpi' : 100,
-          'lines.markersize' : 14,
-#          'axes.formatter.useoffset': False,
-          'axes.formatter.limits' : (-3,3)}
-
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
-
-mpl.rcParams.update(params)
-
-def get_posterior(value, coord,cut = 1e-2):
-    a = copy.deepcopy(coord)
-    a[...,0] = 10**a[...,0]
-    a[...,1] = 10**a[...,1]
-
-    a[...,[1,0]] = np.sort(a[...,[0,1]],axis=2)
-    a[...,2] = a[...,2]
-    a[...,4] = a[...,4]
-    a[...,7] = a[...,7]
-    b = value
-
-    criteria = cut
-
-    count = np.sum((b<criteria).astype(int),axis=1)
-    weight = np.ones(b.shape)/count[:,None]
-    weight[np.invert(np.isfinite(weight))] = 0
-
-    index = np.where((b<criteria))
-    return a[index],b[index],weight[index],index
-
-
-GWTC = np.load('../data/GWTC3_LVK_posterior.npz')
[…]

### 2022-05-23 src/figures/figure4_population.py

@@ -1,67 +1,0 @@
-import numpy as np
-import h5py
-import matplotlib.pyplot as plt
-from scipy.interpolate import interp1d
-from scipy.integrate import cumtrapz
-from scipy.optimize import curve_fit
-import copy
-from astropy.cosmology import Planck18
-
-import matplotlib as mpl
-import corner
-from matplotlib.ticker import (MultipleLocator, FormatStrFormatter,
-                               AutoMinorLocator)
-from matplotlib.lines import Line2D
-from scipy.stats import gaussian_kde
-from scipy.spatial.distance import pdist, cdist
-
-
-params = {'axes.labelsize': 32,
-          'font.family': 'serif',
-          'font.serif': 'Computer Modern Raman',
-          'font.size': 32,
-          'axes.linewidth': 2,
-          'legend.fontsize': 20,
-          'xtick.labelsize': 20,
-          'xtick.top': True,
-          'xtick.direction': "in",
-          'ytick.labelsize': 20,
-          'ytick.right': True,
-          'ytick.direction': "in",
-          'axes.grid' : False,
-          'text.usetex': True,
-          'savefig.dpi' : 100,
-          'lines.markersize' : 14,
-#          'axes.formatter.useoffset': False,
-          'axes.formatter.limits' : (-3,3)}
-
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}'] #for \text command
-
-mpl.rcParams.update(params)
-
-posterior_array = []
-data = h5py.File('../data/GWTC3_mcmc_processed.hdf5','r')
-for i in range(data['coord_array'].shape[0]):
-        evo_params = data['coord_array'][i][6]
-        joint_dist = np.stack([data['m1'][i][()],evo_params],axis=-1)
-        posterior_array.append(joint_dist)
-
-posterior_array = np.array(posterior_array)
-evo_params_array = np.delete(posterior_array,29,axis=0)
-fig, ax = plt.subplots(1,1,figsize=(10,9))
-
-for i in range(data['coord_array'].shape[0]):
-    try:
-        if i != 12: # Event number 12 has significant instability in KL divergence, that it is removed from the analysis. 
[…]

### 2022-05-23 src/scripts/figure1_cornerPlot.py

@@ -1,0 +1,400 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import MultipleLocator, FormatStrFormatter, AutoMinorLocator
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+import paths
+
+params = {
+    "axes.labelsize": 32,
+    "font.family": "serif",
+    "font.serif": "Computer Modern Raman",
+    "font.size": 32,
+    "axes.linewidth": 2,
+    "legend.fontsize": 20,
+    "xtick.labelsize": 20,
+    "xtick.top": True,
+    "xtick.direction": "in",
+    "ytick.labelsize": 20,
+    "ytick.right": True,
+    "ytick.direction": "in",
+    "axes.grid": False,
+    "text.usetex": True,
+    "savefig.dpi": 100,
+    "lines.markersize": 14,
+    "axes.formatter.limits": (-3, 3),
+}
+
+mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
+
+mpl.rcParams.update(params)
+
+
+def fill_corner(
+    fig,
+    data,
+    ndim,
+    axis,
+    color="C0",
+    true_val=None,
+    zorder=1,
+    lw=3,
+    style="-",
+    smooth1d=0.01,
+    smooth2d=0.01,
+    no_fill_contours=False,
+    fill_contours=True,
+    alpha=1,
+    levels=[0.68, 0.95],
+):
+    for i in range(ndim):
+        for j in range(i + 1):
+            ax = fig.axes[np.sum(np.arange(i + 1)) + j]
+            if i == j:
+                ax.hist(
+                    data[i],
+                    bins=axis[i],
+                    histtype="step",
+                    lw=lw,
+                    color=color,
+                    density=True,
+                    zorder=zorder,
+                )
+                ylim = ax.get_ylim()
+                ax.set_xlim(axis[i][0], axis[i][-1])
+                ax.set_ylim(ylim[0], ylim[1])
[…]

### 2022-05-23 src/scripts/figure2_varyAlphaFacc.py

@@ -1,0 +1,86 @@
+import pandas as pd
+import numpy as np
+
+import matplotlib.pyplot as plt
+from matplotlib import colors
+import seaborn as sns
+import paths
+
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["mathtext.fontset"] = "dejavuserif"
+
+models = ["A22_02", "A22", "A22_5", "KW"]
+
+
+def get_data(model):
+    if model == "KW":
+        bpp = pd.read_hdf(paths.data / "{}_rerun.h5".format(model), key="bpp")
+        initC = pd.read_hdf(paths.data / "{}_rerun.h5".format(model), key="initC")
+    else:
+        bpp = pd.read_hdf(paths.data / "{}_rerun.hdf".format(model), key="bpp")
+        initC = pd.read_hdf(paths.data / "{}_rerun.hdf".format(model), key="initC")
+
+    mergers = bpp.loc[(bpp.kstar_1 == 14) & (bpp.kstar_2 == 14) & (bpp.evol_type == 3)]
+    formation = (
+        bpp.loc[
+            (bpp.bin_num.isin(mergers.bin_num))
+            & (bpp.kstar_1 == 14)
+            & (bpp.kstar_2 == 14)
+        ]
+        .groupby("bin_num")
+        .first()
+        .reset_index()
+    )
+    bpp = []
+    return initC.loc[initC.bin_num.isin(mergers.bin_num)], mergers, formation
+
+
+initC_list = []
+merger_list = []
+formation_list = []
+for m in models:
+    dat = get_data(model=m)
+    initC_list.append(dat[0])
+    merger_list.append(dat[1])
+    formation_list.append(dat[2])
+
+labels = [
+    r"$\alpha=0.2$, $f_{\rm{acc}}=0.5$",
+    r"$\alpha=1$, $f_{\rm{acc}}=0.5$",
+    r"$\alpha=5$, $f_{\rm{acc}}=0.5$",
+    r"$\alpha=\rm{var}$, $f_{\rm{acc}}=\rm{var}$",
+]
+
+fig, axes = plt.subplots(nrows=1, ncols=4, figsize=(16, 3))
+
+for ax, initC, label, ii in zip(axes, initC_list, labels, range(len(labels))):
+    p, x_edges, y_edges, im = ax.hist2d(
+        x=initC.mass_1.values,
[…]

### 2022-05-23 src/scripts/figure3_sanityCheck.py

@@ -1,0 +1,143 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import MultipleLocator, FormatStrFormatter, AutoMinorLocator
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+import paths
+
+params = {
+    "axes.labelsize": 32,
+    "font.family": "serif",
+    "font.serif": "Computer Modern Raman",
+    "font.size": 32,
+    "axes.linewidth": 2,
+    "legend.fontsize": 20,
+    "xtick.labelsize": 20,
+    "xtick.top": True,
+    "xtick.direction": "in",
+    "ytick.labelsize": 20,
+    "ytick.right": True,
+    "ytick.direction": "in",
+    "axes.grid": False,
+    "text.usetex": True,
+    "savefig.dpi": 100,
+    "lines.markersize": 14,
+    #          'axes.formatter.useoffset': False,
+    "axes.formatter.limits": (-3, 3),
+}
+
+mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
+
+mpl.rcParams.update(params)
+
+
+def get_posterior(value, coord, cut=1e-2):
+    a = copy.deepcopy(coord)
+    a[..., 0] = 10 ** a[..., 0]
+    a[..., 1] = 10 ** a[..., 1]
+
+    a[..., [1, 0]] = np.sort(a[..., [0, 1]], axis=2)
+    a[..., 2] = a[..., 2]
+    a[..., 4] = a[..., 4]
+    a[..., 7] = a[..., 7]
+    b = value
+
+    criteria = cut
+
+    count = np.sum((b < criteria).astype(int), axis=1)
+    weight = np.ones(b.shape) / count[:, None]
+    weight[np.invert(np.isfinite(weight))] = 0
+
+    index = np.where((b < criteria))
+    return a[index], b[index], weight[index], index
+
+
+GWTC = np.load(paths.data / "GWTC3_LVK_posterior.npz")
[…]

### 2022-05-23 src/scripts/figure4_population.py

@@ -1,0 +1,103 @@
+import numpy as np
+import h5py
+import matplotlib.pyplot as plt
+from scipy.interpolate import interp1d
+from scipy.integrate import cumtrapz
+from scipy.optimize import curve_fit
+import copy
+from astropy.cosmology import Planck18
+
+import matplotlib as mpl
+import corner
+from matplotlib.ticker import MultipleLocator, FormatStrFormatter, AutoMinorLocator
+from matplotlib.lines import Line2D
+from scipy.stats import gaussian_kde
+from scipy.spatial.distance import pdist, cdist
+import paths
+
+params = {
+    "axes.labelsize": 32,
+    "font.family": "serif",
+    "font.serif": "Computer Modern Raman",
+    "font.size": 32,
+    "axes.linewidth": 2,
+    "legend.fontsize": 20,
+    "xtick.labelsize": 20,
+    "xtick.top": True,
+    "xtick.direction": "in",
+    "ytick.labelsize": 20,
+    "ytick.right": True,
+    "ytick.direction": "in",
+    "axes.grid": False,
+    "text.usetex": True,
+    "savefig.dpi": 100,
+    "lines.markersize": 14,
+    #          'axes.formatter.useoffset': False,
+    "axes.formatter.limits": (-3, 3),
+}
+
+mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
+
+mpl.rcParams.update(params)
+
+posterior_array = []
+data = h5py.File(paths.data / "GWTC3_mcmc_processed.hdf5", "r")
+for i in range(data["coord_array"].shape[0]):
+    evo_params = data["coord_array"][i][6]
+    joint_dist = np.stack([data["m1"][i][()], evo_params], axis=-1)
+    posterior_array.append(joint_dist)
+
+posterior_array = np.array(posterior_array)
+evo_params_array = np.delete(posterior_array, 29, axis=0)
+fig, ax = plt.subplots(1, 1, figsize=(10, 9))
+
+for i in range(data["coord_array"].shape[0]):
+    try:
+        if (
+            i != 12
+        ):  # Event number 12 has significant instability in KL divergence, that it is removed from the analysis.
+            corner.hist2d(
+                posterior_array[i, :, 0][posterior_array[i, :, 0] > 0],
+                posterior_array[i, :, 1][posterior_array[i, :, 0] > 0],
+                bins=80,
+                smooth=0.5,
[…]

### 2022-05-23 src/scripts/figure1_cornerPlot.py
disable tex

@@ -29,7 +29,7 @@ params = {
     "ytick.right": True,
     "ytick.direction": "in",
     "axes.grid": False,
-    "text.usetex": True,
+    "text.usetex": False,
     "savefig.dpi": 100,
     "lines.markersize": 14,
     "axes.formatter.limits": (-3, 3),

### 2022-05-23 src/scripts/figure4_population.py

@@ -29,7 +29,7 @@ params = {
     "ytick.right": True,
     "ytick.direction": "in",
     "axes.grid": False,
-    "text.usetex": True,
+    "text.usetex": False,
     "savefig.dpi": 100,
     "lines.markersize": 14,
     #          'axes.formatter.useoffset': False,

### 2022-05-24 src/ms.tex
adding in PB citations

@@
 Merging DCO binaries can also originate from a wide variety of dynamically active environments including 
-triple (or higher multiple) systems \citep[e.g. ][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021}, 
+triple (or higher multiple) systems \citep[e.g.][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021}, 
 as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017, DiCarlo2020, Chattopadhyay2022}, 
@@
 Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions 
-\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black hole formation \textbf{KB add these citations} can lead 
+\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} can lead 
 to merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 

### 2022-05-24 src/tex/ms.tex
figures working

@@
 \newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
+\newcommand{\kb}[1]{\textcolor{pink}{#1}}
 % \setlength{\floatsep}{0cm}
@@
 
+\author{Katelyn Breivik}
+\email{kbreivik@flatironinstitute.org}
+\flatiron
+
+\author{Will M. Farr}
+\email{will.farr@stonybrook.edu}
+\flatiron
+\stonybrook
+
 \date{\today}
@@
 \begin{abstract}
-In this
+One promising way to extract information about stellar astrophysics from
+gravitational wave catalogs is to compare the catalog to the outputs of
+population synthesis modeling with varying physical assumptions.  The parameter
+space of physical assumptions in population synthesis is high-dimensional and
+the choice of parameters that best represents the evolution of a binary system
+may depend in an as-yet-to-be-determined way on the system's properties; indeed,
+measuring the variation of population synthesis parameter settings with binary
+properties is one of the principal ways to extract information about binary
+astrophysics from catalogs. Here we propose an inference pipeline to
+simultaneously infer stellar progenitor parameters (e.g. zero-age main sequence
+masses) and population synthesis parameter settings controlling modeled binary
+evolution from individual gravitational wave observations of merging compact
+binaries.  Our pipeline can efficiently explore the high-dimensional space of
+population synthesis settings and progenitor system properties for each system
+in a catalog of gravitational wave observations.  We apply our pipeline to
+observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog
+(GWTC3).  We showcase the effectiveness of this pipeline with a detailed study
+of the progenitor properties and population synthesis settings that produce
+mergers like the observed GW150914.
 \end{abstract}
@@
 \section{Introduction}
+As the detection rate of gravitation waves (GWs) from merging
+double-compact-object (DCO) binaries increases with the sensitivity of the
+ground-based GW detector network
+\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
+we are beginning to constrain the astrophysical processes which shape the
+evolution of GW progenitor populations. One of the most common ways to study
+progenitor populations of GW mergers is through population synthesis simulations
+for a host of formation environments including isolated binary stars, triple
+systems, globular clusters, nuclear star clusters, AGN disks, and more
+\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
+population is studied through a Monte-Carlo approach.  Using theoretically
+motivated distributions of initial parameters (age, metallicity, masses, orbital
+separations, eccentricities, etc), initial populations are evolved with fixed
+astrophysical assumptions (population synthesis parameter settings) to produce a
+synthetic catalog of mergers which can be compared to observations from
+GW detectors.
 
-As the rate of detection of gravitational-wave (GW) increase with the sensitivity of the GW detector network \kw{},
-understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}.
-A very common way to study the origin of detected GW events is through population synthesis simulation \kw{cite},
-which evolves a population of binaries to form a population of GW events.
-Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
-such as the initial mass function or star-formation history,
-then evolve each individual members according to the physics of interest until some stopping criteria is met.
-This usually mean either a merger event or the binary system stays stable for a Hubble time.
-After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
+%As the rate of detection of gravitational-wave (GW) increases with the sensitivity of the \kb{ground-based} GW detector network \kw{cite},
+%understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}. 
+%\kb{I think we can reword the end of this previous sentence -- I should come back to it.}
+%A very common way to study the origin of detected GW events is through population synthesis simulations \kw{cite} \kb{cite Mandel \& Broekgaarden review},
+%which evolve populations of various kinds of binaries to form populations of GW events. \kb{Do we want to do isolated only here? or other channels?}
+%\kb{The process to forward model a population of GW events relies on random samples }
+%Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
+%such as the initial mass function or star-formation history,
+%then evolve each individual members according to the physics of interest until some stopping criteria is met.
+%This usually mean either a merger event or the binary system stays stable for a Hubble time.
+%After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
 
-\kw{Shorten}
-While forward modelling has been the canonical pathway to simulate binary systems and compare to GW observation,
-it is not necessarily the most efficient way to constrain the underlying physical model.
-Most of the systems do not end in a merger, and the systems ending up as merger events do not necessarily agree with the observation.
-This means if we want to evolve a binary to match the observation by pure trial and error, a large fraction of computation will be wasted in trying initial binary configurations that are not related to the observation.
-To alleviate this inefficiency, the community often takes a population approach --- instead of trying to match binary systems to observations event by event,
-one can draw systems from an initial population, forward model each member in that population then match it to the observed population properties.
-However, this approach is essentially trading one inefficiency with another, since we still need to choose an initial distribution of binary parameters.
-For relatively well-known source properties such as initial masses this is less of a problem,
-but the distribution of some less well-studied properties such as common envelope efficiency is often heuristically chosen to be some simple distribution such as a delta function at some value, i.e.
-a fixed value for the entire population.
-This is a problem when the target of a study is to understand these less studied parameters, since one either need to justify their choice of distribution with other independent studies,
-or they should vary the family of distributions.
-In general, it is unclear how one can cover all possible distribution that may agree with the observation in generating a suite of simulations.
-The problem worsen as we increase the dimensionality of the parameter space.
-\wf{This says what you want to say, but it's a bit wordy right now---I'll return to it, but it would be good to tighten up the language a lot.}
+While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
+it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
+One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
+cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
+(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
+On top of this, these values are often not calibrated to data.
+Even with recent developments in using emulators to speed up the simulation process, training the emulators 
+still requires a significant amount of computation to cover a wide range of uncertainty.
+As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
+run a grid which varies combinations of uncertain physics.
+For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
+this would still require $1024$ simulations.
+Because population synthesis simulations require hundreds to thousands of cpu hours,
+training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
 
-Indeed, one of the most common skepticism population synthesis studies receives is a simulation set usually only cover a small subdomain of the entire parameter space,
-i.e. only vary a number of hyper-parameters while holding all others fixed to some "default" values.
-Even with recently development on using emulators to speed up the simulation process, training the emulators also require a decent amount of samples in the parameter space of interest.
-For example, say we want to emulate a model with 10 parameters, this simplest option for constructing a bank of simulations for training is to lay down a grid.
-Along each axis we put the barely minimum of $2$ simulations, this would still require $1024$ simulations.
-The fastest kind of population synthesis simulation takes about an hour to complete, where longer simulations could take days.
-This means even training an emulator is impossible in high dimensional space.
+\kw{Reorganizing discussion related to Jeff's work such that it doesn't seem like our method is just an extension of DartBoard.}
 
-In this paper, instead of forward modelling a population of binary star to match the observed GW events population,
-we propose a method to "backward" model each GW event to its progenitor state.
-The idea is since we can compute the GW event properties of a guess by forward modelling its progenitor,
-we can define the difference between the GW event properties of this guess and the observed GW event properties.
-By minimizing this difference, i.e. solving a root finding problem, we can hope to recover the progenitor properties.
+This compounds with the fact that there is no guarantee the entire population should share the same fixed hyperparameters, 
+as is often assumed in most population synthesis studies.
+This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
+required to compare data to simulations which span the full set of uncertain physical models.
+However, this issue is not a fundamental issue, but an apparent obstacle which arises from an inefficient approach to the problem.
+Furthermore, each binary system could in principle come from a different formation environment and hence 
+have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
+In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
+for each individual binary system, then perform population in analyses in that space.
+\citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
+to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
+set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
+\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
+However, the method was limited to a single event instead of the population.
+More importantly, the method did not vary the hyperparameters while sampling the posterior density of the ZAMS parameters,
+which limit its capability to be applied to understand a population of binary systems.
 
-This paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
+In this paper, we propose a method to ``backward" model each GW event to its progenitor state 
+while allowing the hyper-parameters to vary across their full range of physical uncertainty
+following a two-stage process.
+First, we solve a root finding problem to obtain guesses that are likely to produce the desired system properties in the joint progenitor--hyper-parameter space.
+We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
+The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
 In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
@@
 
-\section{Method Rewrite}
-\label{sec:method-rewrite}
+\section{Method}
+\label{sec:method}
 
-\wf{Re-writing the methods section with a new viewpoint, gonna try to eventually merge the content.}
-
-We define the parameters that describe the progenitor's properties such as ZAMS
-masses and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
-that control different physical prescriptions such as wind strength and common
-envelope efficiency as hyper-parameters $\bm{\lambda}$, and random variables
-that affect certain stochastic process such as whether natal kick unbinds a
+We define the progenitor's zero age main sequence (ZAMS) properties like
+mass, orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
+that control different uncertain physical prescriptions such as wind strength and common
+envelope efficiency as hyperparameters $\bm{\lambda}$, and random variables
+that affect certain stochastic process such as whether CO natal kick unbinds a
 binary as $\bm{X}$. To avoid clutter in the following derivation, we denote all
@@
 \end{equation}
-The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ is many-to-one.  
-In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
-need to be able to evaluate the likelihood of that data at fixed progenitor
-parameters, 
+The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one
+due to degeneracies in the different physical processes and initial parameters  
+of the GW event progenitors.
+In order to draw inferences about $\bm{\Theta}$ from GW data, we
+must be able to evaluate the likelihood of that data at fixed progenitor
+parameters, namely
 \begin{equation}
-    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right),
+    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right).
 \end{equation}
-because the likelihood depends only on the gravitational waveform generated by
-parameters $\bm{\theta}$ to which the progenitor parameters map.  In principle
-this likelihood could be computed at arbitrary vaues of $\bm{\Theta}$ using the
-same machinery that is used to estimate source parameters in gravitational wave
+The equality holds because the likelihood depends only on the gravitational waveform generated by
+parameters $\bm{\theta}$ to which the progenitor parameters map onto.  In principle
+this likelihood could be computed at arbitrary values of $\bm{\Theta}$ using the
+same machinery that is used to estimate source parameters in GW
 catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
-of progenitor parameters, hyper-parameters, and random variables in
+of progenitor parameters, hyperparameters, and random variables in
 $\bm{\Theta}$ from a posterior density.
 
-However, it is very expensive to couple progenitor evolution to stochastic
-sampling in this way.  Because the gravitational wave likelihood function
-ultimately depends only on the GW parameters $\theta$, and catalogs have already
-produced samples over these parameters from a posterior density 
+Since the GW likelihood function
+ultimately depends only on the GW parameters $\theta$, and GWTC3 already has samples over these parameters from a posterior density 
 \begin{equation}
-    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) p_\mathrm{GW} \left( \bm{\theta} \right),
+    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) \pi_\mathrm{GW} \left( \bm{\theta} \right),
 \end{equation}
-where $p_\mathrm{GW}$ is the prior density used for sampling, it is much more
-efficient to \wf{WIP}.
-
-\section{Method}
-\label{sec:method}
-
-We define the parameters that describe the progenitor's properties such as ZAMS masses and eccentricity as progenitor parameters $\bm{\theta'}$,
-the parameters that control different physical prescriptions such as wind strength and common envelope efficiency as hyper-parameters $\bm{\lambda}$,
-and random variables that affect certain stochastic process such as whether natal kick unbinds a binary as $\bm{X}$.
-To avoid clutter in the following derivation,
-we denote all the parameters related to mapping a particular progenitor system into a GW event collectively as evolutionary parameters $\bm{\Theta}$,
-which includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.
-
-Once all the parameters including a particular draw of all the random variables are known,
-the population synthesis code is a deterministic function that transforms the properties of the progenitors into the GW parameters.
-Hence, the probability of obtaining a GW event with parameters $\bm{\theta}$ given the progenitor parameters $\bm{\Theta}$ is given by
+where $\pi_\mathrm{GW}$ is the prior density used for sampling,
+we can save the computation cost associated with evaluating the GW likelihood by rewriting the posterior 
+density of $\Theta$ in terms of the posterior density of $\theta$, namely
 
 \begin{align}
-    p(\bm{\theta}|\bm{\Theta}) &= \bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta}),
+    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta(\Theta))},
 \end{align}
+which we apply a kernel density estimator to the posterior samples released in GWTC3 to estimate $p(\theta(\Theta)|d)$.
 
-where $\bm{f}$ represents the evolution encoded by the population synthesis model with fixed random parameters $\bm{X}$.
+Including hyperparameters and random variables %in the sampling 
+drastically increases the dimensionality of the problem,
+which makes the sampling process more computationally expensive to converge.
+To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density of a GW event,
+then we use those points to initialize a set of chains in the MCMC process.
 
-While it is straightforward to write down the "forward" probability, the "backward" probability is more complicated since multiple evolutionary parameters can produce the same GW event.
-Using Bayes theorem, the probability of an event with evolutionary parameters given the GW parameters can be derived as following
+%The root-finding procedure is described as follows:
+For each posterior sample point in the GW-observable space,
+we can find the corresponding evolutionary parameters by solving an optimization problem that minimizes 
+the mean square difference between the GW event parameters and evolutionary parameters as:
 
 \begin{align}
-    p(\bm{\Theta}|\bm{\theta}) &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{p(\bm{\theta})}, \nonumber \\
-    &= \frac{p(\bm{\theta}|\bm{\Theta})p(\bm{\Theta})}{\int p(\bm{\theta}|\bm{\Theta}) p(\bm{\Theta}) d\bm{\Theta}}, \nonumber \\
-    &= \frac{\bm{\delta}(\bm{f}(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}{\int d\bm{\theta} \bm{\delta}(f(\bm{\Theta})-\bm{\theta})p(\bm{\Theta})}.
-\label{eq:backward_prob}
-\end{align}
-
-Using the result in eq.\ref{eq:backward_prob},
-the probability of an event having evolutionary parameters given data can be obtained by marginalizing over the GW parameters
-
-\begin{align}
-    p(\bm{\Theta}|d) &= \int p(\bm{\Theta}|\bm{\theta}) p(\bm{\theta}|d) d\bm{\theta}\\
-    &= \frac{1}{N}\sum_{i=1}^{N} p(\bm{\Theta}|\bm{\theta}_i). \label{eq:evolutionary_posterior}
-\end{align}
-
-The term $p(\bm{\theta}|d)$ is the inferred posterior probability for a particular event, which is usually presented by the LVK collaboration in the form of posterior samples.
-In eq.\ref{eq:evolutionary_posterior}, we use importance sampling to turn the integration into a sum over the posterior samples.
-
-For each posterior sample point, we can find the corresponding evolutionary parameters by solving an optimization problem that minimize the mean square difference between the GW event parameters and evolutionary parameters, i.e.
-
-\begin{align}
-\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2
+\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2.
 \label{eq:loss}
@@
 
-In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the observable space.
-However, it is not feasible to achieve such condition in practice, therefore we relax the condition in eq.\ref{eq:loss} to a small acceptance threshold, in our case we picked $10^{-2}$ to be our threshold.
-To make sure we find a reasonably complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses in solving the optimization problem.
-As long as the solution fulfill the acceptance criteria, the root is as valid as all other roots that fulfill the same acceptance criteria, regardless to its initial guess.
-In other word, we have the freedom to choose the set of initial guesses however we think it might benefit the optimization.
-Most of the points in the evolutionary parameter space do not produce mergers, and for systems that do not merge, we set the binary masses to be 0.
-This means the gradient for systems that do not merge is 0, and the root finder would likely to get stuck in such state during the first step.\footnote{On a plateau of constant loss, the root finder performs no better than a random guess.
-In a high dimensional space, this means the root finder will need a long time to get close to a solution.}
-Therefore, it is beneficial to put initial guesses in region of the evolutionary parameter space that does produce merger.
-To construct a list of initial guesses, we evolve a set of points that we uniformly sampled from the evolutionary parameter space, then keep the points that merge within a Hubble time.
+In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the GW-observable space.
+However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.~\ref{eq:loss} 
+to a small acceptance threshold. For this study, we picked a threshold of $10^{-2}$. To make sure we find a reasonably 
+complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses 
+in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all 
+other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
+This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
 
-Ideally we would like to have explicit control over all evolutionary parameters.
-In particular, having explicit control over random variables allows us to marginalize over their contribution, so we can focus on progenitor parameters and hyper-parameters.
-However, in practice most of the population synthesis simulation code (including \texttt{COSMIC} which we use in this study.) define their random variables implicitly,
-i.e. the random variables are drawn within the program, instead of being arguments of the program.
-This means the function we use to evolve the binary is not a deterministic function but a stochastic function instead. 
-Hence, even if the root-finding algorithm performs perfectly,
-forward modelling this set of roots does not guarantee to reproduce the set of posterior samples due to randomness of the evolution function.
-To assure the recovered progenitor and hyper-parameters robustly corresponds to the posterior in the GW observables space,
-we forward model ("Reproject") the recovered evolutionary parameters to the observables space to check whether the reprojected posterior agrees with the posterior given by the LVK collaboration.
-We use KL divergence to measure the agreement between the two posterior distributions, which is given by
+Most of the points in the evolutionary parameter space do not produce DCO mergers in a Hubble time. For systems that do not merge, 
+we set the binary masses to be $0$ such that the gradient for the same systems is also $0$ and leads the root finder to get stuck on 
+its first step.{\footnote{On a plateau of constant loss, the root finder performs no better than a random guess. 
+Because of this, a root finder in a high dimensional space will require a significant amount of computation
+to get close to a solution.}} Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter 
+space that are likely to produce DCO mergers. To construct a list of initial guesses, we evolve a set of 
+ZAMS parameters uniformly sampled from the initial binary parameter space, then keep the systems that 
+merge within a Hubble time.
+
+Retaining explicit control over random variables allows us to marginalize over their contribution, 
+and thus allows us to focus on progenitor parameters and hyperparameters.
+However, in practice most population synthesis codes (including \texttt{COSMIC} which we use in this study) 
+define their random variables implicitly such that the random variables are drawn within the program instead of 
+being passed as arguments. This means the function we use to evolve the binary is not fully deterministic.
+Thus, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not 
+guarantee that the simulated population reproduces exactly the set of posterior samples due to 
+randomness in the evolution of each binary. To assure that the recovered progenitor and hyperparameters 
+robustly correspond to the posterior in the GW-observable space, we forward model (or "reproject") the recovered 
+evolutionary parameters to the GW-observable space to check whether the reprojected posterior agrees 
+with the posterior given by the LVK collaboration. We use KL divergence to measure the agreement between the two posterior 
+distributions as:
 \begin{align}
@@
 
-A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the observable space, hence it is a viable channel for that specific event, otherwise it either means \texttt{COSMIC} can only explain part of the posterior or cannot explain the event at all.
-This is completely expected since \texttt{COSMIC} comes with its own assumptions and is expected to fail in reproducing a subset of the event in GWTC3, such as event with at least one component beyond the pair instability supernova massgap.
-On the other hand, the reprojection could subject to stochasticity in the evolution function we use.
-Instead of reprojecting the posterior only once, we reproject the posterior multiple times with different random seeds and check whether the KL divergence varies significantly.
-A varying KL divergence means this particular event is subjected to randomness in the evolution function, hence one should take extra caution in interpreting the result.
+A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the 
+observable space, and is thus a viable channel for that specific event. Otherwise, it either means \texttt{COSMIC} 
+can only explain part of the posterior or cannot explain the event at all. This is completely expected behavior since 
+\texttt{COSMIC}, and isolated binary evolution more generally, carries its own assumptions and is expected to 
+fail in reproducing a subset of the event in GWTC3. One example of this are GW events with at least one component 
+with mass above the pair instability supernova mass limit \kb{cite Farmer, and others}.
+The reprojection could also be subject to stochasticity in the evolution of each binary
+To account for this, we reproject the posterior multiple times with different random seeds and 
+check whether the KL divergence varies significantly. A varying KL divergence means a particular GW event is subject 
+to randomness in the evolution of the progenitor, and extra caution should be used when interpreting the result.
 
-On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
-The two results agree well with each other.
-This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.less
-\section{Result}
+\section{Results}
 \label{sec:result}
-
-\begin{enumerate}
-\item Raw posterior of successful events.
-\item Trace back the event's $z_{\rm ZAMS}$.
-\item Reproject to observable space
-\item Evidence ratio?
-\item KL divergence?
-\end{enumerate}
 
@@
 
-We apply the method described in section \ref{sec:method} to the first GW events GW150914.
-The parameters involved in the analysis and the range we allow for finding the root are tabulated in table \ref{tab:parameters}.
-\kw{@Katie state why not spin here.}
-Therefore, we solve the rootfinding problem with only the two component masses in the observable space.
-For progenitor parameters, we characterize each guess with five parameters: the two component masses $M_{\rm ZAMS,1}$ and $M_{\rm ZAMS,2}$, the orbital period $t_{\rm orb}$, the eccentricity $e$, the metallicity $Z$ at ZAMS.
-Note that the progenitor formation redshift is not obtained through the root-finding process.
-This is because the formation redshift does not affect the evolution of the binary, so we do not need to include in the optimization.
-Instead, we compute the time it takes for the binary to evolve from formation to the merger, then we add this time back to the lookback time when this event is observed.
-Using the total lookback time, we can compute the redshift of the binary at formation.
-We adopt the cosmological model in \kw{cite Planck 18} to compute the redshift of the binary at formation.
-For hyper-parameters, we choose parameters that affect binary evolution in \texttt{COSMIC} mostly significantly, including the common envelope efficiency $\alpha$, the fraction of accretion during stable mass transfer $f_{\rm acc}$, the critical mass fraction $q_{\rm crit,3}$, and the accretion rate $\sigma$.
-We show a portion of the joint posterior of the progenitor parameters and hyper-parameters of GW150914 in figure \ref{fig:GW150914_posterior}.
+We apply the method described in section \ref{sec:method} to the first GW event GW150914.
+The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
+The spin of BHs formed in isolated binaries depends strongly on the angular 
+momentum transport in massive stars which is still highly uncertain \kb{(see however Bavera+, fuller+)},
+therefore we only consider the two component masses in the observable space in the both the 
+root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with 
+five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$, 
+the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not 
+obtained through the root-finding process because the formation redshift does not intrinsically affect 
+the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from 
+formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total 
+time to merger to be less than $13.7\,\rm{Gyr}$.
+Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
+Planck 2018 cosmological model \citep{Planck2018}.
+The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
+of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
+This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
+by reweighting the posterior distribution according to a particular SFR model,
+\kb{and} therefore do not have to rerun the inference everytime we change the SFR model. This is especially helpful
+since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep{Broekgaarden2021}.
+For hyperparameters, we choose parameters that affect 
+binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 
+efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$, 
+the critical mass ratio which determines whether Roche-overflow mass transfer 
+proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap 
+$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
+employed for CO natal kicks, $\sigma$.
 
-\kw{Recover ratio of the posterior sample goes here.}
+We show a portion of the joint posterior of the progenitor parameters and
+hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
+is colored based on the comparison between progenitor parameters (blue),
+hyperparameters (orange), or a mix of the two (green). We find that the choice
+for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
+unexpected since we use the \citet{Fryer2012} delayed model which reduces the
+natal kick strength at BH formation based on the amount of fallback onto the
+proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
+enough that the fallback reduces the natal kick to zero. Because of this we do
+not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
+not include the ZAMS orbital period and eccentricity, which are correlated with
+one another, but not strongly with the other progenitor parameters or
+hyperparameters. We find strong correlations between the ZAMS masses, as well as
+strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
+$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
+in the upper-right inset of the figure.
 
-The full posterior is available in appendix \kw{Or we should put a link to the dataset?}
+The correlations between the ZAMS masses are similar to those found by A21 with
+the majority of the population preferring primary and secondary masses between
+$60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
+$150\,M_{\odot}$ which is the limit imposed by our assumptions. We explore the
+formation scenarios which lead to successful GW150914-like mergers and those
+which fail to produce GW150914-like mergers below in
+Figure~\ref{fig:compare_fixed_variable}.
+
+The correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm acc}$ illustrate the
+variety of ways which GW150914-like mergers can be produced. For binaries with
+$M_{2,\rm{ZAMS}} < 70\,M_{\odot}$, we find that accretion efficiencies of
+$f_{\rm acc} > 0.5$ are preferred. One reason for this is based on the
+requirement that the total mass in the binary must remain above the total mass
+of GW150914 and strongly non-conservative mass transfer (i.e. $f_{\rm acc} <
+0.5$) reduces the total mass of the system. A compounding factor is that as mass
+leaves the binary due to non-conservative mass transfer, the evolution of the
+binary's orbit is less dramatic and leads to wider binaries on average and thus
+fewer mergers in a Hubble time. For binaries with $M_{2,\rm{ZAMS}} \sim
+80\,M_{\odot}$, $f_{\rm acc}$ is less constrained. This is because of a
+preference for these secondaries to also have primary masses near $
+80\,M_{\odot}$ and thus enter a double-core common envelope evolution in which
+both stars' envelopes are ejected, leaving behind two stripped helium cores in a
+tight orbit. In this case, $f_{\rm acc}$ does not affect the binary evolution
+and is thus unconstrained. This effect can also be seen in the ZAMS mass and
+$q_{\rm crit 3}$ panels for masses near $80\,M_{\odot}$ because double-core
+common-envelope evolution is triggered in \texttt{COSMIC} when the radii of the
+two stars touch due to the rapid expansion of the primary star upon helium
+ignition. In this case, the choice for Roche-overflow mass transfer to proceed
+stably (as prescribed by $f_{\rm{acc}}$) or unstably (as prescribed by $\alpha$)
+using $q_{\rm crit, 3}$ is totally irrelevant, and thus unconstrained. Finally,
+for binaries with $M_{2,\rm{ZAMS}} > 80$, we find that $f_{\rm acc}$ is
+correlated to decrease with increasing mass. This is due to limits on the total
+mass which must be ejected from the binary to produce BH masses that match
+GW150914, the reverse situation to $M_{2,\rm{ZAMS}} < 70 \, M_\odot$. Binaries
+with both component masses near $150\,M_{\odot}$ must either go through a common
+envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very
+non-conservative mass transfer (where $f_{\rm acc}$) is low, to produce BBHs
+with the proper mass.
+
+Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
+though there exist independent trends in each hyper parameter. This is not
+totally unexpected since the physical processes which are described by each
+hyperparameter, i.e. stable Roche-lobe overflow and common envelope evolution
+are two independent channels. Generally, we find that GW150914-like BBHs tend to
+prefer larger accretion efficiencies and have common envelope ejection
+efficiencies peaking near $\alpha=1$. One shortcoming of our method
+implementation is the application of a single hyper parameter for
+$f_{\mathrm{acc}}$ and $\alpha$ for each binary, while both the primary and the
+secondary star could in principle be defined by their own hyperparameters that
+prescribe the outcomes for when each star fills it's Roche lobe. We reserve full
+treatment of this for future work but note that this improvement could reveal
+correlations between hyperparameters that are not present in our current
+analysis.
+%\kb{We need to describe the interpretation of this as well.}
+%\kw{I think we can just hightlight the fact that we have three kinds of correlation, progenitor-hyper, progenitor-hyper, hyper-hyper,
+%since the implication is related to figure @Katie made.}
+%\kb{Need to describe the formation channels and gaps here.}
 
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed.pdf}
+    \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
     \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
@@
     $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-$q_{\rm crit 3}$ are the critical mass during \kw{Katie fill in},
-Note that the redshift is not fitted during the root finding process.
-Once we have find the evolutionary parameters, we add the time it would take the progenitor to merge to the lookback time of the observed posterior sample,
+    $q_{\rm crit 3}$ is the critical mass ratio on the Hertzsprung Gap,
+    Note that the redshift is not fitted during the root finding process or the MCMC process.
+    Once we have find the evolutionary parameters, we add the delay time to the lookback time of the observed posterior sample,
     then from the total lookback time we can compute the redshift at ZAMS.
+    We highlight three panels in the corner plots to show the fine structure of the set of posterior samples in the evolutionary parameter space.
+    We also color the posterior in a particular panel according to the type of parameters involved in the corner.
+    Blue denotes panels that include only progenitor parameters,
+    green denotes panels that include a mix of progenitor and hyperparameters,
+    and orange denotes panels that include only hyperparameters.
     }
     \label{fig:GW150914_posterior}
+    \script{figure1_cornerPlot.py}
 \end{figure*}
 
-To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
-We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
-For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} on average.
-The KL divergence stays stable for reprojecting with different random seeds.
+% \kw{Talk about implication on single event posterior.}
+Once we obtain a set of binaries which successfully map the ZAMS parameters and
+hyperparameters to BBH merger masses, we re-evolve the set of ZAMS parameters
+with the same physical assumptions as A21, but vary the common envelope
+efficiency to explore how keeping a fixed model which only varies one
+hyper-parameter contrasts to our results.
+Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses
+for three models (first through third columns), each with a different $\alpha$
+but the same hyper-parameters as A21, as well as the results of our sampling
+which allows our hyper-parameters to vary (fourth column). We find that holding
+the accretion efficiency, $f_{\rm{acc}}$, and common envelope ejection
+efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space
+that produces GW150914-like mergers. In contrast, by allowing the
+hyper-parameters to fully span the model uncertainty, we find that there are
+distinct ZAMS parameters which produce GW150914-like mergers.
+
+Because we can explore the full evolutionary parameter space, we can also
+determine that there are ZAMS parameters which fail to produce GW150914-like
+mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses
+between $\sim95-135\,M_{\odot}$ and secondary ZAMS masses between
+$\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
+efficiency and common envelope ejection efficiency which produces merging BBHs
+that have masses consistent with GW150914's posteriors. In this region, if BBHs
+with masses consistent with GW150914's masses are produced through a combination
+of $\alpha$ and $f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely,
+in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
+which does produce GW150914-like mergers, the combination of $\alpha$,
+$f_{\rm{acc}}$, and the ZAMS masses and orbital separations balance to produce
+BHs with the correct masses that merge within a Hubble time.
+%\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
+%\kb{done!}
+
+\begin{figure*}
+    \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
+     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce 
+     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
+    \label{fig:compare_fixed_variable}
+    \script{figure2_varyAlphaFacc.py}
+\end{figure*}
+
+% Validity paragraph
+\wf{What if we made this a subsection of ``methods'', and then just one sentence here about the values of the KL divergence for GW150914?}
+To check the validity of the recovered posterior over evolutionary parameters,
+it is essential to check whether it is consistent with the original posterior in
+the observable space. We reproject the recovered posterior to the observable
+space, and compare the KL divergence between the original posterior and the
+reprojected posterior. In particular, we compute KL divergence of the samples
+obtained from root-finding alone and after MCMC sampling. \wf{Clean up wording.}
+For GW150914, the KL divergence between the reprojected posterior and the
+original posterior is \kw{fill later} nats on average if we use the root-finding
+result alone, whereas the MCMC result has a higher KL divergence at \kw{fill}
+nats. The KL divergence stays stable for reprojecting with different random
+seeds. In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected
+posterior from both the root-finding process alone and the MCMC result agrees
+well with the original posterior, even though the agreement between the MCMC
+result and the original posterior is not perfect. This degradation of the
+agreement comes from the fact that we are using a KDE to estimate the posterior
+density in the observable space. Note that the result from root-finding alone
+and the result from MCMC agrees well in the evolutionary parameter space.
+Compared to the result shown in a previous study, our result covers almost all
+the posterior of GW150914 in the masses space, which means our result is more
+consistent with the observation.
 
@@
 \caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
-The blue contour is the posterior reprojected using the samples in the evolutionary parameter space.
-The orange contour is the posterior samples from LVK that have corresponding samples in the evolutionary parameter space.
-The green contour is the posterior plotted using all of LVK posterior samples.
+The blue contour is the reprojected posterior after the root-finding procedure.
+The orange contour is the reprojected posterior after the MCMC procedure.
+The green contour is the posterior plotted using the original LVK posterior samples.
 }
 \label{fig:GW150914_reprojection}
+\script{figure3_sanityCheck.py}
 \end{figure}
 
-We can see the reprojected posterior agrees very well with the original posterior in fig.\ref{fig:GW150914_reprojection}.
-\kw{Qoute the KL divergence.}
+% Paragraph discussing the potential application on the population side.
+To illustrate the potential benefit of our method on the population level, we
+perform the same analysis for all events in GWTC3. Plotted in figure
+\ref{fig:GWTC3_f_acc_mass} is the posterior density in the $m_{1,{\rm
+GW}}-f_{\rm acc}$ space for most of the events in GWTC3. Each contour represents
+the $68\%$ credible interval of the posterior density for that particular event.
+There is a suggestive trend showing $f_{\rm acc}$ could increase as the mass of
+the progenitor increases. Such a trend implies the stable mass transfer (SMT)
+phase of a less massive binary would be preferentially non-conservative, with
+more conservative mass transfer in more massive binary systems.
 
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{figures/GW150914_CE_SMT.pdf}
-% \caption{}
-% \label{fig:GW150914_CE_SMT}
-% \end{figure}
-% \kw{Add figure of stable mass transfer vs common envelope}
+Note that some events did not pass the KL divergence test we proposed in section
+\ref{sec:method}; we do not include these systems in figure
+\ref{fig:GWTC3_f_acc_mass}. Binaries that form low-mass compact objects have
+lower amounts of fallback and tend to have correspondingly larger variance in
+their random natal kicks. Larger kicks can unbind the progenitor binary, or lead
+to wide binaries that do not merge within a Hubble time.  In these cases, the
+extra variance during the reprojection can produce a posterior that may not
+agree with the original posterior, hence yielding a higher KL divergence.
+
+For high-mass binaries, \texttt{COSMIC} struggles to produce events above the
+pair instability supernova mass cutoff $m_{1,{\rm GW}}$, so the posterior in the
+evolutionary parameter space only corresponds to part of the posterior in the
+observable space below this cutoff, and therefore events beyond the PISN mass
+gap also have higher KL divergence.
+
+Events with more extreme mass ratios are hard to produce with \texttt{COSMIC},
+and therefore less likely to be accurately recovered by our method. This could
+be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as
+discussed in Section~\ref{sec:result}.
+
+Merger observations for which \texttt{COSMIC} struggles are likely to be very
+informative about formation channels and binary evolution physics, and are
+therefore likely worthy of close study.  We anticipate that future work will
+follow up these events in detail.
+
+Figure \ref{fig:GWTC3_f_acc_mass} shows that our method could in principle
+reveal the correlation between progenitor parameters and hyperparameters on a
+population level. A careful treatment of all events in the catalog and
+discussion related to the detailed physical implications of figure
+\ref{fig:GWTC3_f_acc_mass} is beyond the scope of this paper.  We defer a
+detailed study of the physics related to the population of GW events to future
+work.
 
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/GW150914_m1m2alpha.png}
-\caption{The posterior samples of GW150914 in the progenitor masses space, colorcoded by the common envelope efficiency $\alpha$ for that sample.}
-\label{fig:GW150914_m1_m2_alpha}
+\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
+\caption{
+    The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
+    Each contour is the $68\%$ credible interval of the posterior density for a particular event. \kw{Check interval number}
+    At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair instability supernova mechanism prevents \texttt{COSMIC} from producing events that are more massive than this cutoff.
+    Therefore, events with the majority of posterior support above this cutoff is not compatible with \texttt{COSMIC}, hence has a large KL divergence and excluded from this figure.
+    On the low mass end, neutron star binaries or neutron star-black hole binaries are subjected to randomness induced by the natal kick,
+    also resulting in a larger and fluctuating KL divergence, therefore they are also excluded from the analysis.
+}
+\label{fig:GWTC3_f_acc_mass}
+\script{figure4_population.py}
 \end{figure}
-
-The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
-We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
-this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
-we would miss the samples that do not match the chosen value of $\alpha$.
-By folding in the hyper-parameters into the root-finding algorithm,
-we can capture the correlation between 
-\kw{Dartboard comparsion} Comparing our result with 
-\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}
 
@@
 
-We present a new pathway to understand binary evolution with GW events in this paper.
-Instead of forward modelling an assumed distribution of initial binary to the observed population,
-we solve a root-finding problem to find the best initial binary parameters that reproduce the posterior distribution for each event. 
-The most obvious advantage our method provides is it returns the joint posterior distribution of progenitor parameters and hyper-parameters for each event.
-This enables a data-driven way to study the distribution of hyper-parameters,
-i.e. once we have a catalog of events, each has their own set of posterior in the hyper-parameter space,
-we can employ common techniques such as hierarchical Bayesian analysis to fit a population model to the distribution of hyper-parameters,
-without making specific assumptions such as fixing the hyper-parameters for all types of event.
+We present a new pathway to understand binary evolution with GW events in this
+paper. Instead of forward modelling an assumed distribution of initial binaries
+to the observed population, we find the corresponding evolutionary parameters
+event-by-event. In this first work, we showcase the power of the proposed method
+with an application to GW150914. We show the joint posterior of the event's
+progenitor parameters and hyperparameters.  Our work is both a more efficient
+way to study GW event progenitors, and also allows the possibility of
+constraining the astrophysics related to binary evolution, especially by
+capturing the correlation between hyperparameters in different systems.
 
-In this work we showcase the power of the proposed method with two events, GW150914 and GW191129.
-We show the joint posterior of the two events' progenitor parameters and hyper-parameters.
-In particular, to the authors knowledge, this is the first time the correlation between progenitor parameters and hyper-parameters is shown.
-Our work is not only a more efficient way to study GW events' progenitors,
-but also allows the possibility of measuring the astrophysics related the binary evolution, especially in capturing the correlation between different hyper-parameters.
+Our results suggest that the accretion efficiency during stable mass transfer
+may depend on the primary black hole mass.  In general, our method returns the
+joint posterior distribution of progenitor parameters and hyperparameters for
+each event, which enables a data-driven way to study the distribution of
+hyperparameters.  That is, once we have a catalog of events, each has their own
+posterior in the hyperparameter space, and we can employ well-known techniques
+such as hierarchical Bayesian analysis to fit a population model to the
+distribution of hyperparameters; this avoids making overly-specific assumptions
+like fixing the hyperparameters for all types of event.
 
 % Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
-Another perk of our method is the "rate" calculation is now independent of assumed initial condition such as a particular star formation rate model,
-which is a dominant factor of uncertainty for rate prediction in the past.
-Once we have pull back the GW event posterior from the observable space to the evolutionary parameter space,
-we can apply the same methods we use to account for the rate in the observable space.
-In the observable space, the overall rate is basically a normalization that is defined in the local universe i.e. the local merger rate. \kw{cite}
-In the evolutionary parameter space, we can instead fit for the production rate of merger events at some redshift.
-\kw{Talk about prior are now in post-processing instead of pre-processing.}
+Working event-by-event as we do permits post-processing application of
+ arbitrarily complicated models of star formation and metalliticy evolution by
+ re-weighting the samples we obtain \wf{Cite at least Lieke here}. Once we have
+ pulled the GW event posterior back from the observable space to the
+ evolutionary parameter space, we can apply the same hierarchical inference
+ methods currently used to infer the compact binary population to our progenitor
+ population.  In particular, we can use our progenitor population to infer the
+ formation rate of compact binary population progenitor systems over cosmic time
+ without needing to make artifical and simplistic assumptions about the delay
+ time distribution or the metallicity-specific star formation rate \wf{again,
+ cite Lieke, and also Vitale, Farr, Ng, Rodriguez?}.  Comparing the formation
+ rate of progenitors to the star formation rate provides an avenue to check our
+ understanding of binary evolution and the relative rate of other formation
+ channels.
 
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyper-parameters space for the first time, there are a number of improvements can be implemented in future studies.
+While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements can be implemented in future studies.
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
@@
 
-\kw{Mention potential complimentary studies of direct sampling, and how all these suggestions are still useful regardless of whether we are doing root-finding or not.}
-
 To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
-We point out hyper-parameters in the usual population synthesis simulation context are not actually parameters related to the population,
+We point out hyperparameters in the usual population synthesis simulation context are not actually parameters related to the population,
 but parameters about the evolutionary function.
@@
 We "back propagate" the posterior in the observable space to the evolutionary parameter space,
-thus allowing us to study hyper-parameters and its correlation with progenitor parameters in a data-driven manner.
-Our method makes less assumption than the traditional forward modelling approach,
-which often fix the hyper-parameters across the entire population.
-Since we are not limited to the fix hyper-parameters assumption, we can explore the behavior of the hyper-parameters across the population much more efficiently.
-While our work lay down a data analysis pathway to understand the population of GW events,
-no physics can be learned without a comprehensive complimentary physical model.
+thus allowing us to study hyperparameters and its correlation with progenitor parameters in a data-driven manner.
+Our method makes less assumptions than the traditional forward modelling approach,
+which often fix the hyperparameters across the entire population.
+Since we are not limited to the fixed hyperparameters assumption, we can explore the behavior of the hyperparameters across the population much more efficiently.
+While our work lays down a data analysis pathway to understand the population of GW events,
+no physics can be learned without a comprehensive physical model.
 We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:

### 2022-05-24 src/scripts/figure2_varyAlphaFacc.py
update figure 2 script

@@ -12,37 +12,14 @@ plt.rcParams["mathtext.fontset"] = "dejavuserif"
 models = ["A22_02", "A22", "A22_5", "KW"]
 
 
-def get_data(model):
-    if model == "KW":
-        bpp = pd.read_hdf(paths.data / "{}_rerun.h5".format(model), key="bpp")
-        initC = pd.read_hdf(paths.data / "{}_rerun.h5".format(model), key="initC")
-    else:
-        bpp = pd.read_hdf(paths.data / "{}_rerun.hdf".format(model), key="bpp")
-        initC = pd.read_hdf(paths.data / "{}_rerun.hdf".format(model), key="initC")
+def get_initC_list(models):
+    initC_list = []
+    for m in models:
+        initC_list.append(pd.read_hdf(paths.data / 'fig_2_dat.h5', key=m))
+    
+    return initC_list
 
-    mergers = bpp.loc[(bpp.kstar_1 == 14) & (bpp.kstar_2 == 14) & (bpp.evol_type == 3)]
-    formation = (
-        bpp.loc[
-            (bpp.bin_num.isin(mergers.bin_num))
-            & (bpp.kstar_1 == 14)
-            & (bpp.kstar_2 == 14)
-        ]
-        .groupby("bin_num")
-        .first()
-        .reset_index()
-    )
-    bpp = []
-    return initC.loc[initC.bin_num.isin(mergers.bin_num)], mergers, formation
-
-
-initC_list = []
-merger_list = []
-formation_list = []
-for m in models:
-    dat = get_data(model=m)
-    initC_list.append(dat[0])
-    merger_list.append(dat[1])
-    formation_list.append(dat[2])
+initC_list = get_initC_list(models)
 
 labels = [
     r"$\alpha=0.2$, $f_{\rm{acc}}=0.5$",

### 2022-05-24 src/tex/ms.tex

@@
 progenitor populations of GW mergers is through population synthesis simulations
-for a host of formation environments including isolated binary stars, triple
-systems, globular clusters, nuclear star clusters, AGN disks, and more
-\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
+for a host of formation environments. In the case of isolated binary evolution, merging 
+double compact objects can form from massive binary stars through the
+standard stable mass transfer or common envelope channels 
+\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022} 
+as well as through chemically homogenous evolution \citep{Mandel2016, Marchant2016, deMink2016} 
+or with population III stars \citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017, Tanikawa2021, Tanikawa2022}.
+Merging DCO binaries can also originate from a wide variety of dynamically active environments including 
+triple (or higher multiple) systems \citep[e.g.][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021}, 
+as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017, DiCarlo2020, Chattopadhyay2022}, 
+globular clusters \citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014, Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019}, 
+nuclear star clusters \citep[][]{Miller2009, Antonini2016}, or a mix of all three \citep[e.g.][]{Mapelli2022}.
+Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions 
+\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} can lead 
+to merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 
+formation envrionment and channels within, see \citet{Mandel2022}. Traditionally in these studies, the
 population is studied through a Monte-Carlo approach.  Using theoretically

### 2022-05-25 src/tex/ms.tex
trying to massage the intro a bit but maybe not very successfully

@@
 \begin{document}
-\title{Backpropagating Gravitational wave events}
+\title{Backpropagating Gravitational wave events \kb{can we come up with something a bit more flashy?}}
 
@@
 the choice of parameters that best represents the evolution of a binary system
-may depend in an as-yet-to-be-determined way on the system's properties; indeed,
+may depend on an as-yet-to-be-determined way on the system's properties. Indeed,
 measuring the variation of population synthesis parameter settings with binary
-properties is one of the principal ways to extract information about binary
+properties is one of the principal ways to extract information about binary evolution
 astrophysics from catalogs. Here we propose an inference pipeline to
-simultaneously infer stellar progenitor parameters (e.g. zero-age main sequence
-masses) and population synthesis parameter settings controlling modeled binary
+simultaneously infer zero-age main sequence properties and population 
+synthesis parameter settings controlling modeled binary
 evolution from individual gravitational wave observations of merging compact
@@
 in a catalog of gravitational wave observations.  We apply our pipeline to
-observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog
-(GWTC3).  We showcase the effectiveness of this pipeline with a detailed study
+observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog.  
+We showcase the effectiveness of this pipeline with a detailed study
 of the progenitor properties and population synthesis settings that produce
@@
 ground-based GW detector network
-\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
+\citep{Aso2013,LIGOScientificCollaboration2015,Acernese2015,Abbott2018,Buikema2020,Tse2019,Acernese2019,Akutsu2021},
 we are beginning to constrain the astrophysical processes which shape the
@@
 progenitor populations of GW mergers is through population synthesis simulations
-for a host of formation environments. In the case of isolated binary evolution, merging 
+of stellar populations from host of formation environments. In the case of isolated binary star evolution, merging 
 double compact objects can form from massive binary stars through the
@@
 Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions 
-\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} can lead 
-to merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 
-formation envrionment and channels within, see \citet{Mandel2022}. Traditionally in these studies, the
-population is studied through a Monte-Carlo approach.  Using theoretically
-motivated distributions of initial parameters (age, metallicity, masses, orbital
-separations, eccentricities, etc), initial populations are evolved with fixed
-astrophysical assumptions (population synthesis parameter settings) to produce a
-synthetic catalog of mergers which can be compared to observations from
-GW detectors.
+\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} 
+can produce merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 
+formation environment and channels within, see \citet{Mandel2022}.
 
-%As the rate of detection of gravitational-wave (GW) increases with the sensitivity of the \kb{ground-based} GW detector network \kw{cite},
-%understanding astrophysical processes related to progenitors of GW events through analyzing a population of GW events has become more popular \kw{cite}. 
-%\kb{I think we can reword the end of this previous sentence -- I should come back to it.}
-%A very common way to study the origin of detected GW events is through population synthesis simulations \kw{cite} \kb{cite Mandel \& Broekgaarden review},
-%which evolve populations of various kinds of binaries to form populations of GW events. \kb{Do we want to do isolated only here? or other channels?}
-%\kb{The process to forward model a population of GW events relies on random samples }
-%Forward modelling of a population of GW events start with sampling a set of initial binary parameters from some distributions related to the progenitor population,
-%such as the initial mass function or star-formation history,
-%then evolve each individual members according to the physics of interest until some stopping criteria is met.
-%This usually mean either a merger event or the binary system stays stable for a Hubble time.
-%After this, one can collect all the GW emitting binaries in the synthetic population and compare to the observed population.
+Traditionally, population synthesis studies simulate merging DCO populations with a 
+Monte-Carlo approach.  Using theoretically motivated distributions of initial parameters 
+like age, metallicity, component mass, orbital separation, and eccentricity, initial 
+populations are evolved with fixed astrophysical assumptions, or population synthesis 
+parameter settings, to produce a synthetic catalog of DCO mergers which can be compared 
+to parameterized models derived from observations. This approach contains several assumptions, 
+both in the simulations and model parameterizations, which add increasing levels of 
+uncertainty in population-wide comparsions of simulations and data. This compounds with the fact 
+that there is no guarantee the entire population should share the same fixed hyperparameters, 
+as is often assumed in most population synthesis studies.
+This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
+required to compare data to simulations which span the full set of uncertain physical models.
+%However, this issue is not a fundamental issue, but instead an apparent obstacle which arises from an inefficient approach to the problem.
+Furthermore, each binary system could in principle come from a different formation environment and hence 
+have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
+Instead, it is likely advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
+for each individual binary system, then perform population in analyses in that space.
 
-While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
-it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
-One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often 
-cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
-On top of this, these values are often not calibrated to data.
-Even with recent developments in using emulators to speed up the simulation process, training the emulators 
-still requires a significant amount of computation to cover a wide range of uncertainty.
+%One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often 
+%cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
+%(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
+%On top of this, these values are often not calibrated to data.
+While forward modelling DCO populations has been the canonical procedure to 
+compare simulations to GW observations, it is also not the most efficient way 
+to constrain the underlying physical processes, or ``hyperparameters", which drive the 
+evolution of the population. Even with recent developments in using emulators to speed 
+up the simulation process \citep[e.g.][]{Wong2021}, training the emulators 
+still requires a significant amount of computation to cover a wide range of uncertainty in the hyperparemeter space.
 As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
-run a grid which varies combinations of uncertain physics.
+run simulations on a grid which varies combinations of uncertain physics.
 For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
@@
 
-\kw{Reorganizing discussion related to Jeff's work such that it doesn't seem like our method is just an extension of DartBoard.}
+%\kw{Reorganizing discussion related to Jeff's work such that it doesn't seem like our method is just an extension of DartBoard.}
 
-This compounds with the fact that there is no guarantee the entire population should share the same fixed hyperparameters, 
-as is often assumed in most population synthesis studies.
-This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
-required to compare data to simulations which span the full set of uncertain physical models.
-However, this issue is not a fundamental issue, but an apparent obstacle which arises from an inefficient approach to the problem.
-Furthermore, each binary system could in principle come from a different formation environment and hence 
-have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
-In this case, it is advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space.
 \citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018} 
-to determine the ZAMS parameters which produce GW150914-like BBH merger based on posterior samples for GW150914 and a fixed 
+to determine the ZAMS parameters which produce GW150914-like BBH merger based on 
+posterior samples for GW150914 and a fixed 
 set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
@@
 following a two-stage process.
-First, we solve a root finding problem to obtain guesses that are likely to produce the desired system properties in the joint progenitor--hyper-parameter space.
-We then sample the posterior in the joint space with Markov chain Monte Carlo (MCMC) algorithm that is initialized using the roots found in the previous stage.
+First, we solve a root finding problem to obtain guesses that are likely to produce the 
+desired system properties in the joint progenitor--hyperparameter space.
+We then sample the posterior in the joint space with Markov chain Monte Carlo 
+(MCMC) algorithm that is initialized using the roots found in the previous stage.
 The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
@@
 define their random variables implicitly such that the random variables are drawn within the program instead of 
-being passed as arguments. This means the function we use to evolve the binary is not fully deterministic.
+being passed as arguments. This means that the process of evolving the binary is not fully deterministic.
 Thus, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not 
@@
 evolutionary parameters to the GW-observable space to check whether the reprojected posterior agrees 
-with the posterior given by the LVK collaboration. We use KL divergence to measure the agreement between the two posterior 
+with the posterior given by the LVK collaboration.
+
+% Validity paragraph
+%\wf{What if we made this a subsection of ``methods'', and then just one sentence here about the values of the KL divergence for GW150914?}
+%To check the validity of the recovered posterior over evolutionary parameters,
+%it is essential to check whether it is consistent with the original posterior in
+%the observable space. We reproject the recovered posterior to the observable
+%space, and compare the KL divergence between the original posterior and the
+%reprojected posterior.
+
+We use KL divergence to measure the agreement between the two posterior 
 distributions as:
@@
 
-A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the 
+\noindent A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the 
 observable space, and is thus a viable channel for that specific event. Otherwise, it either means \texttt{COSMIC} 
@@
 by reweighting the posterior distribution according to a particular SFR model,
-\kb{and} therefore do not have to rerun the inference everytime we change the SFR model. This is especially helpful
+and therefore do not have to rerun the inference everytime we change the SFR model. This is especially helpful
 since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep{Broekgaarden2021}.
@@
 analysis.
-%\kb{We need to describe the interpretation of this as well.}
-%\kw{I think we can just hightlight the fact that we have three kinds of correlation, progenitor-hyper, progenitor-hyper, hyper-hyper,
-%since the implication is related to figure @Katie made.}
-%\kb{Need to describe the formation channels and gaps here.}
 
@@
 
-% \kw{Talk about implication on single event posterior.}
 Once we obtain a set of binaries which successfully map the ZAMS parameters and
@@
 BHs with the correct masses that merge within a Hubble time.
-%\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}
-%\kb{done!}
 
@@
 
-% Validity paragraph
-\wf{What if we made this a subsection of ``methods'', and then just one sentence here about the values of the KL divergence for GW150914?}
-To check the validity of the recovered posterior over evolutionary parameters,
-it is essential to check whether it is consistent with the original posterior in
-the observable space. We reproject the recovered posterior to the observable
+We reproject the recovered posterior to the observable
 space, and compare the KL divergence between the original posterior and the
-reprojected posterior. In particular, we compute KL divergence of the samples
-obtained from root-finding alone and after MCMC sampling. \wf{Clean up wording.}
+reprojected posterior. We compute KL divergence of the samples
+obtained from root-finding alone as well as after the MCMC sampling procedure. %\wf{Clean up wording.}
 For GW150914, the KL divergence between the reprojected posterior and the
@@
 % Paragraph discussing the potential application on the population side.
-To illustrate the potential benefit of our method on the population level, we
+To illustrate the potential benefit of using our method on the population level, we
 perform the same analysis for all events in GWTC3. Plotted in figure
@@
 the $68\%$ credible interval of the posterior density for that particular event.
-There is a suggestive trend showing $f_{\rm acc}$ could increase as the mass of
-the progenitor increases. Such a trend implies the stable mass transfer (SMT)
+There is a suggestive trend showing that $f_{\rm acc}$ could increase as the mass of
+the progenitor increases. Such a trend implies the stable mass transfer 
 phase of a less massive binary would be preferentially non-conservative, with

### 2022-05-25 src/tex/ms.tex
Minor abstract wording change.

@@
 One promising way to extract information about stellar astrophysics from
-gravitational wave catalogs is to compare the catalog to the outputs of
+gravitational wave catalogs is to compare the catalog to the outputs of stellar
 population synthesis modeling with varying physical assumptions.  The parameter
@@
 measuring the variation of population synthesis parameter settings with binary
-properties is one of the principal ways to extract information about binary evolution
-astrophysics from catalogs. Here we propose an inference pipeline to
-simultaneously infer zero-age main sequence properties and population 
-synthesis parameter settings controlling modeled binary
-evolution from individual gravitational wave observations of merging compact
-binaries.  Our pipeline can efficiently explore the high-dimensional space of
-population synthesis settings and progenitor system properties for each system
-in a catalog of gravitational wave observations.  We apply our pipeline to
-observations in the third LIGO Virgo Kagra Gravitational Wave Transient Catalog.  
+properties is one of the principal ways to extract information about binary
+evolution astrophysics from catalogs. Here we propose an inference pipeline to
+simultaneously infer zero-age main sequence properties and population synthesis
+parameter settings controlling modeled binary evolution from individual
+gravitational wave observations of merging compact binaries.  Our pipeline can
+efficiently explore the high-dimensional space of population synthesis settings
+and progenitor system properties for each system in a catalog of gravitational
+wave observations.  We apply our pipeline to observations in the third LIGO
+Virgo Kagra Gravitational Wave Transient Catalog.  
 We showcase the effectiveness of this pipeline with a detailed study

### 2022-05-25 src/tex/ms.tex
New title.

@@
 \begin{document}
-\title{Backpropagating Gravitational wave events \kb{can we come up with something a bit more flashy?}}
+\title{Gravitaitonal Wave Event Progenitor Synthesis}
 
@@
 properties is one of the principal ways to extract information about binary
-evolution astrophysics from catalogs. Here we propose an inference pipeline to
-simultaneously infer zero-age main sequence properties and population synthesis
-parameter settings controlling modeled binary evolution from individual
-gravitational wave observations of merging compact binaries.  Our pipeline can
-efficiently explore the high-dimensional space of population synthesis settings
-and progenitor system properties for each system in a catalog of gravitational
-wave observations.  We apply our pipeline to observations in the third LIGO
-Virgo Kagra Gravitational Wave Transient Catalog.  
+evolution from catalogs. Here we propose an inference pipeline to simultaneously
+infer zero-age main sequence properties and population synthesis parameter
+settings controlling modeled binary evolution from individual gravitational wave
+observations of merging compact binaries.  Our pipeline can efficiently explore
+the high-dimensional space of population synthesis settings and progenitor
+system properties for each system in a catalog of gravitational wave
+observations.  We apply our pipeline to observations in the third LIGO Virgo
+Kagra Gravitational Wave Transient Catalog.  
 We showcase the effectiveness of this pipeline with a detailed study

### 2022-05-25 src/tex/ms.tex
Will edited the 'graph on re-projection.

@@
 
-We reproject the recovered posterior to the observable
-space, and compare the KL divergence between the original posterior and the
-reprojected posterior. We compute KL divergence of the samples
-obtained from root-finding alone as well as after the MCMC sampling procedure. %\wf{Clean up wording.}
-For GW150914, the KL divergence between the reprojected posterior and the
-original posterior is \kw{fill later} nats on average if we use the root-finding
-result alone, whereas the MCMC result has a higher KL divergence at \kw{fill}
-nats. The KL divergence stays stable for reprojecting with different random
-seeds. In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected
-posterior from both the root-finding process alone and the MCMC result agrees
-well with the original posterior, even though the agreement between the MCMC
-result and the original posterior is not perfect. This degradation of the
-agreement comes from the fact that we are using a KDE to estimate the posterior
-density in the observable space. Note that the result from root-finding alone
-and the result from MCMC agrees well in the evolutionary parameter space.
-Compared to the result shown in a previous study, our result covers almost all
-the posterior of GW150914 in the masses space, which means our result is more
-consistent with the observation.
+It is interesting to re-project the posterior over progenitor ($\bm{\theta}'$)
+and hyperparameters ($\bm{\lambda}$) using fresh random variables ($\bm{X}$) to
+the observable space of GW parameters ($\bm{\theta}$).  Agreement between the
+observed parameters and the re-projected parameters indicates good exploration
+of the random variable space and lack of sensitivity to the details of these
+random parameters.  Figure \ref{fig:GW150914_reprojection} shows good agreement
+between the re-projected root-finding outputs as well as the re-projected MCMC
+draws for our analysis of GW150914.  Unlike \citet{Andrews2021}, we find there
+are no regions of observable space inaccessible to our evolutionary models; this
+difference arises because we allow the evolutionary hyperparameters
+$\bm{\lambda}$ to vary.
 

### 2022-05-25 src/tex/ms.tex
Will cites Lieke and others on redshift evolution of SFR.

@@
 arbitrarily complicated models of star formation and metalliticy evolution by
- re-weighting the samples we obtain \wf{Cite at least Lieke here}. Once we have
- pulled the GW event posterior back from the observable space to the
- evolutionary parameter space, we can apply the same hierarchical inference
- methods currently used to infer the compact binary population to our progenitor
- population.  In particular, we can use our progenitor population to infer the
- formation rate of compact binary population progenitor systems over cosmic time
- without needing to make artifical and simplistic assumptions about the delay
- time distribution or the metallicity-specific star formation rate \wf{again,
- cite Lieke, and also Vitale, Farr, Ng, Rodriguez?}.  Comparing the formation
- rate of progenitors to the star formation rate provides an avenue to check our
+re-weighting the samples we obtain \citep[e.g.][]{VanSon2022}. Once we have
+pulled the GW event posterior back from the observable space to the evolutionary
+parameter space, we can apply the same hierarchical inference methods currently
+used to infer the compact binary population to our progenitor population.  In
+particular, we can use our progenitor population to infer the formation rate of
+compact binary population progenitor systems over cosmic time without needing to
+make artifical and simplistic assumptions about the delay time distribution or
+the metallicity-specific star formation rate
+\citep{Vitale2019,Ng2021,VanSon2022}.  Comparing the formation rate of
+progenitors to the star formation rate provides an avenue to check our
 understanding of binary evolution and the relative rate of other formation

### 2022-05-25 src/tex/ms.tex
Cite Mike re: multiple formation channels.

@@
 For example, event with either of the component mass larger than the PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
-Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3.
-As pointed out in the literature \kw{Cite mike}, it is unlikely that all the observed events in GWTC3 can be explained by a single formation channel.
+Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3 \citep{Zevin2021}.
 As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.

### 2022-05-26 src/tex/ms.tex
Will makes introduction edits.

@@
 progenitor populations of GW mergers is through population synthesis simulations
-of stellar populations from host of formation environments. In the case of isolated binary star evolution, merging 
+of stellar populations from a host of formation environments. In the case of isolated binary star evolution, merging
 double compact objects can form from massive binary stars through the
@@
 can produce merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each
-formation environment and channels within, see \citet{Mandel2022}.
+formation environment and channels within, see \citet{Mandel2022} and references therein.
 
-Traditionally, population synthesis studies simulate merging DCO populations with a 
-Monte-Carlo approach.  Using theoretically motivated distributions of initial parameters 
-like age, metallicity, component mass, orbital separation, and eccentricity, initial 
-populations are evolved with fixed astrophysical assumptions, or population synthesis 
-parameter settings, to produce a synthetic catalog of DCO mergers which can be compared 
-to parameterized models derived from observations. This approach contains several assumptions, 
-both in the simulations and model parameterizations, which add increasing levels of 
-uncertainty in population-wide comparsions of simulations and data. This compounds with the fact 
-that there is no guarantee the entire population should share the same fixed hyperparameters, 
-as is often assumed in most population synthesis studies.
-This suggests that there could be a large (possibly infinite) family of hyper-parameter distributions 
-required to compare data to simulations which span the full set of uncertain physical models.
-%However, this issue is not a fundamental issue, but instead an apparent obstacle which arises from an inefficient approach to the problem.
-Furthermore, each binary system could in principle come from a different formation environment and hence 
-have different progenitor parameters as well as hyperparameters, which define uncertain physics associated with each environment.
-Instead, it is likely advantageous to determine the posterior density in the progenitor and hyper-parameter joint space 
-for each individual binary system, then perform population in analyses in that space.
+Traditionally, population synthesis studies simulate merging DCO populations
+with a Monte-Carlo approach.  Using theoretically motivated distributions of
+initial parameters like age, metallicity, component mass, orbital separation,
+and eccentricity, initial populations are evolved with fixed astrophysical
+assumptions, or population synthesis parameter settings, to produce a synthetic
+catalog of DCO mergers which can be compared to parameterized models derived
+from observations. This approach relies on several assumptions, both in the
+simulations and model parameterizations fitted to the GW data, which are
+unlikely to be fully correct.  For example, there is no reason to suppose that
+the entire population of DCOs should evolve with the same fixed hyperparameters;
+there is no a priori reason the fitted parameterized model should capture the
+relevant features of the observed DCO distributions to correspond to the binary
+physics of interest; etc. A better approach would be to compare progenitor
+parameters and hyperparameter settings directly to DCO observations, treating
+the population synthesis model as a mapping from progenitor to merger
+parameters; merger parameters can then be mapped into observations using a
+waveform family, and parameter inference proceeds in the usual way
+\citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
+hyperparameter settings, was advanced in \citet{Andrews2018,Andrews2021}, as we
+discuss in more detail below.
 
@@
 %On top of this, these values are often not calibrated to data.
-While forward modelling DCO populations has been the canonical procedure to 
-compare simulations to GW observations, it is also not the most efficient way 
-to constrain the underlying physical processes, or ``hyperparameters", which drive the 
-evolution of the population. Even with recent developments in using emulators to speed 
-up the simulation process \citep[e.g.][]{Wong2021}, training the emulators 
-still requires a significant amount of computation to cover a wide range of uncertainty in the hyperparemeter space.
-As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
-run simulations on a grid which varies combinations of uncertain physics.
-For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
-this would still require $1024$ simulations.
-Because population synthesis simulations require hundreds to thousands of cpu hours,
-training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
-
-%\kw{Reorganizing discussion related to Jeff's work such that it doesn't seem like our method is just an extension of DartBoard.}
+Attempting to approach such an improved inference procedure for evolutionary
+hyperparameters using traditional en mass Monte Carlo simulations of entire
+populations of DCO mergers is computationally infeasible. Even with recent
+developments in using emulators to speed up the simulation process
+\citep[e.g.][]{Wong2021}, training the emulators still requires a significant
+amount of computation to cover a wide range of uncertainty in the hyperparemeter
+space. As an example, to emulate a model with $10$ parameters, the simplest way
+to construct a training set of simulations is to run simulations on a grid which
+varies combinations of uncertain physics. For $10$ uncertain physical processes,
+if we consider $2$ variations for each physical process this would still require
+$1024$ simulations. Because population synthesis simulations require hundreds to
+thousands of cpu hours, training an emulator which spans a high dimensional
+uncertainty space remains an impossibility at present.  Thus inference of
+hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
+systems to match a particular observation one- or few-at-a-time.
 
@@
 to determine the ZAMS parameters which produce GW150914-like BBH merger based on
-posterior samples for GW150914 and a fixed 
-set of hyperparameters which define assumptions for how isolated binary-star interactions proceed using 
-\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}.
-However, the method was limited to a single event instead of the population.
-More importantly, the method did not vary the hyperparameters while sampling the posterior density of the ZAMS parameters,
-which limit its capability to be applied to understand a population of binary systems.
+posterior samples for GW150914 and a fixed set of hyperparameters which define
+assumptions for how isolated binary-star interactions proceed using
+\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}. The
+approach is very similar to the one advocated here \emph{except} that
+\citet{Andrews2021} treated the evolutionary hyperparameters for the binary
+physics as fixed within each analysis.  The key extension in this work, which
+will enable population modeling of \emph{both} progenitor parameters and
+hyperparemeter settings, including any possible \emph{dependence} of
+hyperparameter settings on progenitor parameters, is to allow these
+hyperparameters to vary at the same time as intrinsic properties of the
+progenitor system to produce a joint inference over the intrinsic properties of
+the progenitor and the necessary hyperparameter settings to produce the observed
+merger properties.
 
-In this paper, we propose a method to ``backward" model each GW event to its progenitor state 
-while allowing the hyper-parameters to vary across their full range of physical uncertainty
-following a two-stage process.
-First, we solve a root finding problem to obtain guesses that are likely to produce the 
-desired system properties in the joint progenitor--hyperparameter space.
-We then sample the posterior in the joint space with Markov chain Monte Carlo 
-(MCMC) algorithm that is initialized using the roots found in the previous stage.
-The rest of the paper is structured as follows: We detail our method in Sec. ~\ref{sec:method}.
-In Sec. ~\ref{sec:result}, we show the results of applying our method to a number of real events.
-We discuss the implications of this work and future direction in Sec. ~\ref{sec:discussion}.
+In full: we propose a method to ``backward" model each GW event to its
+progenitor state while allowing the hyper-parameters to vary across their full
+range of physical uncertainty following a two-stage process. First, we solve a
+root finding problem to obtain guesses that are likely to produce the desired
+system properties in the joint progenitor--hyperparameter space. We then sample
+the posterior in the joint space that is induced by the DCO merger observations
+using a Markov chain Monte Carlo (MCMC) algorithm that is initialized by the
+roots found in the previous stage.  We demonstrate our algorithm in a realistic
+setting by producing a posterior over progenitor properties and \texttt{COSMIC}
+hyperparameters implied by the observation of GW150914.
+
+We find that the progenitor properties and even ability to produce a
+GW150914-like merger event is strongly correlated with hyperparameter settings;
+different assumptions about the black hole masses at merger in GW150914 imply
+wildly different formation channels and ZAMS progenitor masses for this event
+and some nearby combinations of ZAMS masses are unable, with any reasonable
+hyperparameter settings, to produce a DCO merger like GW150914.  We further
+exhibit preliminary results of an ongoing analysis over the entire GWTC-3
+catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
+transfer may depend non-trivially on the primary black hole mass in merging
+systems; our methodology allows for a systematic study of such dependencies,
+which we will explore more fully in future work.
+
+The rest of the paper is structured as
+follows: We detail our method in Sec. ~\ref{sec:method}. In Sec.
+~\ref{sec:result}, we show the results of applying our method to a number of
+real events. We discuss the implications of this work and future direction in
+Sec. ~\ref{sec:discussion}.
 

### 2022-05-26 src/tex/ms.tex
Will further edits intro for clarity and adds some citations.

@@
 and eccentricity, initial populations are evolved with fixed astrophysical
-assumptions, or population synthesis parameter settings, to produce a synthetic
-catalog of DCO mergers which can be compared to parameterized models derived
-from observations. This approach relies on several assumptions, both in the
-simulations and model parameterizations fitted to the GW data, which are
+assumptions, or population synthesis hyperparameter settings, to produce a
+synthetic catalog of DCO mergers which can be compared to parameterized models
+derived from observations. This approach relies on several assumptions, both in
+the simulations and model parameterizations fitted to the GW data, which are
 unlikely to be fully correct.  For example, there is no reason to suppose that
@@
 parameters; merger parameters can then be mapped into observations using a
-waveform family, and parameter inference proceeds in the usual way
+gravitational waveform family, and parameter inference proceeds in the usual
+way, propagating information back up the mapping chain
 \citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
@@
 hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
-systems to match a particular observation one- or few-at-a-time.
+systems via population synthesis recipes to match a particular observation one-
+or few-at-a-time while simultaneously adjusting the hyperparameters. (Eventually
+it may be even be possible to use more physically-motivated modeling of binary
+physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
 

### 2022-05-26 src/tex/ms.tex

@@
 \end{equation}
-The equality holds because the likelihood depends only on the gravitational waveform generated by
-parameters $\bm{\theta}$ to which the progenitor parameters map onto.  In principle
-this likelihood could be computed at arbitrary values of $\bm{\Theta}$ using the
-same machinery that is used to estimate source parameters in GW
-catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
-of progenitor parameters, hyperparameters, and random variables in
-$\bm{\Theta}$ from a posterior density.
+The equality holds because the likelihood depends only on the gravitational
+waveform generated by parameters $\bm{\theta}$ onto which the progenitor
+parameters map.  In principle this likelihood could be computed at arbitrary
+values of $\bm{\Theta}$ using the same machinery that is used to estimate source
+parameters in GW catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3}
+to develop samples of progenitor parameters, hyperparameters, and random
+variables in $\bm{\Theta}$ from a posterior density.
 
@@
 \end{equation}
-where $\pi_\mathrm{GW}$ is the prior density used for sampling,
-we can save the computation cost associated with evaluating the GW likelihood by rewriting the posterior
-density of $\Theta$ in terms of the posterior density of $\theta$, namely
+where $\pi_\mathrm{GW}$ is the prior density used for sampling, we can save the
+computational cost associated with evaluating the GW likelihood by rewriting the
+posterior density of $\bm{\Theta}$ in terms of the posterior density of
+$\bm{\theta}$, namely
 
 \begin{align}
-    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta(\Theta))},
+    p(\bm{\Theta} | d) = \frac{p(\bm{\theta}(\bm{\Theta})| d) \pi(\bm{\Theta})}{\pi_\mathrm{GW}(\bm{\theta}(\bm{\Theta}))}.
 \end{align}
-which we apply a kernel density estimator to the posterior samples released in GWTC3 to estimate $p(\theta(\Theta)|d)$.
+We apply a kernel density estimator with a Gaussian kernel to the posterior
+samples released in GWTC3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
 
 Including hyperparameters and random variables %in the sampling
-drastically increases the dimensionality of the problem,
-which makes the sampling process more computationally expensive to converge.
-To speed up the convergence, we first solve a root-finding problem to find points that are close to the bulk of the posterior density of a GW event,
-then we use those points to initialize a set of chains in the MCMC process.
+drastically increases the dimensionality of the problem, which can make the
+sampling process much more computationally expensive to converge. To speed up
+the convergence, we first solve a root-finding problem to find points
+$\bm{\Theta}$ that will generate GW parameters $\bm{\theta}$ close to the bulk
+of the posterior density of a GW event.  We then use those points to initialize
+a set of chains in the MCMC process.
 
@@
 
-Retaining explicit control over random variables allows us to marginalize over their contribution,
-and thus allows us to focus on progenitor parameters and hyperparameters.
-However, in practice most population synthesis codes (including \texttt{COSMIC} which we use in this study)
-define their random variables implicitly such that the random variables are drawn within the program instead of
-being passed as arguments. This means that the process of evolving the binary is not fully deterministic.
-Thus, even if the root-finding algorithm performs perfectly, forward modelling a set of roots does not
-guarantee that the simulated population reproduces exactly the set of posterior samples due to
-randomness in the evolution of each binary. To assure that the recovered progenitor and hyperparameters
-robustly correspond to the posterior in the GW-observable space, we forward model (or "reproject") the recovered
-evolutionary parameters to the GW-observable space to check whether the reprojected posterior agrees
-with the posterior given by the LVK collaboration.
+Retaining explicit control over random variables allows us to marginalize over
+their contribution, and thus allows us to focus on progenitor parameters and
+hyperparameters. However, in practice most population synthesis codes (including
+\texttt{COSMIC} which we use in this study) define their random variables
+implicitly such that the random variables are drawn within the program instead
+of being passed as arguments. This means that the process of evolving the binary
+is not fully deterministic. Thus, even if the root-finding algorithm performs
+perfectly, forward modelling a set of roots does not guarantee that the
+simulated population reproduces exactly the set of posterior samples due to
+randomness in the evolution of each binary. To assure that the recovered
+progenitor and hyperparameters robustly correspond to the posterior in the
+GW-observable space, we push forward (or ``reproject'') the recovered
+evolutionary parameters to the GW-observable space to check whether the
+reprojected posterior agrees with the posterior given by the LVK collaboration.
 

### 2022-05-26 src/tex/ms.tex
Take away kaze's comment

@@
     The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
-    Each contour is the $68\%$ credible interval of the posterior density for a particular event. \kw{Check interval number}
+    Each contour is the $68\%$ credible interval of the posterior density for a particular event.
     At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair instability supernova mechanism prevents \texttt{COSMIC} from producing events that are more massive than this cutoff.

### 2022-05-26 src/tex/ms.tex
Add rodrigo to the author list
Add rodrigo's email
Minor updates.

--- 2022-05-26 src/tex/ms.tex
@@
 properties is one of the principal ways to extract information about binary
-evolution from catalogs. Here we propose an inference pipeline to simultaneously
-infer zero-age main sequence properties and population synthesis parameter
-settings controlling modeled binary evolution from individual gravitational wave
+evolution from catalogs. Here we propose a pipeline to simultaneously infer
+zero-age main sequence properties and population synthesis parameter settings
+controlling modeled binary evolution from individual gravitational wave
 observations of merging compact binaries.  Our pipeline can efficiently explore
@@
 observations.  We apply our pipeline to observations in the third LIGO Virgo
-Kagra Gravitational Wave Transient Catalog.
-We showcase the effectiveness of this pipeline with a detailed study
-of the progenitor properties and population synthesis settings that produce
-mergers like the observed GW150914.
+Kagra Gravitational Wave Transient Catalog. We showcase the effectiveness of
+this pipeline with a detailed study of the progenitor properties and population
+synthesis settings that produce mergers like the observed GW150914.
 \end{abstract}

### 2022-05-26 src/tex/ms.tex

@@
 \stonybrook
+
+\author{Rodrigo Luger}
+\email{rluger@flatironinstitute.org}
+\flatiron
 

### 2022-05-26 src/tex/ms.tex
adding in new refs and praying the build works so I dont have to download the data locally

@@
 fail in reproducing a subset of the event in GWTC3. One example of this are GW events with at least one component
-with mass above the pair instability supernova mass limit \kb{cite Farmer, and others}.
+with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2020}.
 The reprojection could also be subject to stochasticity in the evolution of each binary
@@
 The spin of BHs formed in isolated binaries depends strongly on the angular
-momentum transport in massive stars which is still highly uncertain \kb{(see however Bavera+, fuller+)},
+momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
 therefore we only consider the two component masses in the observable space in the both the
@@
 Events with more extreme mass ratios are hard to produce with \texttt{COSMIC},
-and therefore less likely to be accurately recovered by our method. This could
+and therefore less likely to be accurately recovered by our method \citep[e.g.][]{Zevin2020}. This could
 be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as

### 2022-05-26 src/tex/ms.tex
adding in software citations
finally adding in julia
Software citations.
Eliminate seaborn. Also add sentances related to caveat in figure 4

@@
 \begin{document}
-\title{Gravitaitonal Wave Event Progenitor Synthesis}
+\title{Gravitational Wave Event Progenitor Synthesis}
 
@@
 reveal the correlation between progenitor parameters and hyperparameters on a
-population level. A careful treatment of all events in the catalog and
+population level. Here we impose a cut to eliminate event with KL divergence larger
+than 0.1.
+This is a heuristic choice to exclude events that are obviously not compatible with
+\texttt{COSMIC}.
+A careful treatment of all events in the catalog and
 discussion related to the detailed physical implications of figure
@@
     The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
+    Note that $m_{1,{\rm GW}}$ here refers to the reprojected posterior instead of the posterior released by LIGO.
     Each contour is the $68\%$ credible interval of the posterior density for a particular event.
@@
 \section{Acknowledgement}
+The authors are grateful for discussions with the CCA Gravitational Wave Astronomy Group.
+The Flatiron Institute is supported by the Simons Foundation.
+
+\software{
+    \texttt{COSMIC}\ \citep{Breivik2020};
+    \texttt{julia}\ \citep{Julia-2017}
+    \texttt{matplotlib}\ \citep{matplotlib};
+    \texttt{numpy}\ \citep{numpy};
+    \texttt{pandas}\ \citep{mckinney-proc-scipy-2010, reback2020pandas};
+    \texttt{scipy}\ \citep{scipy}
+    \texttt{showyourwork}\ \citep{Luger2021}
+}
 

### 2022-05-26 src/scripts/figure2_varyAlphaFacc.py

@@ -3,7 +3,6 @@ import numpy as np
 
 import matplotlib.pyplot as plt
 from matplotlib import colors
-import seaborn as sns
 import paths
 
 plt.rcParams["font.family"] = "serif"

### 2022-05-26 src/scripts/figure4_population.py

@@ -64,7 +64,7 @@ for i in range(data["coord_array"].shape[0]):
                 label="GWTC3",
                 plot_datapoints=False,
                 plot_density=False,
-                levels=[0.5],
+                levels=[0.68],
                 ax=ax,
                 range=[[0, 45], [0, 1]],
                 contour_kwargs={

### 2022-05-26 src/scripts/figure2_varyAlphaFacc.py
Fine, we will have seaborn in our paper.

@@ -3,6 +3,7 @@ import numpy as np
 
 import matplotlib.pyplot as plt
 from matplotlib import colors
+import seaborn as sns
 import paths
 
 plt.rcParams["font.family"] = "serif"

### 2022-05-26 src/tex/ms.tex
Citing seaborn for colormap we used.

@@
 reveal the correlation between progenitor parameters and hyperparameters on a
-population level. Here we impose a cut to eliminate event with KL divergence larger
-than 0.1.
-This is a heuristic choice to exclude events that are obviously not compatible with
-\texttt{COSMIC}.
-A careful treatment of all events in the catalog and
+population level. A careful treatment of all events in the catalog and
 discussion related to the detailed physical implications of figure
@@
     The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
-    Note that $m_{1,{\rm GW}}$ here refers to the reprojected posterior instead of the posterior released by LIGO.
     Each contour is the $68\%$ credible interval of the posterior density for a particular event.
@@
 \software{
+    \texttt{corner}\ \citep{corner};
     \texttt{COSMIC}\ \citep{Breivik2020};
@@
     \texttt{numpy}\ \citep{numpy};
-    \texttt{pandas}\ \citep{mckinney-proc-scipy-2010, reback2020pandas};
+    \texttt{pandas}\ \citep{McKinneyPandas, reback2020pandas};
     \texttt{scipy}\ \citep{scipy}
+    \texttt{seaborn}\ \citep{Waskom2021}
     \texttt{showyourwork}\ \citep{Luger2021}

### 2022-05-26 src/tex/ms.tex
Update title to our awesome new proposal.
Update ms.tex
Minor abstract update.

--- 2022-05-26 src/tex/ms.tex
@@
 \begin{document}
-\title{Gravitational Wave Event Progenitor Synthesis}
+\title{Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors}
 
@@
 the choice of parameters that best represents the evolution of a binary system
-may depend on an as-yet-to-be-determined way on the system's properties. Indeed,
-measuring the variation of population synthesis parameter settings with binary
-properties is one of the principal ways to extract information about binary
-evolution from catalogs. Here we propose a pipeline to simultaneously infer
-zero-age main sequence properties and population synthesis parameter settings
-controlling modeled binary evolution from individual gravitational wave
-observations of merging compact binaries.  Our pipeline can efficiently explore
-the high-dimensional space of population synthesis settings and progenitor
-system properties for each system in a catalog of gravitational wave
-observations.  We apply our pipeline to observations in the third LIGO Virgo
-Kagra Gravitational Wave Transient Catalog. We showcase the effectiveness of
-this pipeline with a detailed study of the progenitor properties and population
-synthesis settings that produce mergers like the observed GW150914.
+may depend in an as-yet-to-be-determined way on the system's properties.  Here
+we propose a pipeline to simultaneously infer zero-age main sequence properties
+and population synthesis parameter settings controlling modeled binary evolution
+from individual gravitational wave observations of merging compact binaries. Our
+pipeline can efficiently explore the high-dimensional space of population
+synthesis settings and progenitor system properties for each system in a catalog
+of gravitational wave observations.  We apply our pipeline to observations in
+the third LIGO Virgo Kagra Gravitational Wave Transient Catalog. We showcase the
+effectiveness of this pipeline with a detailed study of the progenitor
+properties and population synthesis settings that produce mergers like the
+observed GW150914.  Our pipeline permits a measurement of the variation of
+population synthesis parameter settings with binary properties, if any; we
+present inferences for the recent GWTC-3 transient catalog that suggest that the
+stable mass transfer efficiency parameter may vary with primary black hole mass.  
 \end{abstract}

### 2022-06-04 src/tex/ms.tex
Fix GWTC-3 and hyperparameter

@@
 In full: we propose a method to ``backward" model each GW event to its
-progenitor state while allowing the hyper-parameters to vary across their full
+progenitor state while allowing the hyperparameters to vary across their full
 range of physical uncertainty following a two-stage process. First, we solve a
@@
 
-We find that the progenitor properties and even ability to produce a
+We find that the progenitor properties and even the ability to produce a
 GW150914-like merger event is strongly correlated with hyperparameter settings;
@@
 Since the GW likelihood function
-ultimately depends only on the GW parameters $\theta$, and GWTC3 already has samples over these parameters from a posterior density
+ultimately depends only on the GW parameters $\theta$, and GWTC-3 already has samples over these parameters from a posterior density
 \begin{equation}
@@
 We apply a kernel density estimator with a Gaussian kernel to the posterior
-samples released in GWTC3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
+samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
 
@@
 \texttt{COSMIC}, and isolated binary evolution more generally, carries its own assumptions and is expected to
-fail in reproducing a subset of the event in GWTC3. One example of this are GW events with at least one component
+fail in reproducing a subset of the event in GWTC-3. One example of this are GW events with at least one component
 with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2020}.
@@
     \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
-    \caption{The posterior for GW150914 in both the progenitor parameter space and hyper-parameter space.
+    \caption{The posterior for GW150914 in both the progenitor parameter space and hyperparameter space.
     $M_1$ and $M_2$ are the progenitors' masses. $\log{Z}$ is log metallicity at ZAMS.
@@
 efficiency to explore how keeping a fixed model which only varies one
-hyper-parameter contrasts to our results.
+hyperparameter contrasts to our results.
 Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses
 for three models (first through third columns), each with a different $\alpha$
-but the same hyper-parameters as A21, as well as the results of our sampling
-which allows our hyper-parameters to vary (fourth column). We find that holding
+but the same hyperparameters as A21, as well as the results of our sampling
+which allows our hyperparameters to vary (fourth column). We find that holding
 the accretion efficiency, $f_{\rm{acc}}$, and common envelope ejection
@@
 that produces GW150914-like mergers. In contrast, by allowing the
-hyper-parameters to fully span the model uncertainty, we find that there are
+hyperparameters to fully span the model uncertainty, we find that there are
 distinct ZAMS parameters which produce GW150914-like mergers.
@@
 determine that there are ZAMS parameters which fail to produce GW150914-like
-mergers \emph{regardless of our hyper-parameter choice.} For primary ZAMS masses
+mergers \emph{regardless of our hyperparameter choice.} For primary ZAMS masses
 between $\sim95-135\,M_{\odot}$ and secondary ZAMS masses between
@@
 To illustrate the potential benefit of using our method on the population level, we
-perform the same analysis for all events in GWTC3. Plotted in figure
-\ref{fig:GWTC3_f_acc_mass} is the posterior density in the $m_{1,{\rm
-GW}}-f_{\rm acc}$ space for most of the events in GWTC3. Each contour represents
+perform the same analysis for all events in GWTC-3. Plotted in figure
+\ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
+GW}}-f_{\rm acc}$ space for most of the events in GWTC-3. Each contour represents
 the $68\%$ credible interval of the posterior density for that particular event.
@@
 \ref{sec:method}; we do not include these systems in figure
-\ref{fig:GWTC3_f_acc_mass}. Binaries that form low-mass compact objects have
+\ref{fig:GWTC-3_f_acc_mass}. Binaries that form low-mass compact objects have
 lower amounts of fallback and tend to have correspondingly larger variance in
@@
 
-Figure \ref{fig:GWTC3_f_acc_mass} shows that our method could in principle
+Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
 reveal the correlation between progenitor parameters and hyperparameters on a
@@
 discussion related to the detailed physical implications of figure
-\ref{fig:GWTC3_f_acc_mass} is beyond the scope of this paper.  We defer a
+\ref{fig:GWTC-3_f_acc_mass} is beyond the scope of this paper.  We defer a
 detailed study of the physics related to the population of GW events to future
@@
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
+\includegraphics[width=0.49\textwidth]{figures/GWTC-3_f_acc_mass.pdf}
 \caption{
-    The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC3.
+    The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
     Note that $m_{1,{\rm GW}}$ here refers to the reprojected posterior instead of the posterior released by LIGO.
@@
 }
-\label{fig:GWTC3_f_acc_mass}
+\label{fig:GWTC-3_f_acc_mass}
 \script{figure4_population.py}
@@
 While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements can be implemented in future studies.
-In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC3.
+In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
 For example, event with either of the component mass larger than the PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
-Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC3 \citep{Zevin2021}.
+Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
 As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.

### 2022-06-04 src/tex/ms.tex
Add description to figure 3 and 4, and just plot scritps. I love showyourwork.

@@
 \texttt{COSMIC}, and isolated binary evolution more generally, carries its own assumptions and is expected to
-fail in reproducing a subset of the event in GWTC-3. One example of this are GW events with at least one component
+fail in reproducing a subset of the event in GWTC-3. One example of this is GW events with at least one component
 with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2020}.
-The reprojection could also be subject to stochasticity in the evolution of each binary
+The reprojection could also be subject to stochasticity in the evolution of each binary.
 To account for this, we reproject the posterior multiple times with different random seeds and
@@
     \end{tabular}
-    \caption{A list of parameters we used in this study.}
+    \caption{A list of parameters used in this study.}
     \label{tab:parameters}
@@
 momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space in the both the
+therefore we only consider the two component masses in the observable space in both the
 root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with
 five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
-the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
+the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
 obtained through the root-finding process because the formation redshift does not intrinsically affect
 the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
-formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total
+formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
 time to merger to be less than $13.7\,\rm{Gyr}$.
@@
 by reweighting the posterior distribution according to a particular SFR model,
-and therefore do not have to rerun the inference everytime we change the SFR model. This is especially helpful
+and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
 since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep{Broekgaarden2021}.
 For hyperparameters, we choose parameters that affect
-binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope
+binary evolution in \texttt{COSMIC} most significantly for massive stars, including the common envelope
 efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
@@
 and is thus unconstrained. This effect can also be seen in the ZAMS mass and
-$q_{\rm crit 3}$ panels for masses near $80\,M_{\odot}$ because double-core
+$q_{\rm crit, 3}$ panels for masses near $80\,M_{\odot}$ because double-core
 common-envelope evolution is triggered in \texttt{COSMIC} when the radii of the
@@
 envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very
-non-conservative mass transfer (where $f_{\rm acc}$) is low, to produce BBHs
+non-conservative mass transfer (where $f_{\rm acc}$ is low), to produce BBHs
 with the proper mass.
@@
     $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-    $q_{\rm crit 3}$ is the critical mass ratio on the Hertzsprung Gap,
+    $q_{\rm crit, 3}$ is the critical mass ratio on the Hertzsprung Gap.
+    The contours correspond to the $68\%$ and $95\%$ confidence intervals.
     Note that the redshift is not fitted during the root finding process or the MCMC process.
-    Once we have find the evolutionary parameters, we add the delay time to the lookback time of the observed posterior sample,
+    Once we find the evolutionary parameters, we add the delay time to the lookback time of the observed posterior sample,
     then from the total lookback time we can compute the redshift at ZAMS.
     We highlight three panels in the corner plots to show the fine structure of the set of posterior samples in the evolutionary parameter space.
-    We also color the posterior in a particular panel according to the type of parameters involved in the corner.
+    We also color the posterior in a particular panel according to the type of parameters involved in the corner plot.
     Blue denotes panels that include only progenitor parameters,
-    green denotes panels that include a mix of progenitor and hyperparameters,
+    green denotes panels that include a mix of progenitor parameters and hyperparameters,
     and orange denotes panels that include only hyperparameters.
@@
     \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
-     with fixed set of parameter assumptions matching those of A22 (first three columns) and for binaries which produce
-     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{c,3}$ are allowed to vary, (fourth column).}
+     with a fixed set of parameter assumptions matching those of A21 (first three columns) and for binaries which produce
+     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{crit,3}$ are allowed to vary, (fourth column).}
     \label{fig:compare_fixed_variable}
@@
 
-It is interesting to re-project the posterior over progenitor ($\bm{\theta}'$)
+It is interesting to re-project the posterior over progenitor parameters ($\bm{\theta}'$)
 and hyperparameters ($\bm{\lambda}$) using fresh random variables ($\bm{X}$) to
@@
 \caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
+The contours correspond to the $68\%$ and $95\%$ confidence intervals.
 The blue contour is the reprojected posterior after the root-finding procedure.
@@
 reveal the correlation between progenitor parameters and hyperparameters on a
-population level. Here we impose a cut to eliminate event with KL divergence larger
+population level. Here we impose a cut to eliminate events with KL divergence larger
 than 0.1.

### 2022-06-04 src/scripts/figure3_sanityCheck.py

@@ -98,7 +98,7 @@ observable_array.append(observables[i][index[0]])
 quantile_array.append(np.quantile(a, [0.5 - 0.95 / 2, 0.5, 0.5 + 0.95 / 2], axis=0))
 quantile_array = np.array(quantile_array)
 plt.figure(figsize=(10, 9))
-percentile = [0.7, 0.9]
+percentile = [0.68, 0.95]
 corner.hist2d(
     m1_pred[()],
     m2_pred[()],
@@ -108,6 +108,9 @@ corner.hist2d(
     labels="Root-finding",
     levels=percentile,
     plot_datapoints=False,
+    contour_kwargs={
+    "linewidths": 3,
+    },
 )
 corner.hist2d(
     m1_pred_mcmc[()],
@@ -117,6 +120,9 @@ corner.hist2d(
     labels="MCMC",
     levels=percentile,
     plot_datapoints=False,
+    contour_kwargs={
+    "linewidths": 3,
+    },
 )
 corner.hist2d(
     observables[i][:, 0],
@@ -126,6 +132,9 @@ corner.hist2d(
     labels="Data",
     levels=percentile,
     plot_datapoints=False,
+    contour_kwargs={
+    "linewidths": 3,
+    },
 )
 plt.xlabel(r"$M_1$", fontsize=30)
 plt.ylabel(r"$M_2$", fontsize=30)

### 2022-06-04 src/scripts/figure4_population.py

@@ -75,7 +75,7 @@ for i in range(data["coord_array"].shape[0]):
             ax.collections[-2].set_alpha(0.1)
     except:
         continue
-ax.set_xlabel(r"$m_{\rm 1}$")
+ax.set_xlabel(r"$m_{\rm 1, GW}$")
 ax.set_ylabel(r"$f_{\rm acc}$")
 ax.tick_params(
     axis="both",

### 2022-06-04 src/tex/ms.tex
Grammar edit up to figure 4

@@
     At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair instability supernova mechanism prevents \texttt{COSMIC} from producing events that are more massive than this cutoff.
-    Therefore, events with the majority of posterior support above this cutoff is not compatible with \texttt{COSMIC}, hence has a large KL divergence and excluded from this figure.
+    Therefore, events with the majority of posterior support above this cutoff are not compatible with \texttt{COSMIC}, hence have a large KL divergence and are excluded from this figure.
     On the low mass end, neutron star binaries or neutron star-black hole binaries are subjected to randomness induced by the natal kick,
-    also resulting in a larger and fluctuating KL divergence, therefore they are also excluded from the analysis.
+    also resulting in a larger and fluctuating KL divergences, and therefore are also excluded from the analysis.
 }

### 2022-06-04 src/tex/ms.tex
figure name bug fix
Finish blue edit

@@
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/GWTC-3_f_acc_mass.pdf}
+\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
 \caption{
@@
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements can be implemented in future studies.
+While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements that can be implemented in future studies.
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
-For example, event with either of the component mass larger than the PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
-Alternative channels such as dynamic formation channel will be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
-As GW detectors sensitivity increase, we expect to see more and more events that are unusual in some way.
+For example, event with either of the component mass larger than the lower edge of PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
+Alternative channels such as dynamical formation might be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
+As the sensitivity of GW detectors increases, we expect to see more and more events that are unusual in some way.
 Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
-In this paper, our main focus is to illustrate the concept of "back-propagating" GW events posterior sample in this work,
+In this paper, our main focus is to illustrate the concept of "back-propagating" GW events posterior samples,
 highlighting the capability of our method and motivating the community to build the next generation population synthesis code that can work with our method.
@@
 This introduces significant inefficiency in our root-finding and sampling algorithm.
-The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolution pathway of low mass events such as BNS events.
-The effect of natal kick is suppressed for heavier mass events due to fallback.
+The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolutioary pathway of low mass events such as BNS mergers.
+The effect of the natal kick is suppressed for heavier mass events due to fallback.
 This means events with lighter masses are subject to stochasticity of the function,
-where the sampling process for heavier events behaves as if the evolutionary function we use are deterministic.
+where the sampling process for heavier events behaves as if the evolutionary function we use is deterministic.
 % Indeed, we see this behavior in a couple of case studies.
@@
 
-Another technical note is we use finite differencing to estimate the gradient of the objective function, which could be a significant source of error near transition points in the evolutionary parameter space.
-Also, finite differencing is increasing inefficiency as we increase the dimensionality of the problem.
-To improve the accuracy and efficiency in estimating the gradient of objective function, automatic differentiation is a promising feature that modeler should consider incorporating in their population synthesis code in the future.
+Another technical note is that we use finite differencing to estimate the gradient of the objective function, which could be a significant source of error near transition points in the evolutionary parameter space.
+Also, finite differencing is increasingly inefficient as we increase the dimensionality of the problem.
+To improve the accuracy and efficiency in estimating the gradient of the objective function, automatic differentiation is a promising feature that modelers should consider incorporating in their population synthesis codes in the future.
 
@@
 We "back propagate" the posterior in the observable space to the evolutionary parameter space,
-thus allowing us to study hyperparameters and its correlation with progenitor parameters in a data-driven manner.
+thus allowing us to study hyperparameters and theirs correlations with progenitor parameters in a data-driven manner.
 Our method makes less assumptions than the traditional forward modelling approach,
 which often fix the hyperparameters across the entire population.
-Since we are not limited to the fixed hyperparameters assumption, we can explore the behavior of the hyperparameters across the population much more efficiently.
+Since we are not limited to the fixed hyperparameters assumptions, we can explore the behavior of the hyperparameters across the population much more efficiently.
 While our work lays down a data analysis pathway to understand the population of GW events,
 no physics can be learned without a comprehensive physical model.
-We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:
+We hope this letter will motivate the community to build the next generation population synthesis codes that have the following properties:
 first, they should have explicit control over the random variables so marginalizing over random variables can be done more precisely;
 second, they should be as automatically differentiable as much as possible so exploring the evolutionary parameter space is efficient.
-Combining this work and the next generation population synthesis code,
+Combining this work and the next generation of population synthesis codes,
 we can explore the full parameter space of binary evolution models with the next-generation GW detectors network in the near future.

### 2022-06-05 src/tex/ms.tex
added in COMPAS citations and fixed Farmer2019 citation

@@
 standard stable mass transfer or common envelope channels
-\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022}
+\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Stevenson2017, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022}
 as well as through chemically homogenous evolution \citep{Mandel2016, Marchant2016, deMink2016}
@@
 
-Retaining explicit control over random variables allows us to marginalize over
+Retaining explicit control over random variables, as done in \texttt{COMPAS} 
+a binary population synthesis code \citep{Riley2022}, allows us to marginalize over
 their contribution, and thus allows us to focus on progenitor parameters and
@@
 fail in reproducing a subset of the event in GWTC-3. One example of this is GW events with at least one component
-with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2020}.
+with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2019}.
 The reprojection could also be subject to stochasticity in the evolution of each binary.

### 2022-06-05 src/tex/ms.tex
fixed the majority of the yellow comments, but I need to add in some discussion to the pr

--- 2022-06-04 src/tex/ms.tex
@@
 of gravitational wave observations.  We apply our pipeline to observations in
-the third LIGO Virgo Kagra Gravitational Wave Transient Catalog. We showcase the
+the third third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
 effectiveness of this pipeline with a detailed study of the progenitor
@@
 and eccentricity, initial populations are evolved with fixed astrophysical
-assumptions, or population synthesis hyperparameter settings, to produce a
+assumptions, or population synthesis hyperparameter settings like the stability 
+of Roche-overflow mass transfer, common envelope ejection efficiency, or the 
+strength of compact-object natal kicks, to produce a
 synthetic catalog of DCO mergers which can be compared to parameterized models
@@
 
-We apply the method described in section \ref{sec:method} to the first GW event GW150914.
+We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
 The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
@@
 and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
-since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep{Broekgaarden2021}.
+since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
 For hyperparameters, we choose parameters that affect
@@
 For high-mass binaries, \texttt{COSMIC} struggles to produce events above the
-pair instability supernova mass cutoff $m_{1,{\rm GW}}$, so the posterior in the
+pair instability supernova mass cutoff, so the posterior in the
 evolutionary parameter space only corresponds to part of the posterior in the
-observable space below this cutoff, and therefore events beyond the PISN mass
+observable space below this cutoff, and therefore events beyond the lower edge of the PISN mass
 gap also have higher KL divergence.

### 2022-06-06 src/tex/ms.tex
reworking discussion of kicks

--- 2022-06-05 src/tex/ms.tex
@@
 
-Retaining explicit control over random variables, as done in \texttt{COMPAS} 
-a binary population synthesis code \citep{Riley2022}, allows us to marginalize over
+Retaining explicit control over random variables allows us to marginalize over
 their contribution, and thus allows us to focus on progenitor parameters and
-hyperparameters. However, in practice most population synthesis codes (including
-\texttt{COSMIC} which we use in this study) define their random variables
-implicitly such that the random variables are drawn within the program instead
-of being passed as arguments. This means that the process of evolving the binary
+hyperparameters. The strength and direction of natal kicks for compact objects 
+are directly specified in the \texttt{COMPAS} binary population synthesis code 
+\citep{Riley2022} and can also be specified in \texttt{COSMIC}, however 
+this requires the user to specify the strength and direction of natal kicks as 
+inputs through a user-specified pre-processing script.  In this study, we use 
+\texttt{COSMIC} in it's default state which does not specify the kick strength 
+and direction at run time. %However, in practice most population synthesis codes (including
+%\texttt{COSMIC} which we use in this study) define their random variables
+%implicitly such that the random variables are drawn within the program instead
+%of being passed as arguments by default 
+This means that the process of evolving the binary
 is not fully deterministic. Thus, even if the root-finding algorithm performs

### 2022-06-06 src/tex/ms.tex
rewording some of the next gen pop synth stuff

--- 2022-06-05 src/tex/ms.tex
@@
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
-For example, event with either of the component mass larger than the lower edge of PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
+For example, events with either of the component mass larger than the lower edge of PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
 Alternative channels such as dynamical formation might be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
@@
 Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
-In this paper, our main focus is to illustrate the concept of "back-propagating" GW events posterior samples,
-highlighting the capability of our method and motivating the community to build the next generation population synthesis code that can work with our method.
-To avoid cluttering of focus, we discuss the physical implication of the result presented in this work under our specific assumption (i.e.
-using \texttt{COSMIC} as our evolutionary function) in a companion paper.
+In this paper, our main focus is to illustrate the concept of "back-propagating" GW-event posterior samples,
+highlighting the capability of our method and motivating the benefits of building population synthesis codes that can work seamlessly with our method.
+To avoid cluttering of focus, we discuss the physical implications of the results presented in this work under our specific assumptions (i.e.
+using \texttt{COSMIC} as our evolutionary function) in a future companion paper.
 
@@
 no physics can be learned without a comprehensive physical model.
-We hope this letter will motivate the community to build the next generation population synthesis codes that have the following properties:
-first, they should have explicit control over the random variables so marginalizing over random variables can be done more precisely;
-second, they should be as automatically differentiable as much as possible so exploring the evolutionary parameter space is efficient.
-Combining this work and the next generation of population synthesis codes,
-we can explore the full parameter space of binary evolution models with the next-generation GW detectors network in the near future.
+We hope this letter will motivate the construction of next-generation population synthesis codes that have the following properties:
+first, they should retain explicit control over the random variables so marginalizing over random variables can be done precisely and
+second, they should be as automatically differentiable as possible so that exploring the evolutionary parameter space is efficient.
+By combining the methods presented in this work and future, differentiable population synthesis codes,
+we can explore the full parameter space of binary evolution models with future GW data.
 

### 2022-06-06 src/tex/ms.tex
adding in a couple of refs

@@
 since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
-For hyperparameters, we choose parameters that affect
-binary evolution in \texttt{COSMIC} most significantly for massive stars, including the common envelope
+For hyperparameters, we choose parameters that strongly affect
+binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
 efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
@@
 $q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$.
+employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
+with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
+stars merge since the donor has likely not formed a strong core-envelope boundary 
+\citep{Ivanova2004, Belczynski2008}.
 

### 2022-06-06 src/tex/ms.tex

@@
 their contribution, and thus allows us to focus on progenitor parameters and
-hyperparameters. However, in practice most population synthesis codes (including
-\texttt{COSMIC} which we use in this study) define their random variables
-implicitly such that the random variables are drawn within the program instead
-of being passed as arguments. This means that the process of evolving the binary
+hyperparameters. The strength and direction of natal kicks for compact objects 
+are directly specified in the \texttt{COMPAS} binary population synthesis code 
+\citep{Riley2022} and can also be specified in \texttt{COSMIC}, however 
+this requires the user to specify the strength and direction of natal kicks as 
+inputs through a user-specified pre-processing script.  In this study, we use 
+\texttt{COSMIC} in it's default state which does not specify the kick strength 
+and direction at run time. %However, in practice most population synthesis codes (including
+%\texttt{COSMIC} which we use in this study) define their random variables
+%implicitly such that the random variables are drawn within the program instead
+%of being passed as arguments by default 
+This means that the process of evolving the binary
 is not fully deterministic. Thus, even if the root-finding algorithm performs

### 2022-06-06 src/tex/ms.tex
Add GWOSC citation and posterior sample references

@@
 We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
+We use the "Overall_posterior" posterior samples for GW150914,
+publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
 The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
@@
 % Paragraph discussing the potential application on the population side.
-To illustrate the potential benefit of using our method on the population level, we
-perform the same analysis for all events in GWTC-3. Plotted in figure
+To illustrate the potential benefit of using our method on the population level,
+we perform the same analysis for all events in GWTC-3.
+For events announced in GWTC-1, we use "Overall_posterior", "PublicationSamples","IMRPhenomXPHM_comoving" and "C01:Mixed" posterior samples, respectively.
+All the data are publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}.
+Plotted in figure
 \ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
-GW}}-f_{\rm acc}$ space for most of the events in GWTC-3. Each contour represents
+GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
+We use the  Each contour represents
 the $68\%$ credible interval of the posterior density for that particular event.
@@
 The Flatiron Institute is supported by the Simons Foundation.
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org),
+a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as
+the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS),
+and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council. Virgo is funded, through the European Gravitational Observatory (EGO),
+by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef,
+with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT),
+and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea,
+Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
 

### 2022-06-07 src/tex/ms.tex
Typo correction
Update ms.tex

@@
 We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
-We use the "Overall_posterior" posterior samples for GW150914,
+We use the \texttt{Overall_posterior} posterior samples for GW150914,
 publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
@@
 we perform the same analysis for all events in GWTC-3.
-For events announced in GWTC-1, we use "Overall_posterior", "PublicationSamples","IMRPhenomXPHM_comoving" and "C01:Mixed" posterior samples, respectively.
+For events announced in GWTC-1, we use \texttt{Overall_posterior}, \texttt{PublicationSamples},\texttt{IMRPhenomXPHM_comoving} and \texttt{C01:Mixed} posterior samples, respectively.
 All the data are publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}.
@@
 GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
-We use the  Each contour represents
+Each contour represents
 the $68\%$ credible interval of the posterior density for that particular event.

### 2022-06-07 src/tex/ms.tex
bug fix

@@
 We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
-We use the \texttt{Overall_posterior} posterior samples for GW150914,
+We use the \texttt{Overall\_posterior} posterior samples for GW150914,
 publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
@@
 we perform the same analysis for all events in GWTC-3.
-For events announced in GWTC-1, we use \texttt{Overall_posterior}, \texttt{PublicationSamples},\texttt{IMRPhenomXPHM_comoving} and \texttt{C01:Mixed} posterior samples, respectively.
+For events announced in GWTC-1, we use \texttt{Overall\_posterior}, \texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and \texttt{C01:Mixed} posterior samples, respectively.
 All the data are publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}.

### 2022-10-04 src/tex/ms.tex
Added ack'n of Zevin.
Update ms.tex
Typo in abstract
Add degrees of freedom (Minor point 1)

@@
 of gravitational wave observations.  We apply our pipeline to observations in
-the third third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
+the third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
 effectiveness of this pipeline with a detailed study of the progenitor
-properties and population synthesis settings that produce mergers like the
-observed GW150914.  Our pipeline permits a measurement of the variation of
-population synthesis parameter settings with binary properties, if any; we
-present inferences for the recent GWTC-3 transient catalog that suggest that the
-stable mass transfer efficiency parameter may vary with primary black hole mass.  
+properties and population synthesis settings\textbf{, which represent 10 degrees of
+freedom in total,} that produce mergers like the observed GW150914.  Our pipeline
+permits a measurement of the variation of population synthesis parameter
+settings with binary properties, if any; we present inferences for the recent
+GWTC-3 transient catalog that suggest that the stable mass transfer efficiency
+parameter may vary with primary black hole mass.  
 \end{abstract}
@@
 The authors are grateful for discussions with the CCA Gravitational Wave Astronomy Group.
-The Flatiron Institute is supported by the Simons Foundation.
+The Flatiron Institute is supported by the Simons Foundation.  We thank M. Zevin for helpful comments and suggestions on this manuscript.
 This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org),

### 2022-10-04 src/tex/ms.tex
Minor point 2 nuked

@@
 effectiveness of this pipeline with a detailed study of the progenitor
-properties and population synthesis settings\textbf{, which represent 10 degrees of
-freedom in total,} that produce mergers like the observed GW150914.  Our pipeline
-permits a measurement of the variation of population synthesis parameter
-settings with binary properties, if any; we present inferences for the recent
-GWTC-3 transient catalog that suggest that the stable mass transfer efficiency
-parameter may vary with primary black hole mass.  
+properties and population synthesis settings\textbf{, which represent 10 degrees
+of freedom in total,} that produce mergers like the observed GW150914.  Our
+pipeline permits a measurement of the variation of population synthesis
+parameter settings with binary properties, if any; we \textbf{show that} the
+recent GWTC-3 transient catalog \textbf{could be consistent with a model where
+the stable mass transfer efficiency parameter varies with the primary black hole
+mass.}  
 \end{abstract}

### 2022-10-04 src/tex/ms.tex
Point 3 done

@@
 thousands of cpu hours, training an emulator which spans a high dimensional
-uncertainty space remains an impossibility at present.  Thus inference of
-hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
-systems via population synthesis recipes to match a particular observation one-
-or few-at-a-time while simultaneously adjusting the hyperparameters. (Eventually
-it may be even be possible to use more physically-motivated modeling of binary
-physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
+uncertainty space remains an impossibility at present.  Thus \textbf{it is
+beneficial to infer hyperpameter settings \emph{dynamically}}, generating trial
+DCO systems via population synthesis recipes to match a particular observation
+one- or few-at-a-time while simultaneously adjusting the hyperparameters.
+(Eventually it may be even be possible to use more physically-motivated modeling
+of binary physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
 

### 2022-10-04 src/tex/ms.tex
Skipping point 4 since the referee's point is weird, Addressing point 5

@@
 in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all
-other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
+other roots that fulfill the same acceptance criteria, \textbf{\emph{regardless of the initial guess}}.
 This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.

### 2022-10-04 src/tex/ms.tex
Flag point 5 to will

--- 2022-10-04 src/tex/ms.tex
@@
 in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all
-other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
+other roots that fulfill the same acceptance criteria, \textbf{\emph{regardless of the initial guess used in finding the root}}.
 This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.

### 2022-10-04 src/tex/ms.tex
Point 6 addressed
Point 7 done

@@
 %of being passed as arguments by default 
-This means that the process of evolving the binary
-is not fully deterministic. Thus, even if the root-finding algorithm performs
-perfectly, forward modelling a set of roots does not guarantee that the
+This means that the process of evolving the binary is not fully
+deterministic\textbf{, i.e. we only have access to $\bm{\theta'}$ and
+$\bm{\lambda}$ but not $\bm{X}$}. Thus, even if the root-finding algorithm
+performs perfectly, forward modelling a set of roots does not guarantee that the
 simulated population reproduces exactly the set of posterior samples due to
@@
 reprojected posterior agrees with the posterior given by the LVK collaboration.
+\textbf{In a symbolic form, this means we generate a new $\bm{X}$ for each pair of $\bm{\theta'}$ and
+$\bm{\lambda}$, then compute the new posterior in $\bm{\theta}$.}
 

### 2022-10-04 src/tex/ms.tex
Reorder point 12

@@
 be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as
-discussed in Section~\ref{sec:result}.
-
-Merger observations for which \texttt{COSMIC} struggles are likely to be very
-informative about formation channels and binary evolution physics, and are
-therefore likely worthy of close study.  We anticipate that future work will
-follow up these events in detail.
-
-Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
-reveal the correlation between progenitor parameters and hyperparameters on a
-population level. Here we impose a cut to eliminate events with KL divergence larger
+discussed in Section~\ref{sec:result}. Here we impose a cut to eliminate events with KL divergence larger
 than 0.1.
@@
 \texttt{COSMIC}.
-A careful treatment of all events in the catalog and
+\textbf{A careful treatment of all events in the catalog and
 discussion related to the detailed physical implications of figure
@@
 work.
+
+Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
+reveal the correlation between progenitor parameters and hyperparameters on a
+population level.
+Merger observations for which \texttt{COSMIC} struggles are likely to be very
+informative about formation channels and binary evolution physics, and are
+therefore likely worthy of close study.  We anticipate that future work will
+follow up these events in detail.}
 

### 2022-10-14 src/tex/ms.tex
Add L165
Deal with 388

@@
 and some nearby combinations of ZAMS masses are unable, with any reasonable
-hyperparameter settings, to produce a DCO merger like GW150914.  We further
-exhibit preliminary results of an ongoing analysis over the entire GWTC-3
-catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
-transfer may depend non-trivially on the primary black hole mass in merging
-systems; our methodology allows for a systematic study of such dependencies,
-which we will explore more fully in future work.
+hyperparameter settings, to produce a DCO merger like GW150914.  \textbf{Sensitivity to
+hyperparameter choices has been noted before in the context of common envelope
+efficiency (CITE Kaze & friends), mass transfer efficiency (cite Monica; Simone
+B; etc), and metallicity (cite K.B. & co); <snark>here we can explore all these
+choices and more systematically</snark>.}  We further exhibit preliminary results
+of an ongoing analysis over the entire GWTC-3 catalog \citep{GWTC-3} that
+suggest that the accretion efficiency in stable mass transfer may depend
+non-trivially on the primary black hole mass in merging systems; our methodology
+allows for a systematic study of such dependencies, which we will explore more
+fully in future work.
 
@@
 proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
-enough that the fallback reduces the natal kick to zero. Because of this we do
-not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
-not include the ZAMS orbital period and eccentricity, which are correlated with
-one another, but not strongly with the other progenitor parameters or
-hyperparameters. We find strong correlations between the ZAMS masses, as well as
-strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
-$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
-in the upper-right inset of the figure.
+enough that the fallback reduces the natal kick to zero\textbf{, therefore the inference
+result on $\sigma$ is uninformative}. Because of this we do not include $\sigma$
+in Figure~\ref{fig:GW150914_posterior}. Similarly, we do not include the ZAMS
+orbital period and eccentricity, which are correlated with one another, but not
+strongly with the other progenitor parameters or hyperparameters. We find strong
+correlations between the ZAMS masses, as well as strong correlations between
+$M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and $f_{\rm{acc}}$ and $\alpha$. We show
+scatter plots of each of these combinations in the upper-right inset of the
+figure.
 

### 2022-10-19 src/tex/ms.tex
Fix L497
sorted
Closing out minor comments.

--- 2022-10-04 src/tex/ms.tex
@@
 effectiveness of this pipeline with a detailed study of the progenitor
-properties and population synthesis settings\textbf{, which represent 10 degrees
-of freedom in total,} that produce mergers like the observed GW150914.  Our
-pipeline permits a measurement of the variation of population synthesis
-parameter settings with binary properties, if any; we \textbf{show that} the
-recent GWTC-3 transient catalog \textbf{could be consistent with a model where
-the stable mass transfer efficiency parameter varies with the primary black hole
-mass.}  
+properties and population synthesis settings that produce mergers like the
+observed GW150914.  Our pipeline permits a measurement of the variation of
+population synthesis parameter settings with binary properties, if any; we
+present inferences for the recent GWTC-3 transient catalog that suggest that the
+stable mass transfer efficiency parameter may vary with primary black hole mass.  
 \end{abstract}
@@
 thousands of cpu hours, training an emulator which spans a high dimensional
-uncertainty space remains an impossibility at present.  Thus \textbf{it is
-beneficial to infer hyperpameter settings \emph{dynamically}}, generating trial
-DCO systems via population synthesis recipes to match a particular observation
-one- or few-at-a-time while simultaneously adjusting the hyperparameters.
-(Eventually it may be even be possible to use more physically-motivated modeling
-of binary physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
+uncertainty space remains an impossibility at present.  Thus inference of
+hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
+systems via population synthesis recipes to match a particular observation one-
+or few-at-a-time while simultaneously adjusting the hyperparameters. (Eventually
+it may be even be possible to use more physically-motivated modeling of binary
+physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
 
@@
 in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all
-other roots that fulfill the same acceptance criteria, \textbf{\emph{regardless of the initial guess used in finding the root}}.
+other roots that fulfill the same acceptance criteria, \emph{regardless to its initial guess}.
 This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
@@
 %of being passed as arguments by default 
-This means that the process of evolving the binary is not fully
-deterministic\textbf{, i.e. we only have access to $\bm{\theta'}$ and
-$\bm{\lambda}$ but not $\bm{X}$}. Thus, even if the root-finding algorithm
-performs perfectly, forward modelling a set of roots does not guarantee that the
+This means that the process of evolving the binary
+is not fully deterministic. Thus, even if the root-finding algorithm performs
+perfectly, forward modelling a set of roots does not guarantee that the
 simulated population reproduces exactly the set of posterior samples due to
@@
 reprojected posterior agrees with the posterior given by the LVK collaboration.
-\textbf{In a symbolic form, this means we generate a new $\bm{X}$ for each pair of $\bm{\theta'}$ and
-$\bm{\lambda}$, then compute the new posterior in $\bm{\theta}$.}
 
@@
 To illustrate the potential benefit of using our method on the population level,
-we perform the same analysis for all events in GWTC-3.
-For events announced in GWTC-1, we use \texttt{Overall\_posterior}, \texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and \texttt{C01:Mixed} posterior samples, respectively.
-All the data are publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}.
-Plotted in figure
-\ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
-GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
-Each contour represents
-the $68\%$ credible interval of the posterior density for that particular event.
-There is a suggestive trend showing that $f_{\rm acc}$ could increase as the mass of
-the progenitor increases. Such a trend implies the stable mass transfer
-phase of a less massive binary would be preferentially non-conservative, with
-more conservative mass transfer in more massive binary systems.
+we perform the same analysis for all events in GWTC-3. For events announced in
+GWTC-1, we use \texttt{Overall\_posterior},
+\texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and
+\texttt{C01:Mixed} posterior samples, respectively. All the data are publicly
+available from the Gravitational wave Open Science Center
+\citep{LIGOScientific:2019lzm}. Plotted in figure \ref{fig:GWTC-3_f_acc_mass} is
+the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
+most\footnote{\textbf{We excluded events for which our reconstructed posterior has KL
+divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}}
+of the events in GWTC-3. Each contour represents the $68\%$ credible interval of
+the posterior density for that particular event. There is a suggestive trend
+showing that $f_{\rm acc}$ could increase as the mass of the progenitor
+increases. Such a trend implies the stable mass transfer phase of a less massive
+binary would be preferentially non-conservative, with more conservative mass
+transfer in more massive binary systems.
 
@@
 with an application to GW150914. We show the joint posterior of the event's
-progenitor parameters and hyperparameters.  Our work is both a more efficient
+progenitor parameters and hyperparameters.  Our work is a \textbf{data-driven}
 way to study GW event progenitors, and also allows the possibility of
@@
 pulled the GW event posterior back from the observable space to the evolutionary
-parameter space, we can apply the same hierarchical inference methods currently
-used to infer the compact binary population to our progenitor population.  In
-particular, we can use our progenitor population to infer the formation rate of
-compact binary population progenitor systems over cosmic time without needing to
-make artifical and simplistic assumptions about the delay time distribution or
-the metallicity-specific star formation rate
-\citep{Vitale2019,Ng2021,VanSon2022}.  Comparing the formation rate of
-progenitors to the star formation rate provides an avenue to check our
-understanding of binary evolution and the relative rate of other formation
-channels.
+parameter space, \textbf{and have accounted for selection effects}, we can apply
+the same hierarchical inference methods currently used to infer the compact
+binary population to our progenitor population.  In particular, we can use our
+progenitor population to infer the formation rate of compact binary population
+progenitor systems over cosmic time without needing to make artifical and
+simplistic assumptions about the delay time distribution or the
+metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
+Comparing the formation rate of progenitors to the star formation rate provides
+an avenue to check our understanding of binary evolution and the relative rate
+of other formation channels.
 
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements that can be implemented in future studies.
+While our method allows data-driven exploration of the hyperparameters space, there are a number of improvements that can be implemented in future studies.
 In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
@@
 
-To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
+To summarize, we propose a method to recover the posterior samples in the evolutionary parameter space for each GW event.
 We point out hyperparameters in the usual population synthesis simulation context are not actually parameters related to the population,

### 2022-10-25 src/tex/ms.tex
Adding Zenodo record citation.
add in inference references

--- 2022-10-04 src/tex/ms.tex
@@
 
-\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
-\newcommand{\stonybrook}{\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, USA}}
+\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics,
+Flatiron Institute, New York, NY 10010, USA}}
+\newcommand{\stonybrook}{\affiliation{Department of Physics and Astronomy, Stony
+Brook University, Stony Brook NY 11794, USA}}
 \usepackage{amsmath}
@@
 \newcommand{\kb}[1]{\textcolor{pink}{#1}}
-% \setlength{\floatsep}{0cm}
-% \setlength{\textfloatsep}{0cm}
-% \setlength{\belowcaptionskip}{-5pt}
-% \setlength{\abovecaptionskip}{-5pt}
-% Begin!
+% \setlength{\floatsep}{0cm} \setlength{\textfloatsep}{0cm}
+% \setlength{\belowcaptionskip}{-5pt} \setlength{\abovecaptionskip}{-5pt} Begin!
 \begin{document}
@@
 of gravitational wave observations.  We apply our pipeline to observations in
-the third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
+the third LIGO Virgo Kagra Gravitational Wave Transient Catalog. We showcase the
 effectiveness of this pipeline with a detailed study of the progenitor
-properties and population synthesis settings\textbf{, which represent 10 degrees
-of freedom in total,} that produce mergers like the observed GW150914.  Our
-pipeline permits a measurement of the variation of population synthesis
-parameter settings with binary properties, if any; we \textbf{show that} the
-recent GWTC-3 transient catalog \textbf{could be consistent with a model where
-the stable mass transfer efficiency parameter varies with the primary black hole
-mass.}  
+properties and population synthesis settings that produce mergers like the
+observed GW150914.  Our pipeline permits a measurement of the variation of
+population synthesis parameter settings with binary properties, if any; we
+present inferences for the recent GWTC-3 transient catalog that suggest that the
+stable mass transfer efficiency parameter may vary with primary black hole mass.  
 \end{abstract}
@@
 progenitor populations of GW mergers is through population synthesis simulations
-of stellar populations from a host of formation environments. In the case of isolated binary star evolution, merging
-double compact objects can form from massive binary stars through the
-standard stable mass transfer or common envelope channels
-\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Stevenson2017, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022}
-as well as through chemically homogenous evolution \citep{Mandel2016, Marchant2016, deMink2016}
-or with population III stars \citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017, Tanikawa2021, Tanikawa2022}.
-Merging DCO binaries can also originate from a wide variety of dynamically active environments including
-triple (or higher multiple) systems \citep[e.g.][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021},
-as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017, DiCarlo2020, Chattopadhyay2022},
-globular clusters \citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014, Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019},
-nuclear star clusters \citep[][]{Miller2009, Antonini2016}, or a mix of all three \citep[e.g.][]{Mapelli2022}.
-Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions
-\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black holes \citep[e.g.][]{Bird2016, AliHaimoud2017}
-can produce merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each
-formation environment and channels within, see \citet{Mandel2022} and references therein.
+of stellar populations from a host of formation environments. In the case of
+isolated binary star evolution, merging double compact objects can form from
+massive binary stars through the standard stable mass transfer or common
+envelope channels
+\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022}
+as well as through chemically homogenous evolution \citep{Mandel2016,
+Marchant2016, deMink2016} or with population III stars
+\citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017,
+Tanikawa2021, Tanikawa2022}. Merging DCO binaries can also originate from a wide
+variety of dynamically active environments including triple (or higher multiple)
+systems \citep[e.g.][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021},
+as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017,
+DiCarlo2020, Chattopadhyay2022}, globular clusters
+\citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014,
+Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019}, nuclear star clusters
+\citep[][]{Miller2009, Antonini2016}, or a mix of all three
+\citep[e.g.][]{Mapelli2022}. Finally, more exotic channels like active galactic
+nuclei which combine both gravitational and gas interactions
+\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial
+black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} can produce merging DCOs
+observable with a ground-based detector network. For a discussion of the
+relative rates of each formation environment and channels within, see
+\citet{Mandel2022} and references therein.
 
-Traditionally, population synthesis studies simulate merging DCO populations
-with a Monte-Carlo approach.  Using theoretically motivated distributions of
-initial parameters like age, metallicity, component mass, orbital separation,
-and eccentricity, initial populations are evolved with fixed astrophysical
-assumptions, or population synthesis hyperparameter settings like the stability 
-of Roche-overflow mass transfer, common envelope ejection efficiency, or the 
-strength of compact-object natal kicks, to produce a
+Traditionally, population synthesis studies simulate DCO populations with a
+Monte-Carlo approach. Using theoretically motivated distributions of initial
+parameters like age, metallicity, component mass, orbital separation, and
+eccentricity, initial populations are evolved with fixed astrophysical
+assumptions, or population synthesis hyperparameter settings, to produce a
 synthetic catalog of DCO mergers which can be compared to parameterized models
-derived from observations. This approach relies on several assumptions, both in
-the simulations and model parameterizations fitted to the GW data, which are
-unlikely to be fully correct.  For example, there is no reason to suppose that
-the entire population of DCOs should evolve with the same fixed hyperparameters;
-there is no a priori reason the fitted parameterized model should capture the
-relevant features of the observed DCO distributions to correspond to the binary
-physics of interest; etc. A better approach would be to compare progenitor
-parameters and hyperparameter settings directly to DCO observations, treating
-the population synthesis model as a mapping from progenitor to merger
-parameters; merger parameters can then be mapped into observations using a
-gravitational waveform family, and parameter inference proceeds in the usual
-way, propagating information back up the mapping chain
+derived from observations. \textbf{This method has been applied extensively to
+inference on GW data to inform phenomenological models
+\citep[e.g.][]{Fishbach2017, Wysocki2019, Callister2021, Delfavero2021,
+Farah2022, Delfavero2022}, population synthesis for a single formation
+environment \citep[e.g.][]{Belczynski2016, Stevenson2017b, Taylor2018,
+Barrett2018, Mastrogiovanni2022} or contributions from multiple formation
+environments \citep[e.g][]{Zevin2017, Stevenson2017, Bouffanais2019, Zevin2021,
+Wong2021, Bouffanais2021, Mapelli2022}}. These approaches rely on several
+assumptions, both in the simulations and model parameterizations fitted to the
+GW data, which are unlikely to be fully correct. For example, there is no reason
+to suppose that the entire population of DCOs should evolve with the same fixed
+hyperparameters; there is no a priori reason the fitted parameterized model
+should capture the relevant features of the observed DCO distributions to
+correspond to the binary physics of interest; etc. A better approach would be to
+compare progenitor parameters and hyperparameter settings directly to DCO
+observations, treating the population synthesis model as a mapping from
+progenitor to merger parameters; merger parameters can then be mapped into
+observations using a gravitational waveform family, and parameter inference
+proceeds in the usual way, propagating information back up the mapping chain
 \citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
@@
 
-%One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often
-%cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
-%(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
-%On top of this, these values are often not calibrated to data.
+%One of the most common skepticisms of population synthesis studies is that they
+%employ simulation sets which often cover a small subdomain of the entire
+%parameter space of uncertain physics, or only vary one uncertain process per
+%simulation (e.g. common envelope efficiency) while holding all others fixed to
+%some ``default" values. On top of this, these values are often not calibrated
+%to data.
 Attempting to approach such an improved inference procedure for evolutionary
@@
 
-The rest of the paper is structured as
-follows: We detail our method in Sec. ~\ref{sec:method}. In Sec.
-~\ref{sec:result}, we show the results of applying our method to a number of
-real events. We discuss the implications of this work and future direction in
-Sec. ~\ref{sec:discussion}.
+The rest of the paper is structured as follows: We detail our method in Sec.
+~\ref{sec:method}. In Sec. ~\ref{sec:result}, we show the results of applying
+our method to a number of real events. We discuss the implications of this work
+and future direction in Sec. ~\ref{sec:discussion}.
 
@@
 
-We define the progenitor's zero age main sequence (ZAMS) properties like
-mass, orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the parameters
-that control different uncertain physical prescriptions such as wind strength and common
-envelope efficiency as hyperparameters $\bm{\lambda}$, and random variables
-that affect certain stochastic process such as whether CO natal kick unbinds a
-binary as $\bm{X}$. To avoid clutter in the following derivation, we denote all
-the parameters related to mapping a particular progenitor system into a GW event
-collectively as evolutionary parameters $\bm{\Theta}$, which includes
-$\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.
+We define the progenitor's zero age main sequence (ZAMS) properties like mass,
+orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the
+parameters that control different uncertain physical prescriptions such as wind
+strength and common envelope efficiency as hyperparameters $\bm{\lambda}$, and
+random variables that affect certain stochastic process such as whether CO natal
+kick unbinds a binary as $\bm{X}$. To avoid clutter in the following derivation,
+we denote all the parameters related to mapping a particular progenitor system
+into a GW event collectively as evolutionary parameters $\bm{\Theta}$, which
+includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.
 
@@
 \end{equation}
-The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one
-due to degeneracies in the different physical processes and initial parameters
-of the GW event progenitors.
-In order to draw inferences about $\bm{\Theta}$ from GW data, we
-must be able to evaluate the likelihood of that data at fixed progenitor
-parameters, namely
+The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one due
+to degeneracies in the different physical processes and initial parameters of
+the GW event progenitors. In order to draw inferences about $\bm{\Theta}$ from
+GW data, we must be able to evaluate the likelihood of that data at fixed
+progenitor parameters, namely
 \begin{equation}
@@
 
-Since the GW likelihood function
-ultimately depends only on the GW parameters $\theta$, and GWTC-3 already has samples over these parameters from a posterior density
+Since the GW likelihood function ultimately depends only on the GW parameters
+$\theta$, and GWTC-3 already has samples over these parameters from a posterior
+density
 \begin{equation}
@@
 %The root-finding procedure is described as follows:
-For each posterior sample point in the GW-observable space,
-we can find the corresponding evolutionary parameters by solving an optimization problem that minimizes
-the mean square difference between the GW event parameters and evolutionary parameters as:
+For each posterior sample point in the GW-observable space, we can find the
+corresponding evolutionary parameters by solving an optimization problem that
+minimizes the mean square difference between the GW event parameters and
+evolutionary parameters as:
 
@@
 
-In principle, we should only accept solutions that exactly reproduce the LVK posterior samples in the GW-observable space.
-However, it is not feasible to achieve such a condition in practice, therefore we relax the condition in eq.~\ref{eq:loss}
-to a small acceptance threshold. For this study, we picked a threshold of $10^{-2}$. To make sure we find a reasonably
-complete set of progenitor parameters that corresponds to the posterior sample point, we use 1000 different initial guesses
-in solving the optimization problem. As long as the solution fulfills the acceptance criteria, the root is as valid as all
-other roots that fulfill the same acceptance criteria, \textbf{\emph{regardless of the initial guess used in finding the root}}.
-This means, we have the freedom to choose the set of initial guesses however we think might benefit the optimization.
+In principle, we should only accept solutions that exactly reproduce the LVK
+posterior samples in the GW-observable space. However, it is not feasible to
+achieve such a condition in practice, therefore we relax the condition in
+eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
+threshold of $10^{-2}$. To make sure we find a reasonably complete set of
+progenitor parameters that corresponds to the posterior sample point, we use
+1000 different initial guesses in solving the optimization problem. As long as
+the solution fulfills the acceptance criteria, the root is as valid as all other
+roots that fulfill the same acceptance criteria, \emph{regardless to its initial
+guess}. This means, we have the freedom to choose the set of initial guesses
+however we think might benefit the optimization.
 
-Most of the points in the evolutionary parameter space do not produce DCO mergers in a Hubble time. For systems that do not merge,
-we set the binary masses to be $0$ such that the gradient for the same systems is also $0$ and leads the root finder to get stuck on
-its first step.{\footnote{On a plateau of constant loss, the root finder performs no better than a random guess.
-Because of this, a root finder in a high dimensional space will require a significant amount of computation
-to get close to a solution.}} Therefore, it is beneficial to choose initial guesses in region of the evolutionary parameter
-space that are likely to produce DCO mergers. To construct a list of initial guesses, we evolve a set of
-ZAMS parameters uniformly sampled from the initial binary parameter space, then keep the systems that
-merge within a Hubble time.
+Most of the points in the evolutionary parameter space do not produce DCO
+mergers in a Hubble time. For systems that do not merge, we set the binary
+masses to be $0$ such that the gradient for the same systems is also $0$ and
+leads the root finder to get stuck on its first step.{\footnote{On a plateau of
+constant loss, the root finder performs no better than a random guess. Because
+of this, a root finder in a high dimensional space will require a significant
+amount of computation to get close to a solution.}} Therefore, it is beneficial
+to choose initial guesses in region of the evolutionary parameter space that are
+likely to produce DCO mergers. To construct a list of initial guesses, we evolve
+a set of ZAMS parameters uniformly sampled from the initial binary parameter
+space, then keep the systems that merge within a Hubble time.
 
@@
 their contribution, and thus allows us to focus on progenitor parameters and
-hyperparameters. The strength and direction of natal kicks for compact objects 
-are directly specified in the \texttt{COMPAS} binary population synthesis code 
-\citep{Riley2022} and can also be specified in \texttt{COSMIC}, however 
-this requires the user to specify the strength and direction of natal kicks as 
-inputs through a user-specified pre-processing script.  In this study, we use 
-\texttt{COSMIC} in it's default state which does not specify the kick strength 
-and direction at run time. %However, in practice most population synthesis codes (including
-%\texttt{COSMIC} which we use in this study) define their random variables
-%implicitly such that the random variables are drawn within the program instead
-%of being passed as arguments by default 
-This means that the process of evolving the binary is not fully
-deterministic\textbf{, i.e. we only have access to $\bm{\theta'}$ and
-$\bm{\lambda}$ but not $\bm{X}$}. Thus, even if the root-finding algorithm
-performs perfectly, forward modelling a set of roots does not guarantee that the
+hyperparameters. However, in practice most population synthesis codes (including
+\texttt{COSMIC} which we use in this study) define their random variables
+implicitly such that the random variables are drawn within the program instead
+of being passed as arguments. This means that the process of evolving the binary
+is not fully deterministic. Thus, even if the root-finding algorithm performs
+perfectly, forward modelling a set of roots does not guarantee that the
 simulated population reproduces exactly the set of posterior samples due to
@@
 reprojected posterior agrees with the posterior given by the LVK collaboration.
-\textbf{In a symbolic form, this means we generate a new $\bm{X}$ for each pair of $\bm{\theta'}$ and
-$\bm{\lambda}$, then compute the new posterior in $\bm{\theta}$.}
 
-% Validity paragraph
-%\wf{What if we made this a subsection of ``methods'', and then just one sentence here about the values of the KL divergence for GW150914?}
-%To check the validity of the recovered posterior over evolutionary parameters,
-%it is essential to check whether it is consistent with the original posterior in
-%the observable space. We reproject the recovered posterior to the observable
-%space, and compare the KL divergence between the original posterior and the
-%reprojected posterior.
+% Validity paragraph \wf{What if we made this a subsection of ``methods'', and
+%then just one sentence here about the values of the KL divergence for
+%GW150914?} To check the validity of the recovered posterior over evolutionary
+%parameters, it is essential to check whether it is consistent with the original
+%posterior in the observable space. We reproject the recovered posterior to the
+%observable space, and compare the KL divergence between the original posterior
+%and the reprojected posterior.
 
@@
 
-\noindent A small KL divergence means the reprojected \texttt{COSMIC} posterior is similar to the original posterior in the
-observable space, and is thus a viable channel for that specific event. Otherwise, it either means \texttt{COSMIC}
-can only explain part of the posterior or cannot explain the event at all. This is completely expected behavior since
-\texttt{COSMIC}, and isolated binary evolution more generally, carries its own assumptions and is expected to
-fail in reproducing a subset of the event in GWTC-3. One example of this is GW events with at least one component
-with mass above the pair instability supernova mass limit \citep{Woosley2017, Farmer2019}.
-The reprojection could also be subject to stochasticity in the evolution of each binary.
-To account for this, we reproject the posterior multiple times with different random seeds and
-check whether the KL divergence varies significantly. A varying KL divergence means a particular GW event is subject
-to randomness in the evolution of the progenitor, and extra caution should be used when interpreting the result.
+\noindent A small KL divergence means the reprojected \texttt{COSMIC} posterior
+is similar to the original posterior in the observable space, and is thus a
+viable channel for that specific event. Otherwise, it either means
+\texttt{COSMIC} can only explain part of the posterior or cannot explain the
+event at all. This is completely expected behavior since \texttt{COSMIC}, and
+isolated binary evolution more generally, carries its own assumptions and is
+expected to fail in reproducing a subset of the event in GWTC-3. One example of
+this is GW events with at least one component with mass above the pair
+instability supernova mass limit \citep{Woosley2017, Farmer2020}. The
+reprojection could also be subject to stochasticity in the evolution of each
+binary. To account for this, we reproject the posterior multiple times with
+different random seeds and check whether the KL divergence varies significantly.
+A varying KL divergence means a particular GW event is subject to randomness in
+the evolution of the progenitor, and extra caution should be used when
+interpreting the result.
 
@@
     $\alpha$ & Common envelope efficiency & $[0.1,10]$\\
-    $f_{\rm acc}$ & Fraction of accretion during stable mass transfer & $[0,1]$\\
+    $f_{\rm acc}$ & Fraction of accretion during stable mass transfer &
+    $[0,1]$\\
     $q_{\rm crit,3}$ & Critical mass ratio on the Hertzsprung Gap & $[0.1,10]$\\
-    $\sigma$ & Root mean square of Maxwellian natal kick distribution& $[10,300]\ {\rm km/s}$\\
+    $\sigma$ & Root mean square of Maxwellian natal kick distribution&
+    $[10,300]\ {\rm km/s}$\\
 
@@
 
-We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
-We use the \texttt{Overall\_posterior} posterior samples for GW150914,
-publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
-The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-The spin of BHs formed in isolated binaries depends strongly on the angular
-momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space in both the
-root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with
-five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
-the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
-obtained through the root-finding process because the formation redshift does not intrinsically affect
-the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
-formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
-time to merger to be less than $13.7\,\rm{Gyr}$.
-Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
-Planck 2018 cosmological model \citep{Planck2018}.
-The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
-of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
-This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
-by reweighting the posterior distribution according to a particular SFR model,
-and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
-since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
-For hyperparameters, we choose parameters that strongly affect
-binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
-efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
-the critical mass ratio which determines whether Roche-overflow mass transfer
-proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
-$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
-with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
-stars merge since the donor has likely not formed a strong core-envelope boundary 
-\citep{Ivanova2004, Belczynski2008}.
+We apply the method described in section \ref{sec:method} to the first GW event
+GW150914. The parameters involved in the analysis and the range we allow for the
+inference are tabulated in table \ref{tab:parameters}. The spin of BHs formed in
+isolated binaries depends strongly on the angular momentum transport in massive
+stars which is still highly uncertain (see however \citet{Fuller2019,
+Bavera2020}), therefore we only consider the two component masses in the
+observable space in both the root-finding stage and the MCMC stage. For
+progenitor parameters, we characterize each guess with five parameters: the two
+component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period
+$t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that
+the progenitor formation redshift is not obtained through the root-finding
+process because the formation redshift does not intrinsically affect the
+evolution of the binary. Instead, we compute the time it takes for the binary to
+evolve from formation to the merger and add this time to the lookback time when
+a GW event is observed and limit the total time to merger to be less than
+$13.7\,\rm{Gyr}$. Using the total lookback time, we can compute the redshift of
+the binary at ZAMS formation following the Planck 2018 cosmological model
+\citep{Planck2018}. The main benefit of post-processing the redshift in this way
+is that we do not rely on a particular assumption of star formation rate (SFR)
+distribution when we solve for the posterior distribution in redshift. This
+means we can put the prior on formation redshift (or in general,
+metallicity-redshfit distribution) in by reweighting the posterior distribution
+according to a particular SFR model, and therefore do not have to rerun the
+inference every time we change the SFR model. This is especially helpful since
+SFR models are still highly uncertain but strongly impact the local merger rates
+of BBHs \citep{Broekgaarden2021}. For hyperparameters, we choose parameters that
+affect binary evolution in \texttt{COSMIC} most significantly for massive stars,
+including the common envelope efficiency $\alpha$, the fraction of mass accreted
+onto the accretor during stable mass transfer $f_{\rm acc}$, the critical mass
+ratio which determines whether Roche-overflow mass transfer proceeds stably or
+produces a common envelope when the donor star is on the Hertzsprung gap $q_{\rm
+crit,3}$, and the root-mean-square of the Maxwellian distribution employed for
+CO natal kicks, $\sigma$.
 
@@
     \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
-    \caption{The posterior for GW150914 in both the progenitor parameter space and hyperparameter space.
-    $M_1$ and $M_2$ are the progenitors' masses. $\log{Z}$ is log metallicity at ZAMS.
-    $z$ is the redshift at ZAMS.
-    $\alpha$ is the common envelope efficiency.
-    $f_{\rm acc}$ is the fraction of mass accreted during stable mass transfer.
-    $q_{\rm crit, 3}$ is the critical mass ratio on the Hertzsprung Gap.
-    The contours correspond to the $68\%$ and $95\%$ confidence intervals.
-    Note that the redshift is not fitted during the root finding process or the MCMC process.
-    Once we find the evolutionary parameters, we add the delay time to the lookback time of the observed posterior sample,
-    then from the total lookback time we can compute the redshift at ZAMS.
-    We highlight three panels in the corner plots to show the fine structure of the set of posterior samples in the evolutionary parameter space.
-    We also color the posterior in a particular panel according to the type of parameters involved in the corner plot.
-    Blue denotes panels that include only progenitor parameters,
-    green denotes panels that include a mix of progenitor parameters and hyperparameters,
-    and orange denotes panels that include only hyperparameters.
-    }
+    \caption{The posterior for GW150914 in both the progenitor parameter space
+    and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
+    $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
+    is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
+    accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
+    ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
+    $95\%$ confidence intervals. Note that the redshift is not fitted during the
+    root finding process or the MCMC process. Once we find the evolutionary
+    parameters, we add the delay time to the lookback time of the observed
+    posterior sample, then from the total lookback time we can compute the
+    redshift at ZAMS. We highlight three panels in the corner plots to show the
+    fine structure of the set of posterior samples in the evolutionary parameter
+    space. We also color the posterior in a particular panel according to the
+    type of parameters involved in the corner plot. Blue denotes panels that
+    include only progenitor parameters, green denotes panels that include a mix
+    of progenitor parameters and hyperparameters, and orange denotes panels that
+    include only hyperparameters. }
     \label{fig:GW150914_posterior}
@@
 efficiency to explore how keeping a fixed model which only varies one
-hyperparameter contrasts to our results.
-Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses
-for three models (first through third columns), each with a different $\alpha$
-but the same hyperparameters as A21, as well as the results of our sampling
-which allows our hyperparameters to vary (fourth column). We find that holding
-the accretion efficiency, $f_{\rm{acc}}$, and common envelope ejection
-efficiency, $\alpha$, to fixed values greatly reduces the ZAMS parameter space
-that produces GW150914-like mergers. In contrast, by allowing the
-hyperparameters to fully span the model uncertainty, we find that there are
-distinct ZAMS parameters which produce GW150914-like mergers.
+hyperparameter contrasts to our results. Figure~\ref{fig:compare_fixed_variable}
+shows the distribution of ZAMS masses for three models (first through third
+columns), each with a different $\alpha$ but the same hyperparameters as A21, as
+well as the results of our sampling which allows our hyperparameters to vary
+(fourth column). We find that holding the accretion efficiency, $f_{\rm{acc}}$,
+and common envelope ejection efficiency, $\alpha$, to fixed values greatly
+reduces the ZAMS parameter space that produces GW150914-like mergers. In
+contrast, by allowing the hyperparameters to fully span the model uncertainty,
+we find that there are distinct ZAMS parameters which produce GW150914-like
+mergers.
 
@@
     \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
-    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like mergers for three variations of $\alpha$
-     with a fixed set of parameter assumptions matching those of A21 (first three columns) and for binaries which produce
-     GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and $q_{crit,3}$ are allowed to vary, (fourth column).}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
+     mergers for three variations of $\alpha$ with a fixed set of parameter
+     assumptions matching those of A21 (first three columns) and for binaries
+     which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
+     $q_{crit,3}$ are allowed to vary, (fourth column).}
     \label{fig:compare_fixed_variable}
@@
 
-It is interesting to re-project the posterior over progenitor parameters ($\bm{\theta}'$)
-and hyperparameters ($\bm{\lambda}$) using fresh random variables ($\bm{X}$) to
-the observable space of GW parameters ($\bm{\theta}$).  Agreement between the
-observed parameters and the re-projected parameters indicates good exploration
-of the random variable space and lack of sensitivity to the details of these
-random parameters.  Figure \ref{fig:GW150914_reprojection} shows good agreement
-between the re-projected root-finding outputs as well as the re-projected MCMC
-draws for our analysis of GW150914.  Unlike \citet{Andrews2021}, we find there
-are no regions of observable space inaccessible to our evolutionary models; this
-difference arises because we allow the evolutionary hyperparameters
-$\bm{\lambda}$ to vary.
+It is interesting to re-project the posterior over progenitor parameters
+($\bm{\theta}'$) and hyperparameters ($\bm{\lambda}$) using fresh random
+variables ($\bm{X}$) to the observable space of GW parameters ($\bm{\theta}$).
+Agreement between the observed parameters and the re-projected parameters
+indicates good exploration of the random variable space and lack of sensitivity
+to the details of these random parameters.  Figure
+\ref{fig:GW150914_reprojection} shows good agreement between the re-projected
+root-finding outputs as well as the re-projected MCMC draws for our analysis of
+GW150914.  Unlike \citet{Andrews2021}, we find there are no regions of
+observable space inaccessible to our evolutionary models; this difference arises
+because we allow the evolutionary hyperparameters $\bm{\lambda}$ to vary.
 
@@
 \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
-\caption{Reprojecting the posterior in evolutionary parameter space of GW150914 to observable space.
-The contours correspond to the $68\%$ and $95\%$ confidence intervals.
-The blue contour is the reprojected posterior after the root-finding procedure.
-The orange contour is the reprojected posterior after the MCMC procedure.
-The green contour is the posterior plotted using the original LVK posterior samples.
-}
+\caption{Reprojecting the posterior in evolutionary parameter space of GW150914
+to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
+intervals. The blue contour is the reprojected posterior after the root-finding
+procedure. The orange contour is the reprojected posterior after the MCMC
+procedure. The green contour is the posterior plotted using the original LVK
+posterior samples. }
 \label{fig:GW150914_reprojection}
@@
 To illustrate the potential benefit of using our method on the population level,
-we perform the same analysis for all events in GWTC-3.
-For events announced in GWTC-1, we use \texttt{Overall\_posterior}, \texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and \texttt{C01:Mixed} posterior samples, respectively.
-All the data are publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}.
-Plotted in figure
+we perform the same analysis for all events in GWTC-3. Plotted in figure
 \ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
-GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
-Each contour represents
-the $68\%$ credible interval of the posterior density for that particular event.
-There is a suggestive trend showing that $f_{\rm acc}$ could increase as the mass of
-the progenitor increases. Such a trend implies the stable mass transfer
-phase of a less massive binary would be preferentially non-conservative, with
-more conservative mass transfer in more massive binary systems.
+GW}}-f_{\rm acc}$ space for most of the events in GWTC-3. Each contour
+represents the $68\%$ credible interval of the posterior density for that
+particular event. There is a suggestive trend showing that $f_{\rm acc}$ could
+increase as the mass of the progenitor increases. Such a trend implies the
+stable mass transfer phase of a less massive binary would be preferentially
+non-conservative, with more conservative mass transfer in more massive binary
+systems.
 
@@
 For high-mass binaries, \texttt{COSMIC} struggles to produce events above the
-pair instability supernova mass cutoff, so the posterior in the
+pair instability supernova mass cutoff $m_{1,{\rm GW}}$, so the posterior in the
 evolutionary parameter space only corresponds to part of the posterior in the
-observable space below this cutoff, and therefore events beyond the lower edge of the PISN mass
+observable space below this cutoff, and therefore events beyond the PISN mass
 gap also have higher KL divergence.
@@
 Events with more extreme mass ratios are hard to produce with \texttt{COSMIC},
-and therefore less likely to be accurately recovered by our method \citep[e.g.][]{Zevin2020}. This could
-be due to our method assuming a single value for $f_{\rm acc}$ and $\alpha$ as
-discussed in Section~\ref{sec:result}. Here we impose a cut to eliminate events with KL divergence larger
-than 0.1.
-This is a heuristic choice to exclude events that are obviously not compatible with
-\texttt{COSMIC}.
-\textbf{A careful treatment of all events in the catalog and
-discussion related to the detailed physical implications of figure
+and therefore less likely to be accurately recovered by our method
+\citep[e.g.][]{Zevin2020}. This could be due to our method assuming a single
+value for $f_{\rm acc}$ and $\alpha$ as discussed in Section~\ref{sec:result}.
+
+Merger observations for which \texttt{COSMIC} struggles are likely to be very
+informative about formation channels and binary evolution physics, and are
+therefore likely worthy of close study.  We anticipate that future work will
+follow up these events in detail.
+
+Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
+reveal the correlation between progenitor parameters and hyperparameters on a
+population level. Here we impose a cut to eliminate events with KL divergence
+larger than 0.1. This is a heuristic choice to exclude events that are obviously
+not compatible with \texttt{COSMIC}. A careful treatment of all events in the
+catalog and discussion related to the detailed physical implications of figure
 \ref{fig:GWTC-3_f_acc_mass} is beyond the scope of this paper.  We defer a
@@
 
-Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
-reveal the correlation between progenitor parameters and hyperparameters on a
-population level.
-Merger observations for which \texttt{COSMIC} struggles are likely to be very
-informative about formation channels and binary evolution physics, and are
-therefore likely worthy of close study.  We anticipate that future work will
-follow up these events in detail.}
-
 \begin{figure}
 \includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
-\caption{
-    The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for most of the events in GWTC-3.
-    Note that $m_{1,{\rm GW}}$ here refers to the reprojected posterior instead of the posterior released by LIGO.
-    Each contour is the $68\%$ credible interval of the posterior density for a particular event.
-    At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair instability supernova mechanism prevents \texttt{COSMIC} from producing events that are more massive than this cutoff.
-    Therefore, events with the majority of posterior support above this cutoff are not compatible with \texttt{COSMIC}, hence have a large KL divergence and are excluded from this figure.
-    On the low mass end, neutron star binaries or neutron star-black hole binaries are subjected to randomness induced by the natal kick,
-    also resulting in a larger and fluctuating KL divergences, and therefore are also excluded from the analysis.
-}
+\caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
+    most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
+    reprojected posterior instead of the posterior released by LIGO. Each
+    contour is the $68\%$ credible interval of the posterior density for a
+    particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
+    instability supernova mechanism prevents \texttt{COSMIC} from producing
+    events that are more massive than this cutoff. Therefore, events with the
+    majority of posterior support above this cutoff are not compatible with
+    \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
+    figure. On the low mass end, neutron star binaries or neutron star-black
+    hole binaries are subjected to randomness induced by the natal kick, also
+    resulting in a larger and fluctuating KL divergences, and therefore are also
+    excluded from the analysis. }
 \label{fig:GWTC-3_f_acc_mass}
@@
 
-% Rate analysis is now just counting, like what's done in phenomenological model, instead of forward modelling.
+% Rate analysis is now just counting, like what's done in phenomenological
+% model, instead of forward modelling.
 Working event-by-event as we do permits post-processing application of
@@
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyperparameters space for the first time, there are a number of improvements that can be implemented in future studies.
-In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
-For example, events with either of the component mass larger than the lower edge of PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
-Alternative channels such as dynamical formation might be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
-As the sensitivity of GW detectors increases, we expect to see more and more events that are unusual in some way.
-Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
-In this paper, our main focus is to illustrate the concept of "back-propagating" GW-event posterior samples,
-highlighting the capability of our method and motivating the benefits of building population synthesis codes that can work seamlessly with our method.
-To avoid cluttering of focus, we discuss the physical implications of the results presented in this work under our specific assumptions (i.e.
-using \texttt{COSMIC} as our evolutionary function) in a future companion paper.
+While our method allows data-driven exploration of the hyperparameters space for
+the first time, there are a number of improvements that can be implemented in
+future studies. In this study we use only \texttt{COSMIC} as our evolutionary
+function, which by design cannot explain all the events in GWTC-3. For example,
+event with either of the component mass larger than the lower edge of PISN gap
+such as GW190521 cannot be explained by \texttt{COSMIC}. Alternative channels
+such as dynamical formation might be needed to explain some subset of the events
+in GWTC-3 \citep{Zevin2021}. As the sensitivity of GW detectors increases, we
+expect to see more and more events that are unusual in some way. Therefore,
+having a self-consistence population synthesis code that contains multiple
+formation channels is essential to accommodate the growing catalog of GW events.
+In this paper, our main focus is to illustrate the concept of "back-propagating"
+GW events posterior samples, highlighting the capability of our method and
+motivating the community to build the next generation population synthesis code
+that can work with our method. To avoid cluttering of focus, we discuss the
+physical implication of the result presented in this work under our specific
+assumption (i.e. using \texttt{COSMIC} as our evolutionary function) in a
+companion paper.
 
 % Randomness paragraph
-Due to the implicit definition of random variables in \texttt{COSMIC}, our evolutionary function is stochastic.
-This introduces significant inefficiency in our root-finding and sampling algorithm.
-The main stochasticity in \texttt{COSMIC} comes from natal kick, which significantly affect the evolutioary pathway of low mass events such as BNS mergers.
-The effect of the natal kick is suppressed for heavier mass events due to fallback.
-This means events with lighter masses are subject to stochasticity of the function,
-where the sampling process for heavier events behaves as if the evolutionary function we use is deterministic.
-% Indeed, we see this behavior in a couple of case studies.
-% For heavy event such as GW150914, the reprojected posterior agrees with the data posterior.
-% For light event such as GW170817, although during the root-finding process the convergence criteria is met,
-% the reprojected posterior does not agree with the data posterior.
-Due to computational limitation, we only try 1000 different initial guesses per posterior sample in the root-finding process.
-This means any posterior sample that has a probability of merging rarer than 1 in 1000 could be missed.
-Obviously the problems which comes with the randomness can be alleviated by performing more tries per posterior sample,
-but this is not scalable in practice.
-On top of limitation in efficiency, some formation channels require explicit control of random variables by construction.
-For example, in a dynamical formation scenario such as binaries that form in a globular cluster,
-each binary has some probability of undergoing a multi-body encounter with another member in the cluster.
-These encounter probability distributions are either studied with direct N-body simulations or semi-analytical methods.
-In both cases, each member of the cluster is no longer completely independent of the other members, but coupled through the encounter probability distribution.
-By studying the encounter probability distribution, we can infer the properties of the environment which the binary lives in.
-This can only be done if we have explicit control over the random variables that characterize the encounter probability distribution.
+Due to the implicit definition of random variables in \texttt{COSMIC}, our
+evolutionary function is stochastic. This introduces significant inefficiency in
+our root-finding and sampling algorithm. The main stochasticity in
+\texttt{COSMIC} comes from natal kick, which significantly affect the
+evolutioary pathway of low mass events such as BNS mergers. The effect of the
+natal kick is suppressed for heavier mass events due to fallback. This means
+events with lighter masses are subject to stochasticity of the function, where
+the sampling process for heavier events behaves as if the evolutionary function
+we use is deterministic.
+% Indeed, we see this behavior in a couple of case studies. For heavy event such
+% as GW150914, the reprojected posterior agrees with the data posterior. For
+% light event such as GW170817, although during the root-finding process the
+% convergence criteria is met, the reprojected posterior does not agree with the
+% data posterior.
+Due to computational limitation, we only try 1000 different initial guesses per
+posterior sample in the root-finding process. This means any posterior sample
+that has a probability of merging rarer than 1 in 1000 could be missed.
+Obviously the problems which comes with the randomness can be alleviated by
+performing more tries per posterior sample, but this is not scalable in
+practice. On top of limitation in efficiency, some formation channels require
+explicit control of random variables by construction. For example, in a
+dynamical formation scenario such as binaries that form in a globular cluster,
+each binary has some probability of undergoing a multi-body encounter with
+another member in the cluster. These encounter probability distributions are
+either studied with direct N-body simulations or semi-analytical methods. In
+both cases, each member of the cluster is no longer completely independent of
+the other members, but coupled through the encounter probability distribution.
+By studying the encounter probability distribution, we can infer the properties
+of the environment which the binary lives in. This can only be done if we have
+explicit control over the random variables that characterize the encounter
+probability distribution.
 
@@
 
-% Gradient decending in the evolutionary parameter space is much more efficent than rejection sampling.
-% Potential pitfall of interpreting this result.
-% Need for full autodiff.
+% Gradient decending in the evolutionary parameter space is much more efficent
+% than rejection sampling. Potential pitfall of interpreting this result. Need
+% for full autodiff.
 
-% By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
-% In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
-% In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
-% This greatly benefit the efficiency of the root-finding process.
-% On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
-% One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
-% Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
-% the relatively weighting of these roots becomes ambiguous when the number of roots is large.
-% For example, should we consider roots that are close to each other a single root or multiple roots?
-% If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
-% Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
-% This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
-% In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.
+% By leveraging gradient-based optimization strategy, obtaining posterior
+% samples in the evolutionary parameter space is much more efficient than
+% rejection sampling. In each optimization step during the root-finding process,
+% the current state of the guess and its gradient provide useful information
+% about where the root might be. In contrast to sampling algorithm, which
+% discard the sample whenever it is not accepted. This greatly benefit the
+% efficiency of the root-finding process. On the other hand, this method is not
+% a sampling algorithm, but a coordinate transform of the posterior samples. One
+% should pay extra caution in interpreting the posterior samples in the
+% evolutionary parameter space. Since posterior samples in the observable space
+% can in principle have multiple roots in the evolutionary parameter space, the
+% relatively weighting of these roots becomes ambiguous when the number of roots
+% is large. For example, should we consider roots that are close to each other a
+% single root or multiple roots? If we have two clusters of roots, the relative
+% number of roots within each cluster would determine the relative weight of
+% each mode. Note that since they both map to the same observable posterior
+% sample within the acceptance threshold, all weighting assigned are equally
+% valid from a "matching data" standpoint. This leave the interpretation of the
+% posterior in the evolutionary parameter space ambiguous. In our case, we found
+% empirically each posterior sample may at most only correspond to a handful of
+% roots (most of the posterior samples have a unique root.), therefore we choose
+% to assign equal weight to each root.
 
-% The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
-% We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
-% We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
-% This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
-% We leave a detail comparison between this method and direct sampling to future studies.
+% The potential reason for the good agreement between MCMC and the root finding
+% is due to the behavior of posterior in the evolutionary parameter space are
+% larger determined by the posterior distribution in the observable space
+% instead of multiple possible roots per posterior sample. We have also compared
+% two MCMC results, one is initialized using the roots and another is
+% initialized using samples drawn from a uniform prior. We found empirically the
+% MCMC result using the roots as initialization converges at least an order of
+% magnitude faster than the uniform prior initialization. This makes sense since
+% the roots basically eliminate the long burn-in needed to go from random
+% location in the evolutionary parameters to the most probable region. We leave
+% a detail comparison between this method and direct sampling to future studies.
 
-Another technical note is that we use finite differencing to estimate the gradient of the objective function, which could be a significant source of error near transition points in the evolutionary parameter space.
-Also, finite differencing is increasingly inefficient as we increase the dimensionality of the problem.
-To improve the accuracy and efficiency in estimating the gradient of the objective function, automatic differentiation is a promising feature that modelers should consider incorporating in their population synthesis codes in the future.
+Another technical note is that we use finite differencing to estimate the
+gradient of the objective function, which could be a significant source of error
+near transition points in the evolutionary parameter space. Also, finite
+differencing is increasingly inefficient as we increase the dimensionality of
+the problem. To improve the accuracy and efficiency in estimating the gradient
+of the objective function, automatic differentiation is a promising feature that
+modelers should consider incorporating in their population synthesis codes in
+the future.
 
-To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
-We point out hyperparameters in the usual population synthesis simulation context are not actually parameters related to the population,
-but parameters about the evolutionary function.
-This means the binary evolution functions can be constrained on an individual event basis.
-We "back propagate" the posterior in the observable space to the evolutionary parameter space,
-thus allowing us to study hyperparameters and theirs correlations with progenitor parameters in a data-driven manner.
-Our method makes less assumptions than the traditional forward modelling approach,
-which often fix the hyperparameters across the entire population.
-Since we are not limited to the fixed hyperparameters assumptions, we can explore the behavior of the hyperparameters across the population much more efficiently.
-While our work lays down a data analysis pathway to understand the population of GW events,
-no physics can be learned without a comprehensive physical model.
-We hope this letter will motivate the construction of next-generation population synthesis codes that have the following properties:
-first, they should retain explicit control over the random variables so marginalizing over random variables can be done precisely and
-second, they should be as automatically differentiable as possible so that exploring the evolutionary parameter space is efficient.
-By combining the methods presented in this work and future, differentiable population synthesis codes,
-we can explore the full parameter space of binary evolution models with future GW data.
+To summarize, we propose a novel method to recover the posterior samples in the
+evolutionary parameter space for each GW event. We point out hyperparameters in
+the usual population synthesis simulation context are not actually parameters
+related to the population, but parameters about the evolutionary function. This
+means the binary evolution functions can be constrained on an individual event
+basis. We "back propagate" the posterior in the observable space to the
+evolutionary parameter space, thus allowing us to study hyperparameters and
+theirs correlations with progenitor parameters in a data-driven manner. Our
+method makes less assumptions than the traditional forward modelling approach,
+which often fix the hyperparameters across the entire population. Since we are
+not limited to the fixed hyperparameters assumptions, we can explore the
+behavior of the hyperparameters across the population much more efficiently.
+While our work lays down a data analysis pathway to understand the population of
+GW events, no physics can be learned without a comprehensive physical model. We
+hope this letter will motivate the community to build the next generation
+population synthesis codes that have the following properties: first, they
+should have explicit control over the random variables so marginalizing over
+random variables can be done more precisely; second, they should be as
+automatically differentiable as much as possible so exploring the evolutionary
+parameter space is efficient. Combining this work and the next generation of
+population synthesis codes, we can explore the full parameter space of binary
+evolution models with the next-generation GW detectors network in the near
+future.
 
 \section{Acknowledgement}
-The authors are grateful for discussions with the CCA Gravitational Wave Astronomy Group.
-The Flatiron Institute is supported by the Simons Foundation.  We thank M. Zevin for helpful comments and suggestions on this manuscript.
-This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org),
-a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
-LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as
-the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS),
-and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
-Additional support for Advanced LIGO was provided by the Australian Research Council. Virgo is funded, through the European Gravitational Observatory (EGO),
-by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef,
-with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
-The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT),
-and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea,
-Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
+The authors are grateful for discussions with the CCA Gravitational Wave
+Astronomy Group. The Flatiron Institute is supported by the Simons Foundation.
 
-\software{
-    \texttt{corner}\ \citep{corner};
-    \texttt{COSMIC}\ \citep{Breivik2020};
-    \texttt{julia}\ \citep{Julia-2017}
-    \texttt{matplotlib}\ \citep{matplotlib};
-    \texttt{numpy}\ \citep{numpy};
-    \texttt{pandas}\ \citep{McKinneyPandas, reback2020pandas};
-    \texttt{scipy}\ \citep{scipy}
-    \texttt{seaborn}\ \citep{Waskom2021}
-    \texttt{showyourwork}\ \citep{Luger2021}
-}
+\software{ \texttt{corner}\ \citep{corner}; \texttt{COSMIC}\
+    \citep{Breivik2020}; \texttt{julia}\ \citep{Julia-2017} \texttt{matplotlib}\
+    \citep{matplotlib}; \texttt{numpy}\ \citep{numpy}; \texttt{pandas}\
+    \citep{McKinneyPandas, reback2020pandas}; \texttt{scipy}\ \citep{scipy}
+    \texttt{seaborn}\ \citep{Waskom2021} \texttt{showyourwork}\
+    \citep{Luger2021} }
 

### 2022-10-31 src/tex/ms.tex

@@
 of gravitational wave observations.  We apply our pipeline to observations in
-the third LIGO Virgo Kagra Gravitational Wave Transient Catalog. We showcase the
+the third LIGO--Virgo Gravitational-Wave Transient Catalog. We showcase the
 effectiveness of this pipeline with a detailed study of the progenitor
@@
 
-We apply the method described in section \ref{sec:method} to the first GW event
-GW150914. The parameters involved in the analysis and the range we allow for the
-inference are tabulated in table \ref{tab:parameters}. The spin of BHs formed in
-isolated binaries depends strongly on the angular momentum transport in massive
-stars which is still highly uncertain (see however \citet{Fuller2019,
-Bavera2020}), therefore we only consider the two component masses in the
-observable space in both the root-finding stage and the MCMC stage. For
-progenitor parameters, we characterize each guess with five parameters: the two
-component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period
-$t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that
-the progenitor formation redshift is not obtained through the root-finding
-process because the formation redshift does not intrinsically affect the
-evolution of the binary. Instead, we compute the time it takes for the binary to
-evolve from formation to the merger and add this time to the lookback time when
-a GW event is observed and limit the total time to merger to be less than
-$13.7\,\rm{Gyr}$. Using the total lookback time, we can compute the redshift of
-the binary at ZAMS formation following the Planck 2018 cosmological model
-\citep{Planck2018}. The main benefit of post-processing the redshift in this way
-is that we do not rely on a particular assumption of star formation rate (SFR)
-distribution when we solve for the posterior distribution in redshift. This
-means we can put the prior on formation redshift (or in general,
-metallicity-redshfit distribution) in by reweighting the posterior distribution
-according to a particular SFR model, and therefore do not have to rerun the
-inference every time we change the SFR model. This is especially helpful since
-SFR models are still highly uncertain but strongly impact the local merger rates
-of BBHs \citep{Broekgaarden2021}. For hyperparameters, we choose parameters that
-affect binary evolution in \texttt{COSMIC} most significantly for massive stars,
-including the common envelope efficiency $\alpha$, the fraction of mass accreted
-onto the accretor during stable mass transfer $f_{\rm acc}$, the critical mass
-ratio which determines whether Roche-overflow mass transfer proceeds stably or
-produces a common envelope when the donor star is on the Hertzsprung gap $q_{\rm
-crit,3}$, and the root-mean-square of the Maxwellian distribution employed for
-CO natal kicks, $\sigma$.
+We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
+We use the \texttt{Overall\_posterior} posterior samples for GW150914,
+publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
+The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
+The spin of BHs formed in isolated binaries depends strongly on the angular
+momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
+therefore we only consider the two component masses in the observable space in both the
+root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with
+five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
+the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
+obtained through the root-finding process because the formation redshift does not intrinsically affect
+the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
+formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
+time to merger to be less than $13.7\,\rm{Gyr}$.
+Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
+Planck 2018 cosmological model \citep{Planck2018}.
+The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
+of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
+This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
+by reweighting the posterior distribution according to a particular SFR model,
+and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
+since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
+For hyperparameters, we choose parameters that strongly affect
+binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
+efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
+the critical mass ratio which determines whether Roche-overflow mass transfer
+proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
+$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
+employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
+with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
+stars merge since the donor has likely not formed a strong core-envelope boundary 
+\citep{Ivanova2004, Belczynski2008}.
 
@@
 To illustrate the potential benefit of using our method on the population level,
-we perform the same analysis for all events in GWTC-3. Plotted in figure
-\ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
-GW}}-f_{\rm acc}$ space for most of the events in GWTC-3. Each contour
-represents the $68\%$ credible interval of the posterior density for that
-particular event. There is a suggestive trend showing that $f_{\rm acc}$ could
-increase as the mass of the progenitor increases. Such a trend implies the
-stable mass transfer phase of a less massive binary would be preferentially
-non-conservative, with more conservative mass transfer in more massive binary
-systems.
+we perform the same analysis for all events in GWTC-3. For events announced in
+GWTC-1, we use \texttt{Overall\_posterior},
+\texttt{PublicationSamples},\texttt{IMRPhenomXPHM\_comoving} and
+\texttt{C01:Mixed} posterior samples, respectively. All the data are publicly
+available from the Gravitational wave Open Science Center
+\citep{LIGOScientific:2019lzm}. Plotted in figure \ref{fig:GWTC-3_f_acc_mass} is
+the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
+most\footnote{\textbf{We excluded events for which our reconstructed posterior has KL
+divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}}
+of the events in GWTC-3. Each contour represents the $68\%$ credible interval of
+the posterior density for that particular event. There is a suggestive trend
+showing that $f_{\rm acc}$ could increase as the mass of the progenitor
+increases. Such a trend implies the stable mass transfer phase of a less massive
+binary would be preferentially non-conservative, with more conservative mass
+transfer in more massive binary systems.
 
@@
 pulled the GW event posterior back from the observable space to the evolutionary
-parameter space, we can apply the same hierarchical inference methods currently
-used to infer the compact binary population to our progenitor population.  In
-particular, we can use our progenitor population to infer the formation rate of
-compact binary population progenitor systems over cosmic time without needing to
-make artifical and simplistic assumptions about the delay time distribution or
-the metallicity-specific star formation rate
-\citep{Vitale2019,Ng2021,VanSon2022}.  Comparing the formation rate of
-progenitors to the star formation rate provides an avenue to check our
-understanding of binary evolution and the relative rate of other formation
-channels.
+parameter space, \textbf{and have accounted for selection effects}, we can apply
+the same hierarchical inference methods currently used to infer the compact
+binary population to our progenitor population.  In particular, we can use our
+progenitor population to infer the formation rate of compact binary population
+progenitor systems over cosmic time without needing to make artifical and
+simplistic assumptions about the delay time distribution or the
+metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
+Comparing the formation rate of progenitors to the star formation rate provides
+an avenue to check our understanding of binary evolution and the relative rate
+of other formation channels.
 
 % Single formation channel and fundamental limitation on that.
-While our method allows data-driven exploration of the hyperparameters space for
-the first time, there are a number of improvements that can be implemented in
-future studies. In this study we use only \texttt{COSMIC} as our evolutionary
-function, which by design cannot explain all the events in GWTC-3. For example,
-event with either of the component mass larger than the lower edge of PISN gap
-such as GW190521 cannot be explained by \texttt{COSMIC}. Alternative channels
-such as dynamical formation might be needed to explain some subset of the events
-in GWTC-3 \citep{Zevin2021}. As the sensitivity of GW detectors increases, we
-expect to see more and more events that are unusual in some way. Therefore,
-having a self-consistence population synthesis code that contains multiple
-formation channels is essential to accommodate the growing catalog of GW events.
-In this paper, our main focus is to illustrate the concept of "back-propagating"
-GW events posterior samples, highlighting the capability of our method and
-motivating the community to build the next generation population synthesis code
-that can work with our method. To avoid cluttering of focus, we discuss the
-physical implication of the result presented in this work under our specific
-assumption (i.e. using \texttt{COSMIC} as our evolutionary function) in a
-companion paper.
+While our method allows data-driven exploration of the hyperparameters space, there are a number of improvements that can be implemented in future studies.
+In this study we use only \texttt{COSMIC} as our evolutionary function, which by design cannot explain all the events in GWTC-3.
+For example, events with either of the component mass larger than the lower edge of PISN gap such as GW190521 cannot be explained by \texttt{COSMIC}.
+Alternative channels such as dynamical formation might be needed to explain some subset of the events in GWTC-3 \citep{Zevin2021}.
+As the sensitivity of GW detectors increases, we expect to see more and more events that are unusual in some way.
+Therefore, having a self-consistence population synthesis code that contains multiple formation channels is essential to accommodate the growing catalog of GW events.
+In this paper, our main focus is to illustrate the concept of "back-propagating" GW-event posterior samples,
+highlighting the capability of our method and motivating the benefits of building population synthesis codes that can work seamlessly with our method.
+To avoid cluttering of focus, we discuss the physical implications of the results presented in this work under our specific assumptions (i.e.
+using \texttt{COSMIC} as our evolutionary function) in a future companion paper.
 
@@
 
-To summarize, we propose a novel method to recover the posterior samples in the
-evolutionary parameter space for each GW event. We point out hyperparameters in
-the usual population synthesis simulation context are not actually parameters
-related to the population, but parameters about the evolutionary function. This
-means the binary evolution functions can be constrained on an individual event
-basis. We "back propagate" the posterior in the observable space to the
-evolutionary parameter space, thus allowing us to study hyperparameters and
-theirs correlations with progenitor parameters in a data-driven manner. Our
-method makes less assumptions than the traditional forward modelling approach,
-which often fix the hyperparameters across the entire population. Since we are
-not limited to the fixed hyperparameters assumptions, we can explore the
-behavior of the hyperparameters across the population much more efficiently.
-While our work lays down a data analysis pathway to understand the population of
-GW events, no physics can be learned without a comprehensive physical model. We
-hope this letter will motivate the community to build the next generation
-population synthesis codes that have the following properties: first, they
-should have explicit control over the random variables so marginalizing over
-random variables can be done more precisely; second, they should be as
-automatically differentiable as much as possible so exploring the evolutionary
-parameter space is efficient. Combining this work and the next generation of
-population synthesis codes, we can explore the full parameter space of binary
-evolution models with the next-generation GW detectors network in the near
-future.
+To summarize, we propose a method to recover the posterior samples in the evolutionary parameter space for each GW event.
+We point out hyperparameters in the usual population synthesis simulation context are not actually parameters related to the population,
+but parameters about the evolutionary function.
+This means the binary evolution functions can be constrained on an individual event basis.
+We "back propagate" the posterior in the observable space to the evolutionary parameter space,
+thus allowing us to study hyperparameters and theirs correlations with progenitor parameters in a data-driven manner.
+Our method makes less assumptions than the traditional forward modelling approach,
+which often fix the hyperparameters across the entire population.
+Since we are not limited to the fixed hyperparameters assumptions, we can explore the behavior of the hyperparameters across the population much more efficiently.
+While our work lays down a data analysis pathway to understand the population of GW events,
+no physics can be learned without a comprehensive physical model.
+We hope this letter will motivate the construction of next-generation population synthesis codes that have the following properties:
+first, they should retain explicit control over the random variables so marginalizing over random variables can be done precisely and
+second, they should be as automatically differentiable as possible so that exploring the evolutionary parameter space is efficient.
+By combining the methods presented in this work and future, differentiable population synthesis codes,
+we can explore the full parameter space of binary evolution models with future GW data.
 
 \section{Acknowledgement}
-The authors are grateful for discussions with the CCA Gravitational Wave
-Astronomy Group. The Flatiron Institute is supported by the Simons Foundation.
+The authors are grateful for discussions with the CCA Gravitational Wave Astronomy Group.
+The Flatiron Institute is supported by the Simons Foundation.  We thank M. Zevin for helpful comments and suggestions on this manuscript.
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org),
+a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as
+the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS),
+and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council. Virgo is funded, through the European Gravitational Observatory (EGO),
+by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef,
+with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT),
+and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea,
+Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
+
+Data produced in the process of this work can be found at \cite{wong2022-data}.
 

### 2022-11-12 src/tex/ms.tex
Add kde reference

@@
 \end{align}
-We apply a kernel density estimator with a Gaussian kernel to the posterior
+We apply a kernel density estimator with a Gaussian kernel \footnote{We use the
+implementation in \texttt{scipy} \cite{2020SciPy-NMeth}} to the posterior
 samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.

### 2022-11-26 src/tex/ms.tex
Add validation sentence to line 614

@@
 
-Note that some events did not pass the KL divergence test we proposed in section
-\ref{sec:method}; we do not include these systems in figure
-\ref{fig:GWTC-3_f_acc_mass}. Binaries that form low-mass compact objects have
-lower amounts of fallback and tend to have correspondingly larger variance in
-their random natal kicks. Larger kicks can unbind the progenitor binary, or lead
-to wide binaries that do not merge within a Hubble time.  In these cases, the
-extra variance during the reprojection can produce a posterior that may not
-agree with the original posterior, hence yielding a higher KL divergence.
+\textbf{We have validated the robustness of every events shown in
+\ref{fig:GWTC-3_f_acc_mass} by reprojecting the posterior density in the
+evolutionary parameter space to the observable space.} Note that some events did
+not pass the KL divergence test we proposed in section \ref{sec:method}; we do
+not include these systems in figure \ref{fig:GWTC-3_f_acc_mass}. Binaries that
+form low-mass compact objects have lower amounts of fallback and tend to have
+correspondingly larger variance in their random natal kicks. Larger kicks can
+unbind the progenitor binary, or lead to wide binaries that do not merge within
+a Hubble time.  In these cases, the extra variance during the reprojection can
+produce a posterior that may not agree with the original posterior, hence
+yielding a higher KL divergence.
 

### 2022-11-26 src/tex/ms.tex
Add prior sentence
Add rate clarification

@@
 samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
+\textbf{A desired prior in the evolutionary parameters can be included through
+the prior $\pi(\bm{\Theta})$ in the above equation.}
 
@@
 metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
-Comparing the formation rate of progenitors to the star formation rate provides
-an avenue to check our understanding of binary evolution and the relative rate
-of other formation channels.
+\textbf{In other word, just as we measure the event rate of GW events by
+counting, we can measure the event rate of the progenitor systems of GW events
+in the same way.} Comparing the formation rate of progenitors to the star
+formation rate provides an avenue to check our understanding of binary evolution
+and the relative rate of other formation channels.
 

### 2022-11-29 src/tex/ms.tex
Add symbols to table one

@@
     \hline
-    Observables &\ &\  \\
+    Observables $\bm{\theta}$ &\ &\  \\
     \hline
@@
     \hline
-    Progenitor parameters &\ &\  \\
+    Progenitor parameters $\bm{\theta'}$&\ &\  \\
     \hline
@@
     \hline
-    hyperparameters &\ &\ \\
+    Hyperparameters $\bm{\lambda    }$ &\ &\ \\
     \hline
@@
 momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space in both the
-root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with
+therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
+root-finding stage and the MCMC stage. For progenitor parameters $\bm{\theta'}$, we characterize each guess with
 five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
@@
 since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
-For hyperparameters, we choose parameters that strongly affect
+For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
 binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
@@
 
-\begin{figure*}[h]
-    \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
-    \caption{The posterior for GW150914 in both the progenitor parameter space
-    and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
-    $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
-    is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
-    accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
-    ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
-    $95\%$ confidence intervals. Note that the redshift is not fitted during the
-    root finding process or the MCMC process. Once we find the evolutionary
-    parameters, we add the delay time to the lookback time of the observed
-    posterior sample, then from the total lookback time we can compute the
-    redshift at ZAMS. We highlight three panels in the corner plots to show the
-    fine structure of the set of posterior samples in the evolutionary parameter
-    space. We also color the posterior in a particular panel according to the
-    type of parameters involved in the corner plot. Blue denotes panels that
-    include only progenitor parameters, green denotes panels that include a mix
-    of progenitor parameters and hyperparameters, and orange denotes panels that
-    include only hyperparameters. }
-    \label{fig:GW150914_posterior}
-    \script{figure1_cornerPlot.py}
-\end{figure*}
+% \begin{figure*}[h]
+%     \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
+%     \caption{The posterior for GW150914 in both the progenitor parameter space
+%     and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
+%     $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
+%     is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
+%     accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
+%     ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
+%     $95\%$ confidence intervals. Note that the redshift is not fitted during the
+%     root finding process or the MCMC process. Once we find the evolutionary
+%     parameters, we add the delay time to the lookback time of the observed
+%     posterior sample, then from the total lookback time we can compute the
+%     redshift at ZAMS. We highlight three panels in the corner plots to show the
+%     fine structure of the set of posterior samples in the evolutionary parameter
+%     space. We also color the posterior in a particular panel according to the
+%     type of parameters involved in the corner plot. Blue denotes panels that
+%     include only progenitor parameters, green denotes panels that include a mix
+%     of progenitor parameters and hyperparameters, and orange denotes panels that
+%     include only hyperparameters. }
+%     \label{fig:GW150914_posterior}
+%     \script{figure1_cornerPlot.py}
+% \end{figure*}
 
@@
 
-\begin{figure*}
-    \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
-    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
-     mergers for three variations of $\alpha$ with a fixed set of parameter
-     assumptions matching those of A21 (first three columns) and for binaries
-     which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
-     $q_{crit,3}$ are allowed to vary, (fourth column).}
-    \label{fig:compare_fixed_variable}
-    \script{figure2_varyAlphaFacc.py}
-\end{figure*}
+% \begin{figure*}
+%     \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
+%     \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
+%      mergers for three variations of $\alpha$ with a fixed set of parameter
+%      assumptions matching those of A21 (first three columns) and for binaries
+%      which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
+%      $q_{crit,3}$ are allowed to vary, (fourth column).}
+%     \label{fig:compare_fixed_variable}
+%     \script{figure2_varyAlphaFacc.py}
+% \end{figure*}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
-\caption{Reprojecting the posterior in evolutionary parameter space of GW150914
-to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
-intervals. The blue contour is the reprojected posterior after the root-finding
-procedure. The orange contour is the reprojected posterior after the MCMC
-procedure. The green contour is the posterior plotted using the original LVK
-posterior samples. }
-\label{fig:GW150914_reprojection}
-\script{figure3_sanityCheck.py}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
+% \caption{Reprojecting the posterior in evolutionary parameter space of GW150914
+% to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
+% intervals. The blue contour is the reprojected posterior after the root-finding
+% procedure. The orange contour is the reprojected posterior after the MCMC
+% procedure. The green contour is the posterior plotted using the original LVK
+% posterior samples. }
+% \label{fig:GW150914_reprojection}
+% \script{figure3_sanityCheck.py}
+% \end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
-\caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
-    most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
-    reprojected posterior instead of the posterior released by LIGO. Each
-    contour is the $68\%$ credible interval of the posterior density for a
-    particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
-    instability supernova mechanism prevents \texttt{COSMIC} from producing
-    events that are more massive than this cutoff. Therefore, events with the
-    majority of posterior support above this cutoff are not compatible with
-    \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
-    figure. On the low mass end, neutron star binaries or neutron star-black
-    hole binaries are subjected to randomness induced by the natal kick, also
-    resulting in a larger and fluctuating KL divergences, and therefore are also
-    excluded from the analysis. }
-\label{fig:GWTC-3_f_acc_mass}
-\script{figure4_population.py}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
+% \caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
+%     most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
+%     reprojected posterior instead of the posterior released by LIGO. Each
+%     contour is the $68\%$ credible interval of the posterior density for a
+%     particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
+%     instability supernova mechanism prevents \texttt{COSMIC} from producing
+%     events that are more massive than this cutoff. Therefore, events with the
+%     majority of posterior support above this cutoff are not compatible with
+%     \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
+%     figure. On the low mass end, neutron star binaries or neutron star-black
+%     hole binaries are subjected to randomness induced by the natal kick, also
+%     resulting in a larger and fluctuating KL divergences, and therefore are also
+%     excluded from the analysis. }
+% \label{fig:GWTC-3_f_acc_mass}
+% \script{figure4_population.py}
+% \end{figure}
 

### 2022-11-29 src/tex/ms.tex
Tone down figure 4 discussion.

@@
 observed GW150914.  Our pipeline permits a measurement of the variation of
-population synthesis parameter settings with binary properties, if any; we
-present inferences for the recent GWTC-3 transient catalog that suggest that the
-stable mass transfer efficiency parameter may vary with primary black hole mass.  
+population synthesis parameter settings with binary properties, if any;
+\textbf{we illustrate the possibility of such capability by presenting
+inferences for the recent GWTC-3 transient catalog that suggest that the stable
+mass transfer efficiency parameter may vary with primary black hole mass.}  
 \end{abstract}
@@
 assumptions, both in the simulations and model parameterizations fitted to the
-GW data, which are unlikely to be fully correct. For example, there is no reason
-to suppose that the entire population of DCOs should evolve with the same fixed
-hyperparameters; there is no a priori reason the fitted parameterized model
-should capture the relevant features of the observed DCO distributions to
-correspond to the binary physics of interest; etc. A better approach would be to
-compare progenitor parameters and hyperparameter settings directly to DCO
-observations, treating the population synthesis model as a mapping from
-progenitor to merger parameters; merger parameters can then be mapped into
-observations using a gravitational waveform family, and parameter inference
-proceeds in the usual way, propagating information back up the mapping chain
-\citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
-hyperparameter settings, was advanced in \citet{Andrews2018,Andrews2021}, as we
-discuss in more detail below.
+GW data, which are unlikely to be fully correct. For example, \textbf{without
+knowing whether the parameterization of hyperparameters is sufficiently
+accurate}, there is no reason to suppose that the entire population of DCOs
+should evolve with the same fixed hyperparameters; there is no a priori reason
+the fitted parameterized model should capture the relevant features of the
+observed DCO distributions to correspond to the binary physics of interest; etc.
+A better approach would be to compare progenitor parameters and hyperparameter
+settings directly to DCO observations, treating the population synthesis model
+as a mapping from progenitor to merger parameters; merger parameters can then be
+mapped into observations using a gravitational waveform family, and parameter
+inference proceeds in the usual way, propagating information back up the mapping
+chain \citep[e.g.][]{Veitch2015}.  A similar methodology, though with
+\emph{fixed} hyperparameter settings, was advanced in
+\citet{Andrews2018,Andrews2021}, as we discuss in more detail below.
 
@@
 distribution of hyperparameters; this avoids making overly-specific assumptions
-like fixing the hyperparameters for all types of event.
+like fixing the hyperparameters for all types of event. \textbf{In this work we
+focus on illustrating the possibililty of conducting such analysis with our
+pipeline. Note that there are concerns that need to be addressed before one can
+make statments about the physics, such as the number of channels considered and
+specific choices of parameterization. These concerns are beyond the scope of
+this paper, and we defer a detailed study of the physics related to the
+population of GW events to future work.}
 

### 2022-12-19 src/tex/ms.tex

@@
 
-\section{Method}
+\section{Methods}
 \label{sec:method}
 
+\subsection{\textbf{Statistical method}}
+\label{subsec:stats}
 We define the progenitor's zero age main sequence (ZAMS) properties like mass,
@@
 
-\section{Results}
-\label{sec:result}
+\subsection{\textbf{Binary evolution assumptions}}
+\label{subsec:binaries}
 
@@
 
-We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
+We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
 We use the \texttt{Overall\_posterior} posterior samples for GW150914,
@@
 therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
-root-finding stage and the MCMC stage. For progenitor parameters $\bm{\theta'}$, we characterize each guess with
+root-finding stage and the MCMC stage.
+
+For progenitor parameters $\bm{\theta'}$, we characterize each guess with
 five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
@@
 since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
+
 For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
@@
 $q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
+employed for CO natal kicks, $\sigma$. \textbf{We follow the critical mass ratio prescription 
+described in \cite{Neijssel2019} for donors which fill their Roche lobes during
+ other phases of stellar evolution.} 
+For systems which begin a common envelope 
 with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
 stars merge since the donor has likely not formed a strong core-envelope boundary 
-\citep{Ivanova2004, Belczynski2008}.
+\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
+evolution proceed. In particular, we assume that stars lose mass through stellar winds 
+in a metallicity-dependent 
+fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
+during both stable Roche-overflow mass transfer and common envelopes.
+We further assume that the CO formation mechanism follows the delayed 
+prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
+occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
+and that natal kicks due to supernovae are reduced by the amount 
+of mass which falls back onto the proto-CO
+as described in \citet{Fryer2012}.}
+
+\section{Results}
+\label{sec:result}
 
@@
 for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
-unexpected since we use the \citet{Fryer2012} delayed model which reduces the
+unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
 natal kick strength at BH formation based on the amount of fallback onto the

### 2022-12-19 src/tex/ms.tex
Unlock figures and fix simon stevenson's citation

--- 2022-11-29 src/tex/ms.tex
@@
 
-% \begin{figure*}[h]
-%     \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
-%     \caption{The posterior for GW150914 in both the progenitor parameter space
-%     and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
-%     $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
-%     is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
-%     accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
-%     ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
-%     $95\%$ confidence intervals. Note that the redshift is not fitted during the
-%     root finding process or the MCMC process. Once we find the evolutionary
-%     parameters, we add the delay time to the lookback time of the observed
-%     posterior sample, then from the total lookback time we can compute the
-%     redshift at ZAMS. We highlight three panels in the corner plots to show the
-%     fine structure of the set of posterior samples in the evolutionary parameter
-%     space. We also color the posterior in a particular panel according to the
-%     type of parameters involved in the corner plot. Blue denotes panels that
-%     include only progenitor parameters, green denotes panels that include a mix
-%     of progenitor parameters and hyperparameters, and orange denotes panels that
-%     include only hyperparameters. }
-%     \label{fig:GW150914_posterior}
-%     \script{figure1_cornerPlot.py}
-% \end{figure*}
+\begin{figure*}[h]
+    \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
+    \caption{The posterior for GW150914 in both the progenitor parameter space
+    and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
+    $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
+    is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
+    accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
+    ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
+    $95\%$ confidence intervals. Note that the redshift is not fitted during the
+    root finding process or the MCMC process. Once we find the evolutionary
+    parameters, we add the delay time to the lookback time of the observed
+    posterior sample, then from the total lookback time we can compute the
+    redshift at ZAMS. We highlight three panels in the corner plots to show the
+    fine structure of the set of posterior samples in the evolutionary parameter
+    space. We also color the posterior in a particular panel according to the
+    type of parameters involved in the corner plot. Blue denotes panels that
+    include only progenitor parameters, green denotes panels that include a mix
+    of progenitor parameters and hyperparameters, and orange denotes panels that
+    include only hyperparameters. }
+    \label{fig:GW150914_posterior}
+    \script{figure1_cornerPlot.py}
+\end{figure*}
 
@@
 
-% \begin{figure*}
-%     \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
-%     \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
-%      mergers for three variations of $\alpha$ with a fixed set of parameter
-%      assumptions matching those of A21 (first three columns) and for binaries
-%      which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
-%      $q_{crit,3}$ are allowed to vary, (fourth column).}
-%     \label{fig:compare_fixed_variable}
-%     \script{figure2_varyAlphaFacc.py}
-% \end{figure*}
+\begin{figure*}
+    \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
+    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
+     mergers for three variations of $\alpha$ with a fixed set of parameter
+     assumptions matching those of A21 (first three columns) and for binaries
+     which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
+     $q_{crit,3}$ are allowed to vary, (fourth column).}
+    \label{fig:compare_fixed_variable}
+    \script{figure2_varyAlphaFacc.py}
+\end{figure*}
 
@@
 
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
-% \caption{Reprojecting the posterior in evolutionary parameter space of GW150914
-% to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
-% intervals. The blue contour is the reprojected posterior after the root-finding
-% procedure. The orange contour is the reprojected posterior after the MCMC
-% procedure. The green contour is the posterior plotted using the original LVK
-% posterior samples. }
-% \label{fig:GW150914_reprojection}
-% \script{figure3_sanityCheck.py}
-% \end{figure}
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
+\caption{Reprojecting the posterior in evolutionary parameter space of GW150914
+to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
+intervals. The blue contour is the reprojected posterior after the root-finding
+procedure. The orange contour is the reprojected posterior after the MCMC
+procedure. The green contour is the posterior plotted using the original LVK
+posterior samples. }
+\label{fig:GW150914_reprojection}
+\script{figure3_sanityCheck.py}
+\end{figure}
 
@@
 
-% \begin{figure}
-% \includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
-% \caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
-%     most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
-%     reprojected posterior instead of the posterior released by LIGO. Each
-%     contour is the $68\%$ credible interval of the posterior density for a
-%     particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
-%     instability supernova mechanism prevents \texttt{COSMIC} from producing
-%     events that are more massive than this cutoff. Therefore, events with the
-%     majority of posterior support above this cutoff are not compatible with
-%     \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
-%     figure. On the low mass end, neutron star binaries or neutron star-black
-%     hole binaries are subjected to randomness induced by the natal kick, also
-%     resulting in a larger and fluctuating KL divergences, and therefore are also
-%     excluded from the analysis. }
-% \label{fig:GWTC-3_f_acc_mass}
-% \script{figure4_population.py}
-% \end{figure}
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
+\caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
+    most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
+    reprojected posterior instead of the posterior released by LIGO. Each
+    contour is the $68\%$ credible interval of the posterior density for a
+    particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
+    instability supernova mechanism prevents \texttt{COSMIC} from producing
+    events that are more massive than this cutoff. Therefore, events with the
+    majority of posterior support above this cutoff are not compatible with
+    \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
+    figure. On the low mass end, neutron star binaries or neutron star-black
+    hole binaries are subjected to randomness induced by the natal kick, also
+    resulting in a larger and fluctuating KL divergences, and therefore are also
+    excluded from the analysis. }
+\label{fig:GWTC-3_f_acc_mass}
+\script{figure4_population.py}
+\end{figure}
 

### 2022-12-19 src/tex/ms.tex
Add linenumbers to document class
update

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[linenumbers,twocolumn]{aastex631}
 
@@
 this is GW events with at least one component with mass above the pair
-instability supernova mass limit \citep{Woosley2017, Farmer2020}. The
+instability supernova mass limit \citep{Woosley2017, Farmer2019}. The
 reprojection could also be subject to stochasticity in the evolution of each

### 2023-01-20 src/tex/ms.tex
add in discussion of features in the ZAMS mass figure

@@
 $60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
-$150\,M_{\odot}$ which is the limit imposed by our assumptions. We explore the
+$150\,M_{\odot}$ which is the limit imposed by our assumptions.
+\textbf{The features in the scatter points of the ZAMS masses, most obvious in the region where $M_1,\rm{ZAMS}\sim 140-150\,M_\odot$, are the outcome of the interplay between the progenitor parameters and hyperparameters.
+The dominant cause of regions with fewer data points is that fewer binary progenitors produce BBH mergers with chirp masses and mass ratios consistent with GW150914's priors.
+This is most clearly illustrated for progenitors with primary masses between $142-144\,M_\odot$ and secondary masses below $80\,M_[\odot]$.} We explore the
 formation scenarios which lead to successful GW150914-like mergers and those
-which fail to produce GW150914-like mergers below in
+which fail to produce GW150914-like mergers \textbf{in more detail} below in
 Figure~\ref{fig:compare_fixed_variable}.

### 2023-02-10 src/tex/ms.tex
Add sentence related to previous citation

--- 2022-12-19 src/tex/ms.tex
@@
 We find that the progenitor properties and even the ability to produce a
-GW150914-like merger event is strongly correlated with hyperparameter settings;
+GW150914-like merger event is strongly correlated with hyperparameter settings,
+consistent with results from prior forward modeling studies cited above;
 different assumptions about the black hole masses at merger in GW150914 imply

### 2023-02-10 src/tex/ms.tex
Rearrange paragraphs for case-specific configuration

@@
 setting by producing a posterior over progenitor properties and \texttt{COSMIC}
-hyperparameters implied by the observation of GW150914.
+hyperparameters implied by the observation of
+dependencies:
+  src/scripts/figure1_cornerPlot.py:
+    - src/data/GWTC3_LVK_posterior.npz
+    - src/data/GWTC3_mcmc_processed.hdf5
+
+  src/scripts/figure2_varyAlphaFacc.py:
+    - src/data/fig_2_dat.h5
+
+  src/scripts/figure3_sanityCheck.py:
+    - src/data/GWTC3_LVK_posterior.npz
+    - src/data/GW150914_roots_raw.hdf5
+    - src/data/GW150914_roots_processed.hdf5
+    - src/data/GWTC3_mcmc_processed.hdf5
+
+  src/scripts/figure4_population.py:
+    - src/data/GWTC3_mcmc_processed.hdf5
+
+datasets:
+  10.5281/zenodo.6624911:
+    destination: src/data
+    contents:
+      GW150914_roots_processed.hdf5:
+      GW150914_roots_raw.hdf5:
+      GWTC3_LVK_posterior.npz:
+      GWTC3_mcmc_processed.hdf5:
+      fig_2_dat.h5:
+ GW150914.
 
@@
 
+\textbf{We follow the critical mass ratio prescription 
+described in \cite{Neijssel2019} for donors which fill their Roche lobes during
+ other phases of stellar evolution.} 
+For systems which begin a common envelope 
+with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
+stars merge since the donor has likely not formed a strong core-envelope boundary 
+\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
+evolution proceed. In particular, we assume that stars lose mass through stellar winds 
+in a metallicity-dependent 
+fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
+during both stable Roche-overflow mass transfer and common envelopes.
+We further assume that the CO formation mechanism follows the delayed 
+prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
+occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
+and that natal kicks due to supernovae are reduced by the amount 
+of mass which falls back onto the proto-CO
+as described in \citet{Fryer2012}.}
+
+\subsection{\textbf{Case-specific configuration}}
+\label{subsec:configuration}
+
 We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
@@
 $q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$. \textbf{We follow the critical mass ratio prescription 
-described in \cite{Neijssel2019} for donors which fill their Roche lobes during
- other phases of stellar evolution.} 
-For systems which begin a common envelope 
-with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
-stars merge since the donor has likely not formed a strong core-envelope boundary 
-\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
-evolution proceed. In particular, we assume that stars lose mass through stellar winds 
-in a metallicity-dependent 
-fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
-during both stable Roche-overflow mass transfer and common envelopes.
-We further assume that the CO formation mechanism follows the delayed 
-prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
-occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
-and that natal kicks due to supernovae are reduced by the amount 
-of mass which falls back onto the proto-CO
-as described in \citet{Fryer2012}.}
+employed for CO natal kicks, $\sigma$. 
+%Prior
 

### 2023-02-10 src/tex/ms.tex
Add KDE sentence

@@
 
-We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
-We use the \texttt{Overall\_posterior} posterior samples for GW150914,
-publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
-The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-The spin of BHs formed in isolated binaries depends strongly on the angular
-momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
-root-finding stage and the MCMC stage.
+We apply the method described in section \textbf{\ref{subsec:stats}} to the
+first GW event GW150914 \citep{GW150914}. We use the \texttt{Overall\_posterior}
+posterior samples for GW150914, publicly available from the Gravitational wave
+Open Science Center \citep{LIGOScientific:2019lzm}. The parameters involved in
+the analysis and the range we allow for the inference are tabulated in table
+\ref{tab:parameters}. The spin of BHs formed in isolated binaries depends
+strongly on the angular momentum transport in massive stars which is still
+highly uncertain (see however \citet{Fuller2019, Bavera2020}), therefore we only
+consider the two component masses in the observable space $\bm{\theta}$ in both
+the root-finding stage and the MCMC stage. \textbf{Instead of the sampling the posterior
+in the two component masses space, we choose to perform the KDE in the chirp
+mass and mass ratio space, which are less degenerate and is less prone to error
+from the KDE.}
 
@@
 employed for CO natal kicks, $\sigma$.
-%Prior
 

### 2023-02-10 src/tex/ms.tex
Add edits

--- 2023-02-10 src/tex/ms.tex
@@
 setting by producing a posterior over progenitor properties and \texttt{COSMIC}
-hyperparameters implied by the observation of
-dependencies:
-  src/scripts/figure1_cornerPlot.py:
-    - src/data/GWTC3_LVK_posterior.npz
-    - src/data/GWTC3_mcmc_processed.hdf5
-
-  src/scripts/figure2_varyAlphaFacc.py:
-    - src/data/fig_2_dat.h5
-
-  src/scripts/figure3_sanityCheck.py:
-    - src/data/GWTC3_LVK_posterior.npz
-    - src/data/GW150914_roots_raw.hdf5
-    - src/data/GW150914_roots_processed.hdf5
-    - src/data/GWTC3_mcmc_processed.hdf5
-
-  src/scripts/figure4_population.py:
-    - src/data/GWTC3_mcmc_processed.hdf5
-
-datasets:
-  10.5281/zenodo.6624911:
-    destination: src/data
-    contents:
-      GW150914_roots_processed.hdf5:
-      GW150914_roots_raw.hdf5:
-      GWTC3_LVK_posterior.npz:
-      GWTC3_mcmc_processed.hdf5:
-      fig_2_dat.h5:
- GW150914.
+hyperparameters implied by the observation of GW150914.
 
@@
 of the posterior density of a GW event.  We then use those points to initialize
-a set of chains in the MCMC process.
+a set of chains in the MCMC process. \textbf{We use \textsc{emcee} \cite{emcee}
+to sample the posterior density in the evolutionary parameter space.}
 
@@
 reprojected posterior agrees with the posterior given by the LVK collaboration.
-
-% Validity paragraph \wf{What if we made this a subsection of ``methods'', and
-%then just one sentence here about the values of the KL divergence for
-%GW150914?} To check the validity of the recovered posterior over evolutionary
-%parameters, it is essential to check whether it is consistent with the original
-%posterior in the observable space. We reproject the recovered posterior to the
-%observable space, and compare the KL divergence between the original posterior
-%and the reprojected posterior.
 
@@
 
-For progenitor parameters $\bm{\theta'}$, we characterize each guess with
-five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
-the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
-obtained through the root-finding process because the formation redshift does not intrinsically affect
-the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
-formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
-time to merger to be less than $13.7\,\rm{Gyr}$.
-Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
-Planck 2018 cosmological model \citep{Planck2018}.
-The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
-of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
-This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
-by reweighting the posterior distribution according to a particular SFR model,
-and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
-since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
+For progenitor parameters $\bm{\theta'}$, we characterize each guess with five
+parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the
+orbital period $t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at
+ZAMS. \textbf{We use a log-uniform prior in sampling $M_{\rm 1,ZAMS}$, $M_{\rm
+2,ZAMS}$, $t_{\rm orb}$, and $Z$, and we use a uniform prior for $e$.
+The range follows the range for the optimization range as given by table
+\ref{tab:parameters}.} Note that the progenitor formation redshift is not
+obtained through the root-finding process because the formation redshift does
+not intrinsically affect the evolution of the binary. Instead, we compute the
+time it takes for the binary to evolve from formation to the merger and add this
+time to the lookback time when a GW event is observed and limit the total time
+to merger to be less than $13.7\,\rm{Gyr}$. Using the total lookback time, we
+can compute the redshift of the binary at ZAMS formation following the Planck
+2018 cosmological model \citep{Planck2018}. The main benefit of post-processing
+the redshift in this way is that we do not rely on a particular assumption of
+star formation rate (SFR) distribution when we solve for the posterior
+distribution in redshift. This means we can put the prior on formation redshift
+(or in general, metallicity-redshfit distribution) in by reweighting the
+posterior distribution according to a particular SFR model, and therefore do not
+have to rerun the inference every time we change the SFR model. This is
+especially helpful since SFR models are still highly uncertain but strongly
+impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
 
 For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
-binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
-efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
-the critical mass ratio which determines whether Roche-overflow mass transfer
-proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
-$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$.
+binary evolution in \texttt{COSMIC} for massive stars, including the common
+envelope efficiency $\alpha$, the fraction of mass accreted onto the accretor
+during stable mass transfer $f_{\rm acc}$, the critical mass ratio which
+determines whether Roche-overflow mass transfer proceeds stably or produces a
+common envelope when the donor star is on the Hertzsprung gap $q_{\rm crit,3}$,
+and the root-mean-square of the Maxwellian distribution employed for CO natal
+kicks, $\sigma$. \textbf{We use a uniform prior for $\alpha$, $f_{\rm acc}$, and
+$q_{\rm crit,3}$, and a log-uniform prior for $\sigma$.}
+
+\textbf{We use the same prior for all the events in this study. Note that our
+main goal is to see whether the events are compatible with \textsc{COSMIC} on an
+individual event basis, we do not consider any population analysis here, and
+therefore we do not account for any selection effect.}
 

### 2023-02-14 src/tex/ms.tex
adding in discussion about redshift, van Son ref, and remove some instances of "we find"

@@
 GW150914-like merger event is strongly correlated with hyperparameter settings,
-consistent with results from prior forward modeling studies cited above;
+\textbf{consistent with results from prior forward modeling studies cited above};
 different assumptions about the black hole masses at merger in GW150914 imply
@@
 is colored based on the comparison between progenitor parameters (blue),
-hyperparameters (orange), or a mix of the two (green). We find that the choice
-for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
+hyperparameters (orange), or a mix of the two (green). \textbf{The choice
+for $\sigma$ does not affect the formation of GW150914-like BBHs.} This is not
 unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
@@
 one another, but not strongly with the other progenitor parameters or
-hyperparameters. We find strong correlations between the ZAMS masses, as well as
+hyperparameters.
+
+\textbf{The formation redshift distribution is slightly correlated with the $q_{\rm crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors generally prefer formation redshifts of $z\sim0.1$ (delay times of $\lesssim0.1\,\rm{Gyr}$).
+This correlation is due to the preference for longer delay times between ZAMS formation and merger for progenitors which undergo stable mass transfer at Roche overflow instead of a common envelope.
+This result is in direct agreement with \citet{vanSon2022a} who also find that the common envelope evolution channel produces BBH mergers with shorter delay times than the stable Roche-overflow channel.}
+
+\textbf{The joint posterior exhibits} strong correlations between the ZAMS masses, as well as
 strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
@@
 
-Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
+Finally, $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
 though there exist independent trends in each hyper parameter. This is not

### 2023-02-14 src/tex/ms.tex

@@
 
-We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
-We use the \texttt{Overall\_posterior} posterior samples for GW150914,
-publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
-The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-The spin of BHs formed in isolated binaries depends strongly on the angular
-momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
-root-finding stage and the MCMC stage.
+We apply the method described in section \textbf{\ref{subsec:stats}} to the
+first GW event GW150914 \citep{GW150914}. We use the \texttt{Overall\_posterior}
+posterior samples for GW150914, publicly available from the Gravitational wave
+Open Science Center \citep{LIGOScientific:2019lzm}. The parameters involved in
+the analysis and the range we allow for the inference are tabulated in table
+\ref{tab:parameters}. The spin of BHs formed in isolated binaries depends
+strongly on the angular momentum transport in massive stars which is still
+highly uncertain (see however \citet{Fuller2019, Bavera2020}), therefore we only
+consider the two component masses in the observable space $\bm{\theta}$ in both
+the root-finding stage and the MCMC stage. \textbf{Instead of the sampling the posterior
+in the two component masses space, we choose to perform the KDE in the chirp
+mass and mass ratio space, which are less degenerate and is less prone to error
+from the KDE.}
 
-For progenitor parameters $\bm{\theta'}$, we characterize each guess with
-five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
-the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
-obtained through the root-finding process because the formation redshift does not intrinsically affect
-the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
-formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
-time to merger to be less than $13.7\,\rm{Gyr}$.
-Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
-Planck 2018 cosmological model \citep{Planck2018}.
-The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
-of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
-This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
-by reweighting the posterior distribution according to a particular SFR model,
-and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
-since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
+For progenitor parameters $\bm{\theta'}$, we characterize each guess with five
+parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the
+orbital period $t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at
+ZAMS. \textbf{We use a log-uniform prior in sampling $M_{\rm 1,ZAMS}$, $M_{\rm
+2,ZAMS}$, $t_{\rm orb}$, and $Z$, and we use a uniform prior for $e$.
+The range follows the range for the optimization range as given by table
+\ref{tab:parameters}.} Note that the progenitor formation redshift is not
+obtained through the root-finding process because the formation redshift does
+not intrinsically affect the evolution of the binary. Instead, we compute the
+time it takes for the binary to evolve from formation to the merger and add this
+time to the lookback time when a GW event is observed and limit the total time
+to merger to be less than $13.7\,\rm{Gyr}$. Using the total lookback time, we
+can compute the redshift of the binary at ZAMS formation following the Planck
+2018 cosmological model \citep{Planck2018}. The main benefit of post-processing
+the redshift in this way is that we do not rely on a particular assumption of
+star formation rate (SFR) distribution when we solve for the posterior
+distribution in redshift. This means we can put the prior on formation redshift
+(or in general, metallicity-redshfit distribution) in by reweighting the
+posterior distribution according to a particular SFR model, and therefore do not
+have to rerun the inference every time we change the SFR model. This is
+especially helpful since SFR models are still highly uncertain but strongly
+impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
 
 For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
-binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
-efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
-the critical mass ratio which determines whether Roche-overflow mass transfer
-proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
-$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$. 
-%Prior
+binary evolution in \texttt{COSMIC} for massive stars, including the common
+envelope efficiency $\alpha$, the fraction of mass accreted onto the accretor
+during stable mass transfer $f_{\rm acc}$, the critical mass ratio which
+determines whether Roche-overflow mass transfer proceeds stably or produces a
+common envelope when the donor star is on the Hertzsprung gap $q_{\rm crit,3}$,
+and the root-mean-square of the Maxwellian distribution employed for CO natal
+kicks, $\sigma$. \textbf{We use a uniform prior for $\alpha$, $f_{\rm acc}$, and
+$q_{\rm crit,3}$, and a log-uniform prior for $\sigma$.}
+
+\textbf{We use the same prior for all the events in this study. Note that our
+main goal is to see whether the events are compatible with \textsc{COSMIC} on an
+individual event basis, we do not consider any population analysis here, and
+therefore we do not account for any selection effect.}
 

### 2023-02-16 src/tex/ms.tex
Add emcee citation and unit paragraph

@@
 eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
-threshold of $10^{-2}$. To make sure we find a reasonably complete set of
-progenitor parameters that corresponds to the posterior sample point, we use
-1000 different initial guesses in solving the optimization problem. As long as
-the solution fulfills the acceptance criteria, the root is as valid as all other
-roots that fulfill the same acceptance criteria, \emph{regardless to its initial
-guess}. This means, we have the freedom to choose the set of initial guesses
-however we think might benefit the optimization.
+threshold of $10^{-2}$ \textbf{solar mass. For a more general study that use
+observables that are not with the same unit, the norm should be chosen in a way
+that all parameters are represented on a comparable scale.} To make sure we find
+a reasonably complete set of progenitor parameters that corresponds to the
+posterior sample point, we use 1000 different initial guesses in solving the
+optimization problem. As long as the solution fulfills the acceptance criteria,
+the root is as valid as all other roots that fulfill the same acceptance
+criteria, \emph{regardless to its initial guess}. This means, we have the
+freedom to choose the set of initial guesses however we think might benefit the
+optimization.
 

### 2023-03-20 src/tex/ms.tex

--- 2023-02-14 src/tex/ms.tex
@@
 population synthesis parameter settings with binary properties, if any;
-\textbf{we illustrate the possibility of such capability by presenting
+we illustrate the possibility of such capability by presenting
 inferences for the recent GWTC-3 transient catalog that suggest that the stable
-mass transfer efficiency parameter may vary with primary black hole mass.}  
+mass transfer efficiency parameter may vary with primary black hole mass. 
 \end{abstract}
@@
 synthetic catalog of DCO mergers which can be compared to parameterized models
-derived from observations. \textbf{This method has been applied extensively to
+derived from observations. This method has been applied extensively to
 inference on GW data to inform phenomenological models
@@
 environments \citep[e.g][]{Zevin2017, Stevenson2017, Bouffanais2019, Zevin2021,
-Wong2021, Bouffanais2021, Mapelli2022}}. These approaches rely on several
+Wong2021, Bouffanais2021, Mapelli2022}. These approaches rely on several
 assumptions, both in the simulations and model parameterizations fitted to the
-GW data, which are unlikely to be fully correct. For example, \textbf{without
+GW data, which are unlikely to be fully correct. For example, without
 knowing whether the parameterization of hyperparameters is sufficiently
-accurate}, there is no reason to suppose that the entire population of DCOs
+accurate, there is no reason to suppose that the entire population of DCOs
 should evolve with the same fixed hyperparameters; there is no a priori reason
@@
 GW150914-like merger event is strongly correlated with hyperparameter settings,
-\textbf{consistent with results from prior forward modeling studies cited above};
+consistent with results from prior forward modeling studies cited above;
 different assumptions about the black hole masses at merger in GW150914 imply
@@
 
-\subsection{\textbf{Statistical method}}
+\subsection{Statistical method}
 \label{subsec:stats}
@@
 samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.
-\textbf{A desired prior in the evolutionary parameters can be included through
-the prior $\pi(\bm{\Theta})$ in the above equation.}
+A desired prior in the evolutionary parameters can be included through
+the prior $\pi(\bm{\Theta})$ in the above equation.
 
@@
 of the posterior density of a GW event.  We then use those points to initialize
-a set of chains in the MCMC process.
+a set of chains in the MCMC process. We use \textsc{emcee} \cite{emcee}
+to sample the posterior density in the evolutionary parameter space.
 
@@
 eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
-threshold of $10^{-2}$. To make sure we find a reasonably complete set of
-progenitor parameters that corresponds to the posterior sample point, we use
-1000 different initial guesses in solving the optimization problem. As long as
-the solution fulfills the acceptance criteria, the root is as valid as all other
-roots that fulfill the same acceptance criteria, \emph{regardless to its initial
-guess}. This means, we have the freedom to choose the set of initial guesses
-however we think might benefit the optimization.
+threshold of $10^{-2}$ solar mass. For a more general study that use
+observables that are not with the same unit, the norm should be chosen in a way
+that all parameters are represented on a comparable scale. To make sure we find
+a reasonably complete set of progenitor parameters that corresponds to the
+posterior sample point, we use 1000 different initial guesses in solving the
+optimization problem. As long as the solution fulfills the acceptance criteria,
+the root is as valid as all other roots that fulfill the same acceptance
+criteria, \emph{regardless to its initial guess}. This means, we have the
+freedom to choose the set of initial guesses however we think might benefit the
+optimization.
 
@@
 
-% Validity paragraph \wf{What if we made this a subsection of ``methods'', and
-%then just one sentence here about the values of the KL divergence for
-%GW150914?} To check the validity of the recovered posterior over evolutionary
-%parameters, it is essential to check whether it is consistent with the original
-%posterior in the observable space. We reproject the recovered posterior to the
-%observable space, and compare the KL divergence between the original posterior
-%and the reprojected posterior.
-
 We use KL divergence to measure the agreement between the two posterior
@@
 
-\subsection{\textbf{Binary evolution assumptions}}
+\subsection{Binary evolution assumptions}
 \label{subsec:binaries}
@@
 
-\textbf{We follow the critical mass ratio prescription 
+We follow the critical mass ratio prescription 
 described in \cite{Neijssel2019} for donors which fill their Roche lobes during
- other phases of stellar evolution.} 
+ other phases of stellar evolution.
 For systems which begin a common envelope 
@@
 stars merge since the donor has likely not formed a strong core-envelope boundary 
-\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
+\citep{Ivanova2004, Belczynski2008}. We fix other assumptions for how stellar and binary 
 evolution proceed. In particular, we assume that stars lose mass through stellar winds 
@@
 of mass which falls back onto the proto-CO
-as described in \citet{Fryer2012}.}
+as described in \citet{Fryer2012}.
 
-\subsection{\textbf{Case-specific configuration}}
+\subsection{Case-specific configuration}
 \label{subsec:configuration}
 
-We apply the method described in section \textbf{\ref{subsec:stats}} to the first GW event GW150914 \citep{GW150914}.
-We use the \texttt{Overall\_posterior} posterior samples for GW150914,
-publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
-The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
-The spin of BHs formed in isolated binaries depends strongly on the angular
-momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
-therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
-root-finding stage and the MCMC stage.
+We apply the method described in section \ref{subsec:stats} to the
+first GW event GW150914 \citep{GW150914}. We use the \texttt{Overall\_posterior}
+posterior samples for GW150914, publicly available from the Gravitational wave
+Open Science Center \citep{LIGOScientific:2019lzm}. The parameters involved in
+the analysis and the range we allow for the inference are tabulated in table
+\ref{tab:parameters}. The spin of BHs formed in isolated binaries depends
+strongly on the angular momentum transport in massive stars which is still
+highly uncertain (see however \citet{Fuller2019, Bavera2020}), therefore we only
+consider the two component masses in the observable space $\bm{\theta}$ in both
+the root-finding stage and the MCMC stage. Instead of the sampling the posterior
+in the two component masses space, we choose to perform the KDE in the chirp
+mass and mass ratio space, which are less degenerate and is less prone to error
+from the KDE.
 
-For progenitor parameters $\bm{\theta'}$, we characterize each guess with
-five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
-the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
-obtained through the root-finding process because the formation redshift does not intrinsically affect
-the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
-formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
-time to merger to be less than $13.7\,\rm{Gyr}$.
-Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
-Planck 2018 cosmological model \citep{Planck2018}.
-The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
-of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
-This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
-by reweighting the posterior distribution according to a particular SFR model,
-and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
-since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
+For progenitor parameters $\bm{\theta'}$, we characterize each guess with five
+parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the
+orbital period $t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at
+ZAMS. We use a log-uniform prior in sampling $M_{\rm 1,ZAMS}$, $M_{\rm
+2,ZAMS}$, $t_{\rm orb}$, and $Z$, and we use a uniform prior for $e$.
+The range follows the range for the optimization range as given by table
+\ref{tab:parameters}. Note that the progenitor formation redshift is not
+obtained through the root-finding process because the formation redshift does
+not intrinsically affect the evolution of the binary. Instead, we compute the
+time it takes for the binary to evolve from formation to the merger and add this
+time to the lookback time when a GW event is observed and limit the total time
+to merger to be less than $13.7\,\rm{Gyr}$. Using the total lookback time, we
+can compute the redshift of the binary at ZAMS formation following the Planck
+2018 cosmological model \citep{Planck2018}. The main benefit of post-processing
+the redshift in this way is that we do not rely on a particular assumption of
+star formation rate (SFR) distribution when we solve for the posterior
+distribution in redshift. This means we can put the prior on formation redshift
+(or in general, metallicity-redshfit distribution) in by reweighting the
+posterior distribution according to a particular SFR model, and therefore do not
+have to rerun the inference every time we change the SFR model. This is
+especially helpful since SFR models are still highly uncertain but strongly
+impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
 
 For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
-binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
-efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
-the critical mass ratio which determines whether Roche-overflow mass transfer
-proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
-$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
-employed for CO natal kicks, $\sigma$. 
-%Prior
+binary evolution in \texttt{COSMIC} for massive stars, including the common
+envelope efficiency $\alpha$, the fraction of mass accreted onto the accretor
+during stable mass transfer $f_{\rm acc}$, the critical mass ratio which
+determines whether Roche-overflow mass transfer proceeds stably or produces a
+common envelope when the donor star is on the Hertzsprung gap $q_{\rm crit,3}$,
+and the root-mean-square of the Maxwellian distribution employed for CO natal
+kicks, $\sigma$. We use a uniform prior for $\alpha$, $f_{\rm acc}$, and
+$q_{\rm crit,3}$, and a log-uniform prior for $\sigma$.
+
+We use the same prior for all the events in this study. Note that our
+main goal is to see whether the events are compatible with \textsc{COSMIC} on an
+individual event basis, we do not consider any population analysis here, and
+therefore we do not account for any selection effect.
 
@@
 is colored based on the comparison between progenitor parameters (blue),
-hyperparameters (orange), or a mix of the two (green). \textbf{The choice
-for $\sigma$ does not affect the formation of GW150914-like BBHs.} This is not
-unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
+hyperparameters (orange), or a mix of the two (green). The choice
+for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
+unexpected since the \citet{Fryer2012} delayed model can significantly reduce the
 natal kick strength at BH formation based on the amount of fallback onto the
@@
 
-\textbf{The formation redshift distribution is slightly correlated with the $q_{\rm crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors generally prefer formation redshifts of $z\sim0.1$ (delay times of $\lesssim0.1\,\rm{Gyr}$).
-This correlation is due to the preference for longer delay times between ZAMS formation and merger for progenitors which undergo stable mass transfer at Roche overflow instead of a common envelope.
-This result is in direct agreement with \citet{vanSon2022a} who also find that the common envelope evolution channel produces BBH mergers with shorter delay times than the stable Roche-overflow channel.}
+The formation redshift distribution is slightly correlated with the $q_{\rm
+crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors
+generally prefer formation redshifts of $z\sim0.1$ (delay times of
+$\lesssim0.1\,\rm{Gyr}$). This correlation is due to the preference for longer
+delay times between ZAMS formation and merger for progenitors which undergo
+stable mass transfer at Roche overflow instead of a common envelope. This result
+is in direct agreement with \citet{vanSon2022a} who also find that the common
+envelope evolution channel produces BBH mergers with shorter delay times than
+the stable Roche-overflow channel. \textbf{This result is also consistent with
+literature that doesn't consider the delay times originating from common
+envelope and stable mass transfer separately, but is dominated by
+short-delay-time mergers originating from common envelope
+\citep[e.g.][]{OShaughnessy2010, Dominik2012, Eldridge2016, Mapelli2017,
+Lamberts2018, Broekgaarden2022}.}
 
-\textbf{The joint posterior exhibits} strong correlations between the ZAMS masses, as well as
+The joint posterior exhibits strong correlations between the ZAMS masses, as well as
 strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
@@
 $150\,M_{\odot}$ which is the limit imposed by our assumptions.
-\textbf{The features in the scatter points of the ZAMS masses, most obvious in the region where $M_1,\rm{ZAMS}\sim 140-150\,M_\odot$, are the outcome of the interplay between the progenitor parameters and hyperparameters.
+The features in the scatter points of the ZAMS masses, most obvious in the region where $M_1,\rm{ZAMS}\sim 140-150\,M_\odot$, are the outcome of the interplay between the progenitor parameters and hyperparameters.
 The dominant cause of regions with fewer data points is that fewer binary progenitors produce BBH mergers with chirp masses and mass ratios consistent with GW150914's priors.
-This is most clearly illustrated for progenitors with primary masses between $142-144\,M_\odot$ and secondary masses below $80\,M_[\odot]$.} We explore the
+This is most clearly illustrated for progenitors with primary masses between $142-144\,M_\odot$ and secondary masses below $80\,M_[\odot]$.
+We explore the
 formation scenarios which lead to successful GW150914-like mergers and those
-which fail to produce GW150914-like mergers \textbf{in more detail} below in
+which fail to produce GW150914-like mergers in more detail below in
 Figure~\ref{fig:compare_fixed_variable}.
@@
 the posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
-most\footnote{\textbf{We excluded events for which our reconstructed posterior has KL
-divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}}
+most\footnote{We excluded events for which our reconstructed posterior has KL
+divergence greater than 0.1 nats relative to the GWTC-3 posterior; see below.}
 of the events in GWTC-3. Each contour represents the $68\%$ credible interval of
@@
 
-\textbf{We have validated the robustness of every events shown in
+We have validated the robustness of every events shown in
 \ref{fig:GWTC-3_f_acc_mass} by reprojecting the posterior density in the
-evolutionary parameter space to the observable space.} Note that some events did
+evolutionary parameter space to the observable space. Note that some events did
 not pass the KL divergence test we proposed in section \ref{sec:method}; we do
@@
 with an application to GW150914. We show the joint posterior of the event's
-progenitor parameters and hyperparameters.  Our work is a \textbf{data-driven}
+progenitor parameters and hyperparameters.  Our work is a data-driven
 way to study GW event progenitors, and also allows the possibility of
@@
 distribution of hyperparameters; this avoids making overly-specific assumptions
-like fixing the hyperparameters for all types of event. \textbf{In this work we
+like fixing the hyperparameters for all types of event. In this work we
 focus on illustrating the possibililty of conducting such analysis with our
@@
 this paper, and we defer a detailed study of the physics related to the
-population of GW events to future work.}
+population of GW events to future work.
 
@@
 pulled the GW event posterior back from the observable space to the evolutionary
-parameter space, \textbf{and have accounted for selection effects}, we can apply
+parameter space, and have accounted for selection effects, we can apply
 the same hierarchical inference methods currently used to infer the compact
@@
 metallicity-specific star formation rate \citep{Vitale2019,Ng2021,VanSon2022}.
-\textbf{In other word, just as we measure the event rate of GW events by
+In other words, just as we measure the event rate of GW events by
 counting, we can measure the event rate of the progenitor systems of GW events
-in the same way.} Comparing the formation rate of progenitors to the star
+in the same way. Comparing the formation rate of progenitors to the star
 formation rate provides an avenue to check our understanding of binary evolution

### 2023-03-20 src/tex/ms.tex
Add changes about table 1

@@
 Open Science Center \citep{LIGOScientific:2019lzm}. The parameters involved in
-the analysis and the range we allow for the inference are tabulated in table
+the analysis and the range we allow for the inference are tabulated in Table
 \ref{tab:parameters}. The spin of BHs formed in isolated binaries depends
@@
 2,ZAMS}$, $t_{\rm orb}$, and $Z$, and we use a uniform prior for $e$.
-The range follows the range for the optimization range as given by table
+The range follows the range for the optimization range as given by Table
 \ref{tab:parameters}. Note that the progenitor formation redshift is not

### 2023-03-20 src/scripts/figure3_sanityCheck.py
disable tex to bypass error

@@ -35,7 +35,7 @@ params = {
     "mathtext.fontset": "dejavuserif",
 }
 
-mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
+# mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
 
 mpl.rcParams.update(params)
 

### 2023-03-20 src/scripts/figure4_population.py

@@ -36,7 +36,7 @@ params = {
 }
 
 
-mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
+# mpl.rcParams["text.latex.preamble"] = [r"\usepackage{amsmath}"]  # for \text command
 
 mpl.rcParams.update(params)
 
